### 2022-10-25 records/manuscript_2023.tex
update notes on neuromodulator I

\documentclass[10pt,a4paper,draft]{article}
%\documentclass[10pt,a4paper]{article}

%\usepackage[top=3cm, bottom=0cm, left=3.5cm,right=2cm]{geometry}
\usepackage[top=1in, left=1in ,right=1in, bottom=1in, footskip=0in, marginparwidth=0in]{geometry}

% use Unicode characters - try changing the option if you run into troubles with special characters (e.g. umlauts)
\usepackage[utf8]{inputenc}

\usepackage{fancyhdr}

% clean citations
%\usepackage{cite}
%\usepackage[super,sort&compress,comma]{natbib}
\usepackage[numbers, round, sort&compress, comma]{natbib}
%\usepackage[round]{natbib}

% hyperref makes references clicky. use \url{www.example.com} or \href{www.example.com}{description} to add a clicky url
%\usepackage{nameref,hyperref}

% math
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{bbold}

% line numbers
\usepackage[right]{lineno}

% improves typesetting in LaTeX
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }
\usepackage{enumitem}

% text layout - change as needed
%\raggedright
%\setlength{\parindent}{0.5cm}
%\textwidth 5.25in 
%\textheight 8.75in

% Remove % for double line spacing
%\usepackage{setspace} 
%\doublespacing

% adjust caption style
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,singlelinecheck=off]{caption}

% remove brackets from references
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% use \textcolor{color}{text} for colored text (e.g. highlight to-do areas)
\usepackage{color}

% define custom colors (this one is for figure captions)
\definecolor{Gray}{gray}{.25}

% this is required to include graphics
\usepackage{graphicx}
\usepackage{sidecap}

% hyperlinks
\usepackage{hyperref}

% ########################################################

%\pagestyle{headings}
\pagestyle{myheadings}
\markright{}

\begin{document}

\thispagestyle{empty}

% title goes here:
\begin{flushleft}
{\Large
\textbf\newline{Title}
}
%\newline
%% authors go here:
%\\
%Loreen Hert\"ag\textsuperscript{1,*},
%Claudia Clopath\textsuperscript{1,§}
%\\
%\bigskip
%\bf{1} Bioengineering Department, Imperial College London, London, UK.
%\\
%\bigskip
%* l.hertag@imperial.ac.uk, § c.clopath@imperial.ac.uk

\end{flushleft}

% now start line numbers
%\linenumbers

\begin{abstract}
blahhh blahhh blah
\end{abstract}

\section*{Introduction}

XXX

\section*{Results}
%
XXX

\subsection*{first result}
%

\subsubsection*{second result}

\section*{Discussion}

We solved the brain.

\section*{Models and methods}
%

\subsection*{Network model}
%
Network consists of two subnetworks. Each subnetwork consists of a PE circuit, a memory neuron and a neuron representing the variance. XXX The memory neuron of subnetwork feefdowradly connects to the PE circuit of the second subnetwork.

\subsubsection*{Prediction-error network model}
%
Consider a mean field network in which each population is represented by one representative neuron. The mean-field PE network consists of an excitatory nPE and pPE neuron, as well as two inhibitory PV neurons (one receiving S, the other P), as well as inhibitory SOM and VIP neurons.

Each excitatory pyramidal cell (that is, nPE or pPE neuron) is divided into two coupled compartments, representing the soma and the dendrites, respectively. The dynamics of the firing rate~$r_{\mathrm{E}}$ of the somatic compartment of obeys \citep{wilson1972excitatory}
%
\begin{align}
\tau_E\ \frac{dr_\mathrm{E}}{dt} &= - r_\mathrm{E} + w_\mathrm{ED}\cdot  r_\mathrm{D}  -  w_\mathrm{EP}\cdot r_\mathrm{P} + I_\mathrm{E},
\end{align}
%
where $\tau_\mathrm{E}$ denotes the excitatory rate time constant ($\tau_\mathrm{E}$=60 ms), the weight $w_{\mathrm{ED}}$ describes the connection strength between the dendritic compartment and the soma of the same neuron, and $w_{\mathrm{EP}}$ denotes the strength of somatic inhibition from PV neurons. The overall input $I_\mathrm{E}$ comprises external background and feedforward sensory  inputs (see ``Inputs" below). Firing rates are rectified to ensure positivity.

The dynamics of the activity~$r_\mathrm{D}$ of the dendritic compartment obeys \citep{wilson1972excitatory}
%
\begin{align}
\tau_E\ \frac{dr_\mathrm{D}}{dt} &= - r_\mathrm{D} +  w_\mathrm{DE}\cdot r_\mathrm{E}  - w_\mathrm{DS}\cdot r_\mathrm{S} + I_\mathrm{D},
\end{align}
%
where the weight $w_{\mathrm{DE}}$ denotes the recurrent excitatory connections between PCs, including backpropagating activity from the soma to the dendrites. $w_{\mathrm{DS}}$ represents the strength of dendritic inhibition from SOM neurons. The overall input $I_\mathrm{D}$ comprises fixed, external background inputs and feedback predictions (see ``Inputs" below). We assume that any excess of inhibition in a dendrite does not affect the soma, that is, the dendritic compartment is rectified at zero.

Just as for the excitatory neurons, the firing rate dynamics of each interneuron is modeled by a rectified, linear differential equation \citep{wilson1972excitatory},
%
\begin{align}
\label{eq:RateEqINs}
\tau_I\ \frac{dr_\mathrm{X}}{dt} =& - r_\mathrm{X} + I_{\mathrm{X}} + w_\mathrm{XE}\cdot r_\mathrm{E} - w_\mathrm{XP}\cdot r_\mathrm{P}  - w_\mathrm{XS}\cdot r_\mathrm{S} -  w_\mathrm{XV}\cdot r_\mathrm{V}, 
\end{align}
%
where $r_\mathrm{X}$ denotes the firing rate of neuron type $X$, and the weight matrices $w_\mathrm{XY}$ denote the strength of connection between the presynaptic neuron population $Y$ and the postsynaptic neuron population $X$ ($X, Y\in \lbrace P,S,V\rbrace$). The rate time constant $\tau_I$ was chosen to resemble a fast GABA\textsubscript{A} time constant, and set to 2 ms for all interneuron types included. The overall input $I_\mathrm{X}$ comprises fixed, external background inputs, as well as feedforward sensory inputs and feedback predictions (see ``Inputs" below).

\subsubsection*{Memory and variance neuron}
%
\begin{align}
\tau_m \cdot \frac{dr_\mathrm{M}}{dt} = w_\mathrm{M\leftarrow pPE} \cdot r_\mathrm{pPE} - w_\mathrm{M\leftarrow nPE} \cdot r_\mathrm{nPE} 
\end{align}
%
\begin{align}
\tau_v \cdot \frac{dr_\mathrm{V}}{dt} = -r_\mathrm{V} + (w_\mathrm{V\leftarrow pPE} \cdot r_\mathrm{pPE} - w_\mathrm{V\leftarrow nPE} \cdot r_\mathrm{nPE})^2 
\end{align}

\subsubsection*{Weighted output}
%
\begin{align}
r_\mathrm{out} = \alpha \cdot S + (1-\alpha) \cdot P
\end{align}
%
\begin{align}
\alpha &= \frac{1/r_\mathrm{V1}}{1/r_\mathrm{V1} + 1/r_\mathrm{V2}}\nonumber\\
& = \left( 1 + \frac{r_\mathrm{V1}}{r_\mathrm{V2}} \right)^{-1}
\end{align}

\subsection*{Connectivity}
%
%All neurons are randomly connected with connection probabilities motivated by the experimental literature \citep[e.g.][]{fino2011dense, packer2011dense, pfeffer2013inhibition, lee2013disinhibitory, pi2013cortical, jiang2015principles, jouhanneau2015vivo, pala2015vivo},
%%
%\begin{align}
%p = \begin{pmatrix}
%  p_\mathrm{EE}  & p_\mathrm{ED} & p_\mathrm{EP} & p_\mathrm{ES} & p_\mathrm{EV} \\
%  p_\mathrm{DE}  & p_\mathrm{DD} & p_\mathrm{DP} & p_\mathrm{DS} & p_\mathrm{DV} \\
%  p_\mathrm{PE}  & p_\mathrm{PD} & p_\mathrm{PP} & p_\mathrm{PS} & p_\mathrm{PV} \\
%  p_\mathrm{SE}  & p_\mathrm{SD} & p_\mathrm{SP} & p_\mathrm{SS} & p_\mathrm{SV} \\
%   p_\mathrm{VE} & p_\mathrm{VD} & p_\mathrm{VP} & p_\mathrm{VS} & p_\mathrm{VV} \\
%\end{pmatrix}
%=
%\begin{pmatrix}
%  - & 1 & 0.6 & - & -\\
%  0.1  & - & - & 0.55 & - \\
%  0.45  & - & 0.5 & 0.6 & 0.5 \\
%  0.35  & - & - & - & 0.5 \\
%   0.1 & - & - & 0.45 & - \\
%\end{pmatrix}.
%\label{Mtx:ConnProb}
%\end{align}
%%
%All cells of the same neuron type have the same number of incoming connections. The mean total connection strengths are set to
%%
%\begin{align}
%w = \begin{pmatrix}
%  w_\mathrm{EE}  & w_\mathrm{ED} & w_\mathrm{EP} & w_\mathrm{ES} & w_\mathrm{EV} \\
%  w_\mathrm{DE}  & w_\mathrm{DD} & w_\mathrm{DP} & w_\mathrm{DS} & w_\mathrm{DV} \\
%  w_\mathrm{PE}  & w_\mathrm{PD} & w_\mathrm{PP} & w_\mathrm{PS} & w_\mathrm{PV} \\
%  w_\mathrm{SE}  & w_\mathrm{SD} & w_\mathrm{SP} & w_\mathrm{SS} & w_\mathrm{SV} \\
%   w_\mathrm{VE} & w_\mathrm{VD} & w_\mathrm{VP} & w_\mathrm{VS} & w_\mathrm{VV} \\
%\end{pmatrix}
%=
%\begin{pmatrix}
%  - & 1 & 2^{*} & - & -\\
%  0.5 & - & - & 0.5^{*} & -\\
%  1.2 & - & 1 & 0.3^{*} & 0.3^{*} \\
%  1 & - & - & - & 0.6 \\
%  1 & - & - & 0.7 & - \\
%\end{pmatrix},
%\label{Mtx:ConnStrengths}
%\end{align}
%%
%where $^{*}$ denotes either the weights that are free for optimization to satisfy the equations describing an E/I balance (see Supporting Information), or the initial mean connection strengths that are subject to synaptic plasticity during learning. For plastic networks, the initial connections between neurons are drawn from uniform distributions 
%%
%\begin{align*}
%w_{ij}^{initial} \in \mathcal{U} \left(0.5\ w, 1.5\ w\right),
%\end{align*}
%%
%where $w$ denotes the mean connection strengths given in (\ref{Mtx:ConnStrengths}). Please note that the system is robust to the choice of connection strengths. The connection strengths are merely chosen such that the solutions of the equations describing an E/I balance comply with Dale's principle.
%
%In plastic networks, $w_\mathrm{EP}$ is subdivided into assemblies. While one-third of PCs receive stronger initial inhibition from PV neurons that are driven by sensory input, another third receives stronger initial inhibition from PV neurons that are driven by feedback predictions. More precisely, for two-thirds of the excitatory neurons, half of the connections from PV neurons are strengthened by $1.5$, while the remaining ones are weakened by $0.5$.
%
%All weights are scaled in proportion to the number of existing connections (i.e., the product of the number of presynaptic neurons and the connection probability), so that the results are independent of the population size.

\subsection*{Inputs}
%
%All neurons receive external background input that ensures reasonable baseline firing rates in the absence of sensory inputs and predictions thereof. In the case of non-plastic networks, these inputs were set such that the baseline firing rates are $r_\mathrm{E}=1\, s^{-1}$, $r_\mathrm{P} = r_\mathrm{S}=r_\mathrm{V}=4\, s^{-1}$ and $r_\mathrm{D}=0\, s^{-1}$. In the case of plastic networks, we set the external inputs of all neuron types to $x_\mathrm{E}=x_\mathrm{P}=x_\mathrm{S}=x_\mathrm{V}=5\, s^{-1}$, while the background input to the dendrites is dynamically computed during training such that $r_\mathrm{D}=0\, s^{-1}$.
%
%In addition to the external background inputs, the neurons receive either sensory input~($S$), a prediction thereof ($P$), or both. We distinguish between phases of fully predicted ($P=S>0$), overpredicted ($P>S$) and underpredicted ($P<S$) sensory stimuli, as well as baseline phases ($P=S=0$). During training, the network is exposed to baseline phases and fully predicted sensory inputs (Figs.~\ref{fig:Fig_Plasticity} and \ref{fig:Fig_Experience}), or in addition to over- and underpredicted sensory stimuli (Fig.~\ref{figsupp:Fig_Experience_Predictability}). Stimuli are drawn from a uniform distribution from the interval $[0, 5\, s^{-1}]$. Mean and SD of test stimuli for each simulation are listed below (see ``Simulations").

\subsection*{Simulations}
%
%All simulations were performed in customized Python code written by LH. Differential equations were numerically integrated using a 2\textsuperscript{nd}-order Runge-Kutta method with time steps ranging between $0.1$ and $0.2$ ms. Neurons in the PE circuits were initialized with $r=0/s$. The memory neurons were initialized at the mean of the two distributions (see above), and each prediction neuron was either set to the mean of the distribution it is associated with if the stimulus at $t=0\ \mathrm{ms}$ was drawn from that distribution, or set to zero otherwise. Each stimulus was presented for $1$ second. During training of the PE circuit, we presented 350 stimuli alternated with 350 zero-input (baseline) phases. We made sure that the weights converged to a configuration that satisfied the objective given by our homeostatic plasticity rules (see Eqs.~\ref{eq:Plasticity_I}-\ref{eq:Plasticity_II} in ``Plasticity model"). We defined the steady-state firing rate per stimulus as the activity in the last 500 ms of stimulus presentation. The onset firing rate was computed as the activity of the first 10 ms.\\\\
%%
%\textbf{Figures 1 \& S2:} Test stimulus was set to $5/s$ with a SD of $1/s$. Stimulus to compute total excitatory and inhibitory inputs was set to $1/s$.\\
%%
%\textbf{Figures 2 \& S3:} Test stimulus was set to $3/s$ with a SD of $1/s$. The perturbation stimuli ranged from $-1.5/s$ to $1.5/s$.\\
%%
%\textbf{Figures 3, S4 \& S5:} Test stimulus was set to $5/s$ with a SD of $1.5/s$. 50\% of the PV neurons, 70\% of the SOM neurons and 30\% of the VIP neurons receive the actual sensory input, while the remaining ones of each population received a prediction thereof. Perturbation stimulus was $\pm 2/s$. Panels D \& G of Fig. 3 show the median over all PE neurons and the SEM.\\
%%
%\textbf{Figures 4 \& S6:} Test stimulus was set to $5/s$ with a SD of $1.5/s$. In main figure, square: $w_\mathrm{EP}\in[2,4]$, $w_\mathrm{PS}\in[0.5,1]$, $w_\mathrm{PV}\in[1.5,2.5]$; circle: $w_\mathrm{EP}\in[2.5,8]$, $w_\mathrm{PS}\in[1.5,2.5]$, $w_\mathrm{PV}\in[0.5,1]$; triangle: $w_\mathrm{EP}\in[2.5,8]$, $w_\mathrm{PS}\in[1,2.5]$, $w_\mathrm{PV}\in[0.5,2]$. Half of the PV neurons and all SOM neurons receive the actual sensory input, while the remaining PV and SOM neurons as well as all VIP neurons receive a prediction thereof.
%In supporting figure: Total number of stimuli presented during training was increased, so that the number of fully predicted sensory stimuli was constant at 350. Results were averaged over 5 simulations, mean and SD are shown.\\
%%
%\textbf{Figure 5:} For panel E, the performance error was computed as the squared difference between the activity of the respective line attractor and the stimulus presented. For panel F, the initial weight between the stimulus and the representation neuron was set to $0.5$. The basis learning rate (fixed) was set to $5e^{-4}$. And the initial speed was computed as the derivative of the rate with respect to time, averaged over the first 50 ms. 
%
%Source code to reproduce the simulations, analyses and figures will be available after publication at \url{github.com/lhertaeg/SourceCode_Hertaeg2021}.

\section*{Acknowledgments}
%
%We are grateful to Vezha Boboeva, Douglas Feitosa Tom\'e, J\'ulia Gallinaro and Klara Kaleb for helpful comments on earlier versions of this manuscript, and we want to thank all members of the Clopath lab for discussion and support. This work was supported by BBSRC BB/N013956/1, BB/N019008/1, Wellcome Trust 200790/Z/16/Z, Simons Foundation 564408 and EPSRC EP/R035806/1.

\bibliographystyle{naturemag} %{plainnat}
\bibliography{../References_Mismatch}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% APPENDICES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section*{Supplementary Information}
\appendix

\subsection*{Analysis of simplified network model, effect of time constants}

simplified model: dynamics and steady state of rM and rV, rM and rV as a function of time constants and trial duration etc., weighting, then use those expressions to discuss when weighting goes awry and how long transitions take from one state to another ...

\subsection*{Analysis of requirements, perturbing PE neurons}
%
symmetry of nPE and pPE necessary for memory neuron to represent mean .. shown for uniform, normal, gamma distribution

symmetry and effect = 1 necessary for variance neuron ... shown for uniform distribution, gamma distribution

effect of population of nPE and pPE neurons on mean and variance

effect of BL activity on mean and variance

effect of perturbations of nPE and pPE neurons on mean and variance

\subsection*{Comparison to Kalman filter and Bayes Factor surprise}

Comparison to Kalman filter
Comparison to Bayes Factor surprise

\end{document}

### 2023-06-16 records/manuscript_2023.tex
IN activity for different variabilities

@@
 XXX
+with a special focus on the interneurons and how and hwy those may adapt yhe weighting XXX
 
@@
 
-\subsection*{first result}
+\subsection*{nPE and pPE neurons as the basis for computing mean and variance of sensory stimuli}
+%
+We hypothesis that the unique properties of nPE and pPE neurons put them in a perfect position to support both the computation of the mean and the variance of feedforward sensory stimuli.
+XXX Why is that XXX.
+
+To this end we simulated a rate-based mean-field network model the core of which represents a PE circuit with nPE and pPE neurons and three types of inhibitory interneurons (XXX).
+In addition, the PE circuits connect/project to a memory neuron that is modeled as perfect integrator.
+In accordance with XXX, we assume that the pPE neurons excite the memory neuron while the nPE neuron inhibit the memory neuron (for instance through lateral inhibition not modeled explicitly here).
+With this connectivity/setup/architecture, the memory neuron holds/encodes the mean of the feedforward sensory stimuli that drive the PE circuit.
+The memory neuron, on the other hand, connects to the apical dendrites of the PE neurons and some of the inhibitory interneurons (see methods for more details).
+It therefore serves as a prediction that is dynamically updated when new sensory information is available.
+We furthermore assume that the PE neurons excite another neuron, modeled as a leaky integrator, whose activity may represent that variance of the feedforward stimuli when both nPE and pPE neurons have a net excitatory effect on it.
+
+To show that such a network can indeed represent mean and variance in the respective neurons, we stimulate it with sequence of step-wise constant inputs drawn from xxx.
+As shown in Fig.
+XXX, the memory neuron's activity gradually approaches the mean of the sensory stimuli, while the v neuron approaches the variance of the inputs.
+This is true for a wide range of inputs statistics (Fig.
+XXX).
+Deviations from the true mean occur mainly for larger input variances.
+The estimated variance is fairly independent of the input statistics tested.
+Moreover, the precise input distribution does not have a significant effect on the network's ability to estimate the mean and the variance (Supp Fig.
+XXX).
+
+XXX interneuron activity increases with mean but also with variance, some faster than the other ... XXX
+
+XXX validated that also correct for population network (beyond mean-field network) XXX
+
+XXX assumptions (BL, gain-nPE = gain-pPE = 1) and how important really (see above)
+
+\subsubsection*{Weighting external and internal signals requires two sets of PE neurons}
+%
+Coming back to the question at hand/example ...
+Following up the same ideas, the weighting of internal and external signals requires a higher PE circuit that integrates over the prediction of the lower PE circuit. 
+XXX explain set-up XXX ...
+XXX stimulation protocol XXX
+XXX two limit cases XXX
+XXX validation for a wide range of input statistics XXX
+XXX mentioning the dynamic nature of this estimation (Supp material), if you find differences in time scales of adaptation, report here XXX
+XXX predictions: nPE/pPE BL increase, cognitive load + trial duration XXX
+XXX The network does not require on the squared activation function but it is important that updating is faster in lower than in higher PE circuit XXX
+
+XXX Do we actually propose that predictions are sent up the hierarchy?
+
+\subsubsection*{XXX IN influence, neuromodulators ... mechanisms XXX}
 %
 
-\subsubsection*{second result}
+\subsubsection*{XXX contraction bias}
 
@@
 
+\subsection*{Sensory weight and contraction bias}
+%
+If P is rather constant, the slope in the contraction bias is exactly the sensory weight
+%
+ \begin{align*}
+ r_\mathrm{out} = \alpha_\mathrm{S} \cdot S + \left( 1 -\alpha_\mathrm{S} \right) \cdot P \equiv m \cdot S + n
+ \end{align*}
+%
+However, P is usually/normally a function of $S$. For simplicity, let's assume that P decays exponentially to a new value of $S$:
+%
+\begin{align*}
+P = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau} + f(S) \cdot \left( 1 -   \mathrm{e}^{-t/\tau} \right)
+\end{align*}
+%
+Within each trial with trial duration T, P can be expressed by n sections of length t in which the stimulus is constant and, for the sake of simplicity, drawn from a uniform distribution $U\left( s - \frac{\sigma_\mathrm{S}}{12}, s + \frac{\sigma_\mathrm{S}}{12} \right)$.
+$P_\mathrm{0}$ is drawn from $U\left( \mu - \frac{\sigma_\mathrm{P}}{12}, \mu + \frac{\sigma_\mathrm{P}}{12} \right)$.
+$P_\mathrm{n}$ is then given by
+%
+\begin{align*}
+P_\mathrm{n} = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau}  + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} s_i \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau}
+\end{align*}
+%
+This needs to be averaged over all possible states
+%
+\begin{align*}
+P_\mathrm{n} = \mathrm{e}^{-t/\tau}  \int\limits_{\mu - \frac{\sigma_\mathrm{P}}{12}}^{\mu + \frac{\sigma_\mathrm{P}}{12}} P_\mathrm{0} \ f(P_\mathrm{0})\ dP_0+ \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau} \int\limits_{s - \frac{\sigma_\mathrm{S}}{12}}^{s + \frac{\sigma_\mathrm{S}}{12}} s\ f(s)\ ds
+\end{align*}
+%
+This gives
+%
+\begin{align*}
+P_\mathrm{n} = \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot t/ \tau} \cdot S
+\end{align*}
+%
+By making use of the geometric series, this simplifies to
+%
+\begin{align*}
+P_\mathrm{n} =  \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right) \cdot S
+\end{align*}
+%
+Together, this yields for the weighted output
+%
+\begin{align*}
+ r_\mathrm{out} = \left[ \alpha_\mathrm{S}\  \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right)\right] \cdot S + \left( 1 -\alpha_\mathrm{S} \right)\ \mathrm{e}^{-T/\tau}\ \mu
+\end{align*}
+%
+Hence, the slope is a function of both the sensory weight and the trial duration.
+
+In a prediction-driven input regime ($alpha_\mathrm{S} \sim 0)$, the slope is independent of the sensory weight and only determined by the trial duration, $m \sim \left( 1 -   \mathrm{e}^{-T/\tau} \right)$.
+In a sensory-driven input regime ($alpha_\mathrm{S} \sim 0)$, the contraction bias vanishes ($m \sim 1$).
+
+If the trail duration is short ($T \rightarrow 0$), the slope is given by the sensory weight.
+If the trail duration approaches infinity, the slope would be 1 again (however, this seems rather unrealistic, this would only be true in an ideal system without memory decay or reproduction and accumulation noise ...).
+
+\subsection*{Effect of nPE and pPE gain}
+%
+In the steady state, the averaged effect of nPE and pPE must be equal (so the gain must be equal):
+%
+\begin{align*}
+g_{pPE}\ \langle \mathrm{nPE}\rangle &= g_{nPE}\ \langle \mathrm{pPE}\rangle \\
+g_{pPE} \langle \left[ S-P\right]_+\rangle &= g_{nPE} \langle \left[ P-S\right]_+\rangle \\
+g_{pPE} \int\limits_P^b \left( x-P\right)\ f(x)\ dx &= g_{nPE} \int\limits_a^P \left( P-x\right)\ f(x)\ dx
+\end{align*}
+%
+Example, uniform distribution:
+\begin{align*}
+g_{pPE}\ \left[ \frac{1}{2} \left(b^2 - P^2\right) - P\left(b - P\right)\right] = g_{nPE}\ \left[  P\left(P - a\right) - \frac{1}{2} \left(P^2 - a^2\right)\right] 
+\end{align*}
+%
+which gives
+\begin{align*}
+P = \frac{g_{pPE}\ b - g_{nPE}\ a \pm \sqrt{g_{nPE}\ g_{pPE}} (a-b)}{g_{pPE} - g_{nPE}}
+\end{align*}
+%
+Only if $g_\mathrm{nPE}=g_\mathrm{pPE}$, the prediction $P$ is given by $\frac{a+b}{2}$.
+
+For variance, we need $g_{nPE}=g_{pPE}=1$. In case of a uniform distribution, we get
+%
+\begin{align*}
+V &= \langle (S-P)^2\rangle \\
+   &= g_\mathrm{pPE}\ \langle \langle[ S-P \rangle]_+ \rangle + g_\mathrm{nPE}\ \langle \langle[ P-S \rangle]_+ \\
+   &= \frac{g_\mathrm{pPE}}{b-a} \int\limits_P^b (u-P)^2\ du + \frac{g_\mathrm{nPE}}{b-a} \int\limits_a^P (P-u)^2\ du \\
+   &= \frac{g_\mathrm{pPE}}{3} \cdot \frac{(b-P)^3}{b-a} + \frac{g_\mathrm{nPE}}{3} \cdot \frac{(P-a)^3}{b-a} 
+\end{align*}
+%
+if $P = (b+a)/2$
+%
+\begin{align*}
+V = \frac{g_\mathrm{pPE} + g_\mathrm{nPE}}{24} \cdot (b-a)^2
+\end{align*}
+%
+Only if $g_\mathrm{nPE}=g_\mathrm{pPE}=1$, the V neuron encodes the variance, that for a uniform distribution is given by
+%
+\begin{align*}
+V = \frac{ (b-a)^2}{12}
+\end{align*}
+%
+XXXXXXX add the version where P is given by the g version ...
+%
+Of course, P is also a function of the gains (see above). If we take this into account, V is given by
+%
+\begin{align*}
+V = \frac{(b-a)^2}{3\ (g_\mathrm{pPE} - g_\mathrm{nPE})^3} \cdot \left[ g_\mathrm{nPE} \cdot( g_\mathrm{pPE} \mp \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}})^3 - g_\mathrm{pPE} \cdot (g_\mathrm{nPE} \mp \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}})^3\right]
+\end{align*}
+
+\subsection*{Effect of BL activity on mean an variance}
+%
+\begin{align*}
+\langle pPE \rangle &= \langle nPE \rangle \\
+\langle \left[ S - P\right]_+ + p_0\rangle &= \langle \left[ P - S\right]_+ + n_0\rangle \\
+\int\limits_P^b (x - P)\ f(x)\ dx + p_0 \underbrace{\int\limits_a^b f(x)\ dx}_{=1}  &= \int\limits_a^P (P - x)\ f(x)\ dx + n_0 \underbrace{\int\limits_a^b f(x)\ dx}_{=1} 
+\end{align*}
+%
+In case of a uniform distribution, that is $f(x) = \frac{1}{b-a}$ for $x\in [a,b]$, 0 otherwise, this leads to
+%
+\begin{align*}
+P = \frac{b+a}{2} + \frac{p_0 + n_0}{b-a}
+\end{align*}
+%
+
+Variance as function of bias in mean:
+\begin{align*}
+V &= \frac{1}{n} \sum_i \left( x - \left(\mu \pm \delta\mu\right)\right)^2 \\
+&= \frac{1}{n} \sum_i \lbrace  \left( x_i - \mu \right)^2 + \delta\mu^2 \mp 2\delta\mu (x_i - \mu)\rbrace \\
+&= V_\mathrm{uniform} + \delta\mu^2 \mp 2\ \delta\mu \left( \frac{1}{n} \sum_i x_i- \mu\right) \\
+&= V_\mathrm{uniform} + \delta\mu^2
+\end{align*}
+%
+
+Variance as a function of BL in nPE and pPE:
+%
+\begin{align*}
+V &= \langle \left( \mathrm{pPE} + \mathrm{nPE} \right)^2 \rangle \\
+&= \langle \left[ S-P\right]_+\rangle + \langle \left[ P-S\right]_+\rangle + (p_0 + n_0)^2 + 2\ (p_0 + n_0)\ \left( \langle \left[ S-P\right]_+\rangle + \langle \left[ P-S\right]_+ \rangle\right)
+\end{align*}
+%
+In case of a uniform distribution, this can be expressed by
+%
+\begin{align*}
+V = \frac{1}{3\ (b-a)} \left[ (b-P)^3 + (P-a)^3\right] + (p_0 + n_0)^2 + \frac{(p_0 + n_0)}{b-a} \left[ (b-P)^2 + (a-P)^2\right]
+\end{align*}
+%
+Since $P = \frac{b+a}{2} + \frac{p_0 + n_0}{b-a}$, this yields
+%
+\begin{align*}
+V =& \frac{1}{3\ (b-a)} \left[ \left( \frac{b-a}{2} - \frac{p_0 - n_0}{b-a}\right)^3 + \left( \frac{b-a}{2} + \frac{p_0-n_0}{b-a}\right)^3\right] + (p_0 + n_0)^2 \\
+&+ \frac{(p_0 + n_0)}{b-a} \left[ \left( \frac{b-a}{2} + \frac{p_0-n_0}{b-a}\right)^2 + \left( \frac{b-a}{2} - \frac{p_0 - n_0}{b-a}\right)^2\right]
+\end{align*}
+%
+Simplifying this expression, leads to
+%
+\begin{align*}
+V =  \frac{(b-a)^2}{12} + \frac{(p_0-n_0)^2}{(b-a)^2} \left( 1 + 2\ \frac{p_0+n_0}{b-a}\right) + (p_0 + n_0) \left( p_0 + n_0 + \frac{b-a}{2}\right)
+\end{align*}
+%
+
+\subsection*{Influence of a population of nPE and pPE neurons}
+XXX
+
 \subsection*{Analysis of simplified network model, effect of time constants}
@@
 
-\subsection*{Analysis of requirements, perturbing PE neurons}
+\subsection*{Comparison to Kalman filter and Bayes Factor surprise}
 %
-symmetry of nPE and pPE necessary for memory neuron to represent mean .. shown for uniform, normal, gamma distribution
+Kalman filter. Initialisation
+%
+\begin{align*}
+x_{0|init} &= 0 \\
+P_{0|init} &= \sigma^2\ I
+\end{align*}
+%
+with x being the system state (in my terms the prediction), P is the covariance matrix of the errors of x (in my terms the var of the predictions) and I is the identity matrix.
+%
+Then the "correction" is given by
+%
+\begin{align*}
+K_k &= P_{k|k-1}\ H_k^T\ \left( H_k\ P_{k|k-1}\ H_k^T + R_k \right)^{-1} \\
+x_k &= x_{k|k-1} + K_k\ \left( z_k - H_k\ x_{k|k-1}\right) \\
+P_k &= \left( I - K_k\ H_k\right)\ P_{k|k-1}
+\end{align*}
+%
+with K the kalman gain matrix, H the observation matrix ($z_k = H_k\ x_k + noise$), R the covaraince of the measurement noise and z a new observation.
+The last part of the Kalman filter is the "prediction":
+%
+\begin{align*}
+x_{k|k-1} &= F_{k-1}\ x_{k-1} + B_{k-1}\ u_{k-1} \\
+P_{k|k-1} &= F_{k-1}\ P_{k-1}\ F_{k-1}^T + Q_{k-1}
+\end{align*}
+%
+with F the transition matrix ($x_{k|k-1} = F_{k-1}\ x_{k-1}$, u a deterministic perturbation, B the dynamics of the deterministic perturbation. In our terms
+%
+\begin{align*}
+\alpha = K_k = \frac{P_{k|k-1}}{R_k + P_{k|k-1}}
+\end{align*}
+%
+$P_{k|k-1}$, is however $\sigma_P^2$ in my implementation and $R_k$ is fixed variance of inputs $\sigma_S^2$.
+Hence, my implementation represents (?) the Kalman filter.
+Important to note is, that in my implementation we estimate the variance of inputs dynamically, so it is not set!
+Another nice advantage here is that I don't need a good estimate for P.
+I can basically initiate it as I want.
+Another difference is that I consider the optimal weighting in my "output neuron" and not the prediction itself ...
+.
 
-symmetry and effect = 1 necessary for variance neuron ... shown for uniform distribution, gamma distribution
-
-effect of population of nPE and pPE neurons on mean and variance
-
-effect of BL activity on mean and variance
-
-effect of perturbations of nPE and pPE neurons on mean and variance
-
-\subsection*{Comparison to Kalman filter and Bayes Factor surprise}
-
-Comparison to Kalman filter
-Comparison to Bayes Factor surprise
+XXX Comparison to Bayes Factor surprise
 

### 2023-06-16 scripts/run_fig_3_support_1.py

@@ -22,9 +22,9 @@ dtype = np.float32
 
 # %% Simulate control case for number of input statistics 
 
-run_cell = False
+run_cell = True
 load_only = True
-do_plot = False
+do_plot = True
 
 # define parameters used throughout the file
 mean_trials, m_sd = dtype(5), dtype(0)
@@ -137,13 +137,13 @@ if run_cell:
             
     
     ### plot data
-    plot_impact_para(weight_ctrl, weights_mod_base, para_range_tested=add_inputs,
+    plot_impact_para(weight_ctrl, weights_mod_base, para_range_tested=add_inputs, ms=3,
                      colorbar_title = 'baseline', colorbar_tick_labels = ['low','high'],fs=7) 
     
     
 # %% Impact of baseline activity (only PE neurons)
     
-run_cell = False
+run_cell = True
 plot_only = True
 
 if run_cell:
@@ -166,6 +166,60 @@ if run_cell:
     if not plot_only: # simulate respective network
     
         print('Increase baseline activity for PE neurons only:\n')
+    
+        for k, add_input in enumerate(add_inputs):
+        
+            print('Add input:', add_input)
+            [_, _, _, _, _, _, weight] = simulate_impact_para(mfn_flag, variability_within, variability_across, mean_trials, 
+                                                              m_sd, last_n = last_n, n_trials = n_trials, 
+                                                              add_input=add_input, id_cells_modulated = id_cells_modulated)
+            
+            weights_mod_base[:,k] = weight
+            
+        with open(file_for_data,'wb') as f:
+            pickle.dump([add_inputs, weights_mod_base],f) 
+        
+    else:
+        
+        with open(file_for_data,'rb') as f:
+            [add_inputs, weights_mod_base] = pickle.load(f)
+            
+    
+    ### plot data
+    plot_impact_para(weight_ctrl, weights_mod_base, para_range_tested=add_inputs,
+                     colorbar_title = 'baseline', colorbar_tick_labels = ['low','high'],
+                     fs=7, loc_position=5) 
+    
+
+# %% Impact of baseline activity of nPE OR pPE neuron
+    
+run_cell = True
+plot_only = True
+
+if run_cell:
+
+    ### choose mean-field network to simulate
+    mfn_flag = '10' # valid options are '10', '01', '11
+    
+    ### filename for data ... change PE neuron type to be tested here
[…]

### 2023-06-16 scripts/run_fig_4.py

@@ -16,7 +16,7 @@ import pandas as pd
 
 from src.functions_simulate import simulate_neuromod, simulate_moment_estimation_upon_changes_PE, simulate_neuromod_effect_on_neuron_properties
 from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs, plot_neuromod_per_net, plot_points_of_interest_neuromod, plot_bar_neuromod
-from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_changes_upon_input2PE_neurons
+from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_changes_upon_input2PE_neurons, plot_bar_neuromod_stacked
 
 import warnings
 warnings.filterwarnings("ignore")
@@ -32,7 +32,7 @@ plot_only = True
 if run_cell:
     
     ### choose mean-field network to simulate
-    mfn_flag = '01' # valid options are '10', '01', '11
+    mfn_flag = '10' # valid options are '10', '01', '11
     
     ### define target area and stimulus statistics
     column = 0      # 0: both, 1: lower level PE circuit, 2: higher level PE circuit
@@ -99,16 +99,17 @@ if run_cell:
     
 # %% Plot neuromodulation results in bar plots for VIP-PV and VIP-SOM
 
-run_cell = False
+run_cell = True
 mfn_flag = '10'
 
 if run_cell:
     
-    f, ax = plt.subplots(2, 3, figsize=(12,6))
+    f, ax = plt.subplots(1, 3, figsize=(15,3)) #  plt.subplots(2, 3, figsize=(12,6))
 
-    for flag_what in range(2): # 0: VIP-PV, 1: VIP-SOM
-        for column in range(3):
-            plot_bar_neuromod(column, mfn_flag, flag_what, dgrad = 0.001, axs=ax[flag_what,column])
+    #for flag_what in range(2): # 0: VIP-PV, 1: VIP-SOM
+    for column in range(3):
+        #plot_bar_neuromod(column, mfn_flag, flag_what, dgrad = 0.001, axs=ax[flag_what,column])
+        plot_bar_neuromod_stacked(column, mfn_flag, dgrad = 0.001, axs=ax[column])
     
 
 # %% XXX
@@ -122,7 +123,7 @@ if run_cell:
 run_cell = False
 
 from matplotlib.colors import LinearSegmentedColormap, ListedColormap
[…]

### 2023-06-16 scripts/run_fig_5.py

@@ -17,14 +17,9 @@ import matplotlib.pyplot as plt
 
 dtype = np.float32
 
-# %% Notes 
-
-# If I get interesting results for contraction bias, I should show that figure before neuromod fig 
-# and actually discuss what a neuromod does to the bias!
-
 # %% Scalar variability ....
 
-run_cell = True
+run_cell = False
 plot_only = True
 
 if run_cell:
@@ -293,7 +288,7 @@ if run_cell:
 # slope independent of trial variance
 # mechanism: transient effects (steady state not reached ..., so basically history effect, it depends on previous stimulus)
 
-run_cell = False
+run_cell = True
 plot_only = True
 
 if run_cell:

### 2023-06-16 src/functions_networks.py

@@ -58,8 +58,8 @@ def rate_dynamics_mfn(tau_E, tau_I, tc_var, w_var, U, V, W, rates, mean,
     return [rates, mean[0], var] # [rates, mean, var]
 
 
-def run_mfn_circuit(w_PE_to_P, w_P_to_PE, w_PE_to_PE, tc_var_per_stim, tau_pe, fixed_input, 
-                        stimuli, VS = 1, VV = 0, dt = dtype(1), w_PE_to_V = dtype([1,1]), n=2):
+def run_mfn_circuit(w_PE_to_P, w_P_to_PE, w_PE_to_PE, tc_var_per_stim, tau_pe, fixed_input, stimuli, VS = 1, VV = 0,
+                    dt = dtype(1), w_PE_to_V = dtype([1,1]), n=2, record_interneuron_activity = False):
     
     ### neuron and network parameters
     tau_E, tau_I  = tau_pe
@@ -86,14 +86,19 @@ def run_mfn_circuit(w_PE_to_P, w_P_to_PE, w_PE_to_PE, tc_var_per_stim, tau_pe, f
                                                                    rates_lower[id_stim-1,:], m_neuron[id_stim-1], 
                                                                    v_neuron[id_stim-1], feedforward_input, dt, n)
 
-        
-    return m_neuron, v_neuron, rates_lower[:,:2]
+     
+    ret = (m_neuron, v_neuron, rates_lower[:,:2],)
+    
+    if record_interneuron_activity:
+        ret += (rates_lower[:,4:], )
+    
+    return ret                                                               
 
 
 def run_mfn_circuit_coupled(w_PE_to_P, w_P_to_PE, w_PE_to_PE, v_PE_to_P, v_P_to_PE, v_PE_to_PE, tc_var_per_stim, 
                             tc_var_pred, tau_pe, fixed_input, stimuli, VS = 1, VV = 0, dt = dtype(1), n = 2, 
                             w_PE_to_V = dtype([1,1]), v_PE_to_V  = dtype([1,1]), record_pe_activity = False,
-                            fixed_input_lower = None, fixed_input_higher = None):
+                            fixed_input_lower = None, fixed_input_higher = None, record_interneuron_activity = False):
     
     ### neuron and network parameters
     tau_E, tau_I  = tau_pe
[…]

### 2023-06-16 src/plot_data.py

@@ -48,8 +48,8 @@ color_weighted_output = '#5E0035'
 color_mean_prediction = '#70A9A1' 
 
 cmap_sensory_prediction = LinearSegmentedColormap.from_list(name='cmap_sensory_prediction', 
-                                                            colors=[color_m_neuron,'#fefee3',
-                                                                    color_running_average_stimuli])
+                                                            colors=[color_m_neuron,'#D6D6D6',
+                                                                    color_running_average_stimuli]) # fefee3
 
 # %% plot functions
 
@@ -77,7 +77,8 @@ def lighten_color(color, amount=0.5):
 
 def plot_changes_upon_input2PE_neurons(std_mean = 1, n_std = 1, mfn_flag = '10'):
     
-    color_column = ['#00A9A5', '#D1495B']
+    color_columns_mean = [color_m_neuron, lighten_color(color_m_neuron, amount=0.7)]
+    color_columns_variance = [color_v_neuron, lighten_color(color_v_neuron, amount=0.7)]
     
     f, axs = plt.subplots(2, 2, sharex=True, sharey='row')
     axs[1,0].plot(np.nan, np.nan, color='k')
@@ -93,15 +94,15 @@ def plot_changes_upon_input2PE_neurons(std_mean = 1, n_std = 1, mfn_flag = '10')
             [pert_strength, m_act_lower, v_act_lower, v_act_higher] = pickle.load(f)
             
         
-        axs[0,0].plot(pert_strength, m_act_lower[:,1], color=color_column[column-1])
+        axs[0,0].plot(pert_strength, m_act_lower[:,1], color=color_columns_mean[column-1])
 
-        axs[0,1].plot(pert_strength, m_act_lower[:,0], color=color_column[column-1])
+        axs[0,1].plot(pert_strength, m_act_lower[:,0], color=color_columns_mean[column-1])
         
-        axs[1,0].plot(pert_strength, v_act_lower[:,1], color=color_column[column-1])
-        axs[1,0].plot(pert_strength, v_act_higher[:,1], color=color_column[column-1], ls='--')
+        axs[1,0].plot(pert_strength, v_act_lower[:,1], color=color_columns_variance[column-1])
[…]

### 2023-06-23 scripts/run_fig_3.py
polish Fig 2 and 3, work on Fig 3

@@ -12,8 +12,8 @@ import numpy as np
 import pickle
 
 from src.functions_simulate import simulate_weighting_example, simulate_weighting_exploration, simulate_dynamic_weighting_eg
-from src.functions_simulate import simulate_weighting_interneurons
-from src.plot_data import plot_weighting_limit_case_example, plot_fraction_sensory_heatmap, plot_transitions_examples
+from src.functions_simulate import simulate_activity_neurons
+from src.plot_data import plot_weighting_limit_case_example, plot_fraction_sensory_heatmap, plot_transitions_examples, plot_neuron_activity_lower_higher
 
 import warnings
 warnings.filterwarnings("ignore")
@@ -67,7 +67,7 @@ if run_cell:
                                       plot_legend = False)
     
     
-# %% Systematic exploration
+# %% Systematic exploration - record sensory weight
 
 run_cell = False
 plot_only = True
@@ -104,11 +104,7 @@ if run_cell:
                                 ylabel='variability within trial')
     
 
-# %% Systematic exploration - record interneurons as well 
-
-# are INs decreasing or increasing?
-# does this change with type of variability?
-# Do INs in lower or higher increase faster? Again, does this change with type of variabilty? ...
+# %% Systematic exploration - record neuron activity
 
 run_cell = True
 plot_only = True
@@ -119,8 +115,8 @@ if run_cell:
     mfn_flag = '10' # valid options are '10', '01', '11
     
     ### filename for data
-    file_for_data_within = '../results/data/weighting/data_weighting_activity_interneurons_exploration_within_' + mfn_flag + '.pickle'
-    file_for_data_across = '../results/data/weighting/data_weighting_activity_interneurons_exploration_across_' + mfn_flag + '.pickle'
+    file_for_data_within = '../results/data/weighting/data_activity_neurons_exploration_within_' + mfn_flag + '.pickle'
+    file_for_data_across = '../results/data/weighting/data_activity_neurons_exploration_across_' + mfn_flag + '.pickle'
     seeds = np.array([10,20,30,40,50])
     
     ### get data
@@ -133,15 +129,19 @@ if run_cell:
         mean_trials, m_sd = dtype(5), dtype(0)
         variability_within = np.linspace(0,3,5, dtype=dtype)
         variability_across = np.linspace(0,3,5, dtype=dtype)
-        
-        [_, _, activity_interneurons_lower_within,
-         activity_interneurons_higher_within] = simulate_weighting_interneurons(mfn_flag, variability_within, np.array([0], dtype=dtype), 
[…]

### 2023-06-23 scripts/run_fig_3_support_1.py

@@ -143,7 +143,7 @@ if run_cell:
     
 # %% Impact of baseline activity (only PE neurons)
     
-run_cell = True
+run_cell = False
 plot_only = True
 
 if run_cell:
@@ -193,7 +193,7 @@ if run_cell:
 
 # %% Impact of baseline activity of nPE OR pPE neuron
     
-run_cell = True
+run_cell = False
 plot_only = True
 
 if run_cell:
@@ -202,7 +202,7 @@ if run_cell:
     mfn_flag = '10' # valid options are '10', '01', '11
     
     ### filename for data ... change PE neuron type to be tested here
-    id_cells_modulated = np.array([True,False,False,False,False,False,False,False])
+    id_cells_modulated = np.array([False,True,False,False,False,False,False,False])
     
     if id_cells_modulated[0]:
         file_for_data = '../results/data/weighting/data_weighting_baseline_nPE.pickle'

### 2023-06-23 scripts/run_fig_4.py

@@ -17,6 +17,7 @@ import pandas as pd
 from src.functions_simulate import simulate_neuromod, simulate_moment_estimation_upon_changes_PE, simulate_neuromod_effect_on_neuron_properties
 from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs, plot_neuromod_per_net, plot_points_of_interest_neuromod, plot_bar_neuromod
 from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_changes_upon_input2PE_neurons, plot_bar_neuromod_stacked
+from src.plot_data import plot_neurmod_results
 
 import warnings
 warnings.filterwarnings("ignore")
@@ -99,24 +100,21 @@ if run_cell:
     
 # %% Plot neuromodulation results in bar plots for VIP-PV and VIP-SOM
 
-run_cell = True
+run_cell = False
 mfn_flag = '10'
 
 if run_cell:
     
-    f, ax = plt.subplots(1, 3, figsize=(15,3)) #  plt.subplots(2, 3, figsize=(12,6))
+    #f, ax = plt.subplots(1, 3, figsize=(15,3)) #  plt.subplots(2, 3, figsize=(12,6))
 
     #for flag_what in range(2): # 0: VIP-PV, 1: VIP-SOM
-    for column in range(3):
+    #for column in range(3):
         #plot_bar_neuromod(column, mfn_flag, flag_what, dgrad = 0.001, axs=ax[flag_what,column])
-        plot_bar_neuromod_stacked(column, mfn_flag, dgrad = 0.001, axs=ax[column])
+        #plot_bar_neuromod_stacked(column, mfn_flag, dgrad = 0.001, axs=ax[column])
+        
+    plot_neurmod_results(0)
+    plot_neurmod_results(1)
     
-
-# %% XXX
-
-# sensory weight -- variances
-# variance neuron -- nPE, pPE neuron
-# nPE, pPE neurons -- INs
 
 # %% How is the sensory weight influenced by the variance (illustrate)
 
@@ -155,11 +153,20 @@ if run_cell:
     # reduction in V neuron of first column leads to more sensory driven, while increase leads to more prediction driven
     # reduction in V neuron of second column leads to more prediction driven, while increase leads to more sensory driven
     # diagonal (same change in both columns, additive) remains 0.5 but deviations from it to the left/right are more blurred (fading out) when changes are postive
-   
     
+    
[…]

### 2023-06-23 src/plot_data.py

@@ -80,54 +80,238 @@ def plot_changes_upon_input2PE_neurons(std_mean = 1, n_std = 1, mfn_flag = '10')
     color_columns_mean = [color_m_neuron, lighten_color(color_m_neuron, amount=0.7)]
     color_columns_variance = [color_v_neuron, lighten_color(color_v_neuron, amount=0.7)]
     
-    f, axs = plt.subplots(2, 2, sharex=True, sharey='row')
-    axs[1,0].plot(np.nan, np.nan, color='k')
-    axs[1,0].plot(np.nan, np.nan, color='k', ls='--')
+    f, axs = plt.subplots(1, 4, sharex=True, sharey=True)
+    # axs[1,0].plot(np.nan, np.nan, color='k')
+    # axs[1,0].plot(np.nan, np.nan, color='k', ls='--')
     
-    for column in range(1,3):
+    ### results lower column
+    column = 1
+    identifier = '_column_' + str(column) + '_acrossvar_' + str(std_mean) + '_withinvar_' + str(n_std)
+    file_for_data = '../results/data/neuromod/data_moments_vs_PE_neurons_' + mfn_flag + identifier + '.pickle'
+    
+    with open(file_for_data,'rb') as f:
+        [pert_strength, m_act_lower, v_act_lower, v_act_higher] = pickle.load(f)
         
-        ### filename for data
-        identifier = '_column_' + str(column) + '_acrossvar_' + str(std_mean) + '_withinvar_' + str(n_std)
-        file_for_data = '../results/data/neuromod/data_moments_vs_PE_neurons_' + mfn_flag + identifier + '.pickle'
+    axs[0].plot(pert_strength, m_act_lower[:,0], color=color_columns_mean[column-1])
+    axs[0].plot(pert_strength, v_act_lower[:,0], color=color_columns_variance[column-1])
+    axs[0].plot(pert_strength, v_act_higher[:,0], color=color_columns_variance[column-1], ls='--')
+    
+    axs[1].plot(pert_strength, m_act_lower[:,1], color=color_columns_mean[column-1])
+    axs[1].plot(pert_strength, v_act_lower[:,1], color=color_columns_variance[column-1])
[…]

### 2023-06-30 records/manuscript_2023.tex
polish Fig 2 and 3, work on Fig 3
write Results, part I, make & polish Figures 3-5

--- 2022-10-25 records/manuscript_2023.tex
@@
 \section*{Introduction}
+%
+XXX
+%
+Behavior of intelligent life is context-dependent and requires a constant adaptation to the ever-changing life.
+This must also be reflected in the underlying neural networks and the inputs that drive those.
+For instance, when you walk down your your basement stairs and the lighting is well, your brain might rely entirely on the feedforward sensory input your senses receive to take decisions and adapt your body in the environment/world (Fig 1a).
+However, when you walk down the same basement stairs you walked million of times before but suddenly the light bulb is broken and the visual input is absent, you don't freeze forever but walk the stairs almost as confidently as before.
+Your brain now relies on the prediction it had formed over the last years.
+(Fig.
+1b).
+The former example resonates with the feedforward-centric view that has dominated the neuroscience literature for decades.
+However, in recent years, the importance of internally generated, feedback inputs has increased significantly.
+Not only because it was shown that top-down projects xxx outsize feedforward connections (ref) but also because these connection modulate (ref) or even drive (ref) neuron dynamics.
+These feedback projects have been hypothesized to carry expectations, predictions ...
+and may help us to navigate (in the physical and metaphysical sense) the world even in the absence of sensory input.
+But how do our brains decide whether it follows more the sensory feedforward input or the feedback predictions?
 
-XXX
+For instance, if you hike down a mountain for the first time under very foggy conditions, you brain may receive unreliable sensory input but at the same time can only have a very rough prediction about what to expect.
+In this case, your brain might be better off combining both the sensory input and the prediction thereof (Fig 1c).
+A common hypothesis is that the brain weights those signals according to their reliabilties, that is, a function of the variances of those signals (ref).
+These ideas come from Bayesian integration in which it is believed that xxx (ref).
+A prominent example of this is multisensory integration in which xxx (ref).
+Similarly, the brain must have mechanisms to estimate the variances of both the sensory input and the predictions thereof that allow it to weight those signals accordingly and context-dependent.
+But how this is implemented in the brain is still largely unknown.
+
+Here, we hypothesis that xxx. 
+We show that xxx
 
@@
 %
-XXX
 
-\subsection*{first result}
+\subsection*{nPE and pPE neurons as the basis for computing mean and variance of sensory stimuli}
 %
+We hypothesise that the distinct response patterns of negative and positive prediction-error (nPE/pPE) neurons represent the backbone for estimating the mean and the variance of sensory stimuli.
+nPE neurons only increase their activity relative to a baseline when the sensory input is weaker than predicted, while pPE neurons only increase their activity relative to a baseline when the sensory input is stronger than predicted.
+Moreover, both nPE and pPE neurons remain at their baseline activity when the sensory input is fully predicted (XXX).
+Assuming that the prediction equals the mean of the sensory stimulus, the PE neurons, hence, encode the deviation from the mean.
+Thus, the squared sum of nPE and pPE activity represents the variance of the feedforward input.
 
-\subsubsection*{second result}
+To test our hypothesis, we studied a rate-based mean-field network the core of which is a prediction-error (PE) circuit with excitatory nPE and pPE neurons, as well as inhibitory parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (Fig.
+xxx).
+While the excitatory neurons are simulated as two coupled point compartments to emulate the soma and dendrites of elongated pyramidal cells, respectively, all inhibitory cell types were modeled as point neurons.
+The connectivity of and inputs to the network were chosen such that the excitatory (E) and inhibitory (I) pathways onto the pyramidal cells were balanced because it has been shown that this E/I balance is necessary for nPE and pPE neurons to emerge (XXX, see Methods).
+
+In addition to this core circuit, we model a memory (M) neuron that perfectly integrates the activity of the PE neurons.
+In accordance with XXX, we assume that the pPE neuron excites the memory neuron, while the nPE neuron inhibits this neuron (for instance, through lateral inhibition, here not modeled explicitly).
+The M neuron connects to the apical dendrites of the PE neurons and some of the interneurons (here, VIP and PV neurons, see Methods for more details).
+In this network, the M neuron serves as a prediction that is dynamically updated when new sensory information is available.
+We furthermore simulate a downstream neuron (termed V neuron), modeled as a leaky integrator with a squared activation function, that receives excitatory output synapses from the PE neurons.
+Hence, in this setting, the V neuron encodes the variance of the sensory stimuli.
+
+To show that this network can indeed represent mean and variance in the respective neurons, we stimulate it with a sequence of step-wise constant inputs drawn from a uniform distribution (Fig.
+2aXX), assuming that the sensory stimulus varies over time.
+In line with the distinct response patterns for nPE and pPE neurons, these neurons change only slightly with increasing stimulus mean but increase strongly with input variance (Fig.
+2bXXX).
+This is in contrast to the three interneurons that strongly increase with stimulus mean while they only moderately increase with stimulus variance (Fig.
+2cxxx).
+The activity of the memory neuron M gradually approaches the mean of the sensory inputs (Fig.
+2d, middle), while the activity of the V neuron approaches the variance of the inputs (Fig.
+2e, middle).
+This is true for a wide range of input statistics (Fig.
+XXX) and input distributions (Fig.
+SXXX).
+Small deviations from the true mean occur mainly for larger input variances, while the estimated variance is fairly independent of the input statistics tested.
+
+XXX validated that also correct for population network (beyond mean-field network) XXX
+XXX assumptions (BL, gain-nPE = gain-pPE = 1) and how important really (see above)
+
+XXX Summary paragraph XXX
+
+\subsection*{Weighting external and internal signals requires two sets of PE neurons}
+%
+Following the ideas of Bayesian multisensory integration (XXX), the weighting of sensory stimuli and predictions thereof would require knowledge of their variances.
+As we have shown in the previous section, the variance of the sensory stimulus can be estimated by means of PE neurons.
+We hypothesise that the same principles apply to estimating the variance of the prediction.
+Hence, we augment the network with a \textit{higher} PE circuit that receives output synapses from the M neuron of the \textit{lower} PE circuit (Fig.
+1xxx).
+Both subnetworks are modeled the same, except that the M neuron in the higher PE circuit evolves more slowly than the M neuron in the lower PE circuit.
+
+To test the network's ability to estimate the variances correctly, we stimulated the network with a sequence of inputs.
+In each trial one stimulus is shown to the network.
+To account for the stimulus variance, each stimulus is composed of n constant values drawn from a normal distribution with mean $\mu_\mathrm{stimulus}$ and variance $\sigma_\mathrm{stimulus}^2$, and presented one after the other.
+To account for potential changes in the environment, in each trial, we draw $\mu_\mathrm{stimulus}$ from a uniform distribution (Fig.
+2a).
+Hence, the inputs change on two different time scales, with stimulus variability (faster time scale) and trial variability (slower time scale).
+
+As expected, the neurons' activity increase for both stimulus and trial variances (Fig.
+2b).
+While the neurons in the lower PE circuit increase more strongly with stimulus variability, the neurons in the higher PE circuit increase more strongly with trial variability, indicating that the different subnetworks process different aspects of the inputs.
+We first consider two limit cases.
+In the first limit case, a different but low-variance stimulus is presented in each trial (Fig.
+2c, left).
+In line with the ideas of multisensory integration (XXX), the network should therefore follow the sensory inputs closely and ignore the predictions.
+When we arithmetically calculate the weighted output (Fig.
+2c, middle) based on the feedforward and feedback inputs, and the sensory weight (Fig.
+2c, right), the network correctly represents mostly the sensory input (for more details, see Methods).
+In the second limit case, the same but high-variance stimulus is presented in each trial (Fig.
+2d, left).
+According to theory (XXX), the network should downscale the sensory feedforward inputs and weight the prediction more strongly.
+Indeed, the weighted output of the network shows a clear tendency to the mean of the stimuli (Fig.
+2d, middle), also reflected in the low sensory weight (Fig.
+2d, right).
+
+In a next step, to validate the network responses more broadly, we systematically varied the trial and stimulus variability independently.
+If both variances are similar, the sensory weight approaches 0.5, reflecting equal contribution of sensory inputs and predictions to the weighted output.
+Only if both variances are zero, the network represents the sensory input perfectly.
+In line with the limit case examples above, if the stimulus variance is larger than the trial variance, the network weighs the prediction more strongly than the sensory input.
+This is reversed if the stimulus variance is smaller than the trial variance (Fig.
+2e).
+Because the network dynamically estimates the mean and variances of the sensory inputs and the prediction, the weighted output and the sensory weight changes accordingly when the input statistics changes (Fig.
+SX).
+
+The first limit case (Fig.
+2 c) shows that even in a sensory-driven input regime, the prediction is weighted more at the beginning of a new trial than in the steady state.
+This is further confirmed in simulations in which the trail duration was shortened.
+For those simulations, the prediction even outweighs the sensory input, reflected in a very low sensory weight (Fig.
+2e).
+This suggests that predictions influence neural activity more significantly in experiments that rely on very fast stimulus changes.
+
+For some psychiatric disorders, it has been shown that the weighting of sensory inputs and predictions thereof is impaired (XXX), leading to an overweighting of one of these signals.
+Moreover, it has been hypothesised that factors like stress or cognitive load may also influence the processing of feedforward and feedback inputs (XXX).
+Naturally, we were wondering which factors might influence the estimation of the variances, and, consequently, the weighting of different input streams.
+In our network, the M neuron evolves faster in the lower PE circuit than in the higher PE circuit.
+When we xxxx, the network shows a bias for predictions, while when we xxx, the network shows a bias for the sensory stimuli (Fig.
+SXX).
+This indicates that a non-pathological weighting requires xxx.
+Another factor that may contribute to a distorted weighting is the baseline activity of PE neurons that was set to zero in our model, in line with very low spontaneous firing rates of excitatory neurons in rodent primary sensory areas (XXX).
+Increasing this baseline activity for the nPE neuron (Fig.
+SXXX), pPE neuron (Fig.
+SXXX) or both pushes the network to weight sensory stimuli and predictions more equally.
+We speculate that an increase of baseline activity may be a natural result of an increased cognitive load or stress.
+
+XXX Summary XXX
+% I haven't talked about the different activation function but I don't see where this fits naturally ...
+
+% XXX Do we actually propose that predictions are sent up the hierarchy?
+
+\subsection*{XXX IN influence, neuromodulators ... mechanisms XXX}
+%
+The brain's flexibility and adaptability are not least due to the fact that a plethora of neuromodulators influence the activity of neurons in a variety of ways (XXX).
+A prominent target of neuromodulatory inputs are inhibitory neurons (Cardin 2019, XXX).
+Moreover, distinct interneuron types are differently (in-)activated by those neuromodulators.
+For instance, it has been shown that xxx.
+We were therefore wondering if and how the weighting of sensory inputs and predictions thereof may be biased when neuromodulators activate distinct interneuron types.
+
+To this end, we modeled the presence of neuromodulators by injecting an additional, excitatory input into one or two interneuron types.
+We reasoned that the network effect of a neuromodulator not only depends on the interneuron type it targets but also on the inputs this neuron receives and the connections it makes with other neurons in the network.
+We, therefore, tested three different mean-field networks that differ with respect to the distribution of sensory inputs and predictions onto the interneurons and the underlying connectivity.
+The commonality across those networks is that all of them exhibit an E/I balance of excitatory and inhibitory pathways onto the PE neurons (XXX).
+Across the different mean-field networks tested, activating a SOM or VIP neuron individually forces the networks to weigh both inputs more equally.
+As a consequence, in a sensory-driven input regime, predictions are overrated.
+Similarly, in a prediction-driven regime, sensory inputs are overrated.
+Interestingly, when both interneuron types are activated to the same degree, this effect disappears (Fig.
+5axx, left).
+In contrast, activating PV neurons largely biases the network output towards predictions.
+This is even more pronounced when PV and SOM, or PV and VIP neurons are activated simultaneously (Fig.
+5a, middle and right).
+
+In the previous simulations, we assumed that a neuromodulator acts globally, that is, on the interneurons in both the lower and the higher PE circuit.
+While this is in line with experimental data showing that xxx (XXX), we note that neuromodulators may also act more locally.
+We hence tested how the weighting of sensory stimuli and predictions may change when a neuromodulator activates specific interneurons in the lower or higher PE circuit only.
+The effect of activating an interneuron type in the lower PE circuit on the sensory weight is mostly the opposite of activating the same interneuron in the higher PE circuit (Fig.
+SXXX).
+For instance, the sensory inputs are overrated when VIP neurons are activated in the higher PE circuit, while the prediction is overrated when VIP neurons in the lower PE circuit are activated.
+When VIP and SOM neurons are stimulated equally, the sensory weight remains unchanged, independently of which PE circuit is targeted by the neuromodulator.
+
+What are the mechanisms that give rise to these effects? And how do the combined local changes give rise to the global one observed in our network simulations?
+
+% Mention that for V neuron (higher) the darker line is actually not horizontal (unaffected)!
+% Mention that in BL illustration you need a factor to increase visibility ... (in caption)
+
+\subsection*{XXX contraction bias}
+% 
+The weighting of sensory inputs and predictions thereof manifests in all-day behavior, in the form of a phenomenon called \textit{contraction bias}.
+The contraction bias describes the tendency to overestimate sensory stimuli that are at the lower end of a stimulus distribution and to underestimate stimuli that are at the upper end of the same distribution (Fig.
+5aXX).
+This \textit{bias towards the mean} has been reported in different species and modalities (XXX).
+
+The weighted output of our network can be interpreted as a neural manifestation of the contraction bias (see Methods for a thorough analysis).
+The bias increases with stimulus variance (Fig.
+5bXX), decreasing the slope of the linear fit modeling the relationship between the true and estimated stimuli (Fig.5 b, right XXX, compare with Fig.
+5a).
+In contrast, the bias decreases with trial variance, so that the slope of the linear fit approaches 1 (Fig.
+5b rightXXX).
+
+What are the underlying network factors that contribute to this phenomenon?
+To disentangle the potentially different sources of the bias, we first simulated a network without stimulus variability (variance set to zero) for two different trial variabilities.
+In this case, a contraction bias emerges but is independent of the volatility of the environment (Fig.
+5c, leftXXX).
+We show mathematically that the bias is a result of the recent stimulus history and vanishes if the trial duration approaches infinity (see Methods for more details, and Fig.
+5d).
+We next resume the limit case in which the same but high-variance stimulus is shown in every trial.
+In this case, the weighted output exhibits a contraction bias that is largely independent of the exact stimulus variances tested.
+As shown mathematically (see Methods), the bias is a result of xxx and approaches xxx (Fig.
+5dxxx).
+% Maybe show the limit for both cases and show the theoretical values for each T as well ...
+
+So far, we assumed that the stimulus variance is independent of the trial mean.
+A consequence of this choice is that the bias on either end of the stimulus distribution is the same (but with reversed signs).
+However, behavioral (neural?) data (XXX) shows that the bias increases for stimuli drawn from the upper end of the distribution, a phenomenon usually attributed to \textit{scalar variability}.
+To capture this in the model, we assume that the stimulus standard deviation linearly increases with the trial mean.
+In these simulations, as expected, the bias increases for a stimulus distribution shifted to higher trial means (Fig.
+5eXXX).
+
+XXX Summary XXX
 
@@
 
+\subsection*{Sensory weight and contraction bias}
+%
+If P is rather constant, the slope in the contraction bias is exactly the sensory weight
+%
+ \begin{align*}
+ r_\mathrm{out} = \alpha_\mathrm{S} \cdot S + \left( 1 -\alpha_\mathrm{S} \right) \cdot P \equiv m \cdot S + n
+ \end{align*}
+%
+However, P is usually/normally a function of $S$. For simplicity, let's assume that P decays exponentially to a new value of $S$:
+%
+\begin{align*}
+P = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau} + f(S) \cdot \left( 1 -   \mathrm{e}^{-t/\tau} \right)
+\end{align*}
+%
+Within each trial with trial duration T, P can be expressed by n sections of length t in which the stimulus is constant and, for the sake of simplicity, drawn from a uniform distribution $U\left( s - \frac{\sigma_\mathrm{S}}{12}, s + \frac{\sigma_\mathrm{S}}{12} \right)$.
+$P_\mathrm{0}$ is drawn from $U\left( \mu - \frac{\sigma_\mathrm{P}}{12}, \mu + \frac{\sigma_\mathrm{P}}{12} \right)$.
+$P_\mathrm{n}$ is then given by
+%
+\begin{align*}
+P_\mathrm{n} = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau}  + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} s_i \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau}
+\end{align*}
+%
+This needs to be averaged over all possible states
+%
+\begin{align*}
+P_\mathrm{n} = \mathrm{e}^{-t/\tau}  \int\limits_{\mu - \frac{\sigma_\mathrm{P}}{12}}^{\mu + \frac{\sigma_\mathrm{P}}{12}} P_\mathrm{0} \ f(P_\mathrm{0})\ dP_0+ \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau} \int\limits_{s - \frac{\sigma_\mathrm{S}}{12}}^{s + \frac{\sigma_\mathrm{S}}{12}} s\ f(s)\ ds
+\end{align*}
+%
+This gives
+%
+\begin{align*}
+P_\mathrm{n} = \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot t/ \tau} \cdot S
+\end{align*}
+%
+By making use of the geometric series, this simplifies to
+%
+\begin{align*}
+P_\mathrm{n} =  \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right) \cdot S
+\end{align*}
+%
+Together, this yields for the weighted output
+%
+\begin{align*}
+ r_\mathrm{out} = \left[ \alpha_\mathrm{S}\  \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right)\right] \cdot S + \left( 1 -\alpha_\mathrm{S} \right)\ \mathrm{e}^{-T/\tau}\ \mu
+\end{align*}
+%
+Hence, the slope is a function of both the sensory weight and the trial duration.
+
+In a prediction-driven input regime ($alpha_\mathrm{S} \sim 0)$, the slope is independent of the sensory weight and only determined by the trial duration, $m \sim \left( 1 -   \mathrm{e}^{-T/\tau} \right)$.
+In a sensory-driven input regime ($alpha_\mathrm{S} \sim 0)$, the contraction bias vanishes ($m \sim 1$).
+
+If the trail duration is short ($T \rightarrow 0$), the slope is given by the sensory weight.
+If the trail duration approaches infinity, the slope would be 1 again (however, this seems rather unrealistic, this would only be true in an ideal system without memory decay or reproduction and accumulation noise ...).
+
+\subsection*{Impact of PE neurons' gain on estimating mean and variance}
+%
+Only if the the gain of the nPE neuron ($g_{nPE}$) equals the gain of the pPE neuron ($g_{pPE}$) in the mean-field network, the activity of the M neuron represents the mean of the inputs,
+%
+\begin{align}
+\label{eq:condition_mean_gain_equal}
+g_{pPE}\ \langle \mathrm{nPE}\rangle &= g_{nPE}\ \langle \mathrm{pPE}\rangle \\
+g_{pPE} \langle \left[ S-P\right]_+\rangle &= g_{nPE} \langle \left[ P-S\right]_+\rangle \nonumber \\
+g_{pPE} \int\limits_P^b \left( x-P\right)\ f(x)\ dx &= g_{nPE} \int\limits_a^P \left( P-x\right)\ f(x)\ dx. \nonumber
+\end{align}
+%
+In case of a uniform distribution ($f(x) = 1/(b-a)$ when $x\in [a,b]$ and $0$ otherwise) from which the input values are drawn, this condition yields
+%
+\begin{align}
+\label{eq:condition_mean_gain_equal_4_uniform_dist_1}
+P = \int\limits_a^b x\, f(x)\ dx = \frac{a + b}{2}
+\end{align}
+%
+for $g_{nPE} = g_{pPE} = g$, and 
+%
+\begin{align}
+\label{eq:condition_mean_gain_equal_4_uniform_dist_21}
+g_{pPE}\ \left[ \frac{1}{2} \left(b^2 - P^2\right) - P\left(b - P\right)\right] = g_{nPE}\ \left[  P\left(P - a\right) - \frac{1}{2} \left(P^2 - a^2\right)\right]
+\end{align}
+%
+for $g_{nPE} \neq g_{pPE}$, which can be further summaries by
+%
+\begin{align}
+\label{eq:condition_mean_gain_equal_4_uniform_dist_21}
+P = \frac{g_{pPE}\ b - g_{nPE}\ a \pm \sqrt{g_{nPE}\ g_{pPE}} (a-b)}{g_{pPE} - g_{nPE}}.
+\end{align}
+%
+Hence, estimating the mean correctly requires $g_{nPE} = g_{pPE} = g$.
+For the V neuron to represent the variance of the inputs, this condition must be tightened to  $g_{nPE} = g_{pPE} = 1$.
+The variance is given by
+%
+\begin{align}
+\label{eq:condition_variance_gain}
+V &= \langle (S-P)^2\rangle \\
+   &= g_\mathrm{pPE}\ \langle \left[ S-P \right]_+^2 \rangle + g_\mathrm{nPE}\ \langle \left[ P-S \right]_+^2 \rangle. \nonumber
+\end{align}
+%
+In case of a uniform distribution from which the input values are drawn, this condition yields
+%
+\begin{align}
+\label{eq:condition_variance_gain_uniform_dist}
+V &= \frac{g_\mathrm{pPE}}{b-a} \int\limits_P^b (u-P)^2\ du + \frac{g_\mathrm{nPE}}{b-a} \int\limits_a^P (P-u)^2\ du \\
+   &= \frac{g_\mathrm{pPE}}{3} \cdot \frac{(b-P)^3}{b-a} + \frac{g_\mathrm{nPE}}{3} \cdot \frac{(P-a)^3}{b-a}. \nonumber
+\end{align}
+%
+Only if $g_{nPE} = g_{pPE} = 1$ and $P = \frac{a + b}{2}$, the variance is given by $\frac{(b - a)^2}{12}$, otherwise the V neuron's activity is given by
+%
+\begin{align}
+\label{eq:condition_variance_gain_uniform_dist_1}
+V = \frac{(b-a)^2}{3\ (g_\mathrm{pPE} - g_\mathrm{nPE})^3} \cdot \left[ g_\mathrm{nPE} \cdot( g_\mathrm{pPE} \mp \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}})^3 - g_\mathrm{pPE} \cdot (g_\mathrm{nPE} \mp \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}})^3\right].
+\end{align}
+%
+
+\subsection*{Impact of PE neurons' baseline on estimating mean and variance}
+%
+Only if the the baseline of the nPE neuron ($n_0$) equals the baseline of the pPE neuron ($p_0$) in the mean-field network, the activity of the M neuron represents the mean of the inputs,
+%
+\begin{align}
+\label{eq:condition_baseline_mean}
+\langle \mathrm{pPE} \rangle &= \langle \mathrm{nPE} \rangle \\
+\langle \left[ S - P\right]_+ + p_0\rangle &= \langle \left[ P - S\right]_+ + n_0\rangle \nonumber\\
+\int\limits_P^b (x - P)\ f(x)\ dx + p_0 \underbrace{\int\limits_a^b f(x)\ dx}_{=1}  &= \int\limits_a^P (P - x)\ f(x)\ dx + n_0 \underbrace{\int\limits_a^b f(x)\ dx}_{=1} . \nonumber
+\end{align}
+%
+In case of a uniform distribution ($f(x) = 1/(b-a)$ when $x\in [a,b]$ and $0$ otherwise) from which the input values are drawn, this condition yields
+%
+\begin{align}
+\label{eq:condition_baseline_mean_1}
+P = \frac{b+a}{2} + \frac{p_0 - n_0}{b-a}.
+\end{align}
+%
+Hence, if $p_0 = n_0$, the mean can be estimated correctly. For the V neuron to represent the variance of the inputs, this condition must be tightened to $p_0 = n_0 = 0$. The variance is given by
+%
+\begin{align}
+\label{eq:condition_baseline_variance}
+V &= \langle \left( \mathrm{pPE} + \mathrm{nPE} \right)^2 \rangle \\
+&= \langle \left[ S-P\right]_+^2\rangle + \langle \left[ P-S\right]_+^2\rangle + (p_0 + n_0)^2 + 2\ (p_0 + n_0)\ \left( \langle \left[ S-P\right]_+\rangle + \langle \left[ P-S\right]_+ \rangle\right) \nonumber
+\end{align}
+%
+In case of a uniform distribution from which the inputs to the mean-field network are drawn, this expression simplifies to
+%
+\begin{align}
+\label{eq:condition_baseline_variance_1}
+V = \frac{1}{3\ (b-a)} \left[ (b-P)^3 + (P-a)^3\right] + (p_0 + n_0)^2 + \frac{(p_0 + n_0)}{b-a} \left[ (b-P)^2 + (a-P)^2\right].
+\end{align}
+%
+Inserting the expression for P (by itself modulated by the baseline activities of the PE neurons) yields
+%
+\begin{align}
+V =& \frac{1}{3\ (b-a)} \left[ \left( \frac{b-a}{2} - \frac{p_0 - n_0}{b-a}\right)^3 + \left( \frac{b-a}{2} + \frac{p_0-n_0}{b-a}\right)^3\right] + (p_0 + n_0)^2 \nonumber \\
+&+ \frac{(p_0 + n_0)}{b-a} \left[ \left( \frac{b-a}{2} + \frac{p_0-n_0}{b-a}\right)^2 + \left( \frac{b-a}{2} - \frac{p_0 - n_0}{b-a}\right)^2\right]
+\end{align}
+%
+Simplifying the expression, leads to
+%
+\begin{align}
+\label{eq:condition_baseline_variance_2}
+V =  \frac{(b-a)^2}{12} + \frac{(p_0-n_0)^2}{(b-a)^2} \left( 1 + 2\ \frac{p_0+n_0}{b-a}\right) + (p_0 + n_0) \left( p_0 + n_0 + \frac{b-a}{2}\right)
+\end{align}
+%
+
+\subsection*{Modelling the impact of neuromodulators on the weighting of sensory inputs and predictions thereof}
+
+Assume that neuromodulator act through activating INs.
+INs on the other hand modulate both gain and BL of nPE and pPE neurons ...
+.
+Changes in gain and BL will affect the overall activity and the balance between nPE and pPE neurons => this will affect both the prediction and the variance.
+
+Modulate PE in lower:
+variance in lower is combination of direct changes in PE and indirect through prediction (that is changed through PE mod)
+variance in higher is just because of differences in prediction
+
+Modulation in higher:
+variance in lower should be unaffected
+variance in higher is combination of direct changes in PE and indirect through prediction (that is changed through PE mod)
+
+Variance as function of bias in mean:
+\begin{align*}
+V &= \frac{1}{n} \sum_i \left( x_i - \left(\mu \pm \delta\mu\right)\right)^2 \\
+&= \frac{1}{n} \sum_i \lbrace  \left( x_i - \mu \right)^2 + \delta\mu^2 \mp 2\, \delta\mu\,  (x_i - \mu)\rbrace \\
+&= V_\mathrm{unmod} + \delta\mu^2 \mp 2\ \delta\mu \left( \frac{1}{n} \sum_i x_i- \mu\right) \\
+&= V_\mathrm{unmod} + \delta\mu^2
+\end{align*}
+%
+
+\subsection*{Influence of a population of nPE and pPE neurons}
+XXX
+
 \subsection*{Analysis of simplified network model, effect of time constants}

### 2023-06-30 scripts/plot_fig_4.py

@@ -20,46 +20,121 @@ warnings.filterwarnings("ignore")
 dtype = np.float32
 
 
-# %% for plotting
+# %% import
 
-run_cell = True
-plot_only = True
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import os.path
+import seaborn as sns
 
-if run_cell:
-    
-    ### choose mean-field network to simulate
-    mfn_flag = '10' # valid options are '10', '01', '11
-    
-    ### define target area and stimulus statistics
-    column = 1      # 0: both, 1: lower level PE circuit, 2: higher level PE circuit
-    std_mean = 0   # uncertainty of environement [0, 0.5, 1]
-    n_std = 1       # uncertainty of stimulus [1, 0.5, 0]
-    
-    ### filename for data
-    identifier = '_column_' + str(column) + '_acrossvar_' + str(std_mean) + '_withinvar_' + str(n_std)
-    file_for_data = '../results/data/neuromod/data_weighting_neuromod' + identifier + '.pickle'
-    
-    ### get data
-    if not plot_only: # simulate respective network
+from matplotlib.collections import LineCollection
 
-        nums = 11
-        xp, xs = np.meshgrid(np.linspace(0, 1, nums), np.linspace(0, 1, nums))
-        xv = np.sqrt(1 - xp**2 - xs**2)
-        
-        [_, _, _, alpha_before_pert, alpha_after_pertt] = simulate_neuromod(mfn_flag, std_mean, n_std, column, 
-                                                                            xp, xs, xv, file_for_data = file_for_data)
-        
-    else:
-        
-        with open(file_for_data,'rb') as f:
-            [xp, xs, xv, alpha_before_pert, alpha_after_pert] = pickle.load(f)
-            
-    
-    ### plot data
-    plot_heatmap_neuromod(xp, xs, alpha_before_pert, alpha_after_pert)
-    plot_combination_activation_INs(xp, xs, xv, alpha_before_pert, alpha_after_pert)
-    
[…]

### 2023-06-30 src/functions_simulate.py

@@ -784,6 +784,88 @@ def simulate_neuromod(mfn_flag, std_mean, n_sd, column, xp, xs, xv, mean_trials 
     return [xp, xs, xv, alpha_before_pert, alpha_after_pert]
 
 
+def simulate_neuromod_combos(mfn_flag, std_mean, n_sd, column, xp, xs, xv, mean_trials = dtype(5), m_sd = dtype(0), 
+                             last_n = np.int32(50), seed = np.int32(186), n_trials = np.int32(200), 
+                             trial_duration = np.int32(5000), num_values_per_trial = np.int32(10), 
+                             file_for_data = None, mult_input = dtype(1)):
+    
+    ### load default parameters
+    VS, VV = int(mfn_flag[0]), int(mfn_flag[1])
+    
+    [w_PE_to_P, w_P_to_PE, w_PE_to_PE, w_PE_to_V, 
+      v_PE_to_P, v_P_to_PE, v_PE_to_PE, v_PE_to_V, 
+      tc_var_per_stim, tc_var_pred, tau_pe, fixed_input] = default_para_mfn(mfn_flag, one_column=False)
+    
+    ### define stimuli
+    np.random.seed(seed)
+    n_repeats_per_stim = dtype(trial_duration/num_values_per_trial)
+
+    stimuli = stimuli_moments_from_uniform(n_trials, num_values_per_trial, dtype(mean_trials - np.sqrt(3)*std_mean), 
+                                            dtype(mean_trials + np.sqrt(3)*std_mean), dtype(m_sd), dtype(n_sd))
+    stimuli = np.repeat(stimuli, n_repeats_per_stim)
+    
+    ### run model without perturbation
+    print('Without neuromodulation\n')
+    [_, _, _, _, alpha, _, _] = run_mfn_circuit_coupled(w_PE_to_P, w_P_to_PE, w_PE_to_PE, v_PE_to_P, 
+                                                        v_P_to_PE, v_PE_to_PE, tc_var_per_stim, 
+                                                        tc_var_pred, tau_pe, fixed_input, stimuli, 
+                                                        VS = VS, VV = VV, w_PE_to_V = w_PE_to_V, 
+                                                        v_PE_to_V = v_PE_to_V)
+    
+    alpha_before_pert = np.mean(alpha[(n_trials - last_n) * trial_duration:])
+    
+    ### initialise
+    nums = len(xp)
+    alpha_after_pert = np.zeros(nums, dtype=dtype)
+
[…]

### 2023-06-30 src/plot_data.py

@@ -75,12 +75,291 @@ def lighten_color(color, amount=0.5):
 
 
 
-def plot_changes_upon_input2PE_neurons(std_mean = 1, n_std = 1, mfn_flag = '10'):
+def plot_influence_interneurons_gain_baseline(ax=None, fs=6, s=10, lw=1):
+    
+    ### load data
+    file_for_data = '../results/data/neuromod/data_PE_props_vs_neuromod_10.pickle'
+    with open(file_for_data,'rb') as f:
+            [results_base_nPE_10, results_base_pPE_10, results_gain_nPE_10, results_gain_pPE_10] = pickle.load(f) 
+            
+    file_for_data = '../results/data/neuromod/data_PE_props_vs_neuromod_11.pickle'
+    with open(file_for_data,'rb') as f:
+            [results_base_nPE_11, results_base_pPE_11, results_gain_nPE_11, results_gain_pPE_11] = pickle.load(f) 
+            
+    file_for_data = '../results/data/neuromod/data_PE_props_vs_neuromod_01.pickle'
+    with open(file_for_data,'rb') as f:
+            [results_base_nPE_01, results_base_pPE_01, results_gain_nPE_01, results_gain_pPE_01] = pickle.load(f)
+            
+    ### plot data
+    def create_legend_for_INS(c):
+        
+        if c==0:
+            res = 'PV'
+        elif c==1:
+            res = 'SOM'
+        elif c==2:
+            res = 'VIP'
+            
+        return res
+    
+    colors = [Col_Rate_PVv, Col_Rate_SOM, Col_Rate_VIP]
+    
+    if ax is None:
+        _, ax = plt.subplots(1,1)
+    
+    sc = ax.scatter((results_base_nPE_10[1:] - results_base_nPE_10[0]) + (results_base_pPE_10[1:] - results_base_pPE_10[0]),
+                (results_gain_nPE_10[1:] - results_gain_nPE_10[0]) + (results_gain_pPE_10[1:] - results_gain_pPE_10[0]), 
+                c=colors, marker='o', s=s)
+    
+    ax.scatter((results_base_nPE_11[1:] - results_base_nPE_11[0]) + (results_base_pPE_11[1:] - results_base_pPE_11[0]),
[…]

### 2023-07-07 records/manuscript_2023.tex
polish figures for manuscript & write first version of results

@@
 %
-XXX
-%
-Behavior of intelligent life is context-dependent and requires a constant adaptation to the ever-changing life.
-This must also be reflected in the underlying neural networks and the inputs that drive those.
-For instance, when you walk down your your basement stairs and the lighting is well, your brain might rely entirely on the feedforward sensory input your senses receive to take decisions and adapt your body in the environment/world (Fig 1a).
-However, when you walk down the same basement stairs you walked million of times before but suddenly the light bulb is broken and the visual input is absent, you don't freeze forever but walk the stairs almost as confidently as before.
-Your brain now relies on the prediction it had formed over the last years.
-(Fig.
-1b).
-The former example resonates with the feedforward-centric view that has dominated the neuroscience literature for decades.
-However, in recent years, the importance of internally generated, feedback inputs has increased significantly.
-Not only because it was shown that top-down projects xxx outsize feedforward connections (ref) but also because these connection modulate (ref) or even drive (ref) neuron dynamics.
-These feedback projects have been hypothesized to carry expectations, predictions ...
-and may help us to navigate (in the physical and metaphysical sense) the world even in the absence of sensory input.
-But how do our brains decide whether it follows more the sensory feedforward input or the feedback predictions?
+To survive in an ever-changing environment, animals must flexibly adapt their behavior based on previously encoded and novel information.
+This context-dependency must also be reflected in the information processing of neural networks underlying intelligent behavior.
+For instance, when you walk down some staircase in your fully lit basement, your brain might entirely rely on the feedforward (bottom-up) input your senses receive (Fig 1axxx).
+In contrast, when you walk down the same stairs in complete darkness, your brain might rely entirely on feedback (top-down) signals generated from a staircase model it has formed over previous experiences (Fig.
+1bxx).
 
-For instance, if you hike down a mountain for the first time under very foggy conditions, you brain may receive unreliable sensory input but at the same time can only have a very rough prediction about what to expect.
-In this case, your brain might be better off combining both the sensory input and the prediction thereof (Fig 1c).
-A common hypothesis is that the brain weights those signals according to their reliabilties, that is, a function of the variances of those signals (ref).
-These ideas come from Bayesian integration in which it is believed that xxx (ref).
-A prominent example of this is multisensory integration in which xxx (ref).
-Similarly, the brain must have mechanisms to estimate the variances of both the sensory input and the predictions thereof that allow it to weight those signals accordingly and context-dependent.
-But how this is implemented in the brain is still largely unknown.
+The importance of these feedback inputs has been emphasized by observations showing that top-down projections outnumber feedforward connections (XXX) and that they modulate (XXX) or even entirely drive (XXX) neuron activity.
+But how do neural networks switch between a feedforward-dominated and a feedback-dominated processing mode?
+And how do neural networks combine both input streams wisely?
+For instance, if you hike down a mountain for the first time in very foggy conditions, your brain receives unreliable visual information.
+In addition, it can only draw on a shaky prediction about what to expect (Fig.
+1cxxx).
 
-Here, we hypothesis that xxx. 
-We show that xxx
+A common hypothesis is that the brain weights different inputs according to their reliabilities which are a function of the variances of those inputs (XXX).
+A prominent example of this is Bayesian multisensory integration in which the information from multiple modalities is optimally combined (XXX).
+According to this theory, the representation of an object perceived through various senses is a linear combination of the single-modality estimates, weighted by their reliabilities.
+This theory has been backed up by a number of observations showing that xxx (XXX).
+The same concepts can be employed for the weighting of sensory feedforward inputs and predictions thereof (XXX).
+A central point in this weighting is the estimation of the sensory and prediction variances that underlie the reliability of each input.
+However, how this is implemented in the brain is still largely unknown.
+
+We hypothesised that prediction error (PE) neurons can provide the basis for the neural computation of variances.
+PEs are an integral part of the theory of predictive processing that states that the brain constantly compares incoming sensory information with predictions thereof.
+When those predictions are wrong, the resulting PEs allow the network to update the model of the world, thereby ensuring that the predictions are more accurate (XXX).
+Experimental evidence suggests that these PEs may be represented in the activity of distinct groups of neurons, termed PE neurons (XXX).
+In brain areas with excitatory neurons that exhibit low spontaneous firing rates, these neurons may come in two types (XXX).
+Negative PE (nPE) neurons mainly increase their activity when the prediction is \textit{stronger} than the sensory input, while positive PE (pPE) neurons mainly increase their activity when the prediction is \textit{weaker} than the sensory input.
+Indeed, it has been shown that excitatory neurons in layer 2/3 of rodent primary sensory areas can encode negative or positive PEs (XXX).
+
+% also, mention that PE neurons are formed by inhibition, so that you can mention your work and introduce different types of interneurons!?
+(maybe combine with below section, like as it has been shown that PE neurons depend on inhibition and E/I balance, we reasoned that neuromodulators acting on interneurons can hence bias the weighting and hence investigated  how this is infleunced ....
+something like that?)
+
+Here, we show that
+
+- unique response profiles of nPE and pPE can be the backbone for computing mean and variance of sensory stimuli
+- Estimating both the variance of the sensory inputs and the prediction thereof requires a hierarchy of PE neurons
+- In line with the theory, predictons are weighted more strongly than sensory information when environement is stable and therefor ethe prediction is reliable while the sensory stimuli are noisy
+- moreover, even when sensory information are crystal clear but the environement, that is the stimuli shown, is unstable, leading to outweighing of sensory inputs, right at the beginning of  a trial, predictions kick in shortly
+- neuromodulators acting globally on VIP neurons ..., acting on SOM neurons ... and acting on PV neurons ...
+- we show that this is due to an increase/decrease of gain/baseline of PE neurons and can be predicted from effect of IN on both properties
+- Finally, we show that the contraction bias is a manifestation of this weighting.
+We show that it depends on several factors, trail duration, depends on both the volatility of the environement and the stimulus uncertainty.
+A mixture of recent (last stimulus) if xxx and a more averaged lomger history of stimuli when xxx
+- summary sentence
 
@@
 
-\subsection*{nPE and pPE neurons as the basis for computing mean and variance of sensory stimuli}
+\subsection*{nPE and pPE neurons as the basis for estimating mean and variance of sensory stimuli}
 %
@@
 
-\subsection*{Weighting external and internal signals requires two sets of PE neurons}
+\subsection*{Estimating variances of sensory inputs and predictions requires a hierarchy of PE circuits}
 %
@@
 For some psychiatric disorders, it has been shown that the weighting of sensory inputs and predictions thereof is impaired (XXX), leading to an overweighting of one of these signals.
-Moreover, it has been hypothesised that factors like stress or cognitive load may also influence the processing of feedforward and feedback inputs (XXX).
-Naturally, we were wondering which factors might influence the estimation of the variances, and, consequently, the weighting of different input streams.
+We thus wondered which network properties might bias the estimation of the variances, and, consequently, the weighting of different input streams.
 In our network, the M neuron evolves faster in the lower PE circuit than in the higher PE circuit.
-When we xxxx, the network shows a bias for predictions, while when we xxx, the network shows a bias for the sensory stimuli (Fig.
+We identified the speed at which the M neurons are updated in the different PE circuits as a decisive network factor.
+To show this, we varied the weights from the PE neurons onto the M neurons in either PE circuit.
+When xxxx, the network shows a bias for predictions, while when xxx, the network shows a bias for the sensory stimuli (Fig.
 SXX).
-This indicates that a non-pathological weighting requires xxx.
-Another factor that may contribute to a distorted weighting is the baseline activity of PE neurons that was set to zero in our model, in line with very low spontaneous firing rates of excitatory neurons in rodent primary sensory areas (XXX).
-Increasing this baseline activity for the nPE neuron (Fig.
-SXXX), pPE neuron (Fig.
-SXXX) or both pushes the network to weight sensory stimuli and predictions more equally.
-We speculate that an increase of baseline activity may be a natural result of an increased cognitive load or stress.
+While the speed at which the activity of the M neurons evolve in relation to each other may underlie pathological weighting, the precise activation function of the V neurons is of less importance.
+When we replaced the squared activation function with a purely linear function, the V neurons do not encode the variance but the averaged absolute deviation of the sensory stimuli.
+However, the sensory weight is only slightly shifted to larger values for low trial/high stimulus variability (Fig.
+SXXX).
 
 XXX Summary XXX
-% I haven't talked about the different activation function but I don't see where this fits naturally ...
 
@@
 
-\subsection*{XXX IN influence, neuromodulators ... mechanisms XXX}
+\subsection*{Biasing the weighting of sensory inputs and predictions by neuromodulators acting on PE circuits}
 %
@@
 For instance, it has been shown that xxx.
-We were therefore wondering if and how the weighting of sensory inputs and predictions thereof may be biased when neuromodulators activate distinct interneuron types.
+We therefore wondered if and how the weighting of sensory inputs and predictions thereof may be biased when neuromodulators activate distinct interneuron types.
 
@@
 5axx, left).
-In contrast, activating PV neurons largely biases the network output towards predictions.
+In contrast, stimulating PV neurons largely biases the network output towards predictions.
 This is even more pronounced when PV and SOM, or PV and VIP neurons are activated simultaneously (Fig.
@@
 We hence tested how the weighting of sensory stimuli and predictions may change when a neuromodulator activates specific interneurons in the lower or higher PE circuit only.
-The effect of activating an interneuron type in the lower PE circuit on the sensory weight is mostly the opposite of activating the same interneuron in the higher PE circuit (Fig.
+The effect of stimulating an interneuron type in the lower PE circuit on the sensory weight is mostly the opposite of activating the same interneuron in the higher PE circuit (Fig.
 SXXX).
-For instance, the sensory inputs are overrated when VIP neurons are activated in the higher PE circuit, while the prediction is overrated when VIP neurons in the lower PE circuit are activated.
+For instance, the sensory inputs are overrated when VIP neurons are activated in the higher PE circuit, while the prediction is overrated when VIP neurons in the lower PE circuit are stimulated.
 When VIP and SOM neurons are stimulated equally, the sensory weight remains unchanged, independently of which PE circuit is targeted by the neuromodulator.
 
-What are the mechanisms that give rise to these effects? And how do the combined local changes give rise to the global one observed in our network simulations?
+What are the mechanisms that give rise to these effects?
+And how do the combined local changes give rise to the global one observed in our network simulations?
+The sensory weight is chosen to be a function of the V neurons of the lower and higher PE circuit.
+Hence, any changes to the sensory weight are a consequence of changes to the neurons encoding the variance (Fig.
+5xx).
+In our network, the V neurons only receive excitatory output synapses from both nPE and pPE neurons.
+Hence, any changes in the sensory weights upon activation of interneurons must be due to changes in the PE neurons.
+To disentangle the effect of nPE and pPE neurons, we perturbed those neurons individually in both the lower or higher subnetwork by injecting either an inhibitory or excitatory additional input (Fig.
+5cxxx).
+Stimulating either PE neuron in the lower subnetwork increases the activity of the lower V neuron strongly.
+Moreover, the higher V neuron is also slightly affected.
+At first, this is counterintuitive because the V neuron in the higher subnetwork does not receive direct synapses from the PE neurons in the lower subnetwork.
+However, the activity of the M neuron encoding the prediction increases with an excitatory input onto the pPE neuron and decreases when the same input targets the nPE neuron (the opposite is true for an inhibitory input).
+Because neurons in the higher PE circuit receive synapses from the M neuron of the lower subnetwork, the estimation of the variance in the higher subnetwork is also affected.
+In contrast, stimulating either PE neuron in the higher subnetwork increases the activity of the higher V neuron but leaves the M and V neurons in the lower subnetwork unaffected (Fig.
+5cxxx).
 
+The stimulation of the PE neurons may cause both an increase in the baseline activity and a change in the neuron's gain.
+To disentangle the effect of elevated baseline activity and gain change, we illustrate each contribution separately by means of a mathematically tractable toy model (see Methods for more details).
+The variance estimated in the lower subnetwork increases for both an increase in the baseline and the gain of the PE neurons in the lower subnetwork.
+In contrast, when the gain of those PE neurons decreases, so does the variance.
+As we have seen before, changes in the baseline and the gain of PE neurons in the higher PE circuit do not affect the lower-level variance estimation.
+The variance estimated in the higher subnetwork changes equally for baseline increase or gain change of the PE neurons in the same subnetwork.
+However, in line with the perturbation experiments above, the higher-level variance estimation is affected by changes to the baseline or gain of the lower-level PE neurons as a result of the biased prediction.
+The variance increases slightly with increasing baseline or increasing deviation from a gain equal to 1 (Fig 5dxxx).
+Moreover, the mean of the sensory stimuli is overpredicted when the baseline or the gain for the pPE neuron increases, or when the gain of the nPE neuron decreases.
+In contrast, the mean of the sensory stimuli is underpredicted when the baseline or the gain of the nPE neuron increases, or when the gain of the pPE neuron decreases.
+In summary, any change in interneurons that cause an additional inhibitory input to the PE neurons reduces the gain of those neurons.
+Changes in the inhibitory neurons that cause an additional excitatory input to the PE neurons increases the baseline and potentially the gain of those neurons.
+
+This suggests that to understand the effect of neuromodulators on the sensory weight, we need to investigate the effect of interneuron activation on baseline and gain of PE neurons.
+To this end, we stimulated either PV, SOM, or VIP neurons independently for all three mean-field networks and measured the changes in baseline and gain of both PE neuron types (Fig.
+5exx).
+In all three networks tested, activating PV neurons decreases both quantities, leading to a decrease in the variance estimated (Fig.
+5exx \& Fig.
+5Sxxx).
+Activating SOM or VIP neurons decreases the overall gain but increases the baseline activity of the PE neurons.
+Whether and how much the gain of the nPE or pPE neuron is reduced depends on the inputs onto SOM/VIP neuron and the connectivity to the other neurons in the network.
+Similarly, how much the baseline is elevated depends on the specifics of the mean-field network (Fig.
+5exx \& Fig.
+5Sxxx).
+Hence, whether stimulating SOM or VIP neurons decreases or increases the activity of the V neurons depend on the input statistics.
+For low-mean stimuli, the effect of the elevated baseline dominates, while for high-mean stimuli the effect of the gain changes dominates the combined effect.
+
+XXX Summary ...
+effect of neuromodulators acting on INs on sensory weight best understood as combined effect on the baseline and gain of nPE and pPE neurons that in turn determine the activity of the V neuron.
+Those effects not ony depend on the target IN but also the network in which it is embedded, that is the connectivity and the inputs onto INs.
+Moreover, it depends on whether local or global.
+The effects of activation on lower is mainly the opposite of same in higher.
+However, the efefcts don't cancel each other in global modulation.
+Mostly because 1) V higher additionally changes due to changes in prediction but also 2) effects in higher are stronger !?
+and 3) 1/(1+(Vs + D)/(Vp + D)) ...
+that is it cannot cancel per se.
+Even though the underlying reasons for a change in the V neuron depends on the the network characteristics, there are some consistent changes in sensory weigt across MFN ...
+name
+
+% in caption: you need to make sure that in e you used a perturbationof 1 and the mean and std of stimuli were ...
+the lines only illustrate that it depends on the input statistics.
+Of course, the markers would also change with other perturbaton strengths, would they though?
+Think through
 % Mention that for V neuron (higher) the darker line is actually not horizontal (unaffected)!
@@
 
-\subsection*{XXX contraction bias}
+\subsection*{The contraction bias as a result of the weighting of sensory inputs and predictions thereof}
 % 
@@
 We solved the brain.
+
+% At the end, or in between talk about changes in PE neurons directly that you have currently in Fig. 3 SX ... I copy the part of the last paragraph here as inspiration :
+% For some psychiatric disorders, it has been shown that the weighting of sensory inputs and predictions thereof is impaired (XXX), leading to an overweighting of one of these signals.
+Moreover, it has been hypothesised that factors like stress or cognitive load may also influence the processing of feedforward and feedback inputs (XXX).
+...
+Another factor that may contribute to a distorted weighting is the baseline activity of PE neurons that was set to zero in our model, in line with very low spontaneous firing rates of excitatory neurons in rodent primary sensory areas (XXX).
+Increasing this baseline activity for the nPE neuron (Fig.
+SXXX), pPE neuron (Fig.
+SXXX) or both pushes the network to weight sensory stimuli and predictions more equally.
+We speculate that an increase of baseline activity may be a natural result of an increased cognitive load or stress.
+
+% mention Pakan paper
 

### 2023-07-07 scripts/plot_fig_3.py

@@ -48,6 +48,14 @@ G21 = gridspec.GridSpecFromSubplotSpec(1, 3, subplot_spec=G2[0,0], width_ratios=
 G3 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=G[2,0], width_ratios=[3,1], wspace = 0.3)
 G31 = gridspec.GridSpecFromSubplotSpec(1, 3, subplot_spec=G3[0,0], width_ratios=[0.7,1,1], wspace = 0.7)
 
+ax_A = fig.add_subplot(G1[0,0])
+ax_A.axis('off')
+ax_A.text(-0.45, 1.25, 'A', transform=ax_A.transAxes, fontsize=fs+1)
+
+ax_B = fig.add_subplot(G12[0,:])
+ax_B.axis('off')
+ax_B.text(-0.15, 1.25, 'B', transform=ax_B.transAxes, fontsize=fs+1)
+
 ax_B1 = fig.add_subplot(G12[0,0])
 ax_B1.set_title('Neurons activity increases \nwith trial variability', fontsize=fs, pad=10)
 ax_B2 = fig.add_subplot(G12[0,1])
@@ -55,19 +63,27 @@ ax_B2.set_title('Neuron activity increases \nwith stimulus variability', fontsiz
 ax_B3 = fig.add_subplot(G12[0,2])
 plt.setp(ax_B2.get_yticklabels(), visible=False)
 
-#ax_D1 = fig.add_subplot(G21[0,0])
+ax_C = fig.add_subplot(G21[0,:])
+ax_C.axis('off')
+ax_C.text(-0.15, 1.25, 'C', transform=ax_C.transAxes, fontsize=fs+1)
+
 ax_D2 = fig.add_subplot(G21[0,1])
 ax_D2.set_title('Example in which the network relies more strongly on prediction', fontsize=fs, pad=10)
 ax_D3 = fig.add_subplot(G21[0,2])
 ax_D4 = fig.add_subplot(G2[0,1])
 ax_D4.set_title('Sensory weight for different \ntrial and stimulus variabilities', fontsize=fs, pad=10)
+ax_D4.text(-0.4, 1.25, 'E', transform=ax_D4.transAxes, fontsize=fs+1)
 
-#ax_E1 = fig.add_subplot(G31[0,0])
+ax_D = fig.add_subplot(G31[0,:])
+ax_D.axis('off')
+ax_D.text(-0.15, 1.25, 'D', transform=ax_D.transAxes, fontsize=fs+1)
+
[…]

### 2023-07-07 scripts/plot_fig_3_support_1.py

@@ -6,5 +6,127 @@ Created on Mon Jan  9 09:04:01 2023
 @author: loreen.hertaeg
 """
 
-# %% Import
+# %% import
 
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import os.path
+import seaborn as sns
+
+from matplotlib.collections import LineCollection
+
+from src.plot_data import plot_fraction_sensory_heatmap, plot_transitions_examples
+
+# %% Universal parameters
+
+fs = 6
+inch = 2.54
+
+
+# %% Define files and paths
+
+figure_name = 'Fig_3_S1.png'
+figPath = '../results/figures/final/'
+
+if not os.path.exists(figPath):
+    os.mkdir(figPath)
+
+
+# %% Define figure structure
+
+figsize=(18/inch,3/inch)
+fig = plt.figure(figsize=figsize)
+
+G = gridspec.GridSpec(1, 2, figure=fig, wspace=0.4, width_ratios=[1,4])
+G1 = gridspec.GridSpecFromSubplotSpec(1, 1, subplot_spec=G[0,0])
+G2 = gridspec.GridSpecFromSubplotSpec(1, 4, subplot_spec=G[0,1], wspace=0.3)
+
+ax_A = fig.add_subplot(G1[0,0])
+ax_A.text(-0.4, 1.37, 'A', transform=ax_A.transAxes, fontsize=fs+1)
+
+ax_B = fig.add_subplot(G2[0,:])
+ax_B.axis('off')
+ax_B.text(-0.1, 1.2, 'B', transform=ax_B.transAxes, fontsize=fs+1)
+
+ax_B1 = fig.add_subplot(G2[0,0])
+ax_B1.set_title(r'1 $\longrightarrow$ 2', fontsize=fs, pad=10)
+ax_B2 = fig.add_subplot(G2[0,1])
+plt.setp(ax_B2.get_yticklabels(), visible=False)
+ax_B2.set_title(r'2 $\longrightarrow$ 3', fontsize=fs, pad=10)
+ax_B3 = fig.add_subplot(G2[0,2])
+plt.setp(ax_B3.get_yticklabels(), visible=False)
+ax_B3.set_title(r'3 $\longrightarrow$ 4', fontsize=fs, pad=10)
+ax_B4 = fig.add_subplot(G2[0,3])
+plt.setp(ax_B4.get_yticklabels(), visible=False)
[…]

### 2023-07-07 scripts/plot_fig_4.py

@@ -33,6 +33,7 @@ from matplotlib.collections import LineCollection
 
 from src.plot_data import plot_neuromod_impact_inter, illustrate_sensory_weight_variance, plot_changes_upon_input2PE_neurons
 from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_influence_interneurons_gain_baseline
+from src.plot_data import plot_legend_illustrations
 
 # %% Universal parameters
 
@@ -54,43 +55,54 @@ if not os.path.exists(figPath):
 figsize=(18/inch,18/inch)
 fig = plt.figure(figsize=figsize)
 
-G = gridspec.GridSpec(2, 1, figure=fig, hspace=0.3, height_ratios=[1,2])
+G = gridspec.GridSpec(2, 1, figure=fig, hspace=0.6, height_ratios=[1.2,2])
 
-G1 = gridspec.GridSpecFromSubplotSpec(2, 4, subplot_spec=G[0,0])#, wspace=0.3, width_ratios=[1,3])
-G2 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=G[1,0], width_ratios=[1,3], wspace=0.4)
+G1 = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=G[0,0], width_ratios=[3,1], hspace=0.5)
+G2 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=G[1,0], width_ratios=[2.5,1], wspace=0.4)
 
-ax_A = fig.add_subplot(G1[:,:])
+G11 = gridspec.GridSpecFromSubplotSpec(2,3, subplot_spec=G1[:,0])#, height_ratios=[1,3], hspace=0.7)
+
+ax_A = fig.add_subplot(G11[:,:])
 ax_A.axis('off')
-ax_A.set_title('Neuromodulators acting through activating interneurons can bias the weighting of sensory inputs and predictions', fontsize=fs, pad=10)
+ax_A.set_title('Neuromodulators acting on interneurons can bias the weighting of sensory inputs and predictions', fontsize=fs, pad=10)
[…]

### 2023-07-07 scripts/plot_fig_4_support_1.py

@@ -1,0 +1,116 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jan  9 09:04:01 2023
+
+@author: loreen.hertaeg
+"""
+
+# %% Import
+
+import numpy as np
+import pickle
+
+from src.functions_simulate import simulate_neuromod
+from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs
+
+import warnings
+warnings.filterwarnings("ignore")
+
+dtype = np.float32
+
+
+# %% import
+
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import os.path
+import seaborn as sns
+
+from matplotlib.collections import LineCollection
+
+from src.plot_data import plot_neuromod_impact_inter, illustrate_sensory_weight_variance, plot_changes_upon_input2PE_neurons
+from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_influence_interneurons_gain_baseline
+from src.plot_data import plot_legend_illustrations
+
+# %% Universal parameters
+
+fs = 6
+inch = 2.54
+
+
+# %% Define files and paths
+
+figure_name = 'Fig_4_S1.png'
+figPath = '../results/figures/final/'
+
+if not os.path.exists(figPath):
+    os.mkdir(figPath)
+
+
+# %% Define figure structure
+
+figsize=(18/inch,14/inch)
+fig = plt.figure(figsize=figsize)
+
+G = gridspec.GridSpec(2, 1, figure=fig, hspace=0.6)
+
+G1 = gridspec.GridSpecFromSubplotSpec(2, 3, subplot_spec=G[0,0], hspace=0.3)
+G2 = gridspec.GridSpecFromSubplotSpec(2, 3, subplot_spec=G[1,0], hspace=0.3)
+
+ax_A = fig.add_subplot(G1[:,:])
+ax_A.axis('off')
+ax_A.set_title('Neuromodulators acting on interneurons in lower PE circuit', fontsize=fs, pad=10)
+ax_A.text(-0.1, 1.1, 'A', transform=ax_A.transAxes, fontsize=fs+1)
+
[…]

### 2023-07-07 scripts/plot_fig_4_support_2.py

@@ -1,0 +1,70 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jan  9 09:04:01 2023
+
+@author: loreen.hertaeg
+"""
+
+# %% Import
+
+import numpy as np
+import pickle
+
+from src.functions_simulate import simulate_neuromod
+from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs
+
+import warnings
+warnings.filterwarnings("ignore")
+
+dtype = np.float32
+
+
+# %% import
+
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import os.path
+import seaborn as sns
+
+from matplotlib.collections import LineCollection
+
+from src.plot_data import plot_influence_interneurons_baseline_or_gain
+
+# %% Universal parameters
+
+fs = 6
+inch = 2.54
+
+
+# %% Define files and paths
+
+figure_name = 'Fig_4_S2.png'
+figPath = '../results/figures/final/'
+
+if not os.path.exists(figPath):
+    os.mkdir(figPath)
+
+
+# %% Define figure structure
+
+figsize=(9/inch,5/inch)
+fig = plt.figure(figsize=figsize)
+
+G = gridspec.GridSpec(1, 2, figure=fig, wspace=0.7)
+
+ax_A = fig.add_subplot(G[0,0])
+ax_A.text(-0.4, 1.1, 'A', transform=ax_A.transAxes, fontsize=fs+1)
+ax_B = fig.add_subplot(G[0,1])
+ax_B.text(-0.4, 1.1, 'B', transform=ax_B.transAxes, fontsize=fs+1)
+
+# %% How do the interneurons influence the PE neuron properties?
+
+plot_influence_interneurons_baseline_or_gain(ax=ax_A, plot_annotation=False)
+plot_influence_interneurons_baseline_or_gain(plot_baseline=False, ax=ax_B)
+
+# %% save figure
+
+plt.savefig(figPath + figure_name, bbox_inches='tight', transparent=True, dpi=600)

### 2023-07-07 scripts/run_fig_3_support_1.py

@@ -11,8 +11,9 @@ Created on Mon Jan  9 09:04:01 2023
 import numpy as np
 import pickle
 
-from src.functions_simulate import simulate_sensory_weight_time_course, simulate_impact_para
-from src.plot_data import plot_weight_over_trial, plot_impact_para, plot_schema_inputs_tested
+from src.functions_simulate import simulate_weighting_example, simulate_weighting_exploration, simulate_dynamic_weighting_eg
+from src.functions_simulate import simulate_activity_neurons
+from src.plot_data import plot_weighting_limit_case_example, plot_fraction_sensory_heatmap, plot_transitions_examples, plot_neuron_activity_lower_higher
 
 import warnings
 warnings.filterwarnings("ignore")
@@ -20,334 +21,11 @@ warnings.filterwarnings("ignore")
 dtype = np.float32
 
 
-# %% Simulate control case for number of input statistics 
+# %% Examples: system dynamically adjusts to input statistics and environement
+
+# 5,5,0,0 --> 5,5,0,3 --> 0,10,0,3 --> 0,10,0,0 --> 5,5,0,0
 
 run_cell = True
-load_only = True
-do_plot = True
-
-# define parameters used throughout the file
-mean_trials, m_sd = dtype(5), dtype(0)
-variability_within = np.array([0, 0.75, 1.5, 2.25, 3])
-variability_across = np.array([3, 2.25, 1.5, 0.75, 0])
-last_n = np.int32(50)
-n_trials = np.int32(200)
-
-if run_cell:
-    
-    ### choose mean-field network to simulate
-    mfn_flag = '10' # valid options are '10', '01', '11
-    
-    ### filename for data
-    file_for_data = '../results/data/weighting/data_weighting_control.pickle'
-    
-    ### get data
-    if not load_only: # simulate respective network
-        
-        [stimuli, n, gain_w_PE_to_P,
-         gain_v_PE_to_P, add_input, 
-         id_cells_modulated, weight] = simulate_impact_para(mfn_flag, variability_within, variability_across, mean_trials, 
-                                        m_sd, last_n = last_n, n_trials = n_trials, file_for_data = file_for_data)
-        
-    else:
-        
-        with open(file_for_data,'rb') as f:
-            [stimuli, n_ctrl, gain_w_PE_to_P_ctrl, gain_v_PE_to_P_ctrl, 
-             add_input_ctrl, id_cells_modulated_ctrl, weight_ctrl] = pickle.load(f)
[…]

### 2023-07-07 scripts/run_fig_3_support_2.py

@@ -1,0 +1,389 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jul  3 18:45:22 2023
+
+@author: loreen.hertaeg
+"""
+
+# %% Import
+
+import numpy as np
+import pickle
+
+from src.functions_simulate import simulate_sensory_weight_time_course, simulate_impact_para
+from src.plot_data import plot_weight_over_trial, plot_impact_para, plot_schema_inputs_tested
+
+import warnings
+warnings.filterwarnings("ignore")
+
+dtype = np.float32
+
+
+# %% Simulate control case for number of input statistics 
+
+run_cell = True
+load_only = True
+do_plot = True
+
+# define parameters used throughout the file
+mean_trials, m_sd = dtype(5), dtype(0)
+variability_within = np.array([0, 0.75, 1.5, 2.25, 3])
+variability_across = np.array([3, 2.25, 1.5, 0.75, 0])
+last_n = np.int32(50)
+n_trials = np.int32(200)
+
+if run_cell:
+    
+    ### choose mean-field network to simulate
+    mfn_flag = '10' # valid options are '10', '01', '11
+    
+    ### filename for data
+    file_for_data = '../results/data/weighting/data_weighting_control.pickle'
+    
+    ### get data
+    if not load_only: # simulate respective network
+        
+        [stimuli, n, gain_w_PE_to_P,
+         gain_v_PE_to_P, add_input, 
+         id_cells_modulated, weight] = simulate_impact_para(mfn_flag, variability_within, variability_across, mean_trials, 
+                                        m_sd, last_n = last_n, n_trials = n_trials, file_for_data = file_for_data)
+        
+    else:
+        
+        with open(file_for_data,'rb') as f:
+            [stimuli, n_ctrl, gain_w_PE_to_P_ctrl, gain_v_PE_to_P_ctrl, 
+             add_input_ctrl, id_cells_modulated_ctrl, weight_ctrl] = pickle.load(f)
+            
+    if do_plot:
+        
+        plot_schema_inputs_tested()
+        
+
+# %% Impact of activation function
+
+run_cell = False
+plot_only = True
+
+if run_cell:
+
+    ### choose mean-field network to simulate
+    mfn_flag = '10' # valid options are '10', '01', '11
+    
+    ### filename for data
+    file_for_data = '../results/data/weighting/data_weighting_activation_function.pickle'
+    
+    ### get data
[…]

### 2023-07-07 src/plot_data.py

@@ -15,6 +15,7 @@ import seaborn as sns
 import pandas as pd
 import pickle
 
+from matplotlib.legend_handler import HandlerTuple
 import matplotlib.gridspec as gridspec
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
 from matplotlib.colors import LinearSegmentedColormap, ListedColormap
@@ -26,23 +27,27 @@ from src.functions_simulate import random_binary_from_moments, random_binary_fro
 # %% set colors
 
 Col_Rate_E = '#9E3039'
-Col_Rate_nE = '#955F89'
+Col_Rate_nE = '#A06A94' #'#955F89'
 Col_Rate_nD = '#BF9BB8' 
-Col_Rate_pE = '#CB9173' 
+Col_Rate_pE = '#9E3039' # '#CB9173' 
 Col_Rate_pD = '#DEB8A6' 
 Col_Rate_PVv = '#508FCE' 
 Col_Rate_PVm = '#2B6299' 
-Col_Rate_SOM = '#79AFB9' 
-Col_Rate_VIP = '#39656D' 
+Col_Rate_SOM = '#285C8F' #'#79AFB9' 
+Col_Rate_VIP = '#163350' #'#39656D' 
 
 color_sensory = '#D76A03'
-color_prediction = '#19535F'
+color_prediction = '#39656D' #'#19535F'
 
 ####
-color_stimuli_background = '#FED5AF'
+color_stimuli_background = '#FEDFC2'
 color_running_average_stimuli = '#D76A03'
-color_m_neuron = '#19535F'
-color_v_neuron = '#452144'
+color_m_neuron = '#39656D' #'#19535F'
+color_m_neuron_light = '#83c5be' #'#79AFB9'
+color_m_neuron_dark = '#2A4A50'
+color_v_neuron = '#754E06' #'#452144'
+color_v_neuron_light = '#b08968'#'#E99B0C'
+color_v_neuron_dark = '#3A2703'
 color_mse = '#AF1B3F' 
 color_weighted_output = '#5E0035'
 color_mean_prediction = '#70A9A1' 
@@ -73,6 +78,102 @@ def lighten_color(color, amount=0.5):
     c = colorsys.rgb_to_hls(*mc.to_rgb(c))
     return colorsys.hls_to_rgb(c[0], 1 - amount * (1 - c[1]), c[2])
 
+
+def plot_legend_illustrations(ax, fs=6, mew=2):
+    
[…]

### 2023-07-13 records/manuscript_2023_v0.tex
First draft with Intro, Results, Figures and caption

%\documentclass[10pt,a4paper,draft]{article}
\documentclass[10pt,a4paper]{article}

%\usepackage[top=3cm, bottom=0cm, left=3.5cm,right=2cm]{geometry}
\usepackage[top=1in, left=1in ,right=1in, bottom=1in, footskip=0in, marginparwidth=0in]{geometry}

% use Unicode characters - try changing the option if you run into troubles with special characters (e.g. umlauts)
\usepackage[utf8]{inputenc}

\usepackage{fancyhdr}

% clean citations
%\usepackage{cite}
%\usepackage[super,sort&compress,comma]{natbib}
\usepackage[numbers, round, sort&compress, comma]{natbib}
%\usepackage[round]{natbib}

% hyperref makes references clicky. use \url{www.example.com} or \href{www.example.com}{description} to add a clicky url
%\usepackage{nameref,hyperref}

% math
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{bbold}

% line numbers
\usepackage[right]{lineno}

% improves typesetting in LaTeX
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }
\usepackage{enumitem}

% text layout - change as needed
%\raggedright
%\setlength{\parindent}{0.5cm}
%\textwidth 5.25in 
%\textheight 8.75in

% Remove % for double line spacing
%\usepackage{setspace} 
%\doublespacing

% adjust caption style
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,singlelinecheck=off]{caption}

% remove brackets from references
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% use \textcolor{color}{text} for colored text (e.g. highlight to-do areas)
\usepackage{color}

% define custom colors (this one is for figure captions)
\definecolor{Gray}{gray}{.25}

% this is required to include graphics
\usepackage{graphicx}
\usepackage{sidecap}

% hyperlinks
\usepackage{hyperref}

% ########################################################

%\pagestyle{headings}
\pagestyle{myheadings}
\markright{}

\hyphenation{math-e-mat-i-cal-ly} 
\hyphenation{inter-neurons}

\begin{document}

\thispagestyle{empty}

% title goes here:
\begin{flushleft}
{\Large
\textbf\newline{Title}
}
%\newline
%% authors go here:
%\\
%Loreen Hert\"ag\textsuperscript{1,*},
%Claudia Clopath\textsuperscript{1,§}
%\\
%\bigskip
%\bf{1} Bioengineering Department, Imperial College London, London, UK.
%\\
%\bigskip
%* l.hertag@imperial.ac.uk, § c.clopath@imperial.ac.uk

\end{flushleft}

% now start line numbers
%\linenumbers

\begin{abstract}
blahhh blahhh blah
\end{abstract}

\section*{Introduction}
%
To survive in an ever-changing environment, animals must flexibly adapt their behavior based on previously encoded and novel information. This context-dependency must also be reflected in the information processing of neural networks underlying intelligent behavior. For instance, when you walk down some staircase in your fully lit basement, your brain might entirely rely on the feedforward (bottom-up) input your senses receive (Fig. \ref{fig:Fig_1}, left). In contrast, when you walk down the same stairs in complete darkness, your brain might rely entirely on feedback (top-down) signals generated from a staircase model it has formed over previous experiences (Fig. \ref{fig:Fig_1}, middle).

The importance of these feedback inputs has been emphasized by observations showing that top-down projections outnumber feedforward connections (XXX) and that they modulate (XXX) or even entirely drive (XXX) neuron activity. But how do neural networks switch between a feedforward-dominated and a feedback-dominated processing mode? And how do neural networks combine both input streams wisely? For instance, if you hike down an unexplored mountain in very foggy conditions, your brain receives unreliable visual information. In addition, it can only draw on a shaky prediction about what to expect (Fig. \ref{fig:Fig_1}, right).

A common hypothesis is that the brain weights different inputs according to their reliabilities. A prominent example of this hypothesis is Bayesian multisensory integration (XXX). According to this theory, neural networks represent information from multiple modalities by a linear combination of the uncertainty-weighted single-modality estimates. Multisensory integration is supported by several observations showing that xxx (XXX). It is conceivable that the same concepts can be employed for the weighting of sensory inputs and predictions thereof (XXX). A central point in the weighting of inputs is the estimation of variances as a measure of uncertainty. However, how the variance of both the sensory input and the prediction can be computed on the circuit level is not resolved yet.

We hypothesised that prediction error (PE) neurons provide the basis for the neural computation of variances. PEs are an integral part of the theory of predictive processing which states that the brain constantly compares incoming sensory information with predictions. When those predictions are wrong, the resulting PEs allow the network to revise the model of the world, thereby ensuring that the predictions are more accurate (XXX). Experimental evidence suggests that these PEs may be represented in the activity of distinct groups of neurons, termed PE neurons (XXX). Moreover, these neurons may come in two types when excitatory neurons exhibit near-zero, spontaneous firing rates (XXX). Negative PE (nPE) neurons mainly increase their activity when the prediction is \textit{stronger} than the sensory input, while positive PE (pPE) neurons mainly increase their activity when the prediction is \textit{weaker} than the sensory input. Indeed, it has been shown that excitatory neurons in layer 2/3 of rodent primary sensory areas can encode negative or positive PEs (XXX).

Here, we show that the unique response patterns of nPE and pPE neurons may provide the backbone for computing both the mean and the variance of sensory stimuli. Furthermore, we suggest a network model with a hierarchy of PE circuits to estimate the variance of the prediction, in addition to the variance of the sensory inputs. In line with multisensory integration, predictions are weighted more strongly than the sensory stimuli when the environment is stable (that is, predictable) but the sensory inputs are noisy. Moreover, we show that predictions are integrated more strongly after a change in the environment, even when the new sensory stimulus is reliable. In addition, we unravel the mechanisms underlying a neuromodulator-induced shift in the weighting of sensory inputs and predictions. In our model, these neuromodulators activate groups of inhibitory neurons like parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (XXX). In a computational model, these interneurons have been shown to establish a multi-pathway balance of excitation and inhibition that is the basis for nPE and pPE neurons (XXX). By breaking this balance, the excitatory neurons change their baseline firing rate and gain, leading to a biased variance estimation. Finally, we show that this weighting can be understood as the neural manifestation of the contraction bias (XXX).

%
\begin{figure}[t!]
	\centering
    \includegraphics{../results/figures/final/Fig_1}
\caption{\footnotesize{\bf Neural network model to flexibly integrate sensory information and predictions.\newline} 
{\bf (A)} Example illustration of context-dependent integration of information. Left: When walking down a staircase that is clearly visible, the brain might rely solely on external sensory information. Middle: When walking down the same stairs in the absence of visual information, the brain might rely on predictions formed by previous experience. Right: When climbing down an unexplored mountain in foggy conditions, the brain might need to integrate sensory information and predictions at the same time.
{\bf (B)} Illustration of a prediction-error (PE) circuit with both negative and positive PE (nPE/pPE) neurons that receive inhibition from three different inhibitory interneuron types: parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons. Local excitatory connections are not shown for clarity.
{\bf (C)} Illustration of network model that estimates the mean and variance of the external sensory stimuli. The core of this network model is the PE circuit shown in (B). The lower-level V neuron encodes the variance, while the lower-level M neuron encodes the mean of the sensory input.
{\bf (D)} Same as in (C) but the feedforward input is the activity of the lower-level M neuron.
}
\label{fig:Fig_1}
\end{figure}
%

\section*{Results}
%

\subsection*{nPE and pPE neurons as the basis for estimating mean and variance of sensory stimuli}
%
We hypothesise that the distinct response patterns of negative and positive prediction-error (nPE/pPE) neurons represent the backbone for estimating the mean and the variance of sensory stimuli. nPE neurons only increase their activity relative to a baseline when the sensory input is weaker than predicted, while pPE neurons only increase their activity relative to a baseline when the sensory input is stronger than predicted. Moreover, both nPE and pPE neurons remain at their baseline activity when the sensory input is fully predicted (XXX). Assuming that the prediction equals the mean of the sensory stimulus, the PE neurons, hence, encode the deviation from the mean. Thus, the squared sum of nPE and pPE neuron activity represents the variance of the feedforward input. 
%
\begin{figure}[t!]
	\centering
	%\makebox[\textwidth][c]{\includegraphics{../results/figures/final/Fig_2}}
    \includegraphics[width=1\linewidth]{../results/figures/final/Fig_2}
\caption{\footnotesize{\bf Prediction-error neurons provide the basis for estimating mean and variance of sensory stimuli.\newline} 
{\bf (A)} Illustration of the inputs with which the network shown in \ref{fig:Fig_1}C is stimulated. Network is exposed to a sequence of constant stimuli drawn from a uniform distribution. Stimulus duration is XXX.
{\bf (B)} PE neuron activity hardly changes with stimulus strength (left) but strongly increases with stimulus variability (right).
{\bf (C)} Interneuron activity strongly changes with stimulus strength (left) but hardly changes with stimulus variability (right) .
{\bf (D)} M neuron correctly encodes the mean of the sensory stimuli. Left: Illustration of the input synapses onto the M neuron. Middle: Activity of the M neuron over time for a uniform distribution with mean XXX and standard deviation XXX. Right: Normalised mean-squared error (MSE) between the running average and the M neuron activity for different parametrizations of the stimulus distribution.
{\bf (E)} V neuron correctly encodes the variance of the sensory stimuli. Left: Illustration of the input synapses onto the V neuron. Middle: Activity of the V neuron over time for a uniform distribution with mean XXX and standard deviation XXX. Right: Normalised mean-squared error (MSE) between the instantaneous variance and the V neuron activity for different parametrizations of the stimulus distribution.
}
\label{fig:Fig_2}
\end{figure}
%

To test our hypothesis, we studied a rate-based mean-field network the core of which is a prediction-error (PE) circuit with excitatory nPE and pPE neurons, as well as inhibitory parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (Fig. \ref{fig:Fig_1}B). While the excitatory neurons are simulated as two coupled point compartments to emulate the soma and dendrites of elongated pyramidal cells, respectively, all inhibitory cell types were modeled as point neurons. The connectivity of and inputs to the network were chosen such that the excitatory (E) and inhibitory (I) pathways onto the pyramidal cells were balanced because it has been shown that this E/I balance is necessary for nPE and pPE neurons to emerge (XXX, see Methods).

In addition to this core circuit, we model a memory (M) neuron that perfectly integrates the activity of the PE neurons (Fig. \ref{fig:Fig_1}C). In accordance with XXX, we assume that the pPE neuron excites the memory neuron, while the nPE neuron inhibits this neuron (for instance, through lateral inhibition, here not modeled explicitly). The M neuron connects to the apical dendrites of the PE neurons and some of the interneurons (here, VIP and PV neurons, see Methods for more details). In this network, the M neuron serves as a prediction that is dynamically updated when new sensory information is available. We furthermore simulate a downstream neuron (termed V neuron), modeled as a leaky integrator with a squared activation function, that receives excitatory output synapses from the PE neurons. Hence, in this setting, the V neuron encodes the variance of the sensory stimuli (Fig. \ref{fig:Fig_1}C).

To show that this network can indeed represent mean and variance in the respective neurons, we stimulate it with a sequence of step-wise constant inputs drawn from a uniform distribution (Fig. \ref{fig:Fig_2}A), assuming that the sensory stimulus varies over time. In line with the distinct response patterns for nPE and pPE neurons, these neurons change only slightly with increasing stimulus mean but increase strongly with input variance (Fig. \ref{fig:Fig_2}B). This is in contrast to the three interneurons that strongly increase with stimulus mean while they only moderately increase with stimulus variance (Fig. \ref{fig:Fig_2}C). The activity of the memory neuron M gradually approaches the mean of the sensory inputs (Fig. \ref{fig:Fig_2}D, middle), while the activity of the V neuron approaches the variance of the inputs (Fig. \ref{fig:Fig_2}E, middle). This is true for a wide range of input statistics (Fig. \ref{fig:Fig_2}D-E, right) and input distributions (Fig. \ref{fig:Fig_2_S1}). Small deviations from the true mean occur mainly for larger input variances, while the estimated variance is fairly independent of the input statistics tested.

XXX coming soon: Paragraph on network beyond mean-field XXX

In summary, nPE and pPE neurons can be the basis to estimate the mean and the variance of sensory stimuli that vary over time.

\subsection*{Estimating variances of sensory inputs and predictions requires a hierarchy of PE circuits}
%
Following the ideas of Bayesian multisensory integration (XXX), the weighting of sensory stimuli and predictions thereof would require knowledge of their variances. As we have shown in the previous section, the variance of the sensory stimulus can be estimated using PE neurons. We hypothesise that the same principles apply to computing the variance of the prediction. Hence, we augment the network with a \textit{higher} PE circuit that receives output synapses from the M neuron of the \textit{lower} PE circuit (Fig. \ref{fig:Fig_1}D). Both subnetworks are modeled the same, except that the M neuron in the higher PE circuit evolves more slowly than the one in the lower PE circuit. 
%
\begin{figure}[t!]
	\centering
	%\makebox[\textwidth][c]{\includegraphics{../results/figures/final/Fig_3}}
    \includegraphics[width=1\linewidth]{../results/figures/final/Fig_3}
\caption{\footnotesize{\bf Estimating variances of sensory inputs and predictions with hierarchical PE circuits.\newline} 
{\bf (A)} Illustration of the stimulation protocol. Network is exposed to a sequence of stimuli (one stimulus per trial). To account for stimulus variability, each stimulus is represented by xxx stimulus values drawn from a normal distribution with mean $\mu_\mathrm{stim}$ and $\sigma_\mathrm{stim}^2$. To account for the volatility of the environment, in each trial the stimulus mean $\mu_\mathrm{stim}$ is drawn from a uniform distribution (denoted trial variability). Trial duration = xxx. 
{\bf (B)} Neuron activity increases with both stimulus and trial variability. Neurons in the lower PE circuit increase more strongly with stimulus variability. Neurons in the higher PE circuit increase more strongly with trial variability.
{\bf (C)} Limit case example in which the stimulus variability is low but the trial variability is high. Left: Illustration of the stimulation protocol. Middle: Weighted output follows closely the sensory stimuli. Right: Sensory weight (function of the variances, see text) close to 1, indicating that the network ignores the prediction. Input statistics: XXX.
{\bf (D)} Limit case example in which the stimulus variability is high but the trial variability is low. Left: Illustration of the stimulation protocol. Middle: Weighted output pushed towards the mean of the sensory stimuli. Right: Sensory weight close to zero, indicating that the network ignores the sensory stimuli. Input statistics: XXX.
{\bf (E)} Predictions are weighted more strongly when the stimulus variability is larger than the trial variability.
{\bf (F)} Predictions are weighted more strongly at the beginning of a new trial and quickly changing stimuli.
}
\label{fig:Fig_3}
\end{figure}
%

To test the network's ability to estimate the variances correctly, we stimulated the network with a sequence of inputs. In each trial one stimulus is shown to the network. To account for the stimulus variance, each stimulus is composed of \textit{n} constant values drawn from a normal distribution with mean $\mu_\mathrm{stim}$ and variance $\sigma_\mathrm{stim}^2$, and presented one after the other. To account for potential changes in the environment, in each trial, we draw $\mu_\mathrm{stim}$ from a uniform distribution (Fig. \ref{fig:Fig_3}A). Hence, the inputs change on two different time scales, with stimulus variability (faster time scale) and trial variability (slower time scale).

As expected, the neurons' activity increase for both stimulus and trial variances (Fig. \ref{fig:Fig_3}B). While the neurons in the lower PE circuit increase more strongly with stimulus variability, the neurons in the higher PE circuit increase more strongly with trial variability, indicating that the different subnetworks process different aspects of the inputs. We first consider two limit cases. In the first limit case, a different but low-variance stimulus is presented in each trial (Fig. \ref{fig:Fig_3}C, left). In line with the ideas of multisensory integration (XXX), the network should therefore follow the sensory inputs closely and ignore the predictions. When we arithmetically calculate the weighted output (Fig. \ref{fig:Fig_3}C, middle) based on the feedforward and feedback inputs, and the sensory weight (Fig. \ref{fig:Fig_3}C, right), the network correctly represents mostly the sensory input (for more details, see Methods). In the second limit case, the same but high-variance stimulus is presented in each trial (Fig. \ref{fig:Fig_3}D, left). According to the theory, the network should downscale the sensory feedforward input and weight the prediction more strongly. Indeed, the weighted output of the network shows a clear tendency to the mean of the stimuli (Fig. \ref{fig:Fig_3}D, middle), also reflected in the low sensory weight (Fig. \ref{fig:Fig_3}D, right).

In a next step, to validate the network responses more broadly, we systematically varied the trial and stimulus variability independently. If both variances are similar, the sensory weight approaches \textit{0.5}, reflecting equal contribution of sensory inputs and predictions to the weighted output. Only if both variances are zero, the network represents the sensory input perfectly. In line with the limit case examples above, if the stimulus variance is larger than the trial variance, the network weights the prediction more strongly than the sensory input. This is reversed if the stimulus variance is smaller than the trial variance (Fig. \ref{fig:Fig_3}E). Because the network dynamically estimates the mean and variances of the sensory input and the prediction, the weighted output and the sensory weight changes accordingly when the input statistics changes (Fig. \ref{fig:Fig_3_S1}).

The first limit case (Fig. \ref{fig:Fig_3}C) shows that even in a sensory-driven input regime, the prediction is weighted more at the beginning of a new trial than in the steady state. This is further confirmed in simulations in which the trail duration was shortened. For those simulations, the prediction even outweighs the sensory input, reflected in a very low sensory weight (Fig. \ref{fig:Fig_3}F). This suggests that predictions influence neural activity more significantly in experiments that rely on very fast stimulus changes.

It has been shown [speculated?], that sensory inputs or predictions are overrated in some psychiatric disorders (XXX). We thus wondered which network properties might bias the estimation of the variances, and, consequently, the weighting of different input streams. In our network, the M neuron evolves faster in the lower subnetwork than in the higher one. We identified the speed at which the M neurons are updated with new information as a decisive factor in the integration of inputs. To show this, we varied the weights from the PE neurons onto the lower-level M neuron. If the M neuron evolves too slowly, the prediction is overrated. In contrast, if the M neuron incorporates new information too quickly, the sensory input is overrated (Fig. \ref{fig:Fig_3_S2}A). While the speeds at which the activity of the M neurons evolve may underlie pathological weighting of inputs, the precise activation function of the V neurons is less pivotal. When we replaced the squared activation function with a linear, rectified function, the V neurons do not encode the variance but the averaged absolute deviation of the sensory stimuli. However, the sensory weight is only slightly shifted to larger values for low trial/high stimulus variability (Fig. \ref{fig:Fig_3_S2}B).

In summary, we show that the variances of both the sensory inputs and predictions thereof can be dynamically computed in networks comprising a lower and higher PE circuit. The model shows that predictions are trusted more strongly at the beginning of a new stimulus, and if sensory inputs are noisy on a short time scale while predictable on longer time scales.

% XXX Do we actually propose that predictions are sent up the hierarchy?

\subsection*{Biasing the weighting of sensory inputs and predictions by neuromodulators}
%
The brain's flexibility and adaptability are not least because a plethora of neuromodulators influence the activity of neurons in a variety of ways (XXX). A prominent target of neuromodulatory inputs is inhibitory neurons (Cardin 2019, XXX). Moreover, distinct interneuron types are differently (in-)activated by those neuromodulators. For instance, it has been shown that XXX (XXX). We, therefore, wondered if and how the weighting of sensory inputs and predictions thereof may be biased when neuromodulators activate distinct interneuron types.
%
\begin{figure}[t!]
	\centering
	\makebox[\textwidth][c]{\includegraphics{../results/figures/final/Fig_4}}
    %\includegraphics{../results/figures/final/Fig_4} % [width=1\linewidth]
\caption{\footnotesize{\bf Neuromodulator-based shifts in the weighting of sensory inputs and predictions.
\newline} 
{\bf (A)} Neuromodulators acting on the three interneurons may shift the weighting of sensory inputs and predictions. The changes depend on the type/s of interneurons targeted by the additional excitatory input emulating a neuromodulator (additional input = XXX). Considered are two limit cases (upper row: more sensory-driven before modulation, lower row: more prediction-driven before modulation). The combination of interneurons targeted is illustrated below. The results are shown for three different PE circuits, specified in the main text. 
{\bf (B)} Illustration showing how the sensory weight depends on changes in both the stimulus and trial variability.
{\bf (C)} The M and V neuron activities depend on the PE neuron activities. Hence, perturbing the nPE and pPE neurons must change the estimation of mean and variance. While stimulating the lower PE neurons affects both the lower and higher mean and variance estimation, stimulating the higher PE neurons only affects the V and M neurons in the same subnetwork.
{\bf (D)} Illustration of the mechanisms underlying the biased estimation of mean and variance when PE neurons are perturbed. Both changes in the baseline (left) and gain (right) of PE neurons can contribute to the changes observed in (C). Illustration based on toy model described in Methods.
{\bf (E)} Modulated interneurons change the weighting by changing the overall baseline and the overall gain of PE neurons (sum of changes in nPE and pPE neurons). Whether and how a neuromodulator changes the sensory weight, hence, depends on the interneuron targeted and the effect this interneuron has on baseline and gain of the PE neurons, which in turn does depend on the network it is embedded in.
}
\label{fig:Fig_4}
\end{figure}
%

To this end, we modeled the presence of neuromodulators by injecting an additional excitatory input into one or two interneuron types. We reasoned that the network effect of a neuromodulator not only depends on the interneuron type it targets but also on the inputs this neuron receives and the connections it makes with other neurons in the network. We, therefore, tested three different mean-field networks that differ with respect to the distribution of sensory inputs and predictions onto the interneurons, and the underlying connectivity. The commonality across those networks is that they exhibit an E/I balance of excitatory and inhibitory pathways onto the PE neurons (XXX). Across the different mean-field networks tested, activating a SOM or VIP neuron individually forces the networks to weigh both inputs more equally. As a consequence, predictions are overrated in a sensory-driven input regime. Similarly, sensory inputs are overrated in a prediction-driven input regime. Interestingly, when both interneuron types are activated to the same degree, this effect disappears (Fig. \ref{fig:Fig_4}A, left). In contrast, stimulating PV neurons biases the network's output towards predictions. This effect is even more pronounced when PV and SOM, or PV and VIP neurons are activated simultaneously (Fig. \ref{fig:Fig_4}A, middle and right).

In the previous simulations, we assumed that a neuromodulator acts globally, that is, on the interneurons in both the lower and the higher PE circuit. While this agrees with experimental data showing that XXX (XXX), we note that neuromodulators may also act more locally. The effect of stimulating an interneuron type in the lower PE circuit on the sensory weight is mostly the opposite of activating the same interneuron in the higher PE circuit (Fig. \ref{fig:Fig_4_S1}). For instance, the sensory inputs are overrated when the higher-level VIP neuron is activated, while the prediction is overrated when the lower-level VIP neuron is activated. When VIP and SOM neurons are stimulated equally, the sensory weight remains unchanged, independently of which PE circuit is targeted by the neuromodulator.

What are the mechanisms that give rise to these effects? And how do the combined local changes give rise to the global one observed in our network simulations? The sensory weight is chosen to be a function of the V neurons of the lower and higher PE circuit. Hence, any changes to the sensory weight result from changes to the neurons encoding the variances (Fig. \ref{fig:Fig_4}B). In our network, the V neurons only receive excitatory output synapses from PE neurons. Hence, any changes in the sensory weights upon activation of interneurons must be due to changes in the PE neurons. To disentangle the effect of nPE and pPE neurons, we perturbed those neurons individually in both the lower or higher subnetwork by injecting either an inhibitory or excitatory additional input (Fig. \ref{fig:Fig_4}C). Stimulating either PE neuron in the lower subnetwork increases the activity of the lower-level V neuron strongly. Moreover, the higher-level V neuron is also slightly affected. At first, this is counterintuitive because the V neuron in the higher subnetwork does not receive direct synapses from the PE neurons in the lower subnetwork. However, the activity of the lower-level M neuron encoding the prediction increases with an excitatory input onto the pPE neuron and decreases with an excitatory input onto the nPE neuron (the opposite is true for an inhibitory input). Because neurons in the higher PE circuit receive synapses from the lower-level M neuron, the activity of the higher-level V neuron is also affected. In contrast, stimulating either PE neuron in the higher subnetwork increases the activity of the higher-level V neuron but leaves the lower-level M and V neurons unaffected (Fig. \ref{fig:Fig_4}C).

Stimulating PE neurons may cause both an increase in the baseline activity and a change in the neuron's gain. To disentangle both effects, we illustrate each contribution separately using a mathematically tractable toy model (see Methods for more details). The variance estimated in the lower subnetwork increases with both increasing baseline and gain of the lower-level PE neurons. In contrast, when the gain of those PE neurons decreases, so does the variance. Similarly, the variance estimated in the higher subnetwork is equally influenced by changes in baseline and gain of the higher-level PE neurons. Moreover, changes to the baseline and the gain of the lower-level PE neurons increases the higher-level variance as a result of a biased prediction. Furthermore, the mean of the sensory stimuli is overpredicted (underpredicted) when the baseline or the gain of the lower-level pPE (nPE) neuron increases, or when the gain of the lower-level nPE (pPE) neuron decreases. In summary, if an interneuron causes an additional inhibitory input to a PE neuron, the neuron's gain is reduced. If an interneuron causes an additional disinhibitory input to a PE neuron, the neuron's baseline and gain are increased (Fig. \ref{fig:Fig_4}D).

This suggests that to understand the effect of neuromodulators on the sensory weight, we need to unravel the effect of interneuron activation on baseline and gain of PE neurons. To this end, we stimulated either PV, SOM, or VIP neurons independently for all three mean-field networks and measured the changes to baseline and gain of both PE neurons (Fig. \ref{fig:Fig_4}E). In all three networks tested, activating PV neurons decreases both quantities, leading to a decrease in the estimated variance (Fig. \ref{fig:Fig_4}E \& Fig. \ref{fig:Fig_4_S2}). Stimulating SOM or VIP neurons decreases the overall gain but increases the baseline activity of the PE neurons. Whether and how much the gain of the nPE or pPE neuron is reduced depends on the inputs onto SOM and VIP neurons, and the connectivity they make with other neurons in the network. Similarly, how much the baseline is elevated depends on the specifics of the mean-field network (Fig. \ref{fig:Fig_4}E \& Fig. \ref{fig:Fig_4_S2}). Hence, whether stimulating the SOM or VIP neuron decreases or increases the activity of the V neuron depends on the input statistics: for low-mean stimuli, the elevated baseline activity dominates the changes in the variance, while for high-mean stimuli the changes in the gain dominate.

%XXX Summary ... effect of neuromodulators acting on INs on sensory weight best understood as combined effect on the baseline and gain of nPE and pPE neurons that in turn determine the activity of the V neuron. Those effects not ony depend on the target IN but also the network in which it is embedded, that is the connectivity and the inputs onto INs. Moreover, it depends on whether local or global. The effects of activation on lower is mainly the opposite of same in higher. However, the efefcts don't cancel each other in global modulation. Mostly because 1) V higher additionally changes due to changes in prediction but also 2) effects in higher are stronger !? and 3) 1/(1+(Vs + D)/(Vp + D)) ... that is it cannot cancel per se. Even though the underlying reasons for a change in the V neuron depends on the the network characteristics, there are some consistent changes in sensory weigt across MFN ... name

% in caption: you need to make sure that in e you used a perturbation of 1 and the mean and std of stimuli were ... the lines only illustrate that it depends on the input statistics. Of course, the markers would also change with other perturbaton strengths, would they though? Think through
% Mention that for V neuron (higher) the darker line is actually not horizontal (unaffected)!
% Mention that in BL illustration you need a factor to increase visibility ... (in caption)

\subsection*{The contraction bias as a result of the weighted integration of sensory inputs and predictions}
% 
The weighted integration of sensory inputs and predictions thereof manifests in all-day behavior, in the form of a phenomenon called \textit{contraction bias}. The contraction bias describes the tendency to overestimate sensory stimuli drawn from the lower end of a stimulus distribution and to underestimate stimuli drawn from the upper end of the same distribution (Fig. \ref{fig:Fig_5}A). This \textit{bias towards the mean} has been reported in different species and modalities (XXX). 
%
\begin{figure}[t!]
	\centering
    \includegraphics{../results/figures/final/Fig_5}
\caption{\footnotesize{\bf Mechanisms underlying the contraction bias.\newline} 
{\bf (A)} Illustration of the contraction bias. The estimated input is shifted towards the mean of the input distribution. Hence, a linear curve fitted to the data has a slope below 1 (left). The bias is biggest at the end of the stimulus distribution (right).
{\bf (B)} Contraction bias in the model. Left: Example for two different stimulus variabilities. Right: As a consequence of the sensory weight, the slope decreases with stimulus variability (bias increases) and increases with trial variability (bias decreases).  Stimulus statistics: XXX.
{\bf (C)} The bias is independent of trial variability when the stimulus variability is zero (left). Equally, the bias is independent of the stimulus variability when the trial variability is zero (right). Stimulus statistics: XXX.
{\bf (D)} The slope depends on the trial duration. If the sensory weight is 1, the slope is independent of the trail duration and the bias vanishes (dashed-dotted line). If the sensory weight is 0, the slope depends on the trial duration and only reaches 1 if the trial duration approaches infinity. 
{\bf (E)} To ensure a larger bias for stimuli drawn from the upper end of the stimulus distribution than from the lower end, scalar variability as observed experimentally is needed.
}
\label{fig:Fig_5}
\end{figure}
%

The weighted output of our network can be interpreted as a neural manifestation of the contraction bias (see Methods for a thorough analysis). The bias increases with stimulus variance (Fig. \ref{fig:Fig_5}B), decreasing the slope of the linear fit modeling the relationship between the true and estimated stimuli (Fig.\ref{fig:Fig_5}B, right; compare with Fig. \ref{fig:Fig_5}A). In contrast, the bias decreases with trial variance, so that the slope of the linear fit approaches 1 (Fig. \ref{fig:Fig_5}B right).

What are the underlying network factors that contribute to this phenomenon? To disentangle the potentially different sources of the bias, we first simulated a network without stimulus variability (variance set to zero) for two different trial variabilities. In this case, a contraction bias emerges but is independent of the volatility of the environment (Fig. \ref{fig:Fig_5}C, left). We show mathematically that the bias results from the transient neuron activity before reaching a steady state, and vanishes if the trial duration approaches infinity (see Methods for more details, and Fig. \ref{fig:Fig_5}D). We next resume the limit case in which the same but high-variance stimulus is shown in every trial. In this case, the weighted output exhibits a contraction bias that is largely independent of the stimulus variance (Fig. \ref{fig:Fig_5}C, right and Fig. \ref{fig:Fig_5}D). As shown mathematically (see Methods), the bias results from the finite trial duration and the tendency to weight the prediction more strongly than the sensory inputs. 
% Maybe show the limit for both cases and show the theoretical values for each T as well ... ?

So far, we assumed that the stimulus variance is independent of the trial mean. A consequence of this choice is that the bias on either end of the stimulus distribution is largely the same (but with reversed signs). However, behavioral (neural?) data (XXX) shows that the bias increases for stimuli drawn from the upper end of the distribution, a phenomenon usually attributed to \textit{scalar variability}. To capture this in the model, we assume that the stimulus standard deviation linearly increases with the trial mean. In these simulations, as expected, the bias increases for a stimulus distribution shifted to higher trial means (Fig. \ref{fig:Fig_5}E).

In summary, the weighted integration of sensory inputs and predictions can be interpreted as a neural manifestation of the contraction bias. Both stimulus and trial variability contribute to the contraction bias but the underlying mechanisms differ.

% Enough? This section seems rather short! Should we do more, like making predictions about what would happen with the bias if neuromodulators activate INs? Like going back?

\section*{Supplementary Figures}
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}

\begin{figure}[!h]
	\centering
    \includegraphics[scale=1]{../results/figures/final/Fig_2_S1}% [width=1\linewidth]
\caption{\footnotesize{\bf Estimation of mean and variance for different stimulus distributions.\newline}  
Top: The mean-squared error (MSE) between the running average and the activity of the M neuron decreases to a near-zero level for all stimulus distributions tested. 
Bottom: The MSE between the instantaneous variance and the V neuron decreases to a low level with minor differences between the distributions tested. Zoom-in shows the last half of the trial. Mean of the stimulus distribution = XXX, Variance of the stimulus distribution = XXX.
}
\label{fig:Fig_2_S1}
\end{figure}

\begin{figure}[!h]
	\centering
    \includegraphics{../results/figures/final/Fig_3_S1}% [width=1\linewidth]
\caption{\footnotesize{\bf Dynamic variance estimation allows flexible adaptation to changes in the stimulus statistics and environment. \newline}  
{\bf (A)} Sensory weight for different input statistics (same as in Fig. \ref{fig:Fig_3}E). Numbers denote specific example states. Arrows denote the transitions between those states.
{\bf (B)} The sensory weight over time is shown for all transitions in (A). The switch to a new input statistics occurs at trial 60. Parameters are taken from Fig. \ref{fig:Fig_3}.
}
\label{fig:Fig_3_S1}
\end{figure}

\begin{figure}[!h]
	\centering
    \includegraphics{../results/figures/final/Fig_3_S2}% [width=1\linewidth]
\caption{\footnotesize{\bf Perturbing the weighting of sensory inputs and predictions by altering  network properties. \newline}  
{\bf (A)} The weights from the lower-level PE neurons to the M neuron are scaled by a factor below 1 (here, xxx) or above 1 (here, xxx), leading to a distorted weighting. If the update of the M neuron in the lower subnetwork is too slow ($\blacktriangleleft$), the prediction is overrated. If the update of the M neuron in the lower subnetwork is too fast ($\blacktriangleright$), the sensory input is overrated.
{\bf (B)} The precise activation function for the V neurons does not have a major impact on the sensory weight. Only for inputs with high stimulus variability, the sensory stimulus is slightly overrated when the squared activation function is replaced by a linear, rectified activation function.
}
\label{fig:Fig_3_S2}
\end{figure}

\begin{figure}[!h]
	\centering
    \includegraphics{../results/figures/final/Fig_4_S1}% [width=1\linewidth]
\caption{\footnotesize{\bf Neuromodulators acting locally either on interneurons in the lower or higher PE circuit. \newline}  
{\bf (A)} Sensory weight changes with neuromodulators acting on interneurons in the lower PE circuit.
{\bf (B)} Sensory weight changes with neuromodulators acting on interneurons in the lower PE circuit. Simulation parameters, labels and colors as in Fig. \ref{fig:Fig_4}. 
}
\label{fig:Fig_4_S1}
\end{figure}

\begin{figure}[!h]
	\centering
    \includegraphics{../results/figures/final/Fig_4_S2}% [width=1\linewidth]
\caption{\footnotesize{\bf Perturbing the interneurons changes the baseline and gain of nPE and pPE neurons.\newline}  
{\bf (A)} Changes in the baseline activity of nPE and pPE neurons for different interneurons targeted. 3 different mean-field networks are tested.
{\bf (B)} Same as in (A) but for the gain of nPE and pPE neurons. Simulation parameters, labels, and colors as in Fig. \ref{fig:Fig_4}. 
}
\label{fig:Fig_4_S2}
\end{figure}

% To continue, please "uncomment" the entire part below.

%\section*{Discussion}
%
%We solved the brain.
%
%% At the end, or in between talk about changes in PE neurons directly that you have currently in Fig. 3 SX ... I copy the part of the last paragraph here as inspiration :
%% For some psychiatric disorders, it has been shown that the weighting of sensory inputs and predictions thereof is impaired (XXX), leading to an overweighting of one of these signals. Moreover, it has been hypothesised that factors like stress or cognitive load may also influence the processing of feedforward and feedback inputs (XXX). ... Another factor that may contribute to a distorted weighting is the baseline activity of PE neurons that was set to zero in our model, in line with very low spontaneous firing rates of excitatory neurons in rodent primary sensory areas (XXX). Increasing this baseline activity for the nPE neuron (Fig. SXXX), pPE neuron (Fig. SXXX) or both pushes the network to weight sensory stimuli and predictions more equally. We speculate that an increase of baseline activity may be a natural result of an increased cognitive load or stress. 
%
%% mention Pakan paper
%
%\section*{Models and methods}
%%
%
%\subsection*{Network model}
%%
%Network consists of two subnetworks. Each subnetwork consists of a PE circuit, a memory neuron and a neuron representing the variance. XXX The memory neuron of subnetwork feefdowradly connects to the PE circuit of the second subnetwork.
%
%\subsubsection*{Prediction-error network model}
%%
%Consider a mean field network in which each population is represented by one representative neuron. The mean-field PE network consists of an excitatory nPE and pPE neuron, as well as two inhibitory PV neurons (one receiving S, the other P), as well as inhibitory SOM and VIP neurons.
%
%Each excitatory pyramidal cell (that is, nPE or pPE neuron) is divided into two coupled compartments, representing the soma and the dendrites, respectively. The dynamics of the firing rate~$r_{\mathrm{E}}$ of the somatic compartment of obeys \citep{wilson1972excitatory}
%%
%\begin{align}
%\tau_E\ \frac{dr_\mathrm{E}}{dt} &= - r_\mathrm{E} + w_\mathrm{ED}\cdot  r_\mathrm{D}  -  w_\mathrm{EP}\cdot r_\mathrm{P} + I_\mathrm{E},
%\end{align}
%%
%where $\tau_\mathrm{E}$ denotes the excitatory rate time constant ($\tau_\mathrm{E}$=60 ms), the weight $w_{\mathrm{ED}}$ describes the connection strength between the dendritic compartment and the soma of the same neuron, and $w_{\mathrm{EP}}$ denotes the strength of somatic inhibition from PV neurons. The overall input $I_\mathrm{E}$ comprises external background and feedforward sensory  inputs (see ``Inputs" below). Firing rates are rectified to ensure positivity.
%
%The dynamics of the activity~$r_\mathrm{D}$ of the dendritic compartment obeys \citep{wilson1972excitatory}
%%
%\begin{align}
%\tau_E\ \frac{dr_\mathrm{D}}{dt} &= - r_\mathrm{D} +  w_\mathrm{DE}\cdot r_\mathrm{E}  - w_\mathrm{DS}\cdot r_\mathrm{S} + I_\mathrm{D},
%\end{align}
%%
%where the weight $w_{\mathrm{DE}}$ denotes the recurrent excitatory connections between PCs, including backpropagating activity from the soma to the dendrites. $w_{\mathrm{DS}}$ represents the strength of dendritic inhibition from SOM neurons. The overall input $I_\mathrm{D}$ comprises fixed, external background inputs and feedback predictions (see ``Inputs" below). We assume that any excess of inhibition in a dendrite does not affect the soma, that is, the dendritic compartment is rectified at zero. 
%
%Just as for the excitatory neurons, the firing rate dynamics of each interneuron is modeled by a rectified, linear differential equation \citep{wilson1972excitatory},
%%
%\begin{align}
%\label{eq:RateEqINs}
%\tau_I\ \frac{dr_\mathrm{X}}{dt} =& - r_\mathrm{X} + I_{\mathrm{X}} + w_\mathrm{XE}\cdot r_\mathrm{E} - w_\mathrm{XP}\cdot r_\mathrm{P}  - w_\mathrm{XS}\cdot r_\mathrm{S} -  w_\mathrm{XV}\cdot r_\mathrm{V}, 
%\end{align}
%%
%where $r_\mathrm{X}$ denotes the firing rate of neuron type $X$, and the weight matrices $w_\mathrm{XY}$ denote the strength of connection between the presynaptic neuron population $Y$ and the postsynaptic neuron population $X$ ($X, Y\in \lbrace P,S,V\rbrace$). The rate time constant $\tau_I$ was chosen to resemble a fast GABA\textsubscript{A} time constant, and set to 2 ms for all interneuron types included. The overall input $I_\mathrm{X}$ comprises fixed, external background inputs, as well as feedforward sensory inputs and feedback predictions (see ``Inputs" below).
%
%\subsubsection*{Memory and variance neuron}
%%
%\begin{align}
%\tau_m \cdot \frac{dr_\mathrm{M}}{dt} = w_\mathrm{M\leftarrow pPE} \cdot r_\mathrm{pPE} - w_\mathrm{M\leftarrow nPE} \cdot r_\mathrm{nPE} 
%\end{align}
%%
%\begin{align}
%\tau_v \cdot \frac{dr_\mathrm{V}}{dt} = -r_\mathrm{V} + (w_\mathrm{V\leftarrow pPE} \cdot r_\mathrm{pPE} - w_\mathrm{V\leftarrow nPE} \cdot r_\mathrm{nPE})^2 
%\end{align}
%
%\subsubsection*{Weighted output}
%%
%\begin{align}
%r_\mathrm{out} = \alpha \cdot S + (1-\alpha) \cdot P
%\end{align}
%%
%\begin{align}
%\alpha &= \frac{1/r_\mathrm{V1}}{1/r_\mathrm{V1} + 1/r_\mathrm{V2}}\nonumber\\
%& = \left( 1 + \frac{r_\mathrm{V1}}{r_\mathrm{V2}} \right)^{-1}
%\end{align}
%
%\subsection*{Connectivity}
%%
%%All neurons are randomly connected with connection probabilities motivated by the experimental literature \citep[e.g.][]{fino2011dense, packer2011dense, pfeffer2013inhibition, lee2013disinhibitory, pi2013cortical, jiang2015principles, jouhanneau2015vivo, pala2015vivo},
%%%
%%\begin{align}
%%p = \begin{pmatrix}
%%  p_\mathrm{EE}  & p_\mathrm{ED} & p_\mathrm{EP} & p_\mathrm{ES} & p_\mathrm{EV} \\
%%  p_\mathrm{DE}  & p_\mathrm{DD} & p_\mathrm{DP} & p_\mathrm{DS} & p_\mathrm{DV} \\
%%  p_\mathrm{PE}  & p_\mathrm{PD} & p_\mathrm{PP} & p_\mathrm{PS} & p_\mathrm{PV} \\
%%  p_\mathrm{SE}  & p_\mathrm{SD} & p_\mathrm{SP} & p_\mathrm{SS} & p_\mathrm{SV} \\
%%   p_\mathrm{VE} & p_\mathrm{VD} & p_\mathrm{VP} & p_\mathrm{VS} & p_\mathrm{VV} \\
%%\end{pmatrix}
%%=
%%\begin{pmatrix}
%%  - & 1 & 0.6 & - & -\\
%%  0.1  & - & - & 0.55 & - \\
%%  0.45  & - & 0.5 & 0.6 & 0.5 \\
%%  0.35  & - & - & - & 0.5 \\
%%   0.1 & - & - & 0.45 & - \\
%%\end{pmatrix}.
%%\label{Mtx:ConnProb}
%%\end{align}
%%%
%%All cells of the same neuron type have the same number of incoming connections. The mean total connection strengths are set to
%%%
%%\begin{align}
%%w = \begin{pmatrix}
%%  w_\mathrm{EE}  & w_\mathrm{ED} & w_\mathrm{EP} & w_\mathrm{ES} & w_\mathrm{EV} \\
%%  w_\mathrm{DE}  & w_\mathrm{DD} & w_\mathrm{DP} & w_\mathrm{DS} & w_\mathrm{DV} \\
%%  w_\mathrm{PE}  & w_\mathrm{PD} & w_\mathrm{PP} & w_\mathrm{PS} & w_\mathrm{PV} \\
%%  w_\mathrm{SE}  & w_\mathrm{SD} & w_\mathrm{SP} & w_\mathrm{SS} & w_\mathrm{SV} \\
%%   w_\mathrm{VE} & w_\mathrm{VD} & w_\mathrm{VP} & w_\mathrm{VS} & w_\mathrm{VV} \\
%%\end{pmatrix}
%%=
%%\begin{pmatrix}
%%  - & 1 & 2^{*} & - & -\\
%%  0.5 & - & - & 0.5^{*} & -\\
%%  1.2 & - & 1 & 0.3^{*} & 0.3^{*} \\
%%  1 & - & - & - & 0.6 \\
%%  1 & - & - & 0.7 & - \\
%%\end{pmatrix},
%%\label{Mtx:ConnStrengths}
%%\end{align}
%%%
%%where $^{*}$ denotes either the weights that are free for optimization to satisfy the equations describing an E/I balance (see Supporting Information), or the initial mean connection strengths that are subject to synaptic plasticity during learning. For plastic networks, the initial connections between neurons are drawn from uniform distributions 
%%%
%%\begin{align*}
%%w_{ij}^{initial} \in \mathcal{U} \left(0.5\ w, 1.5\ w\right),
%%\end{align*}
%%%
%%where $w$ denotes the mean connection strengths given in (\ref{Mtx:ConnStrengths}). Please note that the system is robust to the choice of connection strengths. The connection strengths are merely chosen such that the solutions of the equations describing an E/I balance comply with Dale's principle.
%%
%%In plastic networks, $w_\mathrm{EP}$ is subdivided into assemblies. While one-third of PCs receive stronger initial inhibition from PV neurons that are driven by sensory input, another third receives stronger initial inhibition from PV neurons that are driven by feedback predictions. More precisely, for two-thirds of the excitatory neurons, half of the connections from PV neurons are strengthened by $1.5$, while the remaining ones are weakened by $0.5$.
%%
%%All weights are scaled in proportion to the number of existing connections (i.e., the product of the number of presynaptic neurons and the connection probability), so that the results are independent of the population size.
%
%\subsection*{Inputs}
%%
%%All neurons receive external background input that ensures reasonable baseline firing rates in the absence of sensory inputs and predictions thereof. In the case of non-plastic networks, these inputs were set such that the baseline firing rates are $r_\mathrm{E}=1\, s^{-1}$, $r_\mathrm{P} = r_\mathrm{S}=r_\mathrm{V}=4\, s^{-1}$ and $r_\mathrm{D}=0\, s^{-1}$. In the case of plastic networks, we set the external inputs of all neuron types to $x_\mathrm{E}=x_\mathrm{P}=x_\mathrm{S}=x_\mathrm{V}=5\, s^{-1}$, while the background input to the dendrites is dynamically computed during training such that $r_\mathrm{D}=0\, s^{-1}$.
%%
%%In addition to the external background inputs, the neurons receive either sensory input~($S$), a prediction thereof ($P$), or both. We distinguish between phases of fully predicted ($P=S>0$), overpredicted ($P>S$) and underpredicted ($P<S$) sensory stimuli, as well as baseline phases ($P=S=0$). During training, the network is exposed to baseline phases and fully predicted sensory inputs (Figs.~\ref{fig:Fig_Plasticity} and \ref{fig:Fig_Experience}), or in addition to over- and underpredicted sensory stimuli (Fig.~\ref{figsupp:Fig_Experience_Predictability}). Stimuli are drawn from a uniform distribution from the interval $[0, 5\, s^{-1}]$. Mean and SD of test stimuli for each simulation are listed below (see ``Simulations").
%
%
%\subsection*{Simulations}
%%
%%All simulations were performed in customized Python code written by LH. Differential equations were numerically integrated using a 2\textsuperscript{nd}-order Runge-Kutta method with time steps ranging between $0.1$ and $0.2$ ms. Neurons in the PE circuits were initialized with $r=0/s$. The memory neurons were initialized at the mean of the two distributions (see above), and each prediction neuron was either set to the mean of the distribution it is associated with if the stimulus at $t=0\ \mathrm{ms}$ was drawn from that distribution, or set to zero otherwise. Each stimulus was presented for $1$ second. During training of the PE circuit, we presented 350 stimuli alternated with 350 zero-input (baseline) phases. We made sure that the weights converged to a configuration that satisfied the objective given by our homeostatic plasticity rules (see Eqs.~\ref{eq:Plasticity_I}-\ref{eq:Plasticity_II} in ``Plasticity model"). We defined the steady-state firing rate per stimulus as the activity in the last 500 ms of stimulus presentation. The onset firing rate was computed as the activity of the first 10 ms.\\\\
%%%
%%\textbf{Figures 1 \& S2:} Test stimulus was set to $5/s$ with a SD of $1/s$. Stimulus to compute total excitatory and inhibitory inputs was set to $1/s$.\\
%%%
%%\textbf{Figures 2 \& S3:} Test stimulus was set to $3/s$ with a SD of $1/s$. The perturbation stimuli ranged from $-1.5/s$ to $1.5/s$.\\
%%%
%%\textbf{Figures 3, S4 \& S5:} Test stimulus was set to $5/s$ with a SD of $1.5/s$. 50\% of the PV neurons, 70\% of the SOM neurons and 30\% of the VIP neurons receive the actual sensory input, while the remaining ones of each population received a prediction thereof. Perturbation stimulus was $\pm 2/s$. Panels D \& G of Fig. 3 show the median over all PE neurons and the SEM.\\
%%%
%%\textbf{Figures 4 \& S6:} Test stimulus was set to $5/s$ with a SD of $1.5/s$. In main figure, square: $w_\mathrm{EP}\in[2,4]$, $w_\mathrm{PS}\in[0.5,1]$, $w_\mathrm{PV}\in[1.5,2.5]$; circle: $w_\mathrm{EP}\in[2.5,8]$, $w_\mathrm{PS}\in[1.5,2.5]$, $w_\mathrm{PV}\in[0.5,1]$; triangle: $w_\mathrm{EP}\in[2.5,8]$, $w_\mathrm{PS}\in[1,2.5]$, $w_\mathrm{PV}\in[0.5,2]$. Half of the PV neurons and all SOM neurons receive the actual sensory input, while the remaining PV and SOM neurons as well as all VIP neurons receive a prediction thereof.
%%In supporting figure: Total number of stimuli presented during training was increased, so that the number of fully predicted sensory stimuli was constant at 350. Results were averaged over 5 simulations, mean and SD are shown.\\
%%%
%%\textbf{Figure 5:} For panel E, the performance error was computed as the squared difference between the activity of the respective line attractor and the stimulus presented. For panel F, the initial weight between the stimulus and the representation neuron was set to $0.5$. The basis learning rate (fixed) was set to $5e^{-4}$. And the initial speed was computed as the derivative of the rate with respect to time, averaged over the first 50 ms. 
%%
%%Source code to reproduce the simulations, analyses and figures will be available after publication at \url{github.com/lhertaeg/SourceCode_Hertaeg2021}. 
%
%\section*{Acknowledgments}
%%
%%We are grateful to Vezha Boboeva, Douglas Feitosa Tom\'e, J\'ulia Gallinaro and Klara Kaleb for helpful comments on earlier versions of this manuscript, and we want to thank all members of the Clopath lab for discussion and support. This work was supported by BBSRC BB/N013956/1, BB/N019008/1, Wellcome Trust 200790/Z/16/Z, Simons Foundation 564408 and EPSRC EP/R035806/1.
%
%\bibliographystyle{naturemag} %{plainnat}
%\bibliography{../References_Mismatch}
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% APPENDICES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%\newpage
%\section*{Supplementary Information}
%\appendix
%
%
%\subsection*{Sensory weight and contraction bias}
%%
%If P is rather constant, the slope in the contraction bias is exactly the sensory weight
%%
% \begin{align*}
% r_\mathrm{out} = \alpha_\mathrm{S} \cdot S + \left( 1 -\alpha_\mathrm{S} \right) \cdot P \equiv m \cdot S + n
% \end{align*}
%%
%However, P is usually/normally a function of $S$. For simplicity, let's assume that P decays exponentially to a new value of $S$:
%%
%\begin{align*}
%P = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau} + f(S) \cdot \left( 1 -   \mathrm{e}^{-t/\tau} \right)
%\end{align*}
%%
%Within each trial with trial duration T, P can be expressed by n sections of length t in which the stimulus is constant and, for the sake of simplicity, drawn from a uniform distribution $U\left( s - \frac{\sigma_\mathrm{S}}{12}, s + \frac{\sigma_\mathrm{S}}{12} \right)$. $P_\mathrm{0}$ is drawn from $U\left( \mu - \frac{\sigma_\mathrm{P}}{12}, \mu + \frac{\sigma_\mathrm{P}}{12} \right)$. $P_\mathrm{n}$ is then given by
%%
%\begin{align*}
%P_\mathrm{n} = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau}  + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} s_i \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau}
%\end{align*}
%%
%This needs to be averaged over all possible states
%%
%\begin{align*}
%P_\mathrm{n} = \mathrm{e}^{-t/\tau}  \int\limits_{\mu - \frac{\sigma_\mathrm{P}}{12}}^{\mu + \frac{\sigma_\mathrm{P}}{12}} P_\mathrm{0} \ f(P_\mathrm{0})\ dP_0+ \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau} \int\limits_{s - \frac{\sigma_\mathrm{S}}{12}}^{s + \frac{\sigma_\mathrm{S}}{12}} s\ f(s)\ ds
%\end{align*}
%%
%This gives
%%
%\begin{align*}
%P_\mathrm{n} = \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot t/ \tau} \cdot S
%\end{align*}
%%
%By making use of the geometric series, this simplifies to
%%
%\begin{align*}
%P_\mathrm{n} =  \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right) \cdot S
%\end{align*}
%%
%Together, this yields for the weighted output
%%
%\begin{align*}
% r_\mathrm{out} = \left[ \alpha_\mathrm{S}\  \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right)\right] \cdot S + \left( 1 -\alpha_\mathrm{S} \right)\ \mathrm{e}^{-T/\tau}\ \mu
%\end{align*}
%%
%Hence, the slope is a function of both the sensory weight and the trial duration. 
% 
%In a prediction-driven input regime ($alpha_\mathrm{S} \sim 0)$, the slope is independent of the sensory weight and only determined by the trial duration, $m \sim \left( 1 -   \mathrm{e}^{-T/\tau} \right)$. In a sensory-driven input regime ($alpha_\mathrm{S} \sim 0)$, the contraction bias vanishes ($m \sim 1$). 
%
%If the trail duration is short ($T \rightarrow 0$), the slope is given by the sensory weight. If the trail duration approaches infinity, the slope would be 1 again (however, this seems rather unrealistic, this would only be true in an ideal system without memory decay or reproduction and accumulation noise ...).
%
%\subsection*{Impact of PE neurons' gain on estimating mean and variance}
%%
%Only if the the gain of the nPE neuron ($g_{nPE}$) equals the gain of the pPE neuron ($g_{pPE}$) in the mean-field network, the activity of the M neuron represents the mean of the inputs,
%%
%\begin{align}
%\label{eq:condition_mean_gain_equal}
%g_{pPE}\ \langle \mathrm{nPE}\rangle &= g_{nPE}\ \langle \mathrm{pPE}\rangle \\
%g_{pPE} \langle \left[ S-P\right]_+\rangle &= g_{nPE} \langle \left[ P-S\right]_+\rangle \nonumber \\
%g_{pPE} \int\limits_P^b \left( x-P\right)\ f(x)\ dx &= g_{nPE} \int\limits_a^P \left( P-x\right)\ f(x)\ dx. \nonumber
%\end{align}
%%
%In case of a uniform distribution ($f(x) = 1/(b-a)$ when $x\in [a,b]$ and $0$ otherwise) from which the input values are drawn, this condition yields
%%
%\begin{align}
%\label{eq:condition_mean_gain_equal_4_uniform_dist_1}
%P = \int\limits_a^b x\, f(x)\ dx = \frac{a + b}{2}
%\end{align}
%%
%for $g_{nPE} = g_{pPE} = g$, and 
%%
%\begin{align}
%\label{eq:condition_mean_gain_equal_4_uniform_dist_21}
%g_{pPE}\ \left[ \frac{1}{2} \left(b^2 - P^2\right) - P\left(b - P\right)\right] = g_{nPE}\ \left[  P\left(P - a\right) - \frac{1}{2} \left(P^2 - a^2\right)\right]
%\end{align}
%%
%for $g_{nPE} \neq g_{pPE}$, which can be further summaries by
%%
%\begin{align}
%\label{eq:condition_mean_gain_equal_4_uniform_dist_21}
%P = \frac{g_{pPE}\ b - g_{nPE}\ a \pm \sqrt{g_{nPE}\ g_{pPE}} (a-b)}{g_{pPE} - g_{nPE}}.
%\end{align}
%%
%Hence, estimating the mean correctly requires $g_{nPE} = g_{pPE} = g$. For the V neuron to represent the variance of the inputs, this condition must be tightened to  $g_{nPE} = g_{pPE} = 1$. The variance is given by
%%
%\begin{align}
%\label{eq:condition_variance_gain}
%V &= \langle (S-P)^2\rangle \\
%   &= g_\mathrm{pPE}\ \langle \left[ S-P \right]_+^2 \rangle + g_\mathrm{nPE}\ \langle \left[ P-S \right]_+^2 \rangle. \nonumber
%\end{align}
%%
%In case of a uniform distribution from which the input values are drawn, this condition yields
%%
%\begin{align}
%\label{eq:condition_variance_gain_uniform_dist}
%V &= \frac{g_\mathrm{pPE}}{b-a} \int\limits_P^b (u-P)^2\ du + \frac{g_\mathrm{nPE}}{b-a} \int\limits_a^P (P-u)^2\ du \\
%   &= \frac{g_\mathrm{pPE}}{3} \cdot \frac{(b-P)^3}{b-a} + \frac{g_\mathrm{nPE}}{3} \cdot \frac{(P-a)^3}{b-a}. \nonumber
%\end{align}
%%
%Only if $g_{nPE} = g_{pPE} = 1$ and $P = \frac{a + b}{2}$, the variance is given by $\frac{(b - a)^2}{12}$, otherwise the V neuron's activity is given by
%%
%\begin{align}
%\label{eq:condition_variance_gain_uniform_dist_1}
%V = \frac{(b-a)^2}{3\ (g_\mathrm{pPE} - g_\mathrm{nPE})^3} \cdot \left[ g_\mathrm{nPE} \cdot( g_\mathrm{pPE} \mp \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}})^3 - g_\mathrm{pPE} \cdot (g_\mathrm{nPE} \mp \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}})^3\right].
%\end{align}
%%
%
%
%\subsection*{Impact of PE neurons' baseline on estimating mean and variance}
%%
%Only if the the baseline of the nPE neuron ($n_0$) equals the baseline of the pPE neuron ($p_0$) in the mean-field network, the activity of the M neuron represents the mean of the inputs,
%%
%\begin{align}
%\label{eq:condition_baseline_mean}
%\langle \mathrm{pPE} \rangle &= \langle \mathrm{nPE} \rangle \\
%\langle \left[ S - P\right]_+ + p_0\rangle &= \langle \left[ P - S\right]_+ + n_0\rangle \nonumber\\
%\int\limits_P^b (x - P)\ f(x)\ dx + p_0 \underbrace{\int\limits_a^b f(x)\ dx}_{=1}  &= \int\limits_a^P (P - x)\ f(x)\ dx + n_0 \underbrace{\int\limits_a^b f(x)\ dx}_{=1} . \nonumber
%\end{align}
%%
%In case of a uniform distribution ($f(x) = 1/(b-a)$ when $x\in [a,b]$ and $0$ otherwise) from which the input values are drawn, this condition yields
%%
%\begin{align}
%\label{eq:condition_baseline_mean_1}
%P = \frac{b+a}{2} + \frac{p_0 - n_0}{b-a}.
%\end{align}
%%
%Hence, if $p_0 = n_0$, the mean can be estimated correctly. For the V neuron to represent the variance of the inputs, this condition must be tightened to $p_0 = n_0 = 0$. The variance is given by
%%
%\begin{align}
%\label{eq:condition_baseline_variance}
%V &= \langle \left( \mathrm{pPE} + \mathrm{nPE} \right)^2 \rangle \\
%&= \langle \left[ S-P\right]_+^2\rangle + \langle \left[ P-S\right]_+^2\rangle + (p_0 + n_0)^2 + 2\ (p_0 + n_0)\ \left( \langle \left[ S-P\right]_+\rangle + \langle \left[ P-S\right]_+ \rangle\right) \nonumber
%\end{align}
%%
%In case of a uniform distribution from which the inputs to the mean-field network are drawn, this expression simplifies to
%%
%\begin{align}
%\label{eq:condition_baseline_variance_1}
%V = \frac{1}{3\ (b-a)} \left[ (b-P)^3 + (P-a)^3\right] + (p_0 + n_0)^2 + \frac{(p_0 + n_0)}{b-a} \left[ (b-P)^2 + (a-P)^2\right].
%\end{align}
%%
%Inserting the expression for P (by itself modulated by the baseline activities of the PE neurons) yields
%%
%\begin{align}
%V =& \frac{1}{3\ (b-a)} \left[ \left( \frac{b-a}{2} - \frac{p_0 - n_0}{b-a}\right)^3 + \left( \frac{b-a}{2} + \frac{p_0-n_0}{b-a}\right)^3\right] + (p_0 + n_0)^2 \nonumber \\
%&+ \frac{(p_0 + n_0)}{b-a} \left[ \left( \frac{b-a}{2} + \frac{p_0-n_0}{b-a}\right)^2 + \left( \frac{b-a}{2} - \frac{p_0 - n_0}{b-a}\right)^2\right]
%\end{align}
%%
%Simplifying the expression, leads to
%%
%\begin{align}
%\label{eq:condition_baseline_variance_2}
%V =  \frac{(b-a)^2}{12} + \frac{(p_0-n_0)^2}{(b-a)^2} \left( 1 + 2\ \frac{p_0+n_0}{b-a}\right) + (p_0 + n_0) \left( p_0 + n_0 + \frac{b-a}{2}\right)
%\end{align}
%%
%
%\subsection*{Modelling the impact of neuromodulators on the weighting of sensory inputs and predictions thereof}
%
%Assume that neuromodulator act through activating INs. INs on the other hand modulate both gain and BL of nPE and pPE neurons ... . Changes in gain and BL will affect the overall activity and the balance between nPE and pPE neurons => this will affect both the prediction and the variance. 
%
%Modulate PE in lower:
%variance in lower is combination of direct changes in PE and indirect through prediction (that is changed through PE mod)
%variance in higher is just because of differences in prediction
%
%
%Modulation in higher:
%variance in lower should be unaffected
%variance in higher is combination of direct changes in PE and indirect through prediction (that is changed through PE mod)
%
%Variance as function of bias in mean:
%\begin{align*}
%V &= \frac{1}{n} \sum_i \left( x_i - \left(\mu \pm \delta\mu\right)\right)^2 \\
%&= \frac{1}{n} \sum_i \lbrace  \left( x_i - \mu \right)^2 + \delta\mu^2 \mp 2\, \delta\mu\,  (x_i - \mu)\rbrace \\
%&= V_\mathrm{unmod} + \delta\mu^2 \mp 2\ \delta\mu \left( \frac{1}{n} \sum_i x_i- \mu\right) \\
%&= V_\mathrm{unmod} + \delta\mu^2
%\end{align*}
%%
%
%
%\subsection*{Influence of a population of nPE and pPE neurons}
%XXX
%
%
%\subsection*{Analysis of simplified network model, effect of time constants}
%
%simplified model: dynamics and steady state of rM and rV, rM and rV as a function of time constants and trial duration etc., weighting, then use those expressions to discuss when weighting goes awry and how long transitions take from one state to another ...
%
%
%\subsection*{Comparison to Kalman filter and Bayes Factor surprise}
%%
%Kalman filter. Initialisation
%%
%\begin{align*}
%x_{0|init} &= 0 \\
%P_{0|init} &= \sigma^2\ I
%\end{align*}
%%
%with x being the system state (in my terms the prediction), P is the covariance matrix of the errors of x (in my terms the var of the predictions) and I is the identity matrix.
%%
%Then the "correction" is given by
%%
%\begin{align*}
%K_k &= P_{k|k-1}\ H_k^T\ \left( H_k\ P_{k|k-1}\ H_k^T + R_k \right)^{-1} \\
%x_k &= x_{k|k-1} + K_k\ \left( z_k - H_k\ x_{k|k-1}\right) \\
%P_k &= \left( I - K_k\ H_k\right)\ P_{k|k-1}
%\end{align*}
%%
%with K the kalman gain matrix, H the observation matrix ($z_k = H_k\ x_k + noise$), R the covaraince of the measurement noise and z a new observation. The last part of the Kalman filter is the "prediction":
%%
%\begin{align*}
%x_{k|k-1} &= F_{k-1}\ x_{k-1} + B_{k-1}\ u_{k-1} \\
%P_{k|k-1} &= F_{k-1}\ P_{k-1}\ F_{k-1}^T + Q_{k-1}
%\end{align*}
%%
%with F the transition matrix ($x_{k|k-1} = F_{k-1}\ x_{k-1}$, u a deterministic perturbation, B the dynamics of the deterministic perturbation. In our terms
%%
%\begin{align*}
%\alpha = K_k = \frac{P_{k|k-1}}{R_k + P_{k|k-1}}
%\end{align*}
%%
%$P_{k|k-1}$, is however $\sigma_P^2$ in my implementation and $R_k$ is fixed variance of inputs $\sigma_S^2$. Hence, my implementation represents (?) the Kalman filter. Important to note is, that in my implementation we estimate the variance of inputs dynamically, so it is not set! Another nice advantage here is that I don't need a good estimate for P. I can basically initiate it as I want. Another difference is that I consider the optimal weighting in my "output neuron" and not the prediction itself ... .
%
%XXX Comparison to Bayes Factor surprise

\end{document}

### 2023-07-13 scripts/plot_fig_2_support_1.py

@@ -34,7 +34,7 @@ if not os.path.exists(figPath):
     
 # %% Define figure structure
 
-figsize=(9/inch,10/inch)
+figsize=(9/inch,7/inch)
 fig = plt.figure(figsize=figsize)
 
 G = gridspec.GridSpec(2, 1, figure=fig, hspace=0.4)#, hspace=1.2)
@@ -63,12 +63,17 @@ else:
     
     # plot data
     mean, std = 5, 2
-    sem_mean = np.std(mse_mean,2)/np.sqrt(num_reps)
-    sem_variance = np.std(mse_variance,2)/np.sqrt(num_reps)
+    var_tested = std**2
     
-    plot_mse_test_distributions(np.mean(mse_mean,2), SEM = sem_mean, fs=fs, mean=mean, std=std, dist_types=dist_types, 
+    mse_mean_normalised = mse_mean/mean**2 * 100
+    mse_variance_normalised = mse_variance/var_tested**2 * 100
+    
+    sem_mean = np.std(mse_mean_normalised,2)/np.sqrt(num_reps)
+    sem_variance = np.std(mse_variance_normalised,2)/np.sqrt(num_reps)
+    
+    plot_mse_test_distributions(np.mean(mse_mean_normalised,2), SEM = sem_mean, fs=fs, mean=mean, std=std, dist_types=dist_types, 
                                 inset_steady_state=True, plot_xlabel = False, ax=ax_A)
-    plot_mse_test_distributions(np.mean(mse_variance,2), SEM = sem_variance, fs = fs, inset_steady_state = True, ax=ax_B)
+    plot_mse_test_distributions(np.mean(mse_variance_normalised,2), SEM = sem_variance, fs = fs, inset_steady_state = True, ax=ax_B)
 
 
 # %% save figure

### 2023-07-13 scripts/plot_fig_3.py

@@ -27,7 +27,7 @@ inch = 2.54
 
 # %% Define files and paths
 
-figure_name = 'Fig_3.png'
+figure_name = 'Fig_3_partial.png'
 figPath = '../results/figures/final/'
 
 if not os.path.exists(figPath):
@@ -57,7 +57,7 @@ ax_B.axis('off')
 ax_B.text(-0.15, 1.25, 'B', transform=ax_B.transAxes, fontsize=fs+1)
 
 ax_B1 = fig.add_subplot(G12[0,0])
-ax_B1.set_title('Neurons activity increases \nwith trial variability', fontsize=fs, pad=10)
+ax_B1.set_title('Neuron activity increases \nwith trial variability', fontsize=fs, pad=10)
 ax_B2 = fig.add_subplot(G12[0,1])
 ax_B2.set_title('Neuron activity increases \nwith stimulus variability', fontsize=fs, pad=10)
 ax_B3 = fig.add_subplot(G12[0,2])

### 2023-08-11 records/manuscript_2023.tex
add population network

--- 2023-06-30 records/manuscript_2023.tex
@@
 
+\hyphenation{math-e-mat-i-cal-ly} 
+\hyphenation{inter-neurons}
+
 \begin{document}
@@
 %
-XXX
+To survive in an ever-changing environment, animals must flexibly adapt their behavior based on previously encoded and novel information.
+This context-dependency must also be reflected in the information processing of neural networks underlying intelligent behavior.
+For instance, when you walk down some staircase in your fully lit basement, your brain might entirely rely on the feedforward (bottom-up) input your senses receive (Fig.
+\ref{fig:Fig_1}, left).
+In contrast, when you walk down the same stairs in complete darkness, your brain might rely entirely on feedback (top-down) signals generated from a staircase model it has formed over previous experiences (Fig.
+\ref{fig:Fig_1}, middle).
+
+The importance of these feedback inputs has been emphasized by observations showing that top-down projections outnumber feedforward connections (XXX) and that they modulate (XXX) or even entirely drive (XXX) neuron activity.
+But how do neural networks switch between a feedforward-dominated and a feedback-dominated processing mode?
+And how do neural networks combine both input streams wisely?
+For instance, if you hike down an unexplored mountain in very foggy conditions, your brain receives unreliable visual information.
+In addition, it can only draw on a shaky prediction about what to expect (Fig.
+\ref{fig:Fig_1}, right).
+
+A common hypothesis is that the brain weights different inputs according to their reliabilities.
+A prominent example of this hypothesis is Bayesian multisensory integration (XXX).
+According to this theory, neural networks represent information from multiple modalities by a linear combination of the uncertainty-weighted single-modality estimates.
+Multisensory integration is supported by several observations showing that xxx (XXX).
+It is conceivable that the same concepts can be employed for the weighting of sensory inputs and predictions thereof (XXX).
+A central point in the weighting of inputs is the estimation of variances as a measure of uncertainty.
+However, how the variance of both the sensory input and the prediction can be computed on the circuit level is not resolved yet.
+
+We hypothesised that prediction error (PE) neurons provide the basis for the neural computation of variances.
+PEs are an integral part of the theory of predictive processing which states that the brain constantly compares incoming sensory information with predictions.
+When those predictions are wrong, the resulting PEs allow the network to revise the model of the world, thereby ensuring that the predictions are more accurate (XXX).
+Experimental evidence suggests that these PEs may be represented in the activity of distinct groups of neurons, termed PE neurons (XXX).
+Moreover, these neurons may come in two types when excitatory neurons exhibit near-zero, spontaneous firing rates (XXX).
+Negative PE (nPE) neurons mainly increase their activity when the prediction is \textit{stronger} than the sensory input, while positive PE (pPE) neurons mainly increase their activity when the prediction is \textit{weaker} than the sensory input.
+Indeed, it has been shown that excitatory neurons in layer 2/3 of rodent primary sensory areas can encode negative or positive PEs (XXX).
+
+Here, we show that the unique response patterns of nPE and pPE neurons may provide the backbone for computing both the mean and the variance of sensory stimuli.
+Furthermore, we suggest a network model with a hierarchy of PE circuits to estimate the variance of the prediction, in addition to the variance of the sensory inputs.
+In line with multisensory integration, predictions are weighted more strongly than the sensory stimuli when the environment is stable (that is, predictable) but the sensory inputs are noisy.
+Moreover, we show that predictions are integrated more strongly after a change in the environment, even when the new sensory stimulus is reliable.
+In addition, we unravel the mechanisms underlying a neuromodulator-induced shift in the weighting of sensory inputs and predictions.
+In our model, these neuromodulators activate groups of inhibitory neurons like parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (XXX).
+In a computational model, these interneurons have been shown to establish a multi-pathway balance of excitation and inhibition that is the basis for nPE and pPE neurons (XXX).
+By breaking this balance, the excitatory neurons change their baseline firing rate and gain, leading to a biased variance estimation.
+Finally, we show that this weighting can be understood as the neural manifestation of the contraction bias (XXX).
+
 %
-Behavior of intelligent life is context-dependent and requires a constant adaptation to the ever-changing life.
-This must also be reflected in the underlying neural networks and the inputs that drive those.
-For instance, when you walk down your your basement stairs and the lighting is well, your brain might rely entirely on the feedforward sensory input your senses receive to take decisions and adapt your body in the environment/world (Fig 1a).
-However, when you walk down the same basement stairs you walked million of times before but suddenly the light bulb is broken and the visual input is absent, you don't freeze forever but walk the stairs almost as confidently as before.
-Your brain now relies on the prediction it had formed over the last years.
-(Fig.
-1b).
-The former example resonates with the feedforward-centric view that has dominated the neuroscience literature for decades.
-However, in recent years, the importance of internally generated, feedback inputs has increased significantly.
-Not only because it was shown that top-down projects xxx outsize feedforward connections (ref) but also because these connection modulate (ref) or even drive (ref) neuron dynamics.
-These feedback projects have been hypothesized to carry expectations, predictions ...
-and may help us to navigate (in the physical and metaphysical sense) the world even in the absence of sensory input.
-But how do our brains decide whether it follows more the sensory feedforward input or the feedback predictions?
-
-For instance, if you hike down a mountain for the first time under very foggy conditions, you brain may receive unreliable sensory input but at the same time can only have a very rough prediction about what to expect.
-In this case, your brain might be better off combining both the sensory input and the prediction thereof (Fig 1c).
-A common hypothesis is that the brain weights those signals according to their reliabilties, that is, a function of the variances of those signals (ref).
-These ideas come from Bayesian integration in which it is believed that xxx (ref).
-A prominent example of this is multisensory integration in which xxx (ref).
-Similarly, the brain must have mechanisms to estimate the variances of both the sensory input and the predictions thereof that allow it to weight those signals accordingly and context-dependent.
-But how this is implemented in the brain is still largely unknown.
-
-Here, we hypothesis that xxx. 
-We show that xxx
+\begin{figure}[t!]
+	\centering
+    \includegraphics{../results/figures/final/Fig_1}
+\caption{\footnotesize{\bf Neural network model to flexibly integrate sensory information and predictions.\newline} 
+{\bf (A)} Example illustration of context-dependent integration of information.
+Left: When walking down a staircase that is clearly visible, the brain might rely solely on external sensory information.
+Middle: When walking down the same stairs in the absence of visual information, the brain might rely on predictions formed by previous experience.
+Right: When climbing down an unexplored mountain in foggy conditions, the brain might need to integrate sensory information and predictions at the same time.
+{\bf (B)} Illustration of a prediction-error (PE) circuit with both negative and positive PE (nPE/pPE) neurons that receive inhibition from three different inhibitory interneuron types: parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons.
+Local excitatory connections are not shown for clarity.
+{\bf (C)} Illustration of network model that estimates the mean and variance of the external sensory stimuli.
+The core of this network model is the PE circuit shown in (B).
+The lower-level V neuron encodes the variance, while the lower-level M neuron encodes the mean of the sensory input.
+{\bf (D)} Same as in (C) but the feedforward input is the activity of the lower-level M neuron.
+}
+\label{fig:Fig_1}
+\end{figure}
+%
 
@@
 Assuming that the prediction equals the mean of the sensory stimulus, the PE neurons, hence, encode the deviation from the mean.
-Thus, the squared sum of nPE and pPE activity represents the variance of the feedforward input.
+Thus, the squared sum of nPE and pPE neuron activity represents the variance of the feedforward input.
+%
+\begin{figure}[t!]
+	\centering
+	%\makebox[\textwidth][c]{\includegraphics{../results/figures/final/Fig_2}}
+    \includegraphics[width=1\linewidth]{../results/figures/final/Fig_2}
+\caption{\footnotesize{\bf Prediction-error neurons provide the basis for estimating mean and variance of sensory stimuli.\newline} 
+{\bf (A)} Illustration of the inputs with which the network shown in \ref{fig:Fig_1}C is stimulated.
+Network is exposed to a sequence of constant stimuli drawn from a uniform distribution.
+Stimulus duration is XXX.
+{\bf (B)} PE neuron activity hardly changes with stimulus strength (left) but strongly increases with stimulus variability (right).
+{\bf (C)} Interneuron activity strongly changes with stimulus strength (left) but hardly changes with stimulus variability (right) .
+{\bf (D)} M neuron correctly encodes the mean of the sensory stimuli.
+Left: Illustration of the input synapses onto the M neuron.
+Middle: Activity of the M neuron over time for a uniform distribution with mean XXX and standard deviation XXX.
+Right: Normalised mean-squared error (MSE) between the running average and the M neuron activity for different parametrizations of the stimulus distribution.
+{\bf (E)} V neuron correctly encodes the variance of the sensory stimuli.
+Left: Illustration of the input synapses onto the V neuron.
+Middle: Activity of the V neuron over time for a uniform distribution with mean XXX and standard deviation XXX.
+Right: Normalised mean-squared error (MSE) between the instantaneous variance and the V neuron activity for different parametrizations of the stimulus distribution.
+}
+\label{fig:Fig_2}
+\end{figure}
+%
 
 To test our hypothesis, we studied a rate-based mean-field network the core of which is a prediction-error (PE) circuit with excitatory nPE and pPE neurons, as well as inhibitory parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (Fig.
-xxx).
+\ref{fig:Fig_1}B).
 While the excitatory neurons are simulated as two coupled point compartments to emulate the soma and dendrites of elongated pyramidal cells, respectively, all inhibitory cell types were modeled as point neurons.
@@
 
-In addition to this core circuit, we model a memory (M) neuron that perfectly integrates the activity of the PE neurons.
+In addition to this core circuit, we model a memory (M) neuron that perfectly integrates the activity of the PE neurons (Fig.
+\ref{fig:Fig_1}C).
 In accordance with XXX, we assume that the pPE neuron excites the memory neuron, while the nPE neuron inhibits this neuron (for instance, through lateral inhibition, here not modeled explicitly).
@@
 We furthermore simulate a downstream neuron (termed V neuron), modeled as a leaky integrator with a squared activation function, that receives excitatory output synapses from the PE neurons.
-Hence, in this setting, the V neuron encodes the variance of the sensory stimuli.
+Hence, in this setting, the V neuron encodes the variance of the sensory stimuli (Fig.
+\ref{fig:Fig_1}C).
 
 To show that this network can indeed represent mean and variance in the respective neurons, we stimulate it with a sequence of step-wise constant inputs drawn from a uniform distribution (Fig.
-2aXX), assuming that the sensory stimulus varies over time.
+\ref{fig:Fig_2}A), assuming that the sensory stimulus varies over time.
 In line with the distinct response patterns for nPE and pPE neurons, these neurons change only slightly with increasing stimulus mean but increase strongly with input variance (Fig.
-2bXXX).
+\ref{fig:Fig_2}B).
 This is in contrast to the three interneurons that strongly increase with stimulus mean while they only moderately increase with stimulus variance (Fig.
-2cxxx).
+\ref{fig:Fig_2}C).
 The activity of the memory neuron M gradually approaches the mean of the sensory inputs (Fig.
-2d, middle), while the activity of the V neuron approaches the variance of the inputs (Fig.
-2e, middle).
+\ref{fig:Fig_2}D, middle), while the activity of the V neuron approaches the variance of the inputs (Fig.
+\ref{fig:Fig_2}E, middle).
 This is true for a wide range of input statistics (Fig.
-XXX) and input distributions (Fig.
-SXXX).
+\ref{fig:Fig_2}D-E, right) and input distributions (Fig.
+\ref{fig:Fig_2_S1}).
 Small deviations from the true mean occur mainly for larger input variances, while the estimated variance is fairly independent of the input statistics tested.
 
-XXX validated that also correct for population network (beyond mean-field network) XXX
-XXX assumptions (BL, gain-nPE = gain-pPE = 1) and how important really (see above)
+XXX coming soon: Paragraph on network beyond mean-field XXX
 
-XXX Summary paragraph XXX
+In summary, nPE and pPE neurons can be the basis to estimate the mean and the variance of sensory stimuli that vary over time.
 
-\subsection*{Weighting external and internal signals requires two sets of PE neurons}
+\subsection*{Estimating variances of sensory inputs and predictions requires a hierarchy of PE circuits}
 %
 Following the ideas of Bayesian multisensory integration (XXX), the weighting of sensory stimuli and predictions thereof would require knowledge of their variances.
-As we have shown in the previous section, the variance of the sensory stimulus can be estimated by means of PE neurons.
-We hypothesise that the same principles apply to estimating the variance of the prediction.
+As we have shown in the previous section, the variance of the sensory stimulus can be estimated using PE neurons.
+We hypothesise that the same principles apply to computing the variance of the prediction.
 Hence, we augment the network with a \textit{higher} PE circuit that receives output synapses from the M neuron of the \textit{lower} PE circuit (Fig.
-1xxx).
-Both subnetworks are modeled the same, except that the M neuron in the higher PE circuit evolves more slowly than the M neuron in the lower PE circuit.
+\ref{fig:Fig_1}D).
+Both subnetworks are modeled the same, except that the M neuron in the higher PE circuit evolves more slowly than the one in the lower PE circuit.
+%
+\begin{figure}[t!]
+	\centering
+	%\makebox[\textwidth][c]{\includegraphics{../results/figures/final/Fig_3}}
+    \includegraphics[width=1\linewidth]{../results/figures/final/Fig_3}
+\caption{\footnotesize{\bf Estimating variances of sensory inputs and predictions with hierarchical PE circuits.\newline} 
+{\bf (A)} Illustration of the stimulation protocol.
+Network is exposed to a sequence of stimuli (one stimulus per trial).
+To account for stimulus variability, each stimulus is represented by xxx stimulus values drawn from a normal distribution with mean $\mu_\mathrm{stim}$ and $\sigma_\mathrm{stim}^2$.
+To account for the volatility of the environment, in each trial the stimulus mean $\mu_\mathrm{stim}$ is drawn from a uniform distribution (denoted trial variability).
+Trial duration = xxx.
+{\bf (B)} Neuron activity increases with both stimulus and trial variability.
+Neurons in the lower PE circuit increase more strongly with stimulus variability.
+Neurons in the higher PE circuit increase more strongly with trial variability.
+{\bf (C)} Limit case example in which the stimulus variability is low but the trial variability is high.
+Left: Illustration of the stimulation protocol.
+Middle: Weighted output follows closely the sensory stimuli.
+Right: Sensory weight (function of the variances, see text) close to 1, indicating that the network ignores the prediction.
+Input statistics: XXX.
+{\bf (D)} Limit case example in which the stimulus variability is high but the trial variability is low.
+Left: Illustration of the stimulation protocol.
+Middle: Weighted output pushed towards the mean of the sensory stimuli.
+Right: Sensory weight close to zero, indicating that the network ignores the sensory stimuli.
+Input statistics: XXX.
+{\bf (E)} Predictions are weighted more strongly when the stimulus variability is larger than the trial variability.
+{\bf (F)} Predictions are weighted more strongly at the beginning of a new trial and quickly changing stimuli.
+}
+\label{fig:Fig_3}
+\end{figure}
+%
 
@@
 In each trial one stimulus is shown to the network.
-To account for the stimulus variance, each stimulus is composed of n constant values drawn from a normal distribution with mean $\mu_\mathrm{stimulus}$ and variance $\sigma_\mathrm{stimulus}^2$, and presented one after the other.
-To account for potential changes in the environment, in each trial, we draw $\mu_\mathrm{stimulus}$ from a uniform distribution (Fig.
-2a).
+To account for the stimulus variance, each stimulus is composed of \textit{n} constant values drawn from a normal distribution with mean $\mu_\mathrm{stim}$ and variance $\sigma_\mathrm{stim}^2$, and presented one after the other.
+To account for potential changes in the environment, in each trial, we draw $\mu_\mathrm{stim}$ from a uniform distribution (Fig.
+\ref{fig:Fig_3}A).
 Hence, the inputs change on two different time scales, with stimulus variability (faster time scale) and trial variability (slower time scale).
@@
 As expected, the neurons' activity increase for both stimulus and trial variances (Fig.
-2b).
+\ref{fig:Fig_3}B).
 While the neurons in the lower PE circuit increase more strongly with stimulus variability, the neurons in the higher PE circuit increase more strongly with trial variability, indicating that the different subnetworks process different aspects of the inputs.
@@
 In the first limit case, a different but low-variance stimulus is presented in each trial (Fig.
-2c, left).
+\ref{fig:Fig_3}C, left).
 In line with the ideas of multisensory integration (XXX), the network should therefore follow the sensory inputs closely and ignore the predictions.
 When we arithmetically calculate the weighted output (Fig.
-2c, middle) based on the feedforward and feedback inputs, and the sensory weight (Fig.
-2c, right), the network correctly represents mostly the sensory input (for more details, see Methods).
+\ref{fig:Fig_3}C, middle) based on the feedforward and feedback inputs, and the sensory weight (Fig.
+\ref{fig:Fig_3}C, right), the network correctly represents mostly the sensory input (for more details, see Methods).
 In the second limit case, the same but high-variance stimulus is presented in each trial (Fig.
-2d, left).
-According to theory (XXX), the network should downscale the sensory feedforward inputs and weight the prediction more strongly.
+\ref{fig:Fig_3}D, left).
+According to the theory, the network should downscale the sensory feedforward input and weight the prediction more strongly.
 Indeed, the weighted output of the network shows a clear tendency to the mean of the stimuli (Fig.
-2d, middle), also reflected in the low sensory weight (Fig.
-2d, right).
+\ref{fig:Fig_3}D, middle), also reflected in the low sensory weight (Fig.
+\ref{fig:Fig_3}D, right).
 
 In a next step, to validate the network responses more broadly, we systematically varied the trial and stimulus variability independently.
-If both variances are similar, the sensory weight approaches 0.5, reflecting equal contribution of sensory inputs and predictions to the weighted output.
+If both variances are similar, the sensory weight approaches \textit{0.5}, reflecting equal contribution of sensory inputs and predictions to the weighted output.
 Only if both variances are zero, the network represents the sensory input perfectly.
-In line with the limit case examples above, if the stimulus variance is larger than the trial variance, the network weighs the prediction more strongly than the sensory input.
+In line with the limit case examples above, if the stimulus variance is larger than the trial variance, the network weights the prediction more strongly than the sensory input.
 This is reversed if the stimulus variance is smaller than the trial variance (Fig.
-2e).
-Because the network dynamically estimates the mean and variances of the sensory inputs and the prediction, the weighted output and the sensory weight changes accordingly when the input statistics changes (Fig.
-SX).
+\ref{fig:Fig_3}E).
+Because the network dynamically estimates the mean and variances of the sensory input and the prediction, the weighted output and the sensory weight changes accordingly when the input statistics changes (Fig.
+\ref{fig:Fig_3_S1}).
 
 The first limit case (Fig.
-2 c) shows that even in a sensory-driven input regime, the prediction is weighted more at the beginning of a new trial than in the steady state.
+\ref{fig:Fig_3}C) shows that even in a sensory-driven input regime, the prediction is weighted more at the beginning of a new trial than in the steady state.
 This is further confirmed in simulations in which the trail duration was shortened.
 For those simulations, the prediction even outweighs the sensory input, reflected in a very low sensory weight (Fig.
-2e).
+\ref{fig:Fig_3}F).
 This suggests that predictions influence neural activity more significantly in experiments that rely on very fast stimulus changes.
 
-For some psychiatric disorders, it has been shown that the weighting of sensory inputs and predictions thereof is impaired (XXX), leading to an overweighting of one of these signals.
+It has been shown [speculated?], that sensory inputs or predictions are overrated in some psychiatric disorders (XXX).
+We thus wondered which network properties might bias the estimation of the variances, and, consequently, the weighting of different input streams.
+In our network, the M neuron evolves faster in the lower subnetwork than in the higher one.
+We identified the speed at which the M neurons are updated with new information as a decisive factor in the integration of inputs.
+To show this, we varied the weights from the PE neurons onto the lower-level M neuron.
+If the M neuron evolves too slowly, the prediction is overrated.
+In contrast, if the M neuron incorporates new information too quickly, the sensory input is overrated (Fig.
+\ref{fig:Fig_3_S2}A).
+While the speeds at which the activity of the M neurons evolve may underlie pathological weighting of inputs, the precise activation function of the V neurons is less pivotal.
+When we replaced the squared activation function with a linear, rectified function, the V neurons do not encode the variance but the averaged absolute deviation of the sensory stimuli.
+However, the sensory weight is only slightly shifted to larger values for low trial/high stimulus variability (Fig.
+\ref{fig:Fig_3_S2}B).
+
+In summary, we show that the variances of both the sensory inputs and predictions thereof can be dynamically computed in networks comprising a lower and higher PE circuit.
+The model shows that predictions are trusted more strongly at the beginning of a new stimulus, and if sensory inputs are noisy on a short time scale while predictable on longer time scales.
+
+% XXX Do we actually propose that predictions are sent up the hierarchy?
+
+\subsection*{Biasing the weighting of sensory inputs and predictions by neuromodulators}
+%
+The brain's flexibility and adaptability are not least because a plethora of neuromodulators influence the activity of neurons in a variety of ways (XXX).
+A prominent target of neuromodulatory inputs is inhibitory neurons (Cardin 2019, XXX).
+Moreover, distinct interneuron types are differently (in-)activated by those neuromodulators.
+For instance, it has been shown that XXX (XXX).
+We, therefore, wondered if and how the weighting of sensory inputs and predictions thereof may be biased when neuromodulators activate distinct interneuron types.
+%
+\begin{figure}[t!]
+	\centering
+	\makebox[\textwidth][c]{\includegraphics{../results/figures/final/Fig_4}}
+    %\includegraphics{../results/figures/final/Fig_4} % [width=1\linewidth]
+\caption{\footnotesize{\bf Neuromodulator-based shifts in the weighting of sensory inputs and predictions.
+\newline} 
+{\bf (A)} Neuromodulators acting on the three interneurons may shift the weighting of sensory inputs and predictions.
+The changes depend on the type/s of interneurons targeted by the additional excitatory input emulating a neuromodulator (additional input = XXX).
+Considered are two limit cases (upper row: more sensory-driven before modulation, lower row: more prediction-driven before modulation).
+The combination of interneurons targeted is illustrated below.
+The results are shown for three different PE circuits, specified in the main text.
+{\bf (B)} Illustration showing how the sensory weight depends on changes in both the stimulus and trial variability.
+{\bf (C)} The M and V neuron activities depend on the PE neuron activities.
+Hence, perturbing the nPE and pPE neurons must change the estimation of mean and variance.
+While stimulating the lower PE neurons affects both the lower and higher mean and variance estimation, stimulating the higher PE neurons only affects the V and M neurons in the same subnetwork.
+{\bf (D)} Illustration of the mechanisms underlying the biased estimation of mean and variance when PE neurons are perturbed.
+Both changes in the baseline (left) and gain (right) of PE neurons can contribute to the changes observed in (C).
+Illustration based on toy model described in Methods.
+{\bf (E)} Modulated interneurons change the weighting by changing the overall baseline and the overall gain of PE neurons (sum of changes in nPE and pPE neurons).
+Whether and how a neuromodulator changes the sensory weight, hence, depends on the interneuron targeted and the effect this interneuron has on baseline and gain of the PE neurons, which in turn does depend on the network it is embedded in.
+}
+\label{fig:Fig_4}
+\end{figure}
+%
+
+To this end, we modeled the presence of neuromodulators by injecting an additional excitatory input into one or two interneuron types.
+We reasoned that the network effect of a neuromodulator not only depends on the interneuron type it targets but also on the inputs this neuron receives and the connections it makes with other neurons in the network.
+We, therefore, tested three different mean-field networks that differ with respect to the distribution of sensory inputs and predictions onto the interneurons, and the underlying connectivity.
+The commonality across those networks is that they exhibit an E/I balance of excitatory and inhibitory pathways onto the PE neurons (XXX).
+Across the different mean-field networks tested, activating a SOM or VIP neuron individually forces the networks to weigh both inputs more equally.
+As a consequence, predictions are overrated in a sensory-driven input regime.
+Similarly, sensory inputs are overrated in a prediction-driven input regime.
+Interestingly, when both interneuron types are activated to the same degree, this effect disappears (Fig.
+\ref{fig:Fig_4}A, left).
+In contrast, stimulating PV neurons biases the network's output towards predictions.
+This effect is even more pronounced when PV and SOM, or PV and VIP neurons are activated simultaneously (Fig.
+\ref{fig:Fig_4}A, middle and right).
+
+In the previous simulations, we assumed that a neuromodulator acts globally, that is, on the interneurons in both the lower and the higher PE circuit.
+While this agrees with experimental data showing that XXX (XXX), we note that neuromodulators may also act more locally.
+The effect of stimulating an interneuron type in the lower PE circuit on the sensory weight is mostly the opposite of activating the same interneuron in the higher PE circuit (Fig.
+\ref{fig:Fig_4_S1}).
+For instance, the sensory inputs are overrated when the higher-level VIP neuron is activated, while the prediction is overrated when the lower-level VIP neuron is activated.
+When VIP and SOM neurons are stimulated equally, the sensory weight remains unchanged, independently of which PE circuit is targeted by the neuromodulator.
+
+What are the mechanisms that give rise to these effects?
+And how do the combined local changes give rise to the global one observed in our network simulations?
+The sensory weight is chosen to be a function of the V neurons of the lower and higher PE circuit.
+Hence, any changes to the sensory weight result from changes to the neurons encoding the variances (Fig.
+\ref{fig:Fig_4}B).
+In our network, the V neurons only receive excitatory output synapses from PE neurons.
+Hence, any changes in the sensory weights upon activation of interneurons must be due to changes in the PE neurons.
+To disentangle the effect of nPE and pPE neurons, we perturbed those neurons individually in both the lower or higher subnetwork by injecting either an inhibitory or excitatory additional input (Fig.
+\ref{fig:Fig_4}C).
+Stimulating either PE neuron in the lower subnetwork increases the activity of the lower-level V neuron strongly.
+Moreover, the higher-level V neuron is also slightly affected.
+At first, this is counterintuitive because the V neuron in the higher subnetwork does not receive direct synapses from the PE neurons in the lower subnetwork.
+However, the activity of the lower-level M neuron encoding the prediction increases with an excitatory input onto the pPE neuron and decreases with an excitatory input onto the nPE neuron (the opposite is true for an inhibitory input).
+Because neurons in the higher PE circuit receive synapses from the lower-level M neuron, the activity of the higher-level V neuron is also affected.
+In contrast, stimulating either PE neuron in the higher subnetwork increases the activity of the higher-level V neuron but leaves the lower-level M and V neurons unaffected (Fig.
+\ref{fig:Fig_4}C).
+
+Stimulating PE neurons may cause both an increase in the baseline activity and a change in the neuron's gain.
+To disentangle both effects, we illustrate each contribution separately using a mathematically tractable toy model (see Methods for more details).
+The variance estimated in the lower subnetwork increases with both increasing baseline and gain of the lower-level PE neurons.
+In contrast, when the gain of those PE neurons decreases, so does the variance.
+Similarly, the variance estimated in the higher subnetwork is equally influenced by changes in baseline and gain of the higher-level PE neurons.
+Moreover, changes to the baseline and the gain of the lower-level PE neurons increases the higher-level variance as a result of a biased prediction.
+Furthermore, the mean of the sensory stimuli is overpredicted (underpredicted) when the baseline or the gain of the lower-level pPE (nPE) neuron increases, or when the gain of the lower-level nPE (pPE) neuron decreases.
+In summary, if an interneuron causes an additional inhibitory input to a PE neuron, the neuron's gain is reduced.
+If an interneuron causes an additional disinhibitory input to a PE neuron, the neuron's baseline and gain are increased (Fig.
+\ref{fig:Fig_4}D).
+
+This suggests that to understand the effect of neuromodulators on the sensory weight, we need to unravel the effect of interneuron activation on baseline and gain of PE neurons.
+To this end, we stimulated either PV, SOM, or VIP neurons independently for all three mean-field networks and measured the changes to baseline and gain of both PE neurons (Fig.
+\ref{fig:Fig_4}E).
+In all three networks tested, activating PV neurons decreases both quantities, leading to a decrease in the estimated variance (Fig.
+\ref{fig:Fig_4}E \& Fig.
+\ref{fig:Fig_4_S2}).
+Stimulating SOM or VIP neurons decreases the overall gain but increases the baseline activity of the PE neurons.
+Whether and how much the gain of the nPE or pPE neuron is reduced depends on the inputs onto SOM and VIP neurons, and the connectivity they make with other neurons in the network.
+Similarly, how much the baseline is elevated depends on the specifics of the mean-field network (Fig.
+\ref{fig:Fig_4}E \& Fig.
+\ref{fig:Fig_4_S2}).
+Hence, whether stimulating the SOM or VIP neuron decreases or increases the activity of the V neuron depends on the input statistics: for low-mean stimuli, the elevated baseline activity dominates the changes in the variance, while for high-mean stimuli the changes in the gain dominate.
+
+%XXX Summary ...
+effect of neuromodulators acting on INs on sensory weight best understood as combined effect on the baseline and gain of nPE and pPE neurons that in turn determine the activity of the V neuron.
+Those effects not ony depend on the target IN but also the network in which it is embedded, that is the connectivity and the inputs onto INs.
+Moreover, it depends on whether local or global.
+The effects of activation on lower is mainly the opposite of same in higher.
+However, the efefcts don't cancel each other in global modulation.
+Mostly because 1) V higher additionally changes due to changes in prediction but also 2) effects in higher are stronger !?
+and 3) 1/(1+(Vs + D)/(Vp + D)) ...
+that is it cannot cancel per se.
+Even though the underlying reasons for a change in the V neuron depends on the the network characteristics, there are some consistent changes in sensory weigt across MFN ...
+name
+
+% in caption: you need to make sure that in e you used a perturbation of 1 and the mean and std of stimuli were ...
+the lines only illustrate that it depends on the input statistics.
+Of course, the markers would also change with other perturbaton strengths, would they though?
+Think through
+% Mention that for V neuron (higher) the darker line is actually not horizontal (unaffected)!
+% Mention that in BL illustration you need a factor to increase visibility ... (in caption)
+
+\subsection*{The contraction bias as a result of the weighted integration of sensory inputs and predictions}
+% 
+The weighted integration of sensory inputs and predictions thereof manifests in all-day behavior, in the form of a phenomenon called \textit{contraction bias}.
+The contraction bias describes the tendency to overestimate sensory stimuli drawn from the lower end of a stimulus distribution and to underestimate stimuli drawn from the upper end of the same distribution (Fig.
+\ref{fig:Fig_5}A).
+This \textit{bias towards the mean} has been reported in different species and modalities (XXX).
+%
+\begin{figure}[t!]
+	\centering
+    \includegraphics{../results/figures/final/Fig_5}
+\caption{\footnotesize{\bf Mechanisms underlying the contraction bias.\newline} 
+{\bf (A)} Illustration of the contraction bias.
+The estimated input is shifted towards the mean of the input distribution.
+Hence, a linear curve fitted to the data has a slope below 1 (left).
+The bias is biggest at the end of the stimulus distribution (right).
+{\bf (B)} Contraction bias in the model.
+Left: Example for two different stimulus variabilities.
+Right: As a consequence of the sensory weight, the slope decreases with stimulus variability (bias increases) and increases with trial variability (bias decreases).
+Stimulus statistics: XXX.
+{\bf (C)} The bias is independent of trial variability when the stimulus variability is zero (left).
+Equally, the bias is independent of the stimulus variability when the trial variability is zero (right).
+Stimulus statistics: XXX.
+{\bf (D)} The slope depends on the trial duration.
+If the sensory weight is 1, the slope is independent of the trail duration and the bias vanishes (dashed-dotted line).
+If the sensory weight is 0, the slope depends on the trial duration and only reaches 1 if the trial duration approaches infinity.
+{\bf (E)} To ensure a larger bias for stimuli drawn from the upper end of the stimulus distribution than from the lower end, scalar variability as observed experimentally is needed.
+}
+\label{fig:Fig_5}
+\end{figure}
+%
+
+The weighted output of our network can be interpreted as a neural manifestation of the contraction bias (see Methods for a thorough analysis).
+The bias increases with stimulus variance (Fig.
+\ref{fig:Fig_5}B), decreasing the slope of the linear fit modeling the relationship between the true and estimated stimuli (Fig.\ref{fig:Fig_5}B, right; compare with Fig.
+\ref{fig:Fig_5}A).
+In contrast, the bias decreases with trial variance, so that the slope of the linear fit approaches 1 (Fig.
+\ref{fig:Fig_5}B right).
+
+What are the underlying network factors that contribute to this phenomenon?
+To disentangle the potentially different sources of the bias, we first simulated a network without stimulus variability (variance set to zero) for two different trial variabilities.
+In this case, a contraction bias emerges but is independent of the volatility of the environment (Fig.
+\ref{fig:Fig_5}C, left).
+We show mathematically that the bias results from the transient neuron activity before reaching a steady state, and vanishes if the trial duration approaches infinity (see Methods for more details, and Fig.
+\ref{fig:Fig_5}D).
+We next resume the limit case in which the same but high-variance stimulus is shown in every trial.
+In this case, the weighted output exhibits a contraction bias that is largely independent of the stimulus variance (Fig.
+\ref{fig:Fig_5}C, right and Fig.
+\ref{fig:Fig_5}D).
+As shown mathematically (see Methods), the bias results from the finite trial duration and the tendency to weight the prediction more strongly than the sensory inputs.
+% Maybe show the limit for both cases and show the theoretical values for each T as well ... ?
+
+So far, we assumed that the stimulus variance is independent of the trial mean.
+A consequence of this choice is that the bias on either end of the stimulus distribution is largely the same (but with reversed signs).
+However, behavioral (neural?) data (XXX) shows that the bias increases for stimuli drawn from the upper end of the distribution, a phenomenon usually attributed to \textit{scalar variability}.
+To capture this in the model, we assume that the stimulus standard deviation linearly increases with the trial mean.
+In these simulations, as expected, the bias increases for a stimulus distribution shifted to higher trial means (Fig.
+\ref{fig:Fig_5}E).
+
+In summary, the weighted integration of sensory inputs and predictions can be interpreted as a neural manifestation of the contraction bias.
+Both stimulus and trial variability contribute to the contraction bias but the underlying mechanisms differ.
+
+% Enough? This section seems rather short! Should we do more, like making predictions about what would happen with the bias if neuromodulators activate INs? Like going back?
+
+\section*{Discussion}
+
+We solved the brain.
+
+% At the end, or in between talk about changes in PE neurons directly that you have currently in Fig. 3 SX ... I copy the part of the last paragraph here as inspiration :
+% For some psychiatric disorders, it has been shown that the weighting of sensory inputs and predictions thereof is impaired (XXX), leading to an overweighting of one of these signals.
 Moreover, it has been hypothesised that factors like stress or cognitive load may also influence the processing of feedforward and feedback inputs (XXX).
-Naturally, we were wondering which factors might influence the estimation of the variances, and, consequently, the weighting of different input streams.
-In our network, the M neuron evolves faster in the lower PE circuit than in the higher PE circuit.
-When we xxxx, the network shows a bias for predictions, while when we xxx, the network shows a bias for the sensory stimuli (Fig.
-SXX).
-This indicates that a non-pathological weighting requires xxx.
+...
 Another factor that may contribute to a distorted weighting is the baseline activity of PE neurons that was set to zero in our model, in line with very low spontaneous firing rates of excitatory neurons in rodent primary sensory areas (XXX).
@@
 
-XXX Summary XXX
-% I haven't talked about the different activation function but I don't see where this fits naturally ...
-
-% XXX Do we actually propose that predictions are sent up the hierarchy?
-
-\subsection*{XXX IN influence, neuromodulators ... mechanisms XXX}
-%
-The brain's flexibility and adaptability are not least due to the fact that a plethora of neuromodulators influence the activity of neurons in a variety of ways (XXX).
-A prominent target of neuromodulatory inputs are inhibitory neurons (Cardin 2019, XXX).
-Moreover, distinct interneuron types are differently (in-)activated by those neuromodulators.
-For instance, it has been shown that xxx.
-We were therefore wondering if and how the weighting of sensory inputs and predictions thereof may be biased when neuromodulators activate distinct interneuron types.
-
-To this end, we modeled the presence of neuromodulators by injecting an additional, excitatory input into one or two interneuron types.
-We reasoned that the network effect of a neuromodulator not only depends on the interneuron type it targets but also on the inputs this neuron receives and the connections it makes with other neurons in the network.
-We, therefore, tested three different mean-field networks that differ with respect to the distribution of sensory inputs and predictions onto the interneurons and the underlying connectivity.
-The commonality across those networks is that all of them exhibit an E/I balance of excitatory and inhibitory pathways onto the PE neurons (XXX).
-Across the different mean-field networks tested, activating a SOM or VIP neuron individually forces the networks to weigh both inputs more equally.
-As a consequence, in a sensory-driven input regime, predictions are overrated.
-Similarly, in a prediction-driven regime, sensory inputs are overrated.
-Interestingly, when both interneuron types are activated to the same degree, this effect disappears (Fig.
-5axx, left).
-In contrast, activating PV neurons largely biases the network output towards predictions.
-This is even more pronounced when PV and SOM, or PV and VIP neurons are activated simultaneously (Fig.
-5a, middle and right).
-
-In the previous simulations, we assumed that a neuromodulator acts globally, that is, on the interneurons in both the lower and the higher PE circuit.
-While this is in line with experimental data showing that xxx (XXX), we note that neuromodulators may also act more locally.
-We hence tested how the weighting of sensory stimuli and predictions may change when a neuromodulator activates specific interneurons in the lower or higher PE circuit only.
-The effect of activating an interneuron type in the lower PE circuit on the sensory weight is mostly the opposite of activating the same interneuron in the higher PE circuit (Fig.
-SXXX).
-For instance, the sensory inputs are overrated when VIP neurons are activated in the higher PE circuit, while the prediction is overrated when VIP neurons in the lower PE circuit are activated.
-When VIP and SOM neurons are stimulated equally, the sensory weight remains unchanged, independently of which PE circuit is targeted by the neuromodulator.
-
-What are the mechanisms that give rise to these effects? And how do the combined local changes give rise to the global one observed in our network simulations?
-
-% Mention that for V neuron (higher) the darker line is actually not horizontal (unaffected)!
-% Mention that in BL illustration you need a factor to increase visibility ... (in caption)
-
-\subsection*{XXX contraction bias}
-% 
-The weighting of sensory inputs and predictions thereof manifests in all-day behavior, in the form of a phenomenon called \textit{contraction bias}.
-The contraction bias describes the tendency to overestimate sensory stimuli that are at the lower end of a stimulus distribution and to underestimate stimuli that are at the upper end of the same distribution (Fig.
-5aXX).
-This \textit{bias towards the mean} has been reported in different species and modalities (XXX).
-
-The weighted output of our network can be interpreted as a neural manifestation of the contraction bias (see Methods for a thorough analysis).
-The bias increases with stimulus variance (Fig.
-5bXX), decreasing the slope of the linear fit modeling the relationship between the true and estimated stimuli (Fig.5 b, right XXX, compare with Fig.
-5a).
-In contrast, the bias decreases with trial variance, so that the slope of the linear fit approaches 1 (Fig.
-5b rightXXX).
-
-What are the underlying network factors that contribute to this phenomenon?
-To disentangle the potentially different sources of the bias, we first simulated a network without stimulus variability (variance set to zero) for two different trial variabilities.
-In this case, a contraction bias emerges but is independent of the volatility of the environment (Fig.
-5c, leftXXX).
-We show mathematically that the bias is a result of the recent stimulus history and vanishes if the trial duration approaches infinity (see Methods for more details, and Fig.
-5d).
-We next resume the limit case in which the same but high-variance stimulus is shown in every trial.
-In this case, the weighted output exhibits a contraction bias that is largely independent of the exact stimulus variances tested.
-As shown mathematically (see Methods), the bias is a result of xxx and approaches xxx (Fig.
-5dxxx).
-% Maybe show the limit for both cases and show the theoretical values for each T as well ...
-
-So far, we assumed that the stimulus variance is independent of the trial mean.
-A consequence of this choice is that the bias on either end of the stimulus distribution is the same (but with reversed signs).
-However, behavioral (neural?) data (XXX) shows that the bias increases for stimuli drawn from the upper end of the distribution, a phenomenon usually attributed to \textit{scalar variability}.
-To capture this in the model, we assume that the stimulus standard deviation linearly increases with the trial mean.
-In these simulations, as expected, the bias increases for a stimulus distribution shifted to higher trial means (Fig.
-5eXXX).
-
-XXX Summary XXX
-
-\section*{Discussion}
-
-We solved the brain.
+% mention Pakan paper
 
@@
 
-\subsection*{Network model}
+\subsection*{Mean-field network model}
 %
-Network consists of two subnetworks.
-Each subnetwork consists of a PE circuit, a memory neuron and a neuron representing the variance.
-XXX The memory neuron of subnetwork feefdowradly connects to the PE circuit of the second subnetwork.
+The mean-field network model consists of a \textit{lower} and \textit{higher} PE circuit (Fig.
+\ref{fig:Fig_1}).
+Each PE circuit contains an excitatory nPE neuron and pPE neuron ($\mathrm{N}_\mathrm{nPE} = \mathrm{N}_\mathrm{pPE} = 1$), as well as inhibitory neurons.
+The inhibitory neurons comprise PV, SOM and VIP neurons ($\mathrm{N}_\mathrm{S} = \mathrm{N}_\mathrm{V} = 1$, $\mathrm{N}_\mathrm{P} = 2$).
+Each subnetwork also contains a memory neuron $M$ and a variance neuron $V$.
+For the results in Figure 2, only the lower subnetwork is simulated.
 
@@
 %
-Consider a mean field network in which each population is represented by one representative neuron.
-The mean-field PE network consists of an excitatory nPE and pPE neuron, as well as two inhibitory PV neurons (one receiving S, the other P), as well as inhibitory SOM and VIP neurons.
-
 Each excitatory pyramidal cell (that is, nPE or pPE neuron) is divided into two coupled compartments, representing the soma and the dendrites, respectively.
-The dynamics of the firing rate~$r_{\mathrm{E}}$ of the somatic compartment of obeys \citep{wilson1972excitatory}
+The dynamics of the firing rates of the somatic compartments~$r_{\mathrm{nE}}$ (nPE neuron) and~$r_{\mathrm{pE}}$ (pPE neuron) obey \citep{wilson1972excitatory}
 %
 \begin{align}
-\tau_E\ \frac{dr_\mathrm{E}}{dt} &= - r_\mathrm{E} + w_\mathrm{ED}\cdot  r_\mathrm{D}  -  w_\mathrm{EP}\cdot r_\mathrm{P} + I_\mathrm{E},
+\tau_E\ \frac{dr_\mathrm{nE}}{dt} &= - r_\mathrm{nE} + w_\mathrm{nEnD}\cdot  r_\mathrm{nD}  -  w_\mathrm{nEP_1}\cdot r_\mathrm{P_1}  -  w_\mathrm{nEP_2}\cdot r_\mathrm{P_2} + I_\mathrm{nE}, \nonumber\\
+\tau_E\ \frac{dr_\mathrm{pE}}{dt} &= - r_\mathrm{pE} + w_\mathrm{pEpD}\cdot  r_\mathrm{pD}  -  w_\mathrm{pEP_1}\cdot r_\mathrm{P_1}  -  w_\mathrm{pEP_2}\cdot r_\mathrm{P_2} + I_\mathrm{pE}
 \end{align}
 %
-where $\tau_\mathrm{E}$ denotes the excitatory rate time constant ($\tau_\mathrm{E}$=60 ms), the weight $w_{\mathrm{ED}}$ describes the connection strength between the dendritic compartment and the soma of the same neuron, and $w_{\mathrm{EP}}$ denotes the strength of somatic inhibition from PV neurons.
-The overall input $I_\mathrm{E}$ comprises external background and feedforward sensory  inputs (see ``Inputs" below).
+where $\tau_\mathrm{E}$ denotes the excitatory rate time constant ($\tau_\mathrm{E}$=60 ms), the weights $w_{\mathrm{nEnD}}$ and $w_{\mathrm{pEpD}}$ describe the connection strength between the dendritic compartment and the soma of the same neuron, and $w_{\mathrm{nEP_1}}$, $w_{\mathrm{nEP_2}}$, $w_{\mathrm{pEP_1}}$ and $w_{\mathrm{pEP_2}}$ denote the strength of somatic inhibition from PV neurons.
+The overall input $I_\mathrm{nE}$ and $I_\mathrm{pE}$ comprise the external background and feedforward inputs (see ``Inputs" below).
 Firing rates are rectified to ensure positivity.
 
-The dynamics of the activity~$r_\mathrm{D}$ of the dendritic compartment obeys \citep{wilson1972excitatory}
+XXX continue here in the same spirit as above XXX The dynamics of the activity~$r_\mathrm{D}$ of the dendritic compartment obeys \citep{wilson1972excitatory}
 %
@@
 %
+Connectivity within PE circuit chosen from paper xxx. We consider mainly xxx but for supp figure we also simulated xxx and xxx. All connectivities unite that they represent E/I balance (xxx refs).
+
+Connectivity from PE to M and back (differences between lower and upper)
+
+Connectivity from PE to variance neuron.
+
 %All neurons are randomly connected with connection probabilities motivated by the experimental literature \citep[e.g.][]{fino2011dense, packer2011dense, pfeffer2013inhibition, lee2013disinhibitory, pi2013cortical, jiang2015principles, jouhanneau2015vivo, pala2015vivo},
@@
 %
+baseline input
+
+neurons may receive feedforward (while others recieve prediction from M neurons)
+
+feedforward input either sensory input (for lower) or activity of M neuron (for higher)
+
+For most of the results, FF goes to while the others receive activity of M
+But also tried other MFN for which input onto SOM and VIP reversed or the same
+
+For population model for the lower PE circuit, xxx.
+
 %All neurons receive external background input that ensures reasonable baseline firing rates in the absence of sensory inputs and predictions thereof.
@@
 %Source code to reproduce the simulations, analyses and figures will be available after publication at \url{github.com/lhertaeg/SourceCode_Hertaeg2021}.
+
+\section*{Population network model}
+XXXX
 
@@
 
+\section*{Supplementary Figures}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+\begin{figure}[!h]
+	\centering
+    \includegraphics[scale=1]{../results/figures/final/Fig_2_S1}% [width=1\linewidth]
+\caption{\footnotesize{\bf Estimation of mean and variance for different stimulus distributions.\newline}  
+Top: The mean-squared error (MSE) between the running average and the activity of the M neuron decreases to a near-zero level for all stimulus distributions tested. 
+Bottom: The MSE between the instantaneous variance and the V neuron decreases to a low level with minor differences between the distributions tested.
+Zoom-in shows the last half of the trial.
+Mean of the stimulus distribution = XXX, Variance of the stimulus distribution = XXX.
+}
+\label{fig:Fig_2_S1}
+\end{figure}
+
+\begin{figure}[!h]
+	\centering
+    \includegraphics{../results/figures/final/Fig_3_S1}% [width=1\linewidth]
+\caption{\footnotesize{\bf Dynamic variance estimation allows flexible adaptation to changes in the stimulus statistics and environment. \newline}  
+{\bf (A)} Sensory weight for different input statistics (same as in Fig. \ref{fig:Fig_3}E). Numbers denote specific example states. Arrows denote the transitions between those states.
+{\bf (B)} The sensory weight over time is shown for all transitions in (A). The switch to a new input statistics occurs at trial 60. Parameters are taken from Fig. \ref{fig:Fig_3}.
+}
+\label{fig:Fig_3_S1}
+\end{figure}
+
+\begin{figure}[!h]
+	\centering
+    \includegraphics{../results/figures/final/Fig_3_S2}% [width=1\linewidth]
+\caption{\footnotesize{\bf Perturbing the weighting of sensory inputs and predictions by altering  network properties. \newline}  
+{\bf (A)} The weights from the lower-level PE neurons to the M neuron are scaled by a factor below 1 (here, xxx) or above 1 (here, xxx), leading to a distorted weighting.
+If the update of the M neuron in the lower subnetwork is too slow ($\blacktriangleleft$), the prediction is overrated.
+If the update of the M neuron in the lower subnetwork is too fast ($\blacktriangleright$), the sensory input is overrated.
+{\bf (B)} The precise activation function for the V neurons does not have a major impact on the sensory weight.
+Only for inputs with high stimulus variability, the sensory stimulus is slightly overrated when the squared activation function is replaced by a linear, rectified activation function.
+}
+\label{fig:Fig_3_S2}
+\end{figure}
+
+\begin{figure}[!h]
+	\centering
+    \includegraphics{../results/figures/final/Fig_4_S1}% [width=1\linewidth]
+\caption{\footnotesize{\bf Neuromodulators acting locally either on interneurons in the lower or higher PE circuit. \newline}  
+{\bf (A)} Sensory weight changes with neuromodulators acting on interneurons in the lower PE circuit.
+{\bf (B)} Sensory weight changes with neuromodulators acting on interneurons in the lower PE circuit. Simulation parameters, labels and colors as in Fig. \ref{fig:Fig_4}. 
+}
+\label{fig:Fig_4_S1}
+\end{figure}
+
+\begin{figure}[!h]
+	\centering
+    \includegraphics{../results/figures/final/Fig_4_S2}% [width=1\linewidth]
+\caption{\footnotesize{\bf Perturbing the interneurons changes the baseline and gain of nPE and pPE neurons.\newline}  
+{\bf (A)} Changes in the baseline activity of nPE and pPE neurons for different interneurons targeted. 3 different mean-field networks are tested.
+{\bf (B)} Same as in (A) but for the gain of nPE and pPE neurons. Simulation parameters, labels, and colors as in Fig. \ref{fig:Fig_4}. 
+}
+\label{fig:Fig_4_S2}
+\end{figure}
+
 \end{document}

### 2023-08-11 scripts/plot_fig_2_support_2.py

@@ -8,3 +8,133 @@ Created on Mon Jan  9 09:04:01 2023
 
 # %% Import
 
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import os.path
+
+from src.plot_data import plot_M_and_V_for_population_example, plot_deviation_in_population_net
+
+# %% Universal parameters
+
+fs = 6
+inch = 2.54
+
+
+# %% Define files and paths
+
+figure_name = 'Fig_2_S2.png'
+figPath = '../results/figures/final/'
+
+if not os.path.exists(figPath):
+    os.mkdir(figPath)
+    
+    
+# %% Define figure structure
+
+figsize=(15/inch,10/inch)
+fig = plt.figure(figsize=figsize)
+
+G = gridspec.GridSpec(2, 1, figure=fig, hspace=1)
+G1 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=G[0,0], width_ratios=[2,1], wspace=0.5)
+G2 = gridspec.GridSpecFromSubplotSpec(1, 3, subplot_spec=G[1,0], wspace=0.4)
+
+ax_A = fig.add_subplot(G1[0,0])
+ax_A.axis('off')
+ax_A.text(-0.1, 1.2, 'A', transform=ax_A.transAxes, fontsize=fs+1)
+
+ax_B = fig.add_subplot(G1[0,1])
+ax_B.set_title('M and V neuron activity for a population network', fontsize=fs, pad=10)
+ax_B.text(-0.6, 1.2, 'B', transform=ax_B.transAxes, fontsize=fs+1)
+
+ax_D = fig.add_subplot(G2[0,1])
+plt.setp(ax_D.get_yticklabels(), visible=False)
+ax_D.set_title('Deviations for V neuron are \ndependent of correlated changes', fontsize=fs, pad=10)
+ax_D.text(-0.2, 1.3, 'D', transform=ax_D.transAxes, fontsize=fs+1)
+
+ax_C = fig.add_subplot(G2[0,0], sharey=ax_D)
+ax_C.set_title('Deviations are independent \nof uncorrelated changes', fontsize=fs, pad=10)
+ax_C.text(-0.2, 1.3, 'C', transform=ax_C.transAxes, fontsize=fs+1)
+
+ax_E = fig.add_subplot(G2[0,2], sharey=ax_D)
[…]

### 2023-08-11 scripts/run_fig_2_support_2.py

@@ -8,3 +8,490 @@ Created on Mon Jan  9 09:04:01 2023
 
 # %% Import
 
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+
+from src.default_parameters import Neurons, Activity_Zero, Network, Stimulation
+from src.functions_save import load_network_para
+from src.functions_networks import run_population_net
+
+dtype = np.float32
+
+
+# %% Sparsity 
+
+run_cell = False
+plot_only = True
+
+if run_cell:
+    
+    # filename to save relevant data
+    file_data = '../results/data/population/data_population_sparsity.pickle'
+    
+    if not plot_only:
+    
+        # get data
+        with open('../results/data/population/data_population_network_parameters.pickle','rb') as f:
+            [neurons_visual, inp_ext_soma, inp_ext_dend, weight_name, Dict_w, Dict_t] = pickle.load(f)
+        
+        with open('../results/data/population/data_population_network_para4corr_gain.pickle','rb') as f:
+            [_, _, gain_factors_nPE_5, gain_factors_pPE_5, nPE_true, pPE_true] = pickle.load(f)
+        
+        # conditions to be tested
+        p_conns = np.linspace(0.6, 1, 5)
+        num_seeds = 10
+        
+        # initialise arrays
+        M_steady_state = np.zeros((len(p_conns), num_seeds))
+        V_steady_state = np.zeros((len(p_conns), num_seeds))
+        
+        for cond_num, p_conn in enumerate(p_conns):
+            
+            # parametrisation neurons, network and rates
+            NeuPar = Neurons()
+            NetPar = Network(NeuPar, Dict_w, Dict_t, weight_name, neurons_visual, gain_factors_nPE = gain_factors_nPE_5, 
+                             gain_factors_pPE = gain_factors_pPE_5, nPE_true = nPE_true, pPE_true = pPE_true, p_conn=p_conn)
+            
+            for seed in range(num_seeds):
+                
+                print('p_conn = ', p_conn, ', seed = ', seed)
+                
+                # parametrisation rates and stimulation
+                RatePar = Activity_Zero(NeuPar)
+                StimPar = Stimulation(5, 2, inp_ext_soma, inp_ext_dend, neurons_visual, trial_duration = np.int32(100000), 
[…]

### 2023-08-11 scripts/run_pickle_save_new.py

@@ -1,0 +1,48 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Aug  9 08:03:15 2023
+
+@author: loreen.hertaeg
+"""
+
+# %% Import
+
+import pickle
+import numpy as np
+
+from src.default_parameters import Neurons, Activity_Zero, Network, Stimulation
+from src.functions_save import load_network_para
+
+import sys
+sys.path.append("../../PE_Circuits")
+
+# %% load pickle data and save in different format
+
+# load data
+path = '../../PE_Circuits/Results/Data/Plasticity'
+filename = path + '/Data_NetworkParameters_Example_Target_Input_After.pickle'
+
+with open(filename,'rb') as f:
+    NeuPar, NetPar, InPar, StimPar, SimPar, RatePar, LearnPar = pickle.load(f)
+    
+# rename and save differently
+weight_name = NetPar.weight_name
+neurons_visual = InPar.neurons_visual
+inp_ext_soma = InPar.inp_ext_soma
+inp_ext_dend = InPar.inp_ext_dend
+
+Dict_w = {}
+Dict_t = {}
+
+for i in range(25):
+    
+    m,n = np.unravel_index(i,(5,5))
+    exec('Dict_w["' + NetPar.weight_name[m][n] + '"] = NetPar.' + NetPar.weight_name[m][n])
+    exec('Dict_t["T' + NetPar.weight_name[m][n][1:] + '"] = NetPar.T' + NetPar.weight_name[m][n][1:])
+
+
+file_for_data = '../results/data/population/data_population_network_parameters.pickle'
+
+with open(file_for_data,'wb') as f:
+    pickle.dump([neurons_visual, inp_ext_soma, inp_ext_dend, weight_name, Dict_w, Dict_t],f)

### 2023-08-11 src/functions_networks.py

@@ -9,12 +9,14 @@ Created on Mon Jan  9 09:31:58 2023
 # %% Import
 
 import numpy as np
-import pickle
+import os
+from numba import njit
 
 dtype = np.float32
 
 # %% functions
 
+## functions for mean-field network/s
 
 def rate_dynamics_mfn(tau_E, tau_I, tc_var, w_var, U, V, W, rates, mean, 
                       var, feedforward_input, dt, n):
@@ -169,3 +171,168 @@ def run_mfn_circuit_coupled(w_PE_to_P, w_P_to_PE, w_PE_to_PE, v_PE_to_P, v_P_to_
     return ret
 
     
+## functions for population network/s
+
+#@njit(cache=True)
+def drdt(tau_inv_E, tau_inv_I, tau_inv_var, wEP, wED, wDS, wDE, wPE, wPP, wPS, wPV, wSE, wSP, wSS, wSV, wVE, wVP, wVS, wVV,
+         wEM, wDM, wPM, wSM, wVM, wME, wVarE, rE, rD, rP, rS, rV, r_mem, r_var, StimSoma_E, StimSoma_P, StimSoma_S, StimSoma_V, StimDend): 
+    
+    drE = tau_inv_E * (-rE + wED @ rD + wEP @ rP + wEM @ r_mem + StimSoma_E)
+    drD = tau_inv_E * (-rD + wDE @ rE + wDS @ rS + wDM @ r_mem + StimDend)
+    drP = tau_inv_I * (-rP + wPE @ rE + wPP @ rP + wPS @ rS + wPV @ rV + wPM @ r_mem + StimSoma_P)
+    drS = tau_inv_I * (-rS + wSE @ rE + wSP @ rP + wSS @ rS + wSV @ rV + wSM @ r_mem + StimSoma_S)
+    drV = tau_inv_I * (-rV + wVE @ rE + wVP @ rP + wVS @ rS + wVV @ rV + wVM @ r_mem + StimSoma_V)
+    
+    dr_mem = tau_inv_E * (wME @ rE)
+    dr_var = tau_inv_var * (-r_var + (wVarE @ rE)**2)
+    
+    return drE, drD, drP, drS, drV, dr_mem, dr_var
+
+
+
+def rate_dynamics(tau_inv_E, tau_inv_I, tau_inv_var, wEP, wED, wDS, wDE, wPE, wPP, wPS, wPV, wSE, wSP, wSS, wSV, wVE, wVP, wVS, wVV,
[…]

### 2023-08-11 src/functions_save.py

@@ -1,0 +1,26 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jan  9 09:31:58 2023
+
+@author: loreen.hertaeg
+"""
+
+# %% Import
+
+import os
+import pickle
+
+
+# %% functions
+
+
+def load_network_para(folder):
+    
+    path = '../results/data/' + folder
+    filename = path + '/data_population_network_parameters.pickle'
+    
+    with open(filename,'rb') as f:
+        NeuPar, NetPar, InPar, StimPar, SimPar, RatePar, LearnPar = pickle.load(f)
+        
+    return NeuPar, NetPar, InPar, StimPar, SimPar, RatePar, LearnPar

### 2023-08-18 records/manuscript_2023.tex
Add Supp Fig 3 for main figure 2

@@
 \begin{align}
-\tau_E\ \frac{dr_\mathrm{nE}}{dt} &= - r_\mathrm{nE} + w_\mathrm{nEnD}\cdot  r_\mathrm{nD}  -  w_\mathrm{nEP_1}\cdot r_\mathrm{P_1}  -  w_\mathrm{nEP_2}\cdot r_\mathrm{P_2} + I_\mathrm{nE}, \nonumber\\
-\tau_E\ \frac{dr_\mathrm{pE}}{dt} &= - r_\mathrm{pE} + w_\mathrm{pEpD}\cdot  r_\mathrm{pD}  -  w_\mathrm{pEP_1}\cdot r_\mathrm{P_1}  -  w_\mathrm{pEP_2}\cdot r_\mathrm{P_2} + I_\mathrm{pE}
+r_\mathrm{nE} = [h_\mathrm{nE}]_+ \ \mbox{ with }\ \tau_E\ \frac{dh_\mathrm{nE}}{dt} &= - h_\mathrm{nE} + w_\mathrm{nEnD}\cdot  r_\mathrm{nD}  -  w_\mathrm{nEP_1}\cdot r_\mathrm{P_1}  -  w_\mathrm{nEP_2}\cdot r_\mathrm{P_2} + I_\mathrm{nE}, \nonumber\\
+r_\mathrm{pE} = [h_\mathrm{pE}]_+ \ \mbox{ with }\ \tau_E\ \frac{dh_\mathrm{pE}}{dt} &= - h_\mathrm{pE} + w_\mathrm{pEpD}\cdot  r_\mathrm{pD}  -  w_\mathrm{pEP_1}\cdot r_\mathrm{P_1}  -  w_\mathrm{pEP_2}\cdot r_\mathrm{P_2} + I_\mathrm{pE}
 \end{align}
@@
 The overall input $I_\mathrm{nE}$ and $I_\mathrm{pE}$ comprise the external background and feedforward inputs (see ``Inputs" below).
-Firing rates are rectified to ensure positivity.
-
-XXX continue here in the same spirit as above XXX The dynamics of the activity~$r_\mathrm{D}$ of the dendritic compartment obeys \citep{wilson1972excitatory}
-%
-\begin{align}
-\tau_E\ \frac{dr_\mathrm{D}}{dt} &= - r_\mathrm{D} +  w_\mathrm{DE}\cdot r_\mathrm{E}  - w_\mathrm{DS}\cdot r_\mathrm{S} + I_\mathrm{D},
-\end{align}
-%
-where the weight $w_{\mathrm{DE}}$ denotes the recurrent excitatory connections between PCs, including backpropagating activity from the soma to the dendrites.
-$w_{\mathrm{DS}}$ represents the strength of dendritic inhibition from SOM neurons.
-The overall input $I_\mathrm{D}$ comprises fixed, external background inputs and feedback predictions (see ``Inputs" below).
+Firing rates are rectified to ensure positivity ($[\bullet]_+$).
+
+The dynamics of the activity of the dendritic compartments~$r_\mathrm{nD}$ (nPE neuron) and~$r_\mathrm{pD}$ (pPE neuron) obey \citep{wilson1972excitatory}
+%
+\begin{align}
+r_\mathrm{nE} = [h_\mathrm{nD}]_+ \ \mbox{ with }\ \tau_E\ \frac{dh_\mathrm{nD}}{dt} &= - h_\mathrm{nD} +  w_\mathrm{nDnE}\cdot r_\mathrm{nE} +  w_\mathrm{nDpE}\cdot r_\mathrm{pE}  - w_\mathrm{nDS}\cdot r_\mathrm{S} + I_\mathrm{nD}, \nonumber\\
+r_\mathrm{pD} = [h_\mathrm{pD}]_+ \ \mbox{ with }\ \tau_E\ \frac{dh_\mathrm{pD}}{dt} &= - h_\mathrm{pD} +  w_\mathrm{pDnE}\cdot r_\mathrm{pE} +  w_\mathrm{pDpE}\cdot r_\mathrm{pE}  - w_\mathrm{pDS}\cdot r_\mathrm{S} + I_\mathrm{pD}
+\end{align}
+%
+where the weights $w_{\mathrm{nDnE}}$, $w_{\mathrm{nDpE}}$, $w_{\mathrm{pDnE}}$ and $w_{\mathrm{pDpE}}$ denote the recurrent excitatory connections between PCs, including backpropagating activity from the soma to the dendrites.
+$w_{\mathrm{nDS}}$ and $w_{\mathrm{pDS}}$ represent the strength of dendritic inhibition from SOM neurons.
+The overall inputs $I_\mathrm{nD}$ and $I_\mathrm{pD}$ comprise fixed, external background inputs and feedback predictions (see ``Inputs" below).
 We assume that any excess of inhibition in a dendrite does not affect the soma, that is, the dendritic compartment is rectified at zero.
 
-Just as for the excitatory neurons, the firing rate dynamics of each interneuron is modeled by a rectified, linear differential equation \citep{wilson1972excitatory},
+Similarly, the firing rate dynamics of each interneuron is modeled by a rectified, linear differential equation,
 %
@@
 \label{eq:RateEqINs}
-\tau_I\ \frac{dr_\mathrm{X}}{dt} =& - r_\mathrm{X} + I_{\mathrm{X}} + w_\mathrm{XE}\cdot r_\mathrm{E} - w_\mathrm{XP}\cdot r_\mathrm{P}  - w_\mathrm{XS}\cdot r_\mathrm{S} -  w_\mathrm{XV}\cdot r_\mathrm{V},
-\end{align}
-%
-where $r_\mathrm{X}$ denotes the firing rate of neuron type $X$, and the weight matrices $w_\mathrm{XY}$ denote the strength of connection between the presynaptic neuron population $Y$ and the postsynaptic neuron population $X$ ($X, Y\in \lbrace P,S,V\rbrace$).
+r_\mathrm{X} =& [h_\mathrm{X}]_+ \ \mbox{ with }\ \tau_I\ \frac{dh_\mathrm{X}}{dt} = -h_\mathrm{X} + I_{\mathrm{X}} + w_\mathrm{XE}\cdot r_\mathrm{E} - w_\mathrm{XP_1}\cdot r_\mathrm{P_1} - w_\mathrm{XP_2}\cdot r_\mathrm{P_2}  - w_\mathrm{XS}\cdot r_\mathrm{S} -  w_\mathrm{XV}\cdot r_\mathrm{V},
+\end{align}
+%
+where $r_\mathrm{X}$ denotes the firing rate of interneuron type $X$, and the weight $w_\mathrm{XY}$ denotes the strength of connection between the presynaptic neuron $Y$ and the postsynaptic neuron $X$ ($X, Y\in \lbrace P_1, P_2, S, V\rbrace$).
 The rate time constant $\tau_I$ was chosen to resemble a fast GABA\textsubscript{A} time constant, and set to 2 ms for all interneuron types included.
@@
 %
-\begin{align}
-\tau_m \cdot \frac{dr_\mathrm{M}}{dt} = w_\mathrm{M\leftarrow pPE} \cdot r_\mathrm{pPE} - w_\mathrm{M\leftarrow nPE} \cdot r_\mathrm{nPE} 
-\end{align}
-%
-\begin{align}
-\tau_v \cdot \frac{dr_\mathrm{V}}{dt} = -r_\mathrm{V} + (w_\mathrm{V\leftarrow pPE} \cdot r_\mathrm{pPE} - w_\mathrm{V\leftarrow nPE} \cdot r_\mathrm{nPE})^2 
-\end{align}
+In addition to the core PE circuit, we simulate a memory neuron \textit{M} and a variance neuron \textit{V}.
+The memory neuron is modeled as a perfect integrator, receiving the output synapses from the nPE and pPE neurons,
+%
+\begin{align}
+\tau_M \cdot \frac{dr_\mathrm{M}}{dt} = w_\mathrm{MpE} \cdot r_\mathrm{pE} - w_\mathrm{MnE} \cdot r_\mathrm{nE}.
+\end{align}
+%
+$w_\mathrm{MpE}$ denotes the connection strength between the pPE neuron and the memory neuron, and  $w_\mathrm{MnE}$ denotes the connection strength between the nPE neuron and the memory neuron.
+The time constant XXX.
+
+The dynamics of the variance neuron obeys a non-linear differential equation with leak term,
+%
+\begin{align}
+\tau_V \cdot \frac{dr_\mathrm{V}}{dt} = -r_\mathrm{V} + (w_\mathrm{VpE} \cdot r_\mathrm{pE} + w_\mathrm{VnE} \cdot r_\mathrm{nE})^2.
+\end{align}
+%
+The weight $w_\mathrm{VpE}$ represents the connection strength between the pPE neuron and the variance neuron, while  $w_\mathrm{VnE}$ denotes the connection strength between the nPE neuron and the variance neuron.
+XXXX contiue here XXXX We chose a squared activation function ...
+but in xxx we also use a xxx this then is not the variance but xxx.
+the time constant was chosen to be xxx.
 

### 2023-08-18 scripts/run_fig_2_support_3.py

@@ -1,0 +1,249 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Aug 14 10:50:33 2023
+
+@author: loreen.hertaeg
+"""
+
+# %% Import
+
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+import pandas as pd
+
+from src.default_parameters import Neurons, Activity_Zero, Network, Stimulation
+from src.functions_save import load_network_para
+from src.functions_networks import run_population_net
+from src.functions_simulate import simulate_spatial_example
+from src.plot_data import plot_examples_spatial_M, plot_examples_spatial_V, plot_deviation_spatial
+
+dtype = np.float32
+
+
+
+# %% Test several input statistcis
+
+run_cell = False
+plot_only = True
+
+if run_cell:
+    
+    ### choose mean-field network to simulate
+    mfn_flag = '10' # valid options are '10', '01', '11
+    
+    ### filename for data
+    file_for_data = '../results/data/moments/data_spatial_mfn_diff_input_statistics.pickle'
+    file_temporary = '../results/data/moments/data_spatial_mfn_temporary.pickle'
+    
+    ### input statistic parameterisation
+    means_tested = np.linspace(3,6,7)
+    stds_tested = np.linspace(np.sqrt(4),np.sqrt(8), 5)
+    
+    ### initialise
+    deviation_mean = np.zeros((len(means_tested), len(stds_tested)))
+    deviation_std = np.zeros((len(means_tested), len(stds_tested)))
+    
+    ### get data
+    if not plot_only: # simulate respective network
+        
+        for i, mean in enumerate(means_tested):
+            for j, std in enumerate(stds_tested):
+        
+                print('Mean no.: ' , (i+1) , '/' , len(means_tested), ', Std no.: ', (j+1), '/', len(stds_tested))
+                [mean_stimulus, spatial_std, spatial_noise, 
+                 num_time_steps, m_neuron, v_neuron, rates_final] = simulate_spatial_example(mfn_flag, dtype(mean), dtype(std), file_temporary, 
+                                                                                num_sub_nets = np.int32(1000), num_time_steps = np.int32(4000))
+                                                                                             
+                deviation_mean[i,j] = (m_neuron[-1] - mean) / mean
[…]

### 2023-08-18 src/functions_networks.py

@@ -16,6 +16,95 @@ dtype = np.float32
 
 # %% functions
 
+## functions for "selective" uncoupled mean-field nets that share one M and V neuron
+
+def rate_dynamics_spatial_mfn(tau_E, tau_I, tc_var, Q, U, V, W, rates, mean, 
+                              var, feedforward_input, dt):
+    
+    # Initialise
+    rates_new = rates.copy() 
+    mean_new = mean.copy()
+    dr_1 = np.zeros(len(rates_new), dtype=dtype)
+    dr_2 = np.zeros(len(rates_new), dtype=dtype)
+    var_new = var.copy() 
+
+    # RK 2nd order
+    dr_mem_1 = (U @ rates_new) / tau_E
+    
+    dr_var_1 = (-var_new + (Q @ rates_new)**2) / tc_var 
+    #dr_var_1 = (-var_new + (Q @ rates_new**2)) / tc_var  
+    
+    dr_1 = -rates_new + W @ rates_new + V @ np.array([mean_new]) + feedforward_input
+    dr_1[:4] /= tau_E 
+    dr_1[4:] /= tau_I 
+
+    rates_new[:] += dt * dr_1 
+    rates_new[rates_new<0] = 0 # !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+    
+    mean_new += dt * dr_mem_1
+    var_new += dt * dr_var_1
+    
+    dr_mem_2 = (U @ rates_new) / tau_E
+    
+    dr_var_2 = (-var_new + (Q @ rates_new)**2) / tc_var
+    #dr_var_2 = (-var_new + (Q @ rates_new**2)) / tc_var
+    
+    dr_2 = -rates_new + W @ rates_new + V @ mean_new + feedforward_input
+    dr_2[:4] /= tau_E 
+    dr_2[4:] /= tau_I
+    
+    rates[:] += dt/2 * (dr_1 + dr_2)
+    mean += dt/2 * (dr_mem_1 + dr_mem_2)
+    var += dt/2 * (dr_var_1 + dr_var_2)
+    
+    # Rectify
+    rates[rates<0] = 0
+    
+    # print((Q @ rates_new**2))
+    # print(var)
+
+    return [rates, mean[0], var] # [rates, mean, var]
+
+
[…]

### 2023-08-18 src/functions_simulate.py

@@ -13,7 +13,7 @@ import pickle
 import matplotlib.pyplot as plt
 
 from src.default_parameters import default_para_mfn
-from src.functions_networks import run_mfn_circuit, run_mfn_circuit_coupled
+from src.functions_networks import run_mfn_circuit, run_mfn_circuit_coupled, run_spatial_mfn_circuit
 
 dtype = np.float32
 
@@ -84,6 +84,68 @@ def random_uniform_from_moments(mean, sd, num):
     rnd = dtype(np.random.uniform(a, b, size = num))
         
     return rnd
+
+
+
+def simulate_spatial_example(mfn_flag, mean_stimulus, spatial_std, file_for_data, num_sub_nets = np.int32(100), seed = 186, 
+                             num_time_steps = np.int32(1000), dist_type = 'uniform', pa=0.8, M_init = None, V_init = None, rates_init = None):
+    
+    ### load default parameters
+    VS, VV = int(mfn_flag[0]), int(mfn_flag[1])
+    
+    [w_PE_to_P, w_P_to_PE, w_PE_to_PE, w_PE_to_V, 
+     v_PE_to_P, v_P_to_PE, v_PE_to_PE, v_PE_to_V, 
+     tc_var_per_stim, tc_var_pred, tau_pe, fixed_input] = default_para_mfn(mfn_flag)
+    
+    ### create spatial noise
+    np.random.seed(seed)
+
+    if dist_type=='uniform':
+        spatial_noise = random_uniform_from_moments(0, spatial_std, num_sub_nets)
+        scaling = 2/(num_sub_nets * np.sqrt(3))
+    elif dist_type=='normal':
+        spatial_noise = np.random.normal(0, spatial_std, size=num_sub_nets)
+        scaling = 1/num_sub_nets # needs to be computed
+        #scaling = 2/(num_sub_nets * np.sqrt(3))
+    elif dist_type=='lognormal':
+        spatial_noise = random_lognormal_from_moments(0, spatial_std, num_sub_nets)
+        scaling = 1/num_sub_nets # needs to be computed
+    elif dist_type=='gamma':
+        spatial_noise = random_gamma_from_moments(0, spatial_std, num_sub_nets)
+        scaling = 1/num_sub_nets # needs to be computed
+    elif dist_type=='binary_equal_prop':
+        spatial_noise = random_binary_from_moments(0, spatial_std, num_sub_nets)
+        scaling = 1/num_sub_nets # needs to be computed
[…]

### 2023-08-18 src/plot_data.py

@@ -79,6 +79,110 @@ def lighten_color(color, amount=0.5):
     return colorsys.hls_to_rgb(c[0], 1 - amount * (1 - c[1]), c[2])
 
 
+def plot_deviation_spatial(deviation, means_tested, stds_tested, vmin=0, vmax=1, fs=6, show_mean = True, show_xlabel=True, ax=None):
+    
+    if ax is None:
+        _, ax = plt.subplots(1,1)
+        
+    if show_mean:
+        color_cbar = LinearSegmentedColormap.from_list(name='mse_mean', colors=['#FEFAE0', color_m_neuron])
+    else:
+        color_cbar = LinearSegmentedColormap.from_list(name='mse_variance', colors=['#FEFAE0', color_v_neuron])
+    
+    data = pd.DataFrame(abs(deviation)*100, index=np.round(means_tested,1), columns=np.round(stds_tested**2,1))
+    sns.heatmap(data, cmap = color_cbar, vmin=vmin, vmax=vmax, xticklabels=2, yticklabels=2,
+                cbar_kws={'label': '|Deviation| (%)', 'ticks':[vmin, vmax]}, ax=ax)
+    
+    cbar = ax.collections[0].colorbar
+    cbar.ax.tick_params(labelsize=fs)
+    cbar.ax.tick_params(size=2.0)
+    cbar.ax.yaxis.label.set_size(fs)
+    
+    ax.invert_yaxis()
+    if show_xlabel:
+        ax.set_xlabel('Mean', fontsize=fs) 
+    ax.set_ylabel('Variance', fontsize=fs) 
+    
+    ax.tick_params(size=2.0) 
+    ax.tick_params(axis='both', labelsize=fs)
+    ax.locator_params(nbins=3, axis='both')
+    
+    
+
+def plot_examples_spatial_V(num_time_steps, v_neuron_before, v_neuron_after, std_before, std_after, labels, ax=None, fs=6, lw=1):
+    
+    if ax is None:
+        _, ax = plt.subplots(1,1)
+    
+    colors = ['#FFBB33', '#E9724C']
+
+    ndim = v_neuron_after.ndim
+    
+    ax.plot(np.arange(num_time_steps), v_neuron_before, c=color_v_neuron, label=labels[0], lw=lw)
+    ax.axhline(std_before, 0, 0.5, ls=':', c=color_v_neuron, lw=lw)
+    
+    if ndim==1:
[…]

### 2023-08-25 records/manuscript_2023.tex
Writing Methods & Supps

--- 2023-08-11 records/manuscript_2023.tex
@@
 %
-
-\subsection*{Mean-field network model}
-%
+\subsection*{Network model}
 The mean-field network model consists of a \textit{lower} and \textit{higher} PE circuit (Fig.
@@
 Each PE circuit contains an excitatory nPE neuron and pPE neuron ($\mathrm{N}_\mathrm{nPE} = \mathrm{N}_\mathrm{pPE} = 1$), as well as inhibitory neurons.
-The inhibitory neurons comprise PV, SOM and VIP neurons ($\mathrm{N}_\mathrm{S} = \mathrm{N}_\mathrm{V} = 1$, $\mathrm{N}_\mathrm{P} = 2$).
-Each subnetwork also contains a memory neuron $M$ and a variance neuron $V$.
+The inhibitory neurons comprise PV, SOM and VIP neurons ($\mathrm{N}_\mathrm{SOM} = \mathrm{N}_\mathrm{VIP} = 1$, $\mathrm{N}_\mathrm{PV} = 2$).
+In addition to the core PE circuit, each subnetwork also includes one memory neuron $M$ and one variance neuron $V$.
+
+The excitatory neurons in the PE circuit are simulated as two coupled point compartments, representing the soma and the dendrites of elongated pyramidal cells.
+All other neurons are modeled as point neurons.
+The activities of all neurons are represented by a set of differential equations describing the network dynamics.
+The dynamics of the neurons in the lower and higher PE circuits ($\underline{r}_\mathrm{PE}^\mathrm{low}$ and $\underline{r}_\mathrm{PE}^\mathrm{high}$) are given by
+%
+\begin{align}
+\underline{r}_\mathrm{PE}^\mathrm{low} = & \left[ \underline{h}_\mathrm{PE}^\mathrm{low} \right]_+ \nonumber\\
+%
+\underline{r}_\mathrm{PE}^\mathrm{high} = & \left[ \underline{h}_\mathrm{PE}^\mathrm{high} \right]_+ 
+\end{align}
+%
+with
+%
+\begin{align}
+T \cdot \underline{\dot{h}}_\mathrm{PE}^\mathrm{low} =& -\underline{h}_\mathrm{PE}^\mathrm{low} + W_\mathrm{PE\leftarrow PE} \cdot \underline{r}_\mathrm{PE}^\mathrm{low} + \underline{w}_\mathrm{PE\leftarrow M} \cdot r_\mathrm{M}^\mathrm{low} + \underline{w}_\mathrm{PE\leftarrow FF} \cdot s + \underline{I}_\mathrm{PE} \nonumber\\
+%
+T \cdot \underline{\dot{h}}_\mathrm{PE}^\mathrm{high} =& -\underline{h}_\mathrm{PE}^\mathrm{high} + W_\mathrm{PE\leftarrow PE} \cdot \underline{r}_\mathrm{PE}^\mathrm{high} + \underline{w}_\mathrm{PE\leftarrow M} \cdot r_\mathrm{M}^\mathrm{high} + \underline{w}_\mathrm{PE\leftarrow FF} \cdot r_\mathrm{M}^\mathrm{low} + \underline{I}_\mathrm{PE}.
+%
+\end{align}
+%
+We follow the notation that column or row vectors are indicated by small letters with an underscore $\underline{\bullet}$, matrices are denoted by capital letters, and scalars are given by small letters without an underscore.
+Furthermore, a time derivative ($\frac{dx}{d}$) is denoted by a dot above the letter $\dot{x}$.
+The rate vector $\underline{r}_\mathrm{PE}^\mathrm{loc} = \left[r_\mathrm{nE}^\mathrm{loc},\ r_\mathrm{pE}^\mathrm{loc},\ r_\mathrm{nD}^\mathrm{loc},\ r_\mathrm{pD}^\mathrm{loc},\ r_\mathrm{PV_1}^\mathrm{loc}, r_\mathrm{PV_2}^\mathrm{loc},\ r_\mathrm{SOM}^\mathrm{loc}, r_\mathrm{VIP}^\mathrm{loc} \right]$  with $\mathrm{loc} \in [\mathrm{low}, \mathrm{high}]$ contains the activities of all neurons or compartments in the PE circuit (soma of nPE/pPE neurons: nE/pE, dendrites of nPE/pPE neurons: nD/pD).
+The network receives time-dependent stimuli $s$ and neuron/compartment-specific external background input $\underline{I}_\mathrm{PE}$.
+The connection strengths between the presynaptic neuron (population) $\bullet$ and the neurons of the PE circuit are denoted by $W_\mathrm{PE\leftarrow \bullet}$ and $\underline{w}_{PE \leftarrow \bullet}$, respectively.
+The activities of the neurons evolve with time constants contained in $T$.
+
+The activity of the lower and higher M neuron evolve according to a perfect integrator. The memory neuron receives  synapses from both nPE and pPE neurons of the same subnetwork,
+%
+\begin{align}
+\dot{r}_\mathrm{M}^\mathrm{low} =& \underline{w}_\mathrm{M \leftarrow PE}^\mathrm{low} \cdot \underline{r}_\mathrm{PE}^\mathrm{low} = w_\mathrm{M \leftarrow pPE}^\mathrm{low} \cdot r_\mathrm{pPE}^\mathrm{low} - w_\mathrm{M \leftarrow nPE}^\mathrm{low} \cdot r_\mathrm{nPE}^\mathrm{low} \nonumber \\
+%
+\dot{r}_\mathrm{M}^\mathrm{high} =& \underline{w}_\mathrm{M \leftarrow PE} ^\mathrm{high} \cdot \underline{r}_\mathrm{PE}^\mathrm{high} = w_\mathrm{M \leftarrow pPE}^\mathrm{high}  \cdot r_\mathrm{pPE}^\mathrm{high} - w_\mathrm{M \leftarrow nPE}^\mathrm{high}  \cdot r_\mathrm{nPE}^\mathrm{high} .
+\end{align}
+%
+
+The activity of the lower and higher V neuron evolve according to a leaky integrator with squared activation function.
+The variance neuron receives synapses from both nPE and pPE neurons of the same subnetwork,
+%
+\begin{align}
+\tau_\mathrm{V}^\mathrm{low} \cdot \dot{r}_\mathrm{V}^\mathrm{low} =& - r_\mathrm{V}^\mathrm{low} + \left( \underline{w}_\mathrm{V \leftarrow PE} \cdot \underline{r}_\mathrm{PE}^\mathrm{low}\right)^2 = - r_\mathrm{V}^\mathrm{low} + \left( w_\mathrm{V \leftarrow pPE} \cdot r_\mathrm{pPE}^\mathrm{low}\ + w_\mathrm{V \leftarrow nPE} \cdot r_\mathrm{nPE}^\mathrm{low}\right)^2 \nonumber\\
+%
+\tau_\mathrm{V}^\mathrm{high} \cdot \dot{r}_\mathrm{V}^\mathrm{high} =& - r_\mathrm{V}^\mathrm{high} + \left( \underline{w}_\mathrm{V \leftarrow PE} \cdot \underline{r}_\mathrm{PE}^\mathrm{high}\right)^2 = - r_\mathrm{V}^\mathrm{high} + \left( w_\mathrm{V \leftarrow pPE} \cdot r_\mathrm{pPE}^\mathrm{high}\ + w_\mathrm{V \leftarrow nPE} \cdot r_\mathrm{nPE}^\mathrm{high}\right)^2.
+\end{align} 
+%
+All values for neuron and network parameters, details on the model equations, as well as supporting analyses can be found in SI XXX.
+
+\subsection*{Weighting of sensory inputs and predictions}
+%
+We arithmetically calculate the weighted output of sensory inputs and predictions, $r_\mathrm{out}$, based on Bayesian multisensory integration (XXX),
+%
+\begin{align}
+r_\mathrm{out} = \alpha \cdot s + (1-\alpha) \cdot r_\mathrm{M}^\mathrm{low},
+\end{align}
+%
+where $\alpha$ denotes the reliability of the sensory input and is given by 
+%
+\begin{align}
+\alpha &= \left( 1 + \frac{r_\mathrm{V}^\mathrm{low}}{r_\mathrm{V}^\mathrm{high}} \right)^{-1}.
+\end{align}
+
+\subsection*{Inputs}
+%
+The network receives feedforward stimuli of duration XXX that can change over time.
+To account for noise, each stimulus is composed of N$_\mathrm{val}$ values drawn from a normal distribution with mean XXX and standard deviation XXX.
+Each of these values is given to the network as piece-wise constant input, one after the other.
+To account for changes in the environment, the mean of the normal distribution is drawn from a uniform distribution with mean XXX and standard deviation XXX.
+The parameterisation of both distributions varies across the experiments.
+All stimulus/input parameters can be found in SI XXX.
+
+\subsection*{Simulations}
+%
+All simulations were performed in customized Python code written by LH.
+Differential equations were numerically integrated using a 2\textsuperscript{nd}-order Runge-Kutta method with time steps ranging between $xx$ and $xxx$ ms.
+All neurons were initialized with $r=0/s$.
+We defined the steady-state firing rate per stimulus as the activity in the last XXXX ms of stimulus presentation.
+Further details and values for simulation parameters can be found in SI XXX.
+
+\section*{Acknowledgments}
+%
+%We are grateful to Vezha Boboeva, Douglas Feitosa Tom\'e, J\'ulia Gallinaro and Klara Kaleb for helpful comments on earlier versions of this manuscript, and we want to thank all members of the Clopath lab for discussion and support.
+This work was supported by BBSRC BB/N013956/1, BB/N019008/1, Wellcome Trust 200790/Z/16/Z, Simons Foundation 564408 and EPSRC EP/R035806/1.
+
+\bibliographystyle{naturemag} %{plainnat}
+\bibliography{../References_Mismatch}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% APPENDICES
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+\section*{Supplementary Information}
+\appendix
+
+Detailed methods + parameters, supporting analyses
+
+\subsection*{Network model}
+%
+The network model consists of a \textit{lower} and \textit{higher} mean-field PE circuit (Fig.
+\ref{fig:Fig_1}).
+Each PE circuit contains an excitatory nPE neuron and pPE neuron ($\mathrm{N}_\mathrm{nPE} = \mathrm{N}_\mathrm{pPE} = 1$), as well as inhibitory neurons.
+The inhibitory neurons comprise PV, SOM and VIP neurons ($\mathrm{N}_\mathrm{SOM} = \mathrm{N}_\mathrm{VIP} = 1$, $\mathrm{N}_\mathrm{PV} = 2$).
+In addition to the core PE circuit, each subnetwork also includes one memory neuron $M$ and one variance neuron $V$.
+
 For the results in Figure 2, only the lower subnetwork is simulated.
+In Supp Fig XXX, we replaced each mean-field PE circuit with a heterogeneous population model with $N_\mathrm{total} = 200$ neurons ($\mathrm{N}_\mathrm{SOM} = \mathrm{N}_\mathrm{VIP} = \mathrm{N}_\mathrm{PV} = 20$, the remaining neurons are excitatory).
+In Supp Fig.
+XXX, the PE circuit comprises $\mathrm{N}_\mathrm{copies} = XXX$ copies of the mean-field network to account for selectivity.
+
+In the following, we describe the dynamics of the neuron types for the mean-field network.
+The equations for the population PE circuit can be deduced from the mean-field equations and can also be found in XXX.
 
@@
 \begin{align}
-\tau_E\ \frac{dr_\mathrm{nE}}{dt} &= - r_\mathrm{nE} + w_\mathrm{nEnD}\cdot  r_\mathrm{nD}  -  w_\mathrm{nEP_1}\cdot r_\mathrm{P_1}  -  w_\mathrm{nEP_2}\cdot r_\mathrm{P_2} + I_\mathrm{nE}, \nonumber\\
-\tau_E\ \frac{dr_\mathrm{pE}}{dt} &= - r_\mathrm{pE} + w_\mathrm{pEpD}\cdot  r_\mathrm{pD}  -  w_\mathrm{pEP_1}\cdot r_\mathrm{P_1}  -  w_\mathrm{pEP_2}\cdot r_\mathrm{P_2} + I_\mathrm{pE}
-\end{align}
-%
-where $\tau_\mathrm{E}$ denotes the excitatory rate time constant ($\tau_\mathrm{E}$=60 ms), the weights $w_{\mathrm{nEnD}}$ and $w_{\mathrm{pEpD}}$ describe the connection strength between the dendritic compartment and the soma of the same neuron, and $w_{\mathrm{nEP_1}}$, $w_{\mathrm{nEP_2}}$, $w_{\mathrm{pEP_1}}$ and $w_{\mathrm{pEP_2}}$ denote the strength of somatic inhibition from PV neurons.
+r_\mathrm{nE} = [h_\mathrm{nE}]_+ \ \mbox{ with }\ \tau_E\ \frac{dh_\mathrm{nE}}{dt} &= - h_\mathrm{nE} + w_\mathrm{nE\leftarrow nD}\cdot  r_\mathrm{nD}  -  w_\mathrm{nE\leftarrow PV_1}\cdot r_\mathrm{PV_1}  -  w_\mathrm{nE\leftarrow PV_2}\cdot r_\mathrm{PV_2} + I_\mathrm{nE}, \nonumber\\
+r_\mathrm{pE} = [h_\mathrm{pE}]_+ \ \mbox{ with }\ \tau_E\ \frac{dh_\mathrm{pE}}{dt} &= - h_\mathrm{pE} + w_\mathrm{pE\leftarrow pD}\cdot  r_\mathrm{pD}  -  w_\mathrm{pE\leftarrow PV_1}\cdot r_\mathrm{PV_1}  -  w_\mathrm{pE\leftarrow PV_2}\cdot r_\mathrm{PV_2} + I_\mathrm{pE}
+\end{align}
+%
+where $\tau_\mathrm{E}$ denotes the excitatory rate time constant ($\tau_\mathrm{E}$=60 ms), the weights $w_{\mathrm{nE\leftarrow nD}}$ and $w_{\mathrm{pE\leftarrow pD}}$ describe the connection strength between the dendritic compartment and the soma of the same neuron, and $w_{\mathrm{nE\leftarrow PV_1}}$, $w_{\mathrm{nE\leftarrow PV_2}}$, $w_{\mathrm{pE\leftarrow PV_1}}$ and $w_{\mathrm{pE\leftarrow PV_2}}$ denote the strength of somatic inhibition from PV neurons.
 The overall input $I_\mathrm{nE}$ and $I_\mathrm{pE}$ comprise the external background and feedforward inputs (see ``Inputs" below).
-Firing rates are rectified to ensure positivity.
-
-XXX continue here in the same spirit as above XXX The dynamics of the activity~$r_\mathrm{D}$ of the dendritic compartment obeys \citep{wilson1972excitatory}
-%
-\begin{align}
-\tau_E\ \frac{dr_\mathrm{D}}{dt} &= - r_\mathrm{D} +  w_\mathrm{DE}\cdot r_\mathrm{E}  - w_\mathrm{DS}\cdot r_\mathrm{S} + I_\mathrm{D},
-\end{align}
-%
-where the weight $w_{\mathrm{DE}}$ denotes the recurrent excitatory connections between PCs, including backpropagating activity from the soma to the dendrites.
-$w_{\mathrm{DS}}$ represents the strength of dendritic inhibition from SOM neurons.
-The overall input $I_\mathrm{D}$ comprises fixed, external background inputs and feedback predictions (see ``Inputs" below).
+Firing rates are rectified to ensure positivity ($[\bullet]_+$).
+
+The dynamics of the activity of the dendritic compartments~$r_\mathrm{nD}$ (nPE neuron) and~$r_\mathrm{pD}$ (pPE neuron) obey \citep{wilson1972excitatory}
+%
+\begin{align}
+r_\mathrm{nD} = [h_\mathrm{nD}]_+ \ \mbox{ with }\ \tau_E\ \frac{dh_\mathrm{nD}}{dt} =& - h_\mathrm{nD} +  w_\mathrm{nD\leftarrow nE}\cdot r_\mathrm{nE} +  w_\mathrm{nD\leftarrow pE}\cdot r_\mathrm{pE} + w_\mathrm{nD\leftarrow M}\cdot  r_\mathrm{M} \nonumber\\
+& - w_\mathrm{nD\leftarrow SOM}\cdot r_\mathrm{SOM} + I_\mathrm{nD}, \nonumber\\
+r_\mathrm{pD} = [h_\mathrm{pD}]_+ \ \mbox{ with }\ \tau_E\ \frac{dh_\mathrm{pD}}{dt} =& - h_\mathrm{pD} +  w_\mathrm{pD\leftarrow nE}\cdot r_\mathrm{pE} +  w_\mathrm{pD\leftarrow pE}\cdot r_\mathrm{pE}+ w_\mathrm{pD\leftarrow M}\cdot  r_\mathrm{M}  \nonumber\\
+&- w_\mathrm{pD\leftarrow SOM}\cdot r_\mathrm{SOM} + I_\mathrm{pD},
+\end{align}
+%
+where the weights $w_{\mathrm{nD\leftarrow nE}}$, $w_{\mathrm{nD\leftarrow pE}}$, $w_{\mathrm{pD\leftarrow nE}}$ and $w_{\mathrm{pD\leftarrow pE}}$ denote the recurrent excitatory connections between PCs, including backpropagating activity from the soma to the dendrites.
+$w_{\mathrm{nD\leftarrow SOM}}$ and $w_{\mathrm{pD\leftarrow SOM}}$ represent the strength of dendritic inhibition from the SOM neuron.
+$w_{\mathrm{nD\leftarrow M}}$ and $w_{\mathrm{pD\leftarrow M}}$ denote the strength of connection between the memory neuron and the dendrites.
+The overall inputs $I_\mathrm{nD}$ and $I_\mathrm{pD}$ comprise fixed, external background inputs (see ``Inputs" below).
 We assume that any excess of inhibition in a dendrite does not affect the soma, that is, the dendritic compartment is rectified at zero.
 
-Just as for the excitatory neurons, the firing rate dynamics of each interneuron is modeled by a rectified, linear differential equation \citep{wilson1972excitatory},
+Similarly, the firing rate dynamics of each interneuron is modeled by a rectified, linear differential equation,
 %
@@
 \label{eq:RateEqINs}
-\tau_I\ \frac{dr_\mathrm{X}}{dt} =& - r_\mathrm{X} + I_{\mathrm{X}} + w_\mathrm{XE}\cdot r_\mathrm{E} - w_\mathrm{XP}\cdot r_\mathrm{P}  - w_\mathrm{XS}\cdot r_\mathrm{S} -  w_\mathrm{XV}\cdot r_\mathrm{V},
-\end{align}
-%
-where $r_\mathrm{X}$ denotes the firing rate of neuron type $X$, and the weight matrices $w_\mathrm{XY}$ denote the strength of connection between the presynaptic neuron population $Y$ and the postsynaptic neuron population $X$ ($X, Y\in \lbrace P,S,V\rbrace$).
+r_\mathrm{X} = [h_\mathrm{X}]_+ \ \mbox{ with }\ \tau_I\ \frac{dh_\mathrm{X}}{dt} =& -h_\mathrm{X} + I_{\mathrm{X}} + w_\mathrm{X\leftarrow nE}\cdot r_\mathrm{nE} + w_\mathrm{X\leftarrow pE}\cdot r_\mathrm{pE}  +  w_\mathrm{X\leftarrow M}\cdot  r_\mathrm{M} - w_\mathrm{X\leftarrow PV_1}\cdot r_\mathrm{PV_1}  \nonumber\\
+&- w_\mathrm{X\leftarrow PV_2}\cdot r_\mathrm{PV_2}  - w_\mathrm{X\leftarrow SOM}\cdot r_\mathrm{SOM} -  w_\mathrm{X\leftarrow VIP}\cdot r_\mathrm{VIP},
+\end{align}
+%
+where $r_\mathrm{X}$ denotes the firing rate of interneuron type $X$, and the weight $w_\mathrm{X\leftarrow Y}$ denotes the strength of connection between the presynaptic neuron $Y$ and the postsynaptic neuron $X$ ($X \in \lbrace \mathrm{PV_1}, \mathrm{PV_2}, \mathrm{SOM}, \mathrm{VIP}\rbrace$, $Y\in \lbrace \mathrm{nPE}, \mathrm{pPE}, \mathrm{PV_1}, \mathrm{PV_2}, \mathrm{SOM}, \mathrm{VIP}, \mathrm{M}\rbrace$).
 The rate time constant $\tau_I$ was chosen to resemble a fast GABA\textsubscript{A} time constant, and set to 2 ms for all interneuron types included.
-The overall input $I_\mathrm{X}$ comprises fixed, external background inputs, as well as feedforward sensory inputs and feedback predictions (see ``Inputs" below).
+The overall input $I_\mathrm{X}$ comprises fixed, external background inputs and feedforward sensory inputs (see ``Inputs" below).
 
@@
 %
-\begin{align}
-\tau_m \cdot \frac{dr_\mathrm{M}}{dt} = w_\mathrm{M\leftarrow pPE} \cdot r_\mathrm{pPE} - w_\mathrm{M\leftarrow nPE} \cdot r_\mathrm{nPE} 
-\end{align}
-%
-\begin{align}
-\tau_v \cdot \frac{dr_\mathrm{V}}{dt} = -r_\mathrm{V} + (w_\mathrm{V\leftarrow pPE} \cdot r_\mathrm{pPE} - w_\mathrm{V\leftarrow nPE} \cdot r_\mathrm{nPE})^2 
-\end{align}
+In addition to the core PE circuit, we simulate a memory neuron \textit{M} and a variance neuron \textit{V}.
+The memory neuron is modeled as a perfect integrator, receiving the output synapses from the nPE and pPE neurons,
+%
+\begin{align}
+\tau_M \cdot \frac{dr_\mathrm{M}}{dt} = w_\mathrm{M\leftarrow pE} \cdot r_\mathrm{pE} - w_\mathrm{M\leftarrow nE} \cdot r_\mathrm{nE}.
+\end{align}
+%
+$w_\mathrm{M\leftarrow pE}$ denotes the connection strength between the pPE neuron and the memory neuron, and  $w_\mathrm{M\leftarrow nE}$ denotes the connection strength between the nPE neuron and the memory neuron.
+The time constant XXX.
+
+The dynamics of the variance neuron obeys a non-linear differential equation with leak term,
+%
+\begin{align}
+\tau_V \cdot \frac{dr_\mathrm{V}}{dt} = -r_\mathrm{V} + (w_\mathrm{V\leftarrow pE} \cdot r_\mathrm{pE} + w_\mathrm{V\leftarrow nE} \cdot r_\mathrm{nE})^2.
+\end{align}
+%
+The weight $w_\mathrm{V\leftarrow pE}$ represents the connection strength between the pPE neuron and the variance neuron, while  $w_\mathrm{V\leftarrow nE}$ denotes the connection strength between the nPE neuron and the variance neuron.
+To ensure that the V neuron encodes the variance, we chose a squared activation function.
+In Fig.
+XXX, to investigate the necessity of a squared activation function in the context of weighting sensory inputs and predictions, we used a linear activation function.
+The time constant  $\tau_V$ was set to XXX.
 
@@
 %
-\begin{align}
-r_\mathrm{out} = \alpha \cdot S + (1-\alpha) \cdot P
-\end{align}
-%
-\begin{align}
-\alpha &= \frac{1/r_\mathrm{V1}}{1/r_\mathrm{V1} + 1/r_\mathrm{V2}}\nonumber\\
-& = \left( 1 + \frac{r_\mathrm{V1}}{r_\mathrm{V2}} \right)^{-1}
+The weighted output $r_\mathrm{out}(t) $ is a linear combination of the current sensory input $s(t)$ and the activity of the memory neuron, $r_\mathrm{M}(t)$, inspired by Bayesian multisensory integration (XXX)
+%
+\begin{align}
+r_\mathrm{out}(t) = \alpha \cdot s(t) + (1-\alpha) \cdot r_\mathrm{M}(t).
+\end{align}
+%
+How strongly either the sensory input or the prediction thereof contributes to the output is denoted by the sensory weight $\alpha$,
+%
+\begin{align}
+\alpha &= \frac{r_\mathrm{V_{lower}}^{-1}}{r_\mathrm{V_{lower}}^{-1} + r_\mathrm{V_{higher}}^{-1}}\nonumber\\
+& = \left( 1 + \frac{r_\mathrm{V_{lower}}}{r_\mathrm{V_{higher}}} \right)^{-1}
 \end{align}
@@
 %
-Connectivity within PE circuit chosen from paper xxx. We consider mainly xxx but for supp figure we also simulated xxx and xxx. All connectivities unite that they represent E/I balance (xxx refs).
-
-Connectivity from PE to M and back (differences between lower and upper)
-
-Connectivity from PE to variance neuron.
-
-%All neurons are randomly connected with connection probabilities motivated by the experimental literature \citep[e.g.][]{fino2011dense, packer2011dense, pfeffer2013inhibition, lee2013disinhibitory, pi2013cortical, jiang2015principles, jouhanneau2015vivo, pala2015vivo},
-%%
-%\begin{align}
-%p = \begin{pmatrix}
-%  p_\mathrm{EE}  & p_\mathrm{ED} & p_\mathrm{EP} & p_\mathrm{ES} & p_\mathrm{EV} \\
-%  p_\mathrm{DE}  & p_\mathrm{DD} & p_\mathrm{DP} & p_\mathrm{DS} & p_\mathrm{DV} \\
-%  p_\mathrm{PE}  & p_\mathrm{PD} & p_\mathrm{PP} & p_\mathrm{PS} & p_\mathrm{PV} \\
-%  p_\mathrm{SE}  & p_\mathrm{SD} & p_\mathrm{SP} & p_\mathrm{SS} & p_\mathrm{SV} \\
-%   p_\mathrm{VE} & p_\mathrm{VD} & p_\mathrm{VP} & p_\mathrm{VS} & p_\mathrm{VV} \\
-%\end{pmatrix}
-%=
-%\begin{pmatrix}
-%  - & 1 & 0.6 & - & -\\
-%  0.1  & - & - & 0.55 & - \\
-%  0.45  & - & 0.5 & 0.6 & 0.5 \\
-%  0.35  & - & - & - & 0.5 \\
-%   0.1 & - & - & 0.45 & - \\
-%\end{pmatrix}.
-%\label{Mtx:ConnProb}
-%\end{align}
-%%
-%All cells of the same neuron type have the same number of incoming connections. The mean total connection strengths are set to
-%%
-%\begin{align}
-%w = \begin{pmatrix}
-%  w_\mathrm{EE}  & w_\mathrm{ED} & w_\mathrm{EP} & w_\mathrm{ES} & w_\mathrm{EV} \\
-%  w_\mathrm{DE}  & w_\mathrm{DD} & w_\mathrm{DP} & w_\mathrm{DS} & w_\mathrm{DV} \\
-%  w_\mathrm{PE}  & w_\mathrm{PD} & w_\mathrm{PP} & w_\mathrm{PS} & w_\mathrm{PV} \\
-%  w_\mathrm{SE}  & w_\mathrm{SD} & w_\mathrm{SP} & w_\mathrm{SS} & w_\mathrm{SV} \\
-%   w_\mathrm{VE} & w_\mathrm{VD} & w_\mathrm{VP} & w_\mathrm{VS} & w_\mathrm{VV} \\
-%\end{pmatrix}
-%=
-%\begin{pmatrix}
-%  - & 1 & 2^{*} & - & -\\
-%  0.5 & - & - & 0.5^{*} & -\\
-%  1.2 & - & 1 & 0.3^{*} & 0.3^{*} \\
-%  1 & - & - & - & 0.6 \\
-%  1 & - & - & 0.7 & - \\
-%\end{pmatrix},
-%\label{Mtx:ConnStrengths}
-%\end{align}
-%%
-%where $^{*}$ denotes either the weights that are free for optimization to satisfy the equations describing an E/I balance (see Supporting Information), or the initial mean connection strengths that are subject to synaptic plasticity during learning.
-For plastic networks, the initial connections between neurons are drawn from uniform distributions
-%%
-%\begin{align*}
-%w_{ij}^{initial} \in \mathcal{U} \left(0.5\ w, 1.5\ w\right),
-%\end{align*}
-%%
-%where $w$ denotes the mean connection strengths given in (\ref{Mtx:ConnStrengths}).
-Please note that the system is robust to the choice of connection strengths.
-The connection strengths are merely chosen such that the solutions of the equations describing an E/I balance comply with Dale's principle.
-%
-%In plastic networks, $w_\mathrm{EP}$ is subdivided into assemblies.
-While one-third of PCs receive stronger initial inhibition from PV neurons that are driven by sensory input, another third receives stronger initial inhibition from PV neurons that are driven by feedback predictions.
-More precisely, for two-thirds of the excitatory neurons, half of the connections from PV neurons are strengthened by $1.5$, while the remaining ones are weakened by $0.5$.
-%
-%All weights are scaled in proportion to the number of existing connections (i.e., the product of the number of presynaptic neurons and the connection probability), so that the results are independent of the population size.
+%
+\subsubsection*{Connections between neurons of the PE circuit}
+%
+The connectivity between neurons of the PE circuit, both for the mean-field and the population network, was taken from XXX. XXX ADD heatmaps for the connectivity? XXX
+
+\subsubsection*{Connections between the PE circuit and the M neuron}
+%
+While the nPE neurons inhibit the M neuron, the pPE neurons excite it.
+To ensure that the activities of the memory neurons represent the mean of the sensory stimuli in the lower PE circuit and the mean of the prediction in the higher subnetwork, respectively, the net effect of nPE and pPE neurons must cancel in the steady state.
+Hence, the weights need to account for the neurons' potentially different gain factors ($g_\mathrm{nPE}$ and $g_\mathrm{pPE}$, see Table XXX) and the neuron numbers ($N_\mathrm{nPE}$ and $N_\mathrm{pPE}$, see Table XXX):
+%
+\begin{align*}
+w_\mathrm{M\leftarrow nE}\  &=\ \frac{-\lambda}{g_\mathrm{nPE} \cdot N_\mathrm{nPE}} \nonumber\\
+w_\mathrm{M\leftarrow pE}\  &=\ \frac{\lambda}{g_\mathrm{pPE} \cdot N_\mathrm{pPE}}
+\end{align*}
+%
+where $\lambda$ denotes the speed at which the perfect integrator evolves (lower PE circuit: $\lambda$=XXX, higher PE circuit:  $\lambda$=XXX).
+
+XXX Here you must say that this is true for the MFN and the one in which you copied the MFN but in case of the population, each neuron has a different gain, lambda is xxx for allXXX not sure how to show all the different gain factors?
+Also, you must say that the weights were systematically changed in Supp Fig.
+XXX to investigate correlated and uncorrelated changes XXX
+
+The memory neuron connects to a number of neurons in the PE circuit.
+Each connection from \textit{M} to a postsynaptic neuron \textit{X} is set to $w_\mathrm{X\leftarrow M} = 1$.
+In all mean-field networks and the population network, the dendrites of nPE and pPE neurons and one of the two (populations of) PV neurons receive connections from the memory neuron.
+We considered three mean-field networks that differed in terms of the inputs onto the SOM and VIP neurons:
+%
+\begin{table}[h!]
+\centering
+\begin{tabular}{ |c|c|c|c|  }
+\hline
+Network name & MFN 1 & MFN 2 & MFN 3  \\
+\hline
+\hline
+SOM & 0 & 1 & 0\\
+VIP & 1 & 0 & 0 \\
+\hline
+\end{tabular}
+\caption{XXX Table shows $w_\mathrm{X\leftarrow M}$ ... .}
+\end{table}
+%
+
+In the population model, 30\% of the SOM neurons and 70\% of the VIP neurons receive input from the memory neuron.
+
+\subsubsection*{Connections between the PE circuit and the V neuron}
+%
+Both nPE and pPE neurons excite the \textit{V} neuron.
+To ensure that the activity of the V neuron represents the variance of the input, the weights must account for differences in the gains ($g_\mathrm{nPE}$ and $g_\mathrm{pPE}$, see Table XXX) and numbers ($N_\mathrm{nPE}$ and $N_\mathrm{pPE}$) of the PE neurons,
+%
+\begin{align*}
+w_\mathrm{V\leftarrow nE}\  &=\ \frac{\theta}{g_\mathrm{nPE} \cdot N_\mathrm{nPE}} \nonumber\\
+w_\mathrm{V\leftarrow pE}\  &=\ \frac{}{g_\mathrm{pPE} \cdot N_\mathrm{pPE}}.
+\end{align*}
+%
+$\theta$ XXX represents a factor that ensures that for population network V still represents variance, no cross -terms, in mean-field network it is 1, XXX check if for population network, it is ok to ignore the nPE1*nPE2 terms, I mean you have cross-terms that do something with the variance, do they XXX
+
+We assume a quadratic activation function for the V neuron.
+In case of a mean-field network, the choice of the activation function ensures that the activity of the V neuron represents the variance of the input.
+This is only true because nPE and pPE neurons are not active at the same time and have a baseline activity of zero.
+=> cross-terms are zero.
+However, if we increase the number of nPE and pPE neurons, cross-terms of the nature $nPE_1 * nPE_2$ will introduce deviations from the the variance.
+These effects can be mitigated or fully corrected by choosing $\theta$ accordingly.
+$\theta$ depends on the distribution at hand.
+In our simulations, we only considered a uniform distribution.
+
+% XXX here start with nPE and pPE
+\begin{align}
+\left(\frac{\theta}{N} \sum_{i} |s_i - \mu| \right)^2 =& \left( \frac{\theta}{N}  \sum_{s_i \geq \mu}^{N/2} (s_i - \mu) + \frac{\theta}{N}  \sum_{s_i \leq \mu}^{N/2} (\mu - s_i )\right)^2 \nonumber \\
+ =& \frac{\theta^2}{4} \left( \frac{2}{N} \sum_{s_i \geq \mu}^{N/2} s_i - \frac{2}{N} \sum_{s_i \leq \mu}^{N/2} s_i\right)^2\nonumber\\
+ =& \frac{\theta^2}{4} \left( \frac{b+\mu}{2} - \frac{\mu + a}{2} \right)^2 = \frac{(b-a)^2}{16} \cdot \theta^2
+\end{align}
+%
+Hence, to ensure the variance of a uniform distribution given by $\frac{(b-a)^2}{12}$, $\theta = \frac{2}{\sqrt{3}}$.
 
@@
 %
-baseline input
-
-neurons may receive feedforward (while others recieve prediction from M neurons)
-
-feedforward input either sensory input (for lower) or activity of M neuron (for higher)
-
-For most of the results, FF goes to while the others receive activity of M
-But also tried other MFN for which input onto SOM and VIP reversed or the same
-
-For population model for the lower PE circuit, xxx.
+Each neuron (type) receives an overall input $I_i$
+%
+\begin{align*}
+I_i = I_i^{BL} + w_i \cdot I_{i}^{FF}
+\end{align*}
+%
+where 
+Feedforward is either the direct sensory input S for the lower PE circuit or the rM for the higher PE circuit.
+
+baseline input is XXX.
+
+show Wzs in form as table from above (rows different MFN, ...). Mention that population model is row 1, all of the SOM and VIP receive S or P, respectively then. Or simply write 1-wXM???
+%
+\begin{table}[h!]
+\centering
+\begin{tabular}{ |c|c|c|c|  }
+\hline
+Network name & MFN 1 & MFN 2 & MFN 3  \\
+\hline
+\hline
+SOM & 1 & 0 & 1\\
+VIP & 0 & 1 & 1 \\
+\hline
+\end{tabular}
+\caption{XXX Table shows $w_\mathrm{X\leftarrow input}$ ... .}
+\end{table}
+%
+
+Each stimulus was presented for $1$ second.
 
@@
 
-\section*{Population network model}
-XXXX
-
-\section*{Acknowledgments}
-%
-%We are grateful to Vezha Boboeva, Douglas Feitosa Tom\'e, J\'ulia Gallinaro and Klara Kaleb for helpful comments on earlier versions of this manuscript, and we want to thank all members of the Clopath lab for discussion and support.
-This work was supported by BBSRC BB/N013956/1, BB/N019008/1, Wellcome Trust 200790/Z/16/Z, Simons Foundation 564408 and EPSRC EP/R035806/1.
-
-\bibliographystyle{naturemag} %{plainnat}
-\bibliography{../References_Mismatch}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% APPENDICES
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\section*{Supplementary Information}
-\appendix
-
-\subsection*{Sensory weight and contraction bias}
+\subsection*{Supporting analyses}
+
+\subsubsection*{Sensory weight and contraction bias}
 %
@@
 
-\subsection*{Impact of PE neurons' gain on estimating mean and variance}
+\subsubsection*{Impact of PE neurons' gain on estimating mean and variance}
 %
@@
 
-\subsection*{Impact of PE neurons' baseline on estimating mean and variance}
+\subsubsection*{Impact of PE neurons' baseline on estimating mean and variance}
 %
@@
 
-\subsection*{Modelling the impact of neuromodulators on the weighting of sensory inputs and predictions thereof}
+\subsubsection*{Modelling the impact of neuromodulators on the weighting of sensory inputs and predictions thereof}
 
@@
 
-\subsection*{Influence of a population of nPE and pPE neurons}
+\subsubsection*{Influence of a population of nPE and pPE neurons}
 XXX
 
-\subsection*{Analysis of simplified network model, effect of time constants}
+\subsubsection*{Analysis of simplified network model, effect of time constants}
 
@@
 
-\subsection*{Comparison to Kalman filter and Bayes Factor surprise}
+\subsubsection*{Comparison to Kalman filter and Bayes Factor surprise}
 %

### 2023-08-25 scripts/run_fig_2_support_2.py

@@ -21,8 +21,8 @@ dtype = np.float32
 
 # %% Sparsity 
 
-run_cell = False
-plot_only = True
+run_cell = True
+plot_only = False
 
 if run_cell:
     
@@ -113,7 +113,7 @@ if run_cell:
 
 # %% Correlated deviations  
 
-run_cell = False
+run_cell = True
 plot_only = False
 
 if run_cell:
@@ -309,8 +309,8 @@ if run_cell:
     NetPar = Network(NeuPar, Dict_w, Dict_t, weight_name, neurons_visual, gain_factors_nPE = gain_factors_nPE_5, 
                      gain_factors_pPE = gain_factors_pPE_5, nPE_true = nPE_true, pPE_true = pPE_true)
     
-    StimPar = Stimulation(5, 2, inp_ext_soma, inp_ext_dend, neurons_visual, trial_duration = np.int32(100000), 
-                          num_values_per_trial = np.int32(200)) # 20000, 20
+    StimPar = Stimulation(5, 2, inp_ext_soma, inp_ext_dend, neurons_visual, trial_duration = np.int32(200000), 
+                          num_values_per_trial = np.int32(400)) # 20000, 20
     
     # run network
     run_population_net(NeuPar, NetPar, StimPar, RatePar, 0.1, folder='population', fln='Example')

### 2023-08-25 src/functions_networks.py

@@ -304,6 +304,12 @@ def rate_dynamics(tau_inv_E, tau_inv_I, tau_inv_var, wEP, wED, wDS, wDE, wPE, wP
     r_mem0[:] += dt * dr_mem1
     r_var0[:] += dt * dr_var1
     
+    rE0[rE0<0] = 0
+    rP0[rP0<0] = 0
+    rS0[rS0<0] = 0
+    rV0[rV0<0] = 0
+    rD0[rD0<0] = 0
+    
     drE2, drD2, drP2, drS2, drV2, dr_mem2, dr_var2 = drdt(tau_inv_E, tau_inv_I, tau_inv_var, wEP, wED, wDS, wDE, wPE, wPP, wPS, wPV, 
                                                           wSE, wSP, wSS, wSV, wVE, wVP, wVS, wVV, wEM, wDM, wPM, wSM, wVM, wME, wVarE, 
                                                           rE0, rD0, rP0, rS0, rV0, r_mem0, r_var0, StimSoma_E, StimSoma_P, StimSoma_S, 

### 2023-09-08 records/manuscript_2023.tex
housekeeping

@@
 \usepackage{hyperref}
+
+% change name of Table of contents
+\renewcommand*\contentsname{Supporting Information}
 
@@
 The activities of all neurons are represented by a set of differential equations describing the network dynamics.
+
 The dynamics of the neurons in the lower and higher PE circuits ($\underline{r}_\mathrm{PE}^\mathrm{low}$ and $\underline{r}_\mathrm{PE}^\mathrm{high}$) are given by
@@
 %
-We follow the notation that column or row vectors are indicated by small letters with an underscore $\underline{\bullet}$, matrices are denoted by capital letters, and scalars are given by small letters without an underscore.
-Furthermore, a time derivative ($\frac{dx}{d}$) is denoted by a dot above the letter $\dot{x}$.
+We follow the notation that column and row vectors are indicated by small letters with an underscore $\underline{\bullet}$, matrices are denoted by capital letters, and scalars are given by small letters without an underscore.
+Furthermore, a time derivative (e.g., $\frac{dx}{dt}$) is denoted by a dot above the letter (e.g., $\dot{x}$).
 The rate vector $\underline{r}_\mathrm{PE}^\mathrm{loc} = \left[r_\mathrm{nE}^\mathrm{loc},\ r_\mathrm{pE}^\mathrm{loc},\ r_\mathrm{nD}^\mathrm{loc},\ r_\mathrm{pD}^\mathrm{loc},\ r_\mathrm{PV_1}^\mathrm{loc}, r_\mathrm{PV_2}^\mathrm{loc},\ r_\mathrm{SOM}^\mathrm{loc}, r_\mathrm{VIP}^\mathrm{loc} \right]$  with $\mathrm{loc} \in [\mathrm{low}, \mathrm{high}]$ contains the activities of all neurons or compartments in the PE circuit (soma of nPE/pPE neurons: nE/pE, dendrites of nPE/pPE neurons: nD/pD).
 The network receives time-dependent stimuli $s$ and neuron/compartment-specific external background input $\underline{I}_\mathrm{PE}$.
-The connection strengths between the presynaptic neuron (population) $\bullet$ and the neurons of the PE circuit are denoted by $W_\mathrm{PE\leftarrow \bullet}$ and $\underline{w}_{PE \leftarrow \bullet}$, respectively.
+The connection strengths between the \textit{pre}-synaptic neuron (population) and the neurons of the PE circuit are denoted by $W_\mathrm{PE\leftarrow pre}$ and $\underline{w}_{\mathrm{PE} \leftarrow pre}$, respectively.
 The activities of the neurons evolve with time constants contained in $T$.
 
-The activity of the lower and higher M neuron evolve according to a perfect integrator. The memory neuron receives  synapses from both nPE and pPE neurons of the same subnetwork,
+The activities of the lower and higher M neuron evolve according to a perfect integrator. The memory neurons receive synapses from both nPE and pPE neurons of the same subnetwork,
 %
@@
 
-The activity of the lower and higher V neuron evolve according to a leaky integrator with squared activation function.
-The variance neuron receives synapses from both nPE and pPE neurons of the same subnetwork,
+The activities of the lower and higher V neuron evolve according to a leaky integrator with quadratic activation function.
+The variance neurons receive synapses from both nPE and pPE neurons of the same subnetwork,
 %
@@
 %
-The network receives feedforward stimuli of duration XXX that can change over time.
-To account for noise, each stimulus is composed of N$_\mathrm{val}$ values drawn from a normal distribution with mean XXX and standard deviation XXX.
-Each of these values is given to the network as piece-wise constant input, one after the other.
+The network receives feedforward stimuli of duration XXX that can change from trial to trial (one stimulus per trial).
+To account for noise, each stimulus is composed of N$_\mathrm{val}$ piece-wise constant values drawn from a normal distribution with mean XXX and standard deviation XXX.
 To account for changes in the environment, the mean of the normal distribution is drawn from a uniform distribution with mean XXX and standard deviation XXX.
@@
 
-\newpage
-\section*{Supplementary Information}
+\clearpage
 \appendix
 
-Detailed methods + parameters, supporting analyses
-
-\subsection*{Network model}
+\tableofcontents
+\section{Detailed Methods}
+
+XXX
+
+\subsection{Network model}
 %
@@
 
-\subsubsection*{Prediction-error network model}
+\subsubsection{Prediction-error network model}
 %
@@
 
-\subsubsection*{Memory and variance neuron}
+\subsubsection{Memory and variance neuron}
 %
@@
 
-\subsubsection*{Weighted output}
+\subsubsection{Weighted output}
 %
@@
 
-\subsection*{Connectivity}
-%
-%
-\subsubsection*{Connections between neurons of the PE circuit}
+\subsection{Connectivity}
+%
+%
+\subsubsection{Connections between neurons of the PE circuit}
 %
@@
 
-\subsubsection*{Connections between the PE circuit and the M neuron}
+\subsubsection{Connections between the PE circuit and the M neuron}
 %
@@
 
-\subsubsection*{Connections between the PE circuit and the V neuron}
+\subsubsection{Connections between the PE circuit and the V neuron}
 %
@@
 w_\mathrm{V\leftarrow nE}\  &=\ \frac{\theta}{g_\mathrm{nPE} \cdot N_\mathrm{nPE}} \nonumber\\
-w_\mathrm{V\leftarrow pE}\  &=\ \frac{}{g_\mathrm{pPE} \cdot N_\mathrm{pPE}}.
+w_\mathrm{V\leftarrow pE}\  &=\ \frac{\theta}{g_\mathrm{pPE} \cdot N_\mathrm{pPE}}.
 \end{align*}
@@
 
-\subsection*{Inputs}
+\subsection{Inputs}
 %
@@
 
-\subsection*{Simulations}
+\subsection{Simulations}
 %
@@
 
-\subsection*{Supporting analyses}
-
-\subsubsection*{Sensory weight and contraction bias}
-%
-If P is rather constant, the slope in the contraction bias is exactly the sensory weight
-%
- \begin{align*}
- r_\mathrm{out} = \alpha_\mathrm{S} \cdot S + \left( 1 -\alpha_\mathrm{S} \right) \cdot P \equiv m \cdot S + n
- \end{align*}
-%
-However, P is usually/normally a function of $S$. For simplicity, let's assume that P decays exponentially to a new value of $S$:
-%
-\begin{align*}
-P = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau} + f(S) \cdot \left( 1 -   \mathrm{e}^{-t/\tau} \right)
-\end{align*}
-%
-Within each trial with trial duration T, P can be expressed by n sections of length t in which the stimulus is constant and, for the sake of simplicity, drawn from a uniform distribution $U\left( s - \frac{\sigma_\mathrm{S}}{12}, s + \frac{\sigma_\mathrm{S}}{12} \right)$.
-$P_\mathrm{0}$ is drawn from $U\left( \mu - \frac{\sigma_\mathrm{P}}{12}, \mu + \frac{\sigma_\mathrm{P}}{12} \right)$.
-$P_\mathrm{n}$ is then given by
-%
-\begin{align*}
-P_\mathrm{n} = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau}  + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} s_i \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau}
-\end{align*}
-%
-This needs to be averaged over all possible states
-%
-\begin{align*}
-P_\mathrm{n} = \mathrm{e}^{-t/\tau}  \int\limits_{\mu - \frac{\sigma_\mathrm{P}}{12}}^{\mu + \frac{\sigma_\mathrm{P}}{12}} P_\mathrm{0} \ f(P_\mathrm{0})\ dP_0+ \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau} \int\limits_{s - \frac{\sigma_\mathrm{S}}{12}}^{s + \frac{\sigma_\mathrm{S}}{12}} s\ f(s)\ ds
-\end{align*}
-%
-This gives
-%
-\begin{align*}
-P_\mathrm{n} = \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot t/ \tau} \cdot S
-\end{align*}
-%
-By making use of the geometric series, this simplifies to
-%
-\begin{align*}
-P_\mathrm{n} =  \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right) \cdot S
-\end{align*}
-%
-Together, this yields for the weighted output
-%
-\begin{align*}
- r_\mathrm{out} = \left[ \alpha_\mathrm{S}\  \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right)\right] \cdot S + \left( 1 -\alpha_\mathrm{S} \right)\ \mathrm{e}^{-T/\tau}\ \mu
-\end{align*}
-%
-Hence, the slope is a function of both the sensory weight and the trial duration.
-
-In a prediction-driven input regime ($alpha_\mathrm{S} \sim 0)$, the slope is independent of the sensory weight and only determined by the trial duration, $m \sim \left( 1 -   \mathrm{e}^{-T/\tau} \right)$.
-In a sensory-driven input regime ($alpha_\mathrm{S} \sim 0)$, the contraction bias vanishes ($m \sim 1$).
-
-If the trail duration is short ($T \rightarrow 0$), the slope is given by the sensory weight.
-If the trail duration approaches infinity, the slope would be 1 again (however, this seems rather unrealistic, this would only be true in an ideal system without memory decay or reproduction and accumulation noise ...).
-
-\subsubsection*{Impact of PE neurons' gain on estimating mean and variance}
+\section{Supporting analyses}
+
+%3. Sensory weight and contraction bias
+%4. Comparison to Kalman filter and Bayes Factor surprise
+
+\subsection{XXX}
+%0. rM exponential average, rV exponential average (see iPad, currently page 107)
+
+\subsection{Impact of PE neurons' gain on estimating mean and variance}
 %
@@
 
-\subsubsection*{Impact of PE neurons' baseline on estimating mean and variance}
+\subsection{Impact of PE neurons' baseline on estimating mean and variance}
 %
@@
 
-\subsubsection*{Modelling the impact of neuromodulators on the weighting of sensory inputs and predictions thereof}
+\subsection{Modelling the impact of neuromodulators on the sensory weight}
 
@@
 
-\subsubsection*{Influence of a population of nPE and pPE neurons}
-XXX
-
-\subsubsection*{Analysis of simplified network model, effect of time constants}
-
-simplified model: dynamics and steady state of rM and rV, rM and rV as a function of time constants and trial duration etc., weighting, then use those expressions to discuss when weighting goes awry and how long transitions take from one state to another ...
-
-\subsubsection*{Comparison to Kalman filter and Bayes Factor surprise}
+\subsection*{Sensory weight and contraction bias}
+%
+If P is rather constant, the slope in the contraction bias is exactly the sensory weight
+%
+ \begin{align*}
+ r_\mathrm{out} = \alpha_\mathrm{S} \cdot S + \left( 1 -\alpha_\mathrm{S} \right) \cdot P \equiv m \cdot S + n
+ \end{align*}
+%
+However, P is usually/normally a function of $S$. For simplicity, let's assume that P decays exponentially to a new value of $S$:
+%
+\begin{align*}
+P = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau} + f(S) \cdot \left( 1 -   \mathrm{e}^{-t/\tau} \right)
+\end{align*}
+%
+Within each trial with trial duration T, P can be expressed by n sections of length t in which the stimulus is constant and, for the sake of simplicity, drawn from a uniform distribution $U\left( s - \frac{\sigma_\mathrm{S}}{12}, s + \frac{\sigma_\mathrm{S}}{12} \right)$.
+$P_\mathrm{0}$ is drawn from $U\left( \mu - \frac{\sigma_\mathrm{P}}{12}, \mu + \frac{\sigma_\mathrm{P}}{12} \right)$.
+$P_\mathrm{n}$ is then given by
+%
+\begin{align*}
+P_\mathrm{n} = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau}  + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} s_i \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau}
+\end{align*}
+%
+This needs to be averaged over all possible states
+%
+\begin{align*}
+P_\mathrm{n} = \mathrm{e}^{-t/\tau}  \int\limits_{\mu - \frac{\sigma_\mathrm{P}}{12}}^{\mu + \frac{\sigma_\mathrm{P}}{12}} P_\mathrm{0} \ f(P_\mathrm{0})\ dP_0+ \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau} \int\limits_{s - \frac{\sigma_\mathrm{S}}{12}}^{s + \frac{\sigma_\mathrm{S}}{12}} s\ f(s)\ ds
+\end{align*}
+%
+This gives
+%
+\begin{align*}
+P_\mathrm{n} = \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot t/ \tau} \cdot S
+\end{align*}
+%
+By making use of the geometric series, this simplifies to
+%
+\begin{align*}
+P_\mathrm{n} =  \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right) \cdot S
+\end{align*}
+%
+Together, this yields for the weighted output
+%
+\begin{align*}
+ r_\mathrm{out} = \left[ \alpha_\mathrm{S}\  \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right)\right] \cdot S + \left( 1 -\alpha_\mathrm{S} \right)\ \mathrm{e}^{-T/\tau}\ \mu
+\end{align*}
+%
+Hence, the slope is a function of both the sensory weight and the trial duration.
+
+In a prediction-driven input regime ($alpha_\mathrm{S} \sim 0)$, the slope is independent of the sensory weight and only determined by the trial duration, $m \sim \left( 1 -   \mathrm{e}^{-T/\tau} \right)$.
+In a sensory-driven input regime ($alpha_\mathrm{S} \sim 0)$, the contraction bias vanishes ($m \sim 1$).
+
+If the trail duration is short ($T \rightarrow 0$), the slope is given by the sensory weight.
+If the trail duration approaches infinity, the slope would be 1 again (however, this seems rather unrealistic, this would only be true in an ideal system without memory decay or reproduction and accumulation noise ...).
+
+%\subsection{Influence of a population of nPE and pPE neurons}
+%XXX
+%
+%
+%\subsection{Analysis of simplified network model, effect of time constants}
+%
+%simplified model: dynamics and steady state of rM and rV, rM and rV as a function of time constants and trial duration etc., weighting, then use those expressions to discuss when weighting goes awry and how long transitions take from one state to another ...
+
+\subsection{Comparison to Kalman filter and Bayes Factor surprise}
 %
@@
 
-\section*{Supplementary Figures}
+\section{Supplementary Figures}
 \setcounter{figure}{0}

### 2023-10-06 records/manuscript_2023.tex
get all parameters for Methods & SI, retrieve bit in run_fig_4.py

@@
 
+% summary of principal question, hypothesis and our findings
+
+% multisensory integration --> general idea that this can be applied to S and P ... Firth? link to psychiatric disorders
+
+% evidence for the model (PE neurons, connectivity PE --> M, predictions or P neurons exist, P neurons may hold the mean, time constants chnage along hierarchy)
+
+% alternative implementations (P in weights ...
+=> plasticity), also we dont implement weighting explicitely -- has been shown that divisive inhibiton can be done with INs (look in the one theoretical paper on neural implemnetation of Kalman filter)
+
+% Prediction update not weighted by uncertainty, discuss evidence fo that
+
+% neuromodulators, does it fit the current literature Lawson etc.? How is it different from expected and unexpected uncertainty ..., and then give more details on local vs. global neuromodulators
+
+% link to Kalman filter and maybe Bayes surprise factor
+
+% ################
+
 % At the end, or in between talk about changes in PE neurons directly that you have currently in Fig. 3 SX ... I copy the part of the last paragraph here as inspiration :
@@
 The connection strengths between the \textit{pre}-synaptic neuron (population) and the neurons of the PE circuit are denoted by $W_\mathrm{PE\leftarrow pre}$ and $\underline{w}_{\mathrm{PE} \leftarrow pre}$, respectively.
-The activities of the neurons evolve with time constants contained in $T$.
+The activities of the neurons evolve with time constants summarized in $T$.
 
@@
 %
-All values for neuron and network parameters, details on the model equations, as well as supporting analyses can be found in SI XXX.
+All values for neuron and network parameters, details on the model equations, as well as supporting analyses can be found in the supplementary material.
 
@@
 %
-We arithmetically calculate the weighted output of sensory inputs and predictions, $r_\mathrm{out}$, based on Bayesian multisensory integration (XXX),
+We arithmetically calculated the weighted output of sensory inputs and predictions, $r_\mathrm{out}$, based on Bayesian multisensory integration (XXX),
 %
@@
 %
-where $\alpha$ denotes the reliability of the sensory input and is given by 
+where $\alpha$ denotes the sensory weight (that is, the reliability of the sensory input) and is given by 
 %
@@
 %
-The network receives feedforward stimuli of duration XXX that can change from trial to trial (one stimulus per trial).
-To account for noise, each stimulus is composed of N$_\mathrm{val}$ piece-wise constant values drawn from a normal distribution with mean XXX and standard deviation XXX.
-To account for changes in the environment, the mean of the normal distribution is drawn from a uniform distribution with mean XXX and standard deviation XXX.
+The network receives feedforward stimuli $s$ that may vary between trials.
+To account for noise, each stimulus is composed of N$_\mathrm{in}$ piece-wise constant values drawn from a normal distribution with mean $\mu_\mathrm{in}$ and standard deviation $\sigma_\mathrm{in}$.
+To account for changes in the environment, $\mu_\mathrm{in}$ is drawn from a uniform distribution $U(a,b)$ with mean $\mu_\mathrm{trial} = \frac{a+b}{2}$ and standard deviation $\sigma_\mathrm{trial} = \frac{b-a}{\sqrt{12}}$.
 The parameterisation of both distributions varies across the experiments.
-All stimulus/input parameters can be found in SI XXX.
+All stimulus/input parameters can be found in the supplementary material.
 
@@
 All simulations were performed in customized Python code written by LH.
-Differential equations were numerically integrated using a 2\textsuperscript{nd}-order Runge-Kutta method with time steps ranging between $xx$ and $xxx$ ms.
-All neurons were initialized with $r=0/s$.
-We defined the steady-state firing rate per stimulus as the activity in the last XXXX ms of stimulus presentation.
-Further details and values for simulation parameters can be found in SI XXX.
+Source code to reproduce the simulations, analyses and figures will be available after publication at \url{XXXX}.
+Differential equations were numerically integrated using a 2\textsuperscript{nd}-order Runge-Kutta method.
+Neurons were initialized with $r=0/s$.
+Further details and values for simulation parameters can be found in the supplementary material.
 
@@
 \section{Detailed Methods}
-
+%
+In the following, we describe in more detail the equations describing the dynamics of the neurons in the prediction-error circuit, as well as the memory and variance neurons.
+We then provide the connectivity of the network and the inputs to the neurons for both the mean-field and population model.
+Finally, to ensure reproducibility, we give all simulation parameters used for the results shown in the figures.
+
+XXX Please note that xxx model parameters chosen do not change overall results are not particular to get the results xxx.
+Important factors are: 1) long enough to get to steady state (as PE circuits used were learned for steady state, 2) lower M updates faster than higher M, 3) update speed for M and V (see lambda and theta etc.) must be chosen in accordance with trial duration stimulus duration.
 XXX
@@
 
-For the results in Figure 2, only the lower subnetwork is simulated.
-In Supp Fig XXX, we replaced each mean-field PE circuit with a heterogeneous population model with $N_\mathrm{total} = 200$ neurons ($\mathrm{N}_\mathrm{SOM} = \mathrm{N}_\mathrm{VIP} = \mathrm{N}_\mathrm{PV} = 20$, the remaining neurons are excitatory).
-In Supp Fig.
-XXX, the PE circuit comprises $\mathrm{N}_\mathrm{copies} = XXX$ copies of the mean-field network to account for selectivity.
-
-In the following, we describe the dynamics of the neuron types for the mean-field network.
-The equations for the population PE circuit can be deduced from the mean-field equations and can also be found in XXX.
+In Figure \ref{fig:Fig_2} and the corresponding supporting figures, only the lower subnetwork is simulated.
+In Fig.
+\ref{fig:Fig_2_S2}, we replaced this lower mean-field PE circuit with a heterogeneous population model containing $200$ neurons ($\mathrm{N}_\mathrm{SOM} = \mathrm{N}_\mathrm{VIP} = \mathrm{N}_\mathrm{PV} = 20$, $140$ excitatory neurons).
+In Fig.
+\ref{fig:Fig_2_S3}, the lower PE circuit comprises $1000$ copies of the mean-field network to account for selectivity.
+
+In the following, we describe the dynamics of the neurons/compartments in the mean-field network.
+The equations for the population PE circuit (Fig.
+\ref{fig:Fig_2_S2}) are directly deduced from the mean-field equations and can also be found in XXXref your own paperXXX.
 
@@
 In addition to the core PE circuit, we simulate a memory neuron \textit{M} and a variance neuron \textit{V}.
-The memory neuron is modeled as a perfect integrator, receiving the output synapses from the nPE and pPE neurons,
-%
-\begin{align}
-\tau_M \cdot \frac{dr_\mathrm{M}}{dt} = w_\mathrm{M\leftarrow pE} \cdot r_\mathrm{pE} - w_\mathrm{M\leftarrow nE} \cdot r_\mathrm{nE}.
+The memory neuron is modeled as a perfect integrator, receiving synapses from both the nPE and pPE neuron,
+%
+\begin{align}
+\tau_E \cdot \frac{dr_\mathrm{M}}{dt} = w_\mathrm{M\leftarrow pE} \cdot r_\mathrm{pE} - w_\mathrm{M\leftarrow nE} \cdot r_\mathrm{nE}.
 \end{align}
@@
 $w_\mathrm{M\leftarrow pE}$ denotes the connection strength between the pPE neuron and the memory neuron, and  $w_\mathrm{M\leftarrow nE}$ denotes the connection strength between the nPE neuron and the memory neuron.
-The time constant XXX.
+The time constant $\tau_E = 60$ ms.
 
@@
 The weight $w_\mathrm{V\leftarrow pE}$ represents the connection strength between the pPE neuron and the variance neuron, while  $w_\mathrm{V\leftarrow nE}$ denotes the connection strength between the nPE neuron and the variance neuron.
-To ensure that the V neuron encodes the variance, we chose a squared activation function.
+To ensure that the V neuron encodes the variance, we chose a quadratic activation function.
 In Fig.
-XXX, to investigate the necessity of a squared activation function in the context of weighting sensory inputs and predictions, we used a linear activation function.
-The time constant  $\tau_V$ was set to XXX.
+\ref{fig:Fig_3_S2}, we used a linear activation function to investigate the impact of the input-output transfer function on the weighting of sensory inputs and predictions.
+The time constant $\tau_V = 5$ s.
 
@@
 \alpha &= \frac{r_\mathrm{V_{lower}}^{-1}}{r_\mathrm{V_{lower}}^{-1} + r_\mathrm{V_{higher}}^{-1}}\nonumber\\
-& = \left( 1 + \frac{r_\mathrm{V_{lower}}}{r_\mathrm{V_{higher}}} \right)^{-1}
+& = \left( 1 + \frac{r_\mathrm{V_{lower}}}{r_\mathrm{V_{higher}}} \right)^{-1}.
 \end{align}
@@
 %
-The connectivity between neurons of the PE circuit, both for the mean-field and the population network, was taken from XXX. XXX ADD heatmaps for the connectivity? XXX
+The connectivity between neurons of the PE circuit, both for the mean-field and the population network, was taken from XXX.
+XXX ADD heatmaps for the connectivity?
+or simply refer to the 2022 paper (maybe rather that)
 
@@
 While the nPE neurons inhibit the M neuron, the pPE neurons excite it.
-To ensure that the activities of the memory neurons represent the mean of the sensory stimuli in the lower PE circuit and the mean of the prediction in the higher subnetwork, respectively, the net effect of nPE and pPE neurons must cancel in the steady state.
+To ensure that the activities of the memory neurons represent the mean of the sensory stimuli in the lower PE circuit and the mean of the prediction in the higher subnetwork, respectively, the net effect of nPE and pPE neurons must cancel in the steady state (see Analysis in \ref{sec:gain_impact}).
 Hence, the weights need to account for the neurons' potentially different gain factors ($g_\mathrm{nPE}$ and $g_\mathrm{pPE}$, see Table XXX) and the neuron numbers ($N_\mathrm{nPE}$ and $N_\mathrm{pPE}$, see Table XXX):
@@
 %
-where $\lambda$ denotes the speed at which the perfect integrator evolves (lower PE circuit: $\lambda$=XXX, higher PE circuit:  $\lambda$=XXX).
+where $\lambda$ denotes the speed at which the perfect integrator evolves (lower PE circuit: $\lambda=3\cdot 10^{-3}$  in Fig.
+\ref{fig:Fig_2} and $\lambda=4.5\cdot 10^{-2}$ otherwise, higher PE circuit:  $\lambda = 7\cdot 10^{-4}$).
 
@@
 
-The memory neuron connects to a number of neurons in the PE circuit.
-Each connection from \textit{M} to a postsynaptic neuron \textit{X} is set to $w_\mathrm{X\leftarrow M} = 1$.
+The memory neuron \textit{M} connects to the post-synaptic neurons \textit{X} in the PE circuit with the connection strength $w_\mathrm{X\leftarrow M}$.
+If a connection exists, we assume $w_\mathrm{X\leftarrow M} = 1$.
 In all mean-field networks and the population network, the dendrites of nPE and pPE neurons and one of the two (populations of) PV neurons receive connections from the memory neuron.
@@
 \caption{XXX Table shows $w_\mathrm{X\leftarrow M}$ ... .}
+\label{tab:wXM}
 \end{table}
@@
 Both nPE and pPE neurons excite the \textit{V} neuron.
-To ensure that the activity of the V neuron represents the variance of the input, the weights must account for differences in the gains ($g_\mathrm{nPE}$ and $g_\mathrm{pPE}$, see Table XXX) and numbers ($N_\mathrm{nPE}$ and $N_\mathrm{pPE}$) of the PE neurons,
+To ensure that the activity of the V neuron represents the variance of the input (see Analysis in \ref{sec:gain_impact}), the weights must account for differences in the gains ($g_\mathrm{nPE}$ and $g_\mathrm{pPE}$, see Table XXX) and numbers ($N_\mathrm{nPE}$ and $N_\mathrm{pPE}$) of the PE neurons,
 %
@@
 %
-Each neuron (type) receives an overall input $I_i$
+Each neuron (type) receives an overall input $I_i$,
 %
@@
 %
-where 
-Feedforward is either the direct sensory input S for the lower PE circuit or the rM for the higher PE circuit.
-
-baseline input is XXX.
-
-show Wzs in form as table from above (rows different MFN, ...). Mention that population model is row 1, all of the SOM and VIP receive S or P, respectively then. Or simply write 1-wXM???
-%
-\begin{table}[h!]
-\centering
-\begin{tabular}{ |c|c|c|c|  }
-\hline
-Network name & MFN 1 & MFN 2 & MFN 3  \\
-\hline
-\hline
-SOM & 1 & 0 & 1\\
-VIP & 0 & 1 & 1 \\
-\hline
-\end{tabular}
-\caption{XXX Table shows $w_\mathrm{X\leftarrow input}$ ... .}
-\end{table}
-%
-
-Each stimulus was presented for $1$ second.
-
-%All neurons receive external background input that ensures reasonable baseline firing rates in the absence of sensory inputs and predictions thereof.
-In the case of non-plastic networks, these inputs were set such that the baseline firing rates are $r_\mathrm{E}=1\, s^{-1}$, $r_\mathrm{P} = r_\mathrm{S}=r_\mathrm{V}=4\, s^{-1}$ and $r_\mathrm{D}=0\, s^{-1}$.
-In the case of plastic networks, we set the external inputs of all neuron types to $x_\mathrm{E}=x_\mathrm{P}=x_\mathrm{S}=x_\mathrm{V}=5\, s^{-1}$, while the background input to the dendrites is dynamically computed during training such that $r_\mathrm{D}=0\, s^{-1}$.
-%
-%In addition to the external background inputs, the neurons receive either sensory input~($S$), a prediction thereof ($P$), or both.
-We distinguish between phases of fully predicted ($P=S>0$), overpredicted ($P>S$) and underpredicted ($P<S$) sensory stimuli, as well as baseline phases ($P=S=0$).
-During training, the network is exposed to baseline phases and fully predicted sensory inputs (Figs.~\ref{fig:Fig_Plasticity} and \ref{fig:Fig_Experience}), or in addition to over- and underpredicted sensory stimuli (Fig.~\ref{figsupp:Fig_Experience_Predictability}).
-Stimuli are drawn from a uniform distribution from the interval $[0, 5\, s^{-1}]$.
-Mean and SD of test stimuli for each simulation are listed below (see ``Simulations").
+where $I_{i}^{FF}$ denotes a feedforward input and $I_i^{BL}$ represents an external background input that ensures reasonable baseline firing rates in the absence of sensory inputs and predictions thereof.
+In the case of the mean-field network, these inputs were set such that the baseline firing rates are $r_\mathrm{pE}=r_\mathrm{pD}=r_\mathrm{nE}=r_\mathrm{nD}=0\, s^{-1}$ and $r_\mathrm{P} = r_\mathrm{S}=r_\mathrm{V}=4\, s^{-1}$.
+In the case of the population network, we set the external inputs of all neuron types to $5\, s^{-1}$, while the background inputs to the dendrites were computed such that the dendrites are inactive during baseline.
+
+The feedforward input is either the direct sensory input $s$ for the lower PE circuit, or the activity of the M neuron $r_\mathrm{M}$ for the higher PE circuit.
+In general, for the three mean-field networks tested, we chose $w_i = 1 - w_\mathrm{X\leftarrow M}$  (see Table \ref{tab:wXM}).
+In the population network, 70\% of the SOM neurons and 30\% of the VIP neurons receive the feedforward input.
 
@@
 %
-%All simulations were performed in customized Python code written by LH.
-Differential equations were numerically integrated using a 2\textsuperscript{nd}-order Runge-Kutta method with time steps ranging between $0.1$ and $0.2$ ms.
-Neurons in the PE circuits were initialized with $r=0/s$.
-The memory neurons were initialized at the mean of the two distributions (see above), and each prediction neuron was either set to the mean of the distribution it is associated with if the stimulus at $t=0\ \mathrm{ms}$ was drawn from that distribution, or set to zero otherwise.
-Each stimulus was presented for $1$ second.
-During training of the PE circuit, we presented 350 stimuli alternated with 350 zero-input (baseline) phases.
-We made sure that the weights converged to a configuration that satisfied the objective given by our homeostatic plasticity rules (see Eqs.~\ref{eq:Plasticity_I}-\ref{eq:Plasticity_II} in ``Plasticity model").
-We defined the steady-state firing rate per stimulus as the activity in the last 500 ms of stimulus presentation.
-The onset firing rate was computed as the activity of the first 10 ms.\\\\
-%%
-%\textbf{Figures 1 \& S2:} Test stimulus was set to $5/s$ with a SD of $1/s$. Stimulus to compute total excitatory and inhibitory inputs was set to $1/s$.\\
-%%
-%\textbf{Figures 2 \& S3:} Test stimulus was set to $3/s$ with a SD of $1/s$. The perturbation stimuli ranged from $-1.5/s$ to $1.5/s$.\\
-%%
-%\textbf{Figures 3, S4 \& S5:} Test stimulus was set to $5/s$ with a SD of $1.5/s$.
-50\% of the PV neurons, 70\% of the SOM neurons and 30\% of the VIP neurons receive the actual sensory input, while the remaining ones of each population received a prediction thereof.
-Perturbation stimulus was $\pm 2/s$.
-Panels D \& G of Fig.
-3 show the median over all PE neurons and the SEM.\\
-%%
-%\textbf{Figures 4 \& S6:} Test stimulus was set to $5/s$ with a SD of $1.5/s$.
-In main figure, square: $w_\mathrm{EP}\in[2,4]$, $w_\mathrm{PS}\in[0.5,1]$, $w_\mathrm{PV}\in[1.5,2.5]$; circle: $w_\mathrm{EP}\in[2.5,8]$, $w_\mathrm{PS}\in[1.5,2.5]$, $w_\mathrm{PV}\in[0.5,1]$; triangle: $w_\mathrm{EP}\in[2.5,8]$, $w_\mathrm{PS}\in[1,2.5]$, $w_\mathrm{PV}\in[0.5,2]$.
-Half of the PV neurons and all SOM neurons receive the actual sensory input, while the remaining PV and SOM neurons as well as all VIP neurons receive a prediction thereof.
-%In supporting figure: Total number of stimuli presented during training was increased, so that the number of fully predicted sensory stimuli was constant at 350.
-Results were averaged over 5 simulations, mean and SD are shown.\\
-%%
-%\textbf{Figure 5:} For panel E, the performance error was computed as the squared difference between the activity of the respective line attractor and the stimulus presented.
-For panel F, the initial weight between the stimulus and the representation neuron was set to $0.5$.
-The basis learning rate (fixed) was set to $5e^{-4}$.
-And the initial speed was computed as the derivative of the rate with respect to time, averaged over the first 50 ms.
-%
-%Source code to reproduce the simulations, analyses and figures will be available after publication at \url{github.com/lhertaeg/SourceCode_Hertaeg2021}.
+All simulations were performed in customized Python code written by LH.
+Source code to reproduce the simulations, analyses and figures will be available after publication at \url{XXXX}.
+Differential equations were numerically integrated using a 2\textsuperscript{nd}-order Runge-Kutta method.
+Neurons were initialized with $r=0/s$.
+
+In the following, we give all figure-specific parameters not directly visible or mentioned in the Figures and captions.
+Please note that the qualitative results are robust to the choice of the simulation parameters and are here stated merely to ensure the reproducibility of all figures.
+Furthermore, to increase readability, we do not include units for the parameters.
+All units can be deduced from the equations above.\newline\\
+%
+\textbf{Figure 1:} Simulation time steps = $10^5$, constant prediction (fixed) = $5$, input mean $\in [0,10]$, input standard deviation $0$.\\
+%
+\textbf{Figure 2:} Simulation time steps = $10^5$, $200$ constant values (each $500$ time steps long) drawn from a uniform distribution, B and C: input standard deviation fixed at $5$ when mean is varied, input mean fixed at $4.5$ when input variance is varied.\\
+%
+\textbf{Figure 2 Supplementary Fig.
+1:} Simulation time steps = $10^5$, $200$ constant values (each $500$ time steps long) drawn from a uniform distribution with mean of $5$ and variance of $4$.
+Number of repetitions with different seeds: 20.\\
+%
+\textbf{Figure 2 Supplementary Fig.
+2:} Number of time steps were $10^5$ (C-E) or $2\cdot 10^5$ (B) with $400$ (C-E) or $200$ (B) constant values drawn from a uniform distribution with mean of $5$ and variance of $4$.
+Time step was $0.1$.
+The connections from the PE neurons to the M or V neuron were altered by a factor $\gamma$ drawn from a normal distribution.
+If not stated otherwise (in the Figure), the mean of this normal distribution was $1$ and the variance $0$, while the connection probability was $1$.
+Number of repetitions with different seeds: 10.\\
+%
+\textbf{Figure 2 Supplementary Fig. 3:} Number of identical mean-field networks: $1000$.  Simulation time steps = $10^5$.\\
+%
+\textbf{Figure 3:} $100$ trials, each simulated for $5000$ time steps.
+In a trail, $10$ constant values were drawn from a normal distribution (each $500$ time steps long).
+The variance of this distribution was $0$ (C, F)  and $5.3$ XXX (D).
+In each trial, the stimulus mean was drawn from an uniform distribution $U(a, b)$ (C: $U(1,9)$, D: $U(5,5)$).
+In F, $U$ was parameterized such that the trail variability was $3$.\\
+%
+\textbf{Figure 3 Supplementary Fig.
+1:}  $120$ trials, each simulated for $5000$ time steps.
+In a trail, $10$ constant values were drawn from a normal distribution $N(\mu_\mathrm{in}, \sigma_\mathrm{in})$, each $500$ time steps long.
+In each trial, the stimulus mean was drawn from an uniform distribution $U(a, b)$.
+Switch of input statistics occurs after 60 trails.
+State 1: stimulus $\in N(\mu_\mathrm{in}, 0)$ with $\mu_\mathrm{in} \in U(5,5)$, State 2:  stimulus $\in N(\mu_\mathrm{in}, 3)$ with $\mu_\mathrm{in} \in U(5,5)$, State 3: stimulus $\in N(\mu_\mathrm{in}, 3)$ with $\mu_\mathrm{in} \in U(0,10)$, State 4: stimulus $\in N(\mu_\mathrm{in}, 0)$ with $\mu_\mathrm{in} \in U(0,10)$.\\
+%
+\textbf{Figure 3 Supplementary Fig.
+2:} $200$ trials, each simulated for $5000$ time steps.
+In a trail, $10$ constant values were drawn from a normal distribution with variance $\sigma^2_\mathrm{in}$, each $500$ time steps long.
+In each trial, the stimulus mean was drawn from an uniform distribution with mean of $5$ and a variance of $\sigma^2_\mathrm{trial}$.
+$\sigma^2_\mathrm{trial} \in \lbrace0, 0.75, 1.5, 2.25, 3\rbrace$, $\sigma^2_\mathrm{in} \in \lbrace 3, 2.25, 1.5,0.75, 0\rbrace$.
+A: scaling factors of $w_\mathrm{M\leftarrow PE}$ were 0.3 and 6.7 XXX.\\
+%
+\textbf{Figure 4:} $100-200$ trials, each simulated for $5000$ time steps.
+In a trail, $10$ constant values were drawn from a normal distribution with variance $\sigma^2_\mathrm{in}$, each $500$ time steps long.
+In each trial, the stimulus mean was drawn from an uniform distribution with mean of $5$ and a variance of $\sigma^2_\mathrm{trial}$.
+A, top: $\sigma^2_\mathrm{in} = 0$ and $\sigma^2_\mathrm{trial} = 1$.
+A, bottom: $\sigma^2_\mathrm{in} = 1$ and $\sigma^2_\mathrm{trial} = 0$.
+C: $\sigma^2_\mathrm{in} = 1$ and $\sigma^2_\mathrm{trial} = 1$.
+D: $\sigma^2_\mathrm{in} = 5.3 XXX$ and $\sigma^2_\mathrm{trial} = 0$.
+Additional input (perturbation) was either fixed at $1$ (A, D) or systematically varied between $-1$ and $1$ (C).
+To estimate changes in baseline and gain of nPE and pPE neurons, we fitted a linear function to the PE neuron activity for the input range $[0, 2.5]$.\\
+%
+\textbf{Figure 4 Supplementary Fig. 1:} All parameters are from Figure 4 A.\\
+%
+\textbf{Figure 4 Supplementary Fig.
+2:} In the default setting baseline was $0$ and gain of PE neurons was $1$ (see Eqs.
+XXX).
+The results (left) were computed for baselines $\in [0, 3]$, while the results (right) were computed for gains $\in [0.5, 1.5]$.
+XXX Here refer to equations XXX.
+\\
+%
+\textbf{Figure 4 Supplementary Fig. 3:} All parameters are from Figure 4 D.\\
+%
+\textbf{Figure 5:} $200$ trials, each simulated for $5000$ time steps (except in panel E).
+In a trail, $10$ constant values were drawn from a normal distribution (each $500$ time steps long) with standard deviation $\sigma_\mathrm{in}$.
+In each trial, the stimulus mean $\mu_\mathrm{in}$ was drawn from an uniform distribution $U(a, b)$.
+A: $\sigma_\mathrm{in} \in \lbrace1,7\rbrace$, $\mu_\mathrm{in} \in U(15, 25)$.
+B: $\sigma_\mathrm{in} \in [0,8]$, $\mu_\mathrm{in} \in U(15, 25)$, and $\sigma_\mathrm{in} = 5$, $\mu_\mathrm{in} \in U(15, b)$ with $b\in [20, 48]$.
+C: $\sigma_\mathrm{in} \in \lbrace2,5\rbrace$, $\mu_\mathrm{in} \in U(15, 15)$.
+D: $\sigma_\mathrm{in} = 0$, $\mu_\mathrm{in} \in U(15, 25)$ or $U(10, 30)$.
+E: $\sigma_\mathrm{in} = 0$, $\mu_\mathrm{in} \in U(15, 25)$, or $\sigma_\mathrm{in} = 5$, $\mu_\mathrm{in} \in U(15, 15)$,  Time steps per trail increased from $5000$ to $10^4$.\\
+%
+\textbf{Figure 5 Supplementary Fig.
+1:} $200$ trials, each simulated for $5000$ time steps.
+In a trail, $10$ constant values were drawn from a normal distribution (each $500$ time steps long) with standard deviation $\sigma_\mathrm{in}$.
+In each trial, the stimulus mean $\mu_\mathrm{in}$ was drawn from an uniform distribution $U(a,b)$.
+We used two different uniform distributions $U(15, 25)$ and  $U(25, 35)$, and introduced scalar variability so that $\sigma_\mathrm{in}$ is a linear function of $\mu_\mathrm{in}$.
+Specifically, we chose $\sigma_\mathrm{in} = \mu_\mathrm{in} -14$.
 
@@
 
-%3. Sensory weight and contraction bias
-%4. Comparison to Kalman filter and Bayes Factor surprise
-
-\subsection{XXX}
-%0. rM exponential average, rV exponential average (see iPad, currently page 107)
-
-\subsection{Impact of PE neurons' gain on estimating mean and variance}
-%
-Only if the the gain of the nPE neuron ($g_{nPE}$) equals the gain of the pPE neuron ($g_{pPE}$) in the mean-field network, the activity of the M neuron represents the mean of the inputs,
+We first describe a simplified model and show that the M neuron represents the mean, while the V neuron represents the variance of the feedforward input.
+We then investigate the impact of the gain and baseline of PE neurons on estimating the mean and variance.
+Furthermore, we use the simplified model to discuss the effect of neuromodulators in our network.
+Finally, we reveal the connection between the sensory weight and the contraction bias, and compare our network with the Kalman filter as well as Bayes factor surprise.
+XXX
+
+\subsection{Activity of M and V neuron in a simplified model}\label{sec:toy}
+%
+To show that the M neuron encodes the mean, while the V neuron encodes the variance of the feedforward input, we resume a toy model in which the activity of the nPE and pPE neuron is replaced by its ideal output
+%
+\begin{align}
+r_\mathrm{nE} = \left[ r_\mathrm{M} - s_\mathrm{FF}\right]_+ \nonumber \\
+r_\mathrm{pE} = \left[ s_\mathrm{FF} - r_\mathrm{M} \right]_+
+\end{align}
+%
+with $s_\mathrm{FF}$ denoting the time-dependent feedforward input. The activity of the M neuron can then be described as
+%
+\begin{align}
+\tau_M \cdot \frac{dr_\mathrm{M}}{dt} = r_\mathrm{pE} - r_\mathrm{nE}
+\end{align}
+%
+If $r_\mathrm{M} \geq s_\mathrm{FF}$, we get
+%
+\begin{align}
+\tau_M \cdot \frac{dr_\mathrm{M}}{dt} = -r_\mathrm{nE} = -r_\mathrm{M} + s_\mathrm{FF}.
+\end{align}
+%
+If $r_\mathrm{M} \leq s_\mathrm{FF}$, we also get
+%
+\begin{align}
+\tau_M \cdot \frac{dr_\mathrm{M}}{dt} = r_\mathrm{pE} = -r_\mathrm{M} + s_\mathrm{FF}.\nonumber
+\end{align}
+%
+Hence, the activity of $r_\mathrm{M}$ is given by
+%
+\begin{align}
+r_\mathrm{M} = \frac{1}{\tau_M} \int\limits_0^t e^{-(t-x)/\tau_M}\cdot s_\mathrm{FF}(x)\ dx
+\end{align}  
+%
+for zero activity at time $t=0$. In the limit of $t\rightarrow \infty$ (steady state), this is exactly the exponential average of the feedforward input, $E(s_\mathrm{FF})$.
+
+With the simplified activity of the nPE and pPE neuron, the activity of the V neuron can then be described as
+%
+\begin{align}
+\tau_V \cdot \frac{dr_\mathrm{V}}{dt} = -r_\mathrm{V} + (r_\mathrm{pE} +  r_\mathrm{nE})^2 = -r_\mathrm{V}  + (r_\mathrm{M} -  s_\mathrm{FF})^2,
+\end{align}
+%
+leading to the time-dependent solution
+%
+\begin{align}
+r_\mathrm{V} = \frac{1}{\tau_V} \int\limits_0^t e^{-(t-x)/\tau_V}\cdot \left[r_\mathrm{M}(x) -  s_\mathrm{FF}(x)\right]^2\ dx.
+\end{align}  
+%
+In the limit of $t\rightarrow \infty$, $r_\mathrm{V}$ approaches $E(s_\mathrm{FF} - E(s_\mathrm{FF}))$.
+
+\subsection{Impact of PE neurons' gain on estimating mean and variance}\label{sec:impact_gain}\label{sec:gain_impact}
+%
+The gains of the PE neurons, if not equal between nPE and pPE neuron on average, can bias the activity of both the M and V neuron. To show this, we resume the toy model from section \ref{sec:toy}. 
 %
@@
 \label{eq:condition_mean_gain_equal}
-g_{pPE}\ \langle \mathrm{nPE}\rangle &= g_{nPE}\ \langle \mathrm{pPE}\rangle \\
-g_{pPE} \langle \left[ S-P\right]_+\rangle &= g_{nPE} \langle \left[ P-S\right]_+\rangle \nonumber \\
-g_{pPE} \int\limits_P^b \left( x-P\right)\ f(x)\ dx &= g_{nPE} \int\limits_a^P \left( P-x\right)\ f(x)\ dx. \nonumber
-\end{align}
-%
-In case of a uniform distribution ($f(x) = 1/(b-a)$ when $x\in [a,b]$ and $0$ otherwise) from which the input values are drawn, this condition yields
-%
-\begin{align}
-\label{eq:condition_mean_gain_equal_4_uniform_dist_1}
-P = \int\limits_a^b x\, f(x)\ dx = \frac{a + b}{2}
-\end{align}
-%
-for $g_{nPE} = g_{pPE} = g$, and 
-%
-\begin{align}
-\label{eq:condition_mean_gain_equal_4_uniform_dist_21}
-g_{pPE}\ \left[ \frac{1}{2} \left(b^2 - P^2\right) - P\left(b - P\right)\right] = g_{nPE}\ \left[  P\left(P - a\right) - \frac{1}{2} \left(P^2 - a^2\right)\right]
-\end{align}
-%
-for $g_{nPE} \neq g_{pPE}$, which can be further summaries by
-%
-\begin{align}
-\label{eq:condition_mean_gain_equal_4_uniform_dist_21}
-P = \frac{g_{pPE}\ b - g_{nPE}\ a \pm \sqrt{g_{nPE}\ g_{pPE}} (a-b)}{g_{pPE} - g_{nPE}}.
-\end{align}
-%
-Hence, estimating the mean correctly requires $g_{nPE} = g_{pPE} = g$.
-For the V neuron to represent the variance of the inputs, this condition must be tightened to  $g_{nPE} = g_{pPE} = 1$.
-The variance is given by
+g_\mathrm{pPE}\ \langle r_\mathrm{nPE}\rangle &= g_\mathrm{nPE}\ \langle r_\mathrm{pPE}\rangle \\
+g_\mathrm{pPE}\ \langle \left[ s_\mathrm{FF}-P\right]_+\rangle &= g_\mathrm{nPE}\ \langle \left[ P-s_\mathrm{FF}\right]_+\rangle \nonumber\\
+g_\mathrm{pPE} \int\limits_P^{\infty} \left( x-P\right)\ f(x)\ dx &= g_\mathrm{nPE} \int\limits_{-\infty}^P \left( P-x\right)\ f(x)\ dx. \nonumber
+\end{align}
+%
+Here, $P$ denotes the prediction encoded in the M neuron, and $f(x)$ is the distribution of feedforward inputs.
+In case of a uniform distribution, $f(x) = 1/(b-a)$ for $x\in [a,b]$ and $0$ otherwise, we get
+%
+\begin{equation}
+\label{eq:prediction_gain}
+    P=
+    \begin{cases}
+      \frac{a + b}{2} & \text{if}\ g_\mathrm{nPE} = g_\mathrm{pPE} = g \\
+      \frac{g_\mathrm{pPE}\cdot b - g_\mathrm{nPE}\cdot a \pm \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}} (a-b)}{g_\mathrm{pPE} - g_\mathrm{nPE}} & \text{otherwise.}
+    \end{cases}
+\end{equation}
+%
+XXX Maybe mention the SuppFig. in which you show the function XXX and explain what you see for gnPE > gpPE or vice versa ...
+
+Similarly, the variance is affected by the gain of the nPE and pPE neuron,
 %
@@
 \label{eq:condition_variance_gain}
-V &= \langle (S-P)^2\rangle \\
-   &= g_\mathrm{pPE}\ \langle \left[ S-P \right]_+^2 \rangle + g_\mathrm{nPE}\ \langle \left[ P-S \right]_+^2 \rangle. \nonumber
-\end{align}
-%
-In case of a uniform distribution from which the input values are drawn, this condition yields
-%
-\begin{align}
-\label{eq:condition_variance_gain_uniform_dist}
-V &= \frac{g_\mathrm{pPE}}{b-a} \int\limits_P^b (u-P)^2\ du + \frac{g_\mathrm{nPE}}{b-a} \int\limits_a^P (P-u)^2\ du \\
+V &= g_\mathrm{pPE}\ \langle \left[ s_\mathrm{FF}-P \right]_+^2 \rangle + g_\mathrm{nPE}\ \langle \left[ P-s_\mathrm{FF} \right]_+^2 \rangle.
+\end{align}
+%
+In case of a uniform distribution, we get
+%
+\begin{align}
+V &= \frac{g_\mathrm{pPE}}{b-a} \int\limits_P^b (x-P)^2\ dx + \frac{g_\mathrm{nPE}}{b-a} \int\limits_a^P (P-x)^2\ dx \\
    &= \frac{g_\mathrm{pPE}}{3} \cdot \frac{(b-P)^3}{b-a} + \frac{g_\mathrm{nPE}}{3} \cdot \frac{(P-a)^3}{b-a}. \nonumber
@@
 %
-Only if $g_{nPE} = g_{pPE} = 1$ and $P = \frac{a + b}{2}$, the variance is given by $\frac{(b - a)^2}{12}$, otherwise the V neuron's activity is given by
-%
-\begin{align}
-\label{eq:condition_variance_gain_uniform_dist_1}
-V = \frac{(b-a)^2}{3\ (g_\mathrm{pPE} - g_\mathrm{nPE})^3} \cdot \left[ g_\mathrm{nPE} \cdot( g_\mathrm{pPE} \mp \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}})^3 - g_\mathrm{pPE} \cdot (g_\mathrm{nPE} \mp \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}})^3\right].
-\end{align}
-%
+Inserting eqs. (\ref{eq:prediction_gain}) yields
+%
+\begin{equation}
+\label{eq:variance_gain}
+    V=
+    \begin{cases}
+      \frac{(b - a)^2}{12} & \text{if}\ g_\mathrm{nPE} = g_\mathrm{pPE} = 1 \\
+      \frac{(b-a)^2}{3\ (g_\mathrm{pPE} - g_\mathrm{nPE})^3} \cdot \left[ g_\mathrm{nPE} \cdot( g_\mathrm{pPE} \mp \gamma)^3 - g_\mathrm{pPE} \cdot (g_\mathrm{nPE} \mp \gamma)^3\right] & \text{otherwise.}
+    \end{cases}
+\end{equation}
+%
+with $\gamma = \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}}$. 
+XXX As above, discuss cases from the Supp Figs. XXX Important: for P and V please check with sign is correct!
 
@@
 %
-Only if the the baseline of the nPE neuron ($n_0$) equals the baseline of the pPE neuron ($p_0$) in the mean-field network, the activity of the M neuron represents the mean of the inputs,
+The baselines of the PE neurons, if not equal between nPE and pPE neuron on average, can bias the activity of both the M and V neuron.
+By means of the toy model from section \ref{sec:toy}, we can write
 %
@@
 \label{eq:condition_baseline_mean}
-\langle \mathrm{pPE} \rangle &= \langle \mathrm{nPE} \rangle \\
-\langle \left[ S - P\right]_+ + p_0\rangle &= \langle \left[ P - S\right]_+ + n_0\rangle \nonumber\\
-\int\limits_P^b (x - P)\ f(x)\ dx + p_0 \underbrace{\int\limits_a^b f(x)\ dx}_{=1}  &= \int\limits_a^P (P - x)\ f(x)\ dx + n_0 \underbrace{\int\limits_a^b f(x)\ dx}_{=1} . \nonumber
-\end{align}
-%
-In case of a uniform distribution ($f(x) = 1/(b-a)$ when $x\in [a,b]$ and $0$ otherwise) from which the input values are drawn, this condition yields
+\langle r_\mathrm{pPE} \rangle &= \langle r_\mathrm{nPE} \rangle \\
+\langle \left[s_\mathrm{FF} - P\right]_+ + p_0\rangle &= \langle \left[P - s_\mathrm{FF}\right]_+ + n_0\rangle \nonumber\\
+\int\limits_P^\infty (x - P)\ f(x)\ dx + p_0 \underbrace{\int\limits_a^b f(x)\ dx}_{=1}  &= \int\limits_{-\infty}^P (P - x)\ f(x)\ dx + n_0 \underbrace{\int\limits_a^b f(x)\ dx}_{=1} . \nonumber
+\end{align}
+%
+$n_0$ and $p_0$ denote the baseline activity of the nPE and pPE neuron, respectively.
+In case of a uniform distribution (c.f. \ref{sec:impact_gain}), we get 
+%
 %
@@
 %
-Hence, if $p_0 = n_0$, the mean can be estimated correctly. For the V neuron to represent the variance of the inputs, this condition must be tightened to $p_0 = n_0 = 0$. The variance is given by
+Thus, the M neuron encodes the true mean of the feedforward input only if $p_0 = n_0$. Otherwise, XXX refer to Suppfig and state briefly the cases n > p or n<p XXX.
+
+With non-zero baseline activities, the steady state activity of the V neuron is given by
 %
@@
 \label{eq:condition_baseline_variance}
-V &= \langle \left( \mathrm{pPE} + \mathrm{nPE} \right)^2 \rangle \\
-&= \langle \left[ S-P\right]_+^2\rangle + \langle \left[ P-S\right]_+^2\rangle + (p_0 + n_0)^2 + 2\ (p_0 + n_0)\ \left( \langle \left[ S-P\right]_+\rangle + \langle \left[ P-S\right]_+ \rangle\right) \nonumber
-\end{align}
-%
-In case of a uniform distribution from which the inputs to the mean-field network are drawn, this expression simplifies to
+V &= \langle \left( r_\mathrm{pPE} + r_\mathrm{nPE} \right)^2 \rangle \\
+&= \langle \left[ s_\mathrm{FF}-P\right]_+^2\rangle + \langle \left[ P-s_\mathrm{FF}\right]_+^2\rangle + (p_0 + n_0)^2 + 2\ (p_0 + n_0)\ \left( \langle \left[ s_\mathrm{FF}-P\right]_+\rangle + \langle \left[ P-s_\mathrm{FF}\right]_+ \rangle\right) \nonumber
+\end{align}
+%
+In case of a uniform distribution, this expression yields
 %
@@
 %
-Inserting the expression for P (by itself modulated by the baseline activities of the PE neurons) yields
-%
-\begin{align}
-V =& \frac{1}{3\ (b-a)} \left[ \left( \frac{b-a}{2} - \frac{p_0 - n_0}{b-a}\right)^3 + \left( \frac{b-a}{2} + \frac{p_0-n_0}{b-a}\right)^3\right] + (p_0 + n_0)^2 \nonumber \\
-&+ \frac{(p_0 + n_0)}{b-a} \left[ \left( \frac{b-a}{2} + \frac{p_0-n_0}{b-a}\right)^2 + \left( \frac{b-a}{2} - \frac{p_0 - n_0}{b-a}\right)^2\right]
-\end{align}
-%
-Simplifying the expression, leads to
+Inserting the expression for P (Eg. \ref{eq:condition_baseline_mean_1}) which is itself a function of the baseline activities, gives
 %
@@
 \label{eq:condition_baseline_variance_2}
-V =  \frac{(b-a)^2}{12} + \frac{(p_0-n_0)^2}{(b-a)^2} \left( 1 + 2\ \frac{p_0+n_0}{b-a}\right) + (p_0 + n_0) \left( p_0 + n_0 + \frac{b-a}{2}\right)
-\end{align}
-%
+V =  \frac{(b-a)^2}{12} + \frac{(p_0-n_0)^2}{(b-a)^2} \left( 1 + 2\ \frac{p_0+n_0}{b-a}\right) + (p_0 + n_0) \left( p_0 + n_0 + \frac{b-a}{2}\right).
+\end{align}
+%
+Thus, for the V neuron to encode the variance unbiased, $n_0 = p_0 = 0$. If XXX discuss what you see in the supplementary figures ... .
 
 \subsection{Modelling the impact of neuromodulators on the sensory weight}
+
+- also note here that while prediction and variance might be biased, no necessarilty lead to changes in sensory weight ... explain why
 
@@
 
-\subsection*{Sensory weight and contraction bias}
-%
-If P is rather constant, the slope in the contraction bias is exactly the sensory weight
+\subsection{Sensory weight and contraction bias}
+%
+The contraction bias can be formalized through the slope in a linear equation describing the relationship between the estimated and true sensory stimulus.
+If the slope is close to 1, the bias is small.
+In contrast, if the slope is close to zero, the bias is large.
+
+If the prediction encoded in the activity of the M neuron changes very slowly, the bias is well-captured by the sensory weight, 
+%
+ \begin{align}
+ r_\mathrm{out} = \alpha_\mathrm{S} \cdot s_\mathrm{FF} + \mathrm{constant}
+ \end{align}
+%
+However, the prediction is normally a function of the history of sensory stimuli the network received, and cannot be approximated by a constant.
+To investigate how the slope (describing the contraction bias) depends on the sensory weight and other factors, let us resume an approximation in which the prediction decays exponentially to a presented constant stimulus value, $s$,
+%
+\begin{align}
+P = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau_M} +  s \cdot \left( 1 -   \mathrm{e}^{-t/\tau} \right)
+\end{align}
+%
+with $P_0$ describing the initial value at time $t=0$.
+Let us assume that within a trial with trial duration $T$, the stimulus value changes $n$ times ($T = n\cdot \Delta t$).
+The prediction during the presentation of the $n$th stimulus value can be expressed as
+%
+\begin{align}
+P_\mathrm{n} = P_\mathrm{0} \cdot \mathrm{e}^{-\Delta t/\tau_M}  + \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} s_i \cdot \mathrm{e}^{-(n-1)\cdot \Delta t/ \tau_M}
+\end{align}
+%
+To obtain an estimate for the prediction at the end of a trial, $P_\mathrm{n}$ must be averaged over the stimulus distribution, $\langle P_\mathrm{n} \rangle_s$.
+For the sake of simplicity, let us assume the stimulus values are drawn from a uniform distribution $U\left( s - \frac{\sigma_\mathrm{S}}{12}, s + \frac{\sigma_\mathrm{S}}{12} \right)$.
+Moreover, we assume that the initial state, $P_0$, at the beginning of a new trial is drawn from a uniform distribution $U\left( \mu - \frac{\sigma_\mathrm{P}}{12}, \mu + \frac{\sigma_\mathrm{P}}{12} \right)$.
+With these assumptions, $\langle P_\mathrm{n} \rangle_s$ is given by
+%
+\begin{align}
+\langle P_\mathrm{n} \rangle_s = \mathrm{e}^{-\Delta t/\tau_M}  \int\limits_{\mu - \frac{\sigma_\mathrm{P}}{12}}^{\mu + \frac{\sigma_\mathrm{P}}{12}} P_\mathrm{0} \ f(P_\mathrm{0})\ dP_0+ \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} \cdot \mathrm{e}^{-(n-1)\cdot \Delta t/ \tau_M} \int\limits_{s - \frac{\sigma_\mathrm{S}}{12}}^{s + \frac{\sigma_\mathrm{S}}{12}} x\ f(x)\ dx
+\end{align}
+%
+Solving the integrals yield
+%
+\begin{align}
+\langle P_\mathrm{n} \rangle_s = \mu \cdot \mathrm{e}^{-T/\tau_M} + \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot \Delta t/ \tau_M} \cdot s
+\end{align}
+%
+Making use of the geometric series, the expression simplifies to
 %
 \begin{align*}
- r_\mathrm{out} = \alpha_\mathrm{S} \cdot S + \left( 1 -\alpha_\mathrm{S} \right) \cdot P \equiv m \cdot S + n
+\langle P_\mathrm{n} \rangle_s =  \mu \cdot \mathrm{e}^{-T/\tau_M} + \left( 1 -   \mathrm{e}^{-T/\tau_M} \right) \cdot s
 \end{align*}
 %
-However, P is usually/normally a function of $S$. For simplicity, let's assume that P decays exponentially to a new value of $S$:
+Inserting the expression in the equation for the weighted output yields
 %
 \begin{align*}
-P = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau} + f(S) \cdot \left( 1 -   \mathrm{e}^{-t/\tau} \right)
+ r_\mathrm{out} = \left[ \alpha_\mathrm{S}\  \mathrm{e}^{-T/\tau_M} + \left( 1 -   \mathrm{e}^{-T/\tau_M} \right)\right] \cdot s + \left( 1 -\alpha_\mathrm{S} \right)\ \mathrm{e}^{-T/\tau_M}\ \mu
 \end{align*}
 %
-Within each trial with trial duration T, P can be expressed by n sections of length t in which the stimulus is constant and, for the sake of simplicity, drawn from a uniform distribution $U\left( s - \frac{\sigma_\mathrm{S}}{12}, s + \frac{\sigma_\mathrm{S}}{12} \right)$.
-$P_\mathrm{0}$ is drawn from $U\left( \mu - \frac{\sigma_\mathrm{P}}{12}, \mu + \frac{\sigma_\mathrm{P}}{12} \right)$.
-$P_\mathrm{n}$ is then given by
-%
-\begin{align*}
-P_\mathrm{n} = P_\mathrm{0} \cdot \mathrm{e}^{-t/\tau}  + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} s_i \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau}
-\end{align*}
-%
-This needs to be averaged over all possible states
-%
-\begin{align*}
-P_\mathrm{n} = \mathrm{e}^{-t/\tau}  \int\limits_{\mu - \frac{\sigma_\mathrm{P}}{12}}^{\mu + \frac{\sigma_\mathrm{P}}{12}} P_\mathrm{0} \ f(P_\mathrm{0})\ dP_0+ \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \cdot \mathrm{e}^{-(n-1)\cdot t/ \tau} \int\limits_{s - \frac{\sigma_\mathrm{S}}{12}}^{s + \frac{\sigma_\mathrm{S}}{12}} s\ f(s)\ ds
-\end{align*}
-%
-This gives
-%
-\begin{align*}
-P_\mathrm{n} = \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-t/\tau} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot t/ \tau} \cdot S
-\end{align*}
-%
-By making use of the geometric series, this simplifies to
-%
-\begin{align*}
-P_\mathrm{n} =  \mu \cdot \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right) \cdot S
-\end{align*}
-%
-Together, this yields for the weighted output
-%
-\begin{align*}
- r_\mathrm{out} = \left[ \alpha_\mathrm{S}\  \mathrm{e}^{-T/\tau} + \left( 1 -   \mathrm{e}^{-T/\tau} \right)\right] \cdot S + \left( 1 -\alpha_\mathrm{S} \right)\ \mathrm{e}^{-T/\tau}\ \mu
-\end{align*}
-%
 Hence, the slope is a function of both the sensory weight and the trial duration.
 
+XXX check the derivation above and then polish the below XXX Also, it might be important to state that alpha itself is a function of the trial duration!? xx
 In a prediction-driven input regime ($alpha_\mathrm{S} \sim 0)$, the slope is independent of the sensory weight and only determined by the trial duration, $m \sim \left( 1 -   \mathrm{e}^{-T/\tau} \right)$.

### 2023-10-06 scripts/run_fig_2_support_2.py

@@ -22,7 +22,7 @@ dtype = np.float32
 # %% Sparsity 
 
 run_cell = True
-plot_only = False
+plot_only = True
 
 if run_cell:
     
@@ -113,7 +113,7 @@ if run_cell:
 
 # %% Correlated deviations  
 
-run_cell = True
+run_cell = False
 plot_only = False
 
 if run_cell:
@@ -203,7 +203,7 @@ if run_cell:
 # %% Uncorrelated deviations  
 
 run_cell = False
-plot_only = True
+plot_only = False
 
 if run_cell:
     

### 2023-10-06 scripts/run_fig_3.py

@@ -104,6 +104,8 @@ if run_cell:
                                 ylabel='variability within trial')
     
 
+# %% Probably all below can be erased ....
+
 # %% Systematic exploration - record neuron activity
 
 run_cell = True

### 2023-10-06 scripts/run_fig_4.py

@@ -91,140 +91,59 @@ if run_cell:
     
     plot_neuromod_impact_inter(column, std_mean, n_std, s=30)
 
-                                                                      
-# %% ################################## OLD ################################### 
-# #############################################################################
 
-# %% Activate fractions of IN neurons in lower/higher PE circuit or both for a specified input statistics
+# %% How are the variance neurons influenced by changes in nPE and pPE neurons in the lower and higher PE circuit?
+
+ # changes in nPE and pPE neurons in lower PE circuit will lead to changes in the variance neuron in the lower PE circuit 
+ # AND the higher PE circuit
+ # the latter changes occur due to changes in the prediction neuron of the lower PE circuit (see calculations & illustration)
+ # changes in PE neurons in higher PE circuit will lead to changes in the variance of the higher column
+ # changes in both will be superimposed of the first two ...
 
 run_cell = False
 plot_only = True
 
 if run_cell:
-    
+
     ### choose mean-field network to simulate
     mfn_flag = '10' # valid options are '10', '01', '11
-    
+
     ### define target area and stimulus statistics
-    column = 0      # 0: both, 1: lower level PE circuit, 2: higher level PE circuit
-    std_mean = 1   # uncertainty of environement [0, 0.5, 1]
-    n_std = 0       # uncertainty of stimulus [1, 0.5, 0]
-    
+    column = 2      # 1: lower level PE circuit, 2: higher level PE circuit
+    std_mean = 1   
+    n_std = 1       
+
     ### filename for data
     identifier = '_column_' + str(column) + '_acrossvar_' + str(std_mean) + '_withinvar_' + str(n_std)
-    file_for_data = '../results/data/neuromod/data_weighting_neuromod_' + mfn_flag + identifier + '.pickle'
-    
+    file_for_data = '../results/data/neuromod/data_moments_vs_PE_neurons_' + mfn_flag + identifier + '.pickle'
+
     ### get data
     if not plot_only: # simulate respective network
 
         nums = 11
[…]

### 2023-10-10 records/manuscript_2023.tex
Polish Code for Figs 1 - 2 S2

@@
 {\Large
-\textbf\newline{Title}
+\textbf\newline{Weighting sensory inputs and predictions with prediction-error neurons}
 }
-%\newline
-%% authors go here:
-%\\
-%Loreen Hert\"ag\textsuperscript{1,*},
-%Claudia Clopath\textsuperscript{1,§}
-%\\
-%\bigskip
-%\bf{1} Bioengineering Department, Imperial College London, London, UK.
-%\\
-%\bigskip
-%* l.hertag@imperial.ac.uk, § c.clopath@imperial.ac.uk
+\newline
+% authors go here:
+\\
+Loreen Hert\"ag\textsuperscript{1,*},
+Katharina A. Wilmes\textsuperscript{2},
+Claudia Clopath\textsuperscript{3}
+\\
+\bigskip
+1 Modelling of Cognitive Processes, TU Berlin, Berlin, Germany.\\
+2 Department of Physiology, University of Bern, Switzerland.\\
+3 Bioengineering Department, Imperial College London, London, UK.
+\\
+\bigskip
+* loreen.hertaeg@tu-berlin.de
 
@@
 %
-In the following, we describe in more detail the equations describing the dynamics of the neurons in the prediction-error circuit, as well as the memory and variance neurons.
+In the following, we describe in more detail the equations for the dynamics of the neurons in the prediction-error circuit, as well as the memory and variance neurons.
 We then provide the connectivity of the network and the inputs to the neurons for both the mean-field and population model.
@@
 Important factors are: 1) long enough to get to steady state (as PE circuits used were learned for steady state, 2) lower M updates faster than higher M, 3) update speed for M and V (see lambda and theta etc.) must be chosen in accordance with trial duration stimulus duration.
-XXX
+XXX This could also go to the section Simulations xxx
 
@@
 %
-The connectivity between neurons of the PE circuit, both for the mean-field and the population network, was taken from XXX.
-XXX ADD heatmaps for the connectivity?
-or simply refer to the 2022 paper (maybe rather that)
+The connectivity between neurons of the PE circuit, both for the mean-field and the population network, were taken from \ref{XXX}.
+We considered three mean-field networks (see Table \ref{tab:wXM}) that differed in terms of the inputs (feedforward vs.
+feedback) onto the SOM and VIP neurons, and, hence, in their connectivity that established an E/I balance in the excitatory neurons.
 
@@
 To ensure that the activities of the memory neurons represent the mean of the sensory stimuli in the lower PE circuit and the mean of the prediction in the higher subnetwork, respectively, the net effect of nPE and pPE neurons must cancel in the steady state (see Analysis in \ref{sec:gain_impact}).
-Hence, the weights need to account for the neurons' potentially different gain factors ($g_\mathrm{nPE}$ and $g_\mathrm{pPE}$, see Table XXX) and the neuron numbers ($N_\mathrm{nPE}$ and $N_\mathrm{pPE}$, see Table XXX):
+Hence, the weights need to account for the neurons' potentially different gain factors ($g_\mathrm{nPE}$ and $g_\mathrm{pPE}$) and the neuron numbers ($N_\mathrm{nPE}$ and $N_\mathrm{pPE}$):
 %
@@
 
-XXX Here you must say that this is true for the MFN and the one in which you copied the MFN but in case of the population, each neuron has a different gain, lambda is xxx for allXXX not sure how to show all the different gain factors?
-Also, you must say that the weights were systematically changed in Supp Fig.
-XXX to investigate correlated and uncorrelated changes XXX
-
-The memory neuron \textit{M} connects to the post-synaptic neurons \textit{X} in the PE circuit with the connection strength $w_\mathrm{X\leftarrow M}$.
-If a connection exists, we assume $w_\mathrm{X\leftarrow M} = 1$.
-In all mean-field networks and the population network, the dendrites of nPE and pPE neurons and one of the two (populations of) PV neurons receive connections from the memory neuron.
-We considered three mean-field networks that differed in terms of the inputs onto the SOM and VIP neurons:
+For the mean-field networks ($ N_\mathrm{nPE} = N_\mathrm{pPE} = 1 $), the gain factors are given in Table~\ref{tab:gain_factors_MFN}.
+For the population network, the gain factors for all PE neurons are shown in Fig.
+\ref{fig:Fig_gains}.
 %
@@
 \hline
-Network name & MFN 1 & MFN 2 & MFN 3  \\
+ & \textbf{MFN 1} & \textbf{MFN 2} & \textbf{MFN 3}  \\
+\textbf{Network} & FF $\rightarrow$ SOM  & FB $\rightarrow$ SOM  & FF $\rightarrow$ SOM  \\
+ & FB $\rightarrow$ VIP  & FF $\rightarrow$ VIP  & FF $\rightarrow$ VIP  \\
+\hline
+\hline
+nPE & 1 & 1.7 & 2.5\\
+pPE & 1 & 1.7 & 2.5 \\
+\hline
+\end{tabular}
+\caption{\footnotesize{Gain factors for nPE and pPE neurons in three different mean-field networks (MFN).
+Each MFN differs with respect to the inputs onto SOM and VIP neurons.
+The interneurons either receive the feedforward (FF) or feedback (FB) input.
+All numbers are rounded to the first digit.}}
+\label{tab:gain_factors_MFN}
+\end{table}
+%
+%
+\begin{figure}[h!]
+	\centering
+    \includegraphics{../results/figures/final/Figure_gains}
+\caption{\footnotesize{\bf Gain factors of nPE and pPE neurons in the population model.\newline}
+{The logarithm of the gain factors of nPE (top) and pPE (bottom) neurons in the population model from \ref{XX}.
+The network contains $67$ nPE neurons and $66$ pPE neurons.
+The remaining excitatory neurons were not classified as PE neurons and were not connected to the $M$ neuron.
+}}
+\label{fig:Fig_gains}
+\end{figure}
+%
+%
+
+The memory neuron \textit{M} connects to the post-synaptic neurons \textit{X} in the PE circuit with the connection strength $w_\mathrm{X\leftarrow M}$.
+If a connection exists, we assume $w_\mathrm{X\leftarrow M} = 1$.
+In all mean-field networks and the population network, the dendrites of nPE and pPE neurons and one of the two (populations of) PV neurons receive connections from the memory neuron.
+Furthermore, we assume that the \textit{M} neuron does not excite the soma of PCs.
+Whether the SOM or VIP neurons are the target of the feedback projections depend on the specific mean-field network (see Table \ref{tab:wXM}).
+In the population model, 30\% of the SOM neurons and 70\% of the VIP neurons receive input from the memory neuron.
+%
+\begin{table}[h!]
+\centering
+\begin{tabular}{ |c|c|c|c|  }
+\hline
+ & \textbf{MFN 1} & \textbf{MFN 2} & \textbf{MFN 3}  \\
+\textbf{Network} & FF $\rightarrow$ SOM  & FB $\rightarrow$ SOM  & FF $\rightarrow$ SOM  \\
+ & FB $\rightarrow$ VIP  & FF $\rightarrow$ VIP  & FF $\rightarrow$ VIP  \\
 \hline
@@
 \end{tabular}
-\caption{XXX Table shows $w_\mathrm{X\leftarrow M}$ ... .}
+\caption{\footnotesize{$w_\mathrm{X\leftarrow M}$ for the post-synaptic SOM and VIP neurons in all three mean-field networks considered.}}
 \label{tab:wXM}
@@
 
-In the population model, 30\% of the SOM neurons and 70\% of the VIP neurons receive input from the memory neuron.
-
 \subsubsection{Connections between the PE circuit and the V neuron}
@@
 Both nPE and pPE neurons excite the \textit{V} neuron.
-To ensure that the activity of the V neuron represents the variance of the input (see Analysis in \ref{sec:gain_impact}), the weights must account for differences in the gains ($g_\mathrm{nPE}$ and $g_\mathrm{pPE}$, see Table XXX) and numbers ($N_\mathrm{nPE}$ and $N_\mathrm{pPE}$) of the PE neurons,
+To ensure that the activity of the V neuron represents the variance of the input (see Analysis in \ref{sec:gain_impact}), the weights must account for differences in the gains ($g_\mathrm{nPE}$ and $g_\mathrm{pPE}$, see Table \ref{tab:gain_factors_MFN} and Fig.
+\ref{fig:Fig_gains}) and numbers ($N_\mathrm{nPE}$ and $N_\mathrm{pPE}$) of the PE neurons,
 %
@@
 %
-$\theta$ XXX represents a factor that ensures that for population network V still represents variance, no cross -terms, in mean-field network it is 1, XXX check if for population network, it is ok to ignore the nPE1*nPE2 terms, I mean you have cross-terms that do something with the variance, do they XXX
-
-We assume a quadratic activation function for the V neuron.
-In case of a mean-field network, the choice of the activation function ensures that the activity of the V neuron represents the variance of the input.
-This is only true because nPE and pPE neurons are not active at the same time and have a baseline activity of zero.
-=> cross-terms are zero.
-However, if we increase the number of nPE and pPE neurons, cross-terms of the nature $nPE_1 * nPE_2$ will introduce deviations from the the variance.
-These effects can be mitigated or fully corrected by choosing $\theta$ accordingly.
-$\theta$ depends on the distribution at hand.
-In our simulations, we only considered a uniform distribution.
-
-% XXX here start with nPE and pPE
-\begin{align}
-\left(\frac{\theta}{N} \sum_{i} |s_i - \mu| \right)^2 =& \left( \frac{\theta}{N}  \sum_{s_i \geq \mu}^{N/2} (s_i - \mu) + \frac{\theta}{N}  \sum_{s_i \leq \mu}^{N/2} (\mu - s_i )\right)^2 \nonumber \\
- =& \frac{\theta^2}{4} \left( \frac{2}{N} \sum_{s_i \geq \mu}^{N/2} s_i - \frac{2}{N} \sum_{s_i \leq \mu}^{N/2} s_i\right)^2\nonumber\\
- =& \frac{\theta^2}{4} \left( \frac{b+\mu}{2} - \frac{\mu + a}{2} \right)^2 = \frac{(b-a)^2}{16} \cdot \theta^2
-\end{align}
-%
-Hence, to ensure the variance of a uniform distribution given by $\frac{(b-a)^2}{12}$, $\theta = \frac{2}{\sqrt{3}}$.
+We introduce the factor $\theta$ to compensate for cross-terms that are the result of a quadratic activation function but are not in line with the definition of the variance.
+
+In the mean-field network, $\theta = 1$ because nPE and pPE neuron activity is mutually exclusive, and, hence, the cross-term $\mathrm{nPE}\cdot\mathrm{pPE}$ would be zero (under the assumption that they have a negligible baseline activity).
+This is also true for the population model, in which each nPE neuron only contributes a small fraction to the overall, mean nPE, and each pPE neuron only contributes a small fraction to the overall, mean pPE.
+
+However, in Supp Fig.
+\ref{fig:Fig_2_S3} XXX, each mean-field network receives a stimulus $s_i$ drawn from a distribution.
+In this case, $\theta$ must be chosen such that deviations from the true variance can be mitigated or fully corrected.
+The true $\theta$ depends on the distribution at hand.
+In our simulations, we used a uniform distribution $U(a,b)$, in which case $\theta$ can be derived from
+%
+\begin{align}
+\left( \sum_{i} r_\mathrm{nE,i} + \sum_{i} r_\mathrm{pE,i}\right)^2 \overset{(i)}{=} & \left( \frac{\theta}{N}  \sum_{s_i \geq \mu}^{N/2} (s_i - \mu) + \frac{\theta}{N}  \sum_{s_i \leq \mu}^{N/2} (\mu - s_i )\right)^2 \nonumber \\
+%
+=&\ \frac{\theta^2}{4} \left( \frac{2}{N} \sum_{s_i \geq \mu}^{N/2} s_i - \frac{2}{N} \sum_{s_i \leq \mu}^{N/2} s_i\right)^2\nonumber\\
+%
+\overset{(ii)}{=}&\ \frac{\theta^2}{4} \left( \frac{b+\mu}{2} - \frac{\mu + a}{2} \right)^2 = \frac{(b-a)^2}{16} \cdot \theta^2,
+\label{eq:theta}
+\end{align}
+%
+where we assumed that (i) the number of nPE and pPE neurons is equal, and (ii) this number goes to infinity.
+Comparing eq.
+(\ref{eq:theta}) with the equation for the variance of a uniform distribution, $\frac{(b-a)^2}{12}$, we get $\theta = \frac{2}{\sqrt{3}}$.
 
@@
 
-The feedforward input is either the direct sensory input $s$ for the lower PE circuit, or the activity of the M neuron $r_\mathrm{M}$ for the higher PE circuit.
+The feedforward input is either the direct sensory input $s$ for the lower PE circuit, or the activity of the M neuron, $r_\mathrm{M}$, for the higher PE circuit.
 In general, for the three mean-field networks tested, we chose $w_i = 1 - w_\mathrm{X\leftarrow M}$  (see Table \ref{tab:wXM}).
@@
 %
-\textbf{Figure 2:} Simulation time steps = $10^5$, $200$ constant values (each $500$ time steps long) drawn from a uniform distribution, B and C: input standard deviation fixed at $5$ when mean is varied, input mean fixed at $4.5$ when input variance is varied.\\
+\textbf{Figure 2:} Simulation time steps = $10^5$, $200$ constant values (each $500$ time steps long) drawn from a uniform distribution, B and C: input standard deviation fixed at $4.5$ when mean is varied, input mean fixed at $4.5$ when input variance is varied.\\
 %
@@
       \frac{a + b}{2} & \text{if}\ g_\mathrm{nPE} = g_\mathrm{pPE} = g \\
-      \frac{g_\mathrm{pPE}\cdot b - g_\mathrm{nPE}\cdot a \pm \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}} (a-b)}{g_\mathrm{pPE} - g_\mathrm{nPE}} & \text{otherwise.}
+      \frac{g_\mathrm{pPE}\cdot b - g_\mathrm{nPE}\cdot a + \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}}\ (a-b)}{g_\mathrm{pPE} - g_\mathrm{nPE}} & \text{otherwise.}
     \end{cases}
@@
 %
-XXX Maybe mention the SuppFig. in which you show the function XXX and explain what you see for gnPE > gpPE or vice versa ...
+Hence, the mean of the feedforward input is overpredicted when $g_\mathrm{nPE} < g_\mathrm{pPE}$.
+Similarly, the mean of the feedforward input is underpredicted when $g_\mathrm{nPE} > g_\mathrm{pPE}$ (Fig.
+\ref{fig:Fig_4_S2}).
 
@@
 \label{eq:condition_variance_gain}
-V &= g_\mathrm{pPE}\ \langle \left[ s_\mathrm{FF}-P \right]_+^2 \rangle + g_\mathrm{nPE}\ \langle \left[ P-s_\mathrm{FF} \right]_+^2 \rangle.
-\end{align}
-%
-In case of a uniform distribution, we get
-%
-\begin{align}
-V &= \frac{g_\mathrm{pPE}}{b-a} \int\limits_P^b (x-P)^2\ dx + \frac{g_\mathrm{nPE}}{b-a} \int\limits_a^P (P-x)^2\ dx \\
-   &= \frac{g_\mathrm{pPE}}{3} \cdot \frac{(b-P)^3}{b-a} + \frac{g_\mathrm{nPE}}{3} \cdot \frac{(P-a)^3}{b-a}. \nonumber
+V &= \langle \left(r_\mathrm{pPE} + r_\mathrm{nPE}\right)^2 \rangle \overset{(i)}{=} \langle r_\mathrm{pPE}^2 \rangle + \langle r_\mathrm{nPE}^2 \rangle \nonumber\\
+&=g_\mathrm{pPE}^2\ \langle \left[ s_\mathrm{FF}-P \right]_+^2 \rangle + g_\mathrm{nPE}^2\ \langle \left[ P-s_\mathrm{FF} \right]_+^2 \rangle,
+\end{align}
+%
+where we assume (i) that both the nPE and pPE neuron have a zero baseline activity. In case of a uniform distribution, we get
+%
+\begin{align}
+V &= \frac{g_\mathrm{pPE}^2}{b-a} \int\limits_P^b (x-P)^2\ dx + \frac{g_\mathrm{nPE}^2}{b-a} \int\limits_a^P (P-x)^2\ dx \\
+   &= \frac{g_\mathrm{pPE}^2}{3} \cdot \frac{(b-P)^3}{b-a} + \frac{g_\mathrm{nPE}^2}{3} \cdot \frac{(P-a)^3}{b-a}. \nonumber
 \end{align}
@@
       \frac{(b - a)^2}{12} & \text{if}\ g_\mathrm{nPE} = g_\mathrm{pPE} = 1 \\
-      \frac{(b-a)^2}{3\ (g_\mathrm{pPE} - g_\mathrm{nPE})^3} \cdot \left[ g_\mathrm{nPE} \cdot( g_\mathrm{pPE} \mp \gamma)^3 - g_\mathrm{pPE} \cdot (g_\mathrm{nPE} \mp \gamma)^3\right] & \text{otherwise.}
+      \frac{(b-a)^2}{3\ (g_\mathrm{pPE} - g_\mathrm{nPE})^3} \cdot \left[ g_\mathrm{nPE}^2 \cdot( g_\mathrm{pPE} - \gamma)^3 - g_\mathrm{pPE}^2 \cdot (g_\mathrm{nPE} - \gamma)^3\right] & \text{otherwise.}
     \end{cases}
@@
 with $\gamma = \sqrt{g_\mathrm{nPE}\ g_\mathrm{pPE}}$.
-XXX As above, discuss cases from the Supp Figs. XXX Important: for P and V please check with sign is correct!
-
-\subsection{Impact of PE neurons' baseline on estimating mean and variance}
+Hence, the variance of the feedforward input is overpredicted when $g_\mathrm{nPE}>1$ or $g_\mathrm{pPE}>1$.
+Similarly, the variance of the feedforward input is underpredicted when $g_\mathrm{nPE} < 1$ or $g_\mathrm{pPE} < 1$ (Fig.
+\ref{fig:Fig_4_S2}).
+
+\subsection{Impact of PE neurons' baseline on estimating mean and variance}\label{sec:impact_baseline} 
 %
@@
 %
-Thus, the M neuron encodes the true mean of the feedforward input only if $p_0 = n_0$. Otherwise, XXX refer to Suppfig and state briefly the cases n > p or n<p XXX.
+Thus, the M neuron encodes the true mean of the feedforward input only if $p_0 = n_0$.
+Hence, the mean is overpredicted if $p_0 > n_0$.
+Likewise, the mean is underpredicted if $p_0 < n_0$ (see Fig.
+\ref{fig:Fig_4_S2}).
 
@@
 %
-In case of a uniform distribution, this expression yields
+In case of a uniform distribution $U(a,b)$, this expression yields
 %
@@
 %
-Inserting the expression for P (Eg. \ref{eq:condition_baseline_mean_1}) which is itself a function of the baseline activities, gives
+Inserting the expression for P (Eq. \ref{eq:condition_baseline_mean_1}) which is itself a function of the baseline activities, gives
 %
@@
 %
-Thus, for the V neuron to encode the variance unbiased, $n_0 = p_0 = 0$. If XXX discuss what you see in the supplementary figures ... .
+Thus, for the V neuron to encode the variance unbiased, $n_0 = p_0 = 0$. The variance is overpredicted if either $n_0 > 0$ or $p_0 > 0$ (see Fig. \ref{fig:Fig_4_S2}).
 
 \subsection{Modelling the impact of neuromodulators on the sensory weight}
-
-- also note here that while prediction and variance might be biased, no necessarilty lead to changes in sensory weight ... explain why
-
-Assume that neuromodulator act through activating INs.
-INs on the other hand modulate both gain and BL of nPE and pPE neurons ...
-.
-Changes in gain and BL will affect the overall activity and the balance between nPE and pPE neurons => this will affect both the prediction and the variance.
-
-Modulate PE in lower:
-variance in lower is combination of direct changes in PE and indirect through prediction (that is changed through PE mod)
-variance in higher is just because of differences in prediction
-
-Modulation in higher:
-variance in lower should be unaffected
-variance in higher is combination of direct changes in PE and indirect through prediction (that is changed through PE mod)
-
-Variance as function of bias in mean:
+%
+We modeled the presence of a neuromodulator by simulating an additive excitatory input onto (groups of) interneurons.
+These interneurons, in turn, modulate the gain and baseline of PE neurons.
+As shown in sections \ref{sec:impact_gain} and \ref{sec:impact_baseline}, changes in the input-output transfer function of the PE neurons may bias the variance estimation in the network, and, hence, the sensory weight.
+Thus, understanding changes in the sensory weight requires an understanding whether and how different types of interneurons change the PE neurons.
+
+If a neuromodulator only acts on interneurons of the lower-level subnetwork, the sensory weight changes as a consequence of the modulated firing rates of the lower-level and higher-level $V$ neurons.
+The lower-level $V$ neuron is directly affected by the changes in the lower-level PE neurons and indirectly affected by changes in the $M$ neuron of the same network.
+The higher-level $V$ neuron is also affected by a neuromodulator acting in the lower-level subnetwork because the lower-level $M$ neuron projects onto the neurons in the higher-level PE circuit.
+Hence, if the lower-level $M$ neuron represents a biased mean $\mu \pm \delta\mu$, the variance estimation will be biased as well.
+This can be seen directly from the definition of the variance,
+%
 \begin{align*}
@@
 
+In contrast, if a neuromodulator only acts on interneurons of the higher-level subnetwork, the sensory weight changes as a consequence of the modulated firing rates of the higher-level $V$ neuron.
+The higher-level $V$ neuron is directly affected by the changes in the higher-level PE neurons and indirectly affected by changes in the $M$ neuron of the same network.
+
+Together, this suggests that whether the sensory weight decreases, increases or remains the same in the presence of a neuromodulator depends on a number of factors:
+%
+\begin{itemize}
+\item Does the neuromodulator act on the lower-level or higher-level subnetwork (that is, local impact), or does the neuromodulator act on both to the same degree (that is, global impact)?
+\item Which interneuron type/s is/are affected by the neuromodulator? And are these interneurons inhibited or excited by the neuromodulator?
+\item How are these interneurons embedded in the network, that is, what is the connectivity and the inputs to those neurons?
+\end{itemize}
+%
+Hence, different neuromodulators may have the same effect on the sensory weight or the same neuromodulator may have different effects depending on brain area, species etc..
+
 \subsection{Sensory weight and contraction bias}
@@
 However, the prediction is normally a function of the history of sensory stimuli the network received, and cannot be approximated by a constant.
-To investigate how the slope (describing the contraction bias) depends on the sensory weight and other factors, let us resume an approximation in which the prediction decays exponentially to a presented constant stimulus value, $s$,
+To investigate how the slope (describing the contraction bias) depends on the sensory weight and other factors, let us resume a toy model in which we assume that the prediction decays exponentially with time constant $\tau$ to a presented constant stimulus value, $s$,
 %
@@
 \begin{align}
-P_\mathrm{n} = P_\mathrm{0} \cdot \mathrm{e}^{-\Delta t/\tau_M}  + \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} s_i \cdot \mathrm{e}^{-(n-1)\cdot \Delta t/ \tau_M}
+P_\mathrm{n} = P_\mathrm{0} \cdot \mathrm{e}^{-\Delta t/\tau_M}  + \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} s_i \cdot \mathrm{e}^{-(n-i)\cdot \Delta t/ \tau_M}
 \end{align}
@@
 \begin{align}
-\langle P_\mathrm{n} \rangle_s = \mathrm{e}^{-\Delta t/\tau_M}  \int\limits_{\mu - \frac{\sigma_\mathrm{P}}{12}}^{\mu + \frac{\sigma_\mathrm{P}}{12}} P_\mathrm{0} \ f(P_\mathrm{0})\ dP_0+ \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} \cdot \mathrm{e}^{-(n-1)\cdot \Delta t/ \tau_M} \int\limits_{s - \frac{\sigma_\mathrm{S}}{12}}^{s + \frac{\sigma_\mathrm{S}}{12}} x\ f(x)\ dx
+\langle P_\mathrm{n} \rangle_s = \mathrm{e}^{-\Delta t/\tau_M}  \int\limits_{\mu - \frac{\sigma_\mathrm{P}}{12}}^{\mu + \frac{\sigma_\mathrm{P}}{12}} P_\mathrm{0} \ f(P_\mathrm{0})\ dP_0+ \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot \Delta t/ \tau_M} \int\limits_{s - \frac{\sigma_\mathrm{S}}{12}}^{s + \frac{\sigma_\mathrm{S}}{12}} x\ f(x)\ dx
 \end{align}
@@
 %
-Hence, the slope is a function of both the sensory weight and the trial duration.
-
-XXX check the derivation above and then polish the below XXX Also, it might be important to state that alpha itself is a function of the trial duration!? xx
-In a prediction-driven input regime ($alpha_\mathrm{S} \sim 0)$, the slope is independent of the sensory weight and only determined by the trial duration, $m \sim \left( 1 -   \mathrm{e}^{-T/\tau} \right)$.
-In a sensory-driven input regime ($alpha_\mathrm{S} \sim 0)$, the contraction bias vanishes ($m \sim 1$).
-
-If the trail duration is short ($T \rightarrow 0$), the slope is given by the sensory weight.
-If the trail duration approaches infinity, the slope would be 1 again (however, this seems rather unrealistic, this would only be true in an ideal system without memory decay or reproduction and accumulation noise ...).
-
-%\subsection{Influence of a population of nPE and pPE neurons}
-%XXX
-%
-%
-%\subsection{Analysis of simplified network model, effect of time constants}
-%
-%simplified model: dynamics and steady state of rM and rV, rM and rV as a function of time constants and trial duration etc., weighting, then use those expressions to discuss when weighting goes awry and how long transitions take from one state to another ...
+Hence, the slope $  m = \left[ \alpha_\mathrm{S}\  \mathrm{e}^{-T/\tau_M} + \left( 1 -   \mathrm{e}^{-T/\tau_M} \right)\right] $ depends on the sensory weight $\alpha_\mathrm{S}$, the trial duration $T$ and time constant $\tau$.
+Please note that the sensory weight is a function of the trial duration itself (see Fig.
+\ref{fig:Fig_3}F).
+However, for illustration purposes, we take $\alpha_\mathrm{S}$ to be constant here.
+
+In this toy model, if the variance of the prediction is zero (that is, in a prediction-driven input regime), $\alpha_\mathrm{S} \approx 0$, and, hence, the slope is $\left( 1 -   \mathrm{e}^{-T/\tau_M} \right)$.
+Thus, decreasing or increasing the stimulus variance does not have an effect on the bias (see Fig.
+\ref{fig:Fig_5} D).
+
+Likewise, if the variance of the sensory stimulus is zero (that is, in a stimulus-driven input regime), $\alpha_\mathrm{S} \approx 1$, and, hence, the slope is $1$.
+Thus, decreasing or increasing the trail variance does not have an effect on the bias (see Fig.
+\ref{fig:Fig_5} C).
+
+Moreover, the toy model also reveals that if $T\rightarrow \infty$, the slope $m$ goes to 1, suggesting that without further noise in the system, the network can perfectly represent the sensory stimulus (the prediction would approach the current stimulus).
 

### 2023-10-10 scripts/plot_fig_1.py

@@ -32,20 +32,15 @@ if not os.path.exists(figPath):
 
 # %% Define figure structure
 
-figsize=(5.5/inch,2/inch)
+figsize=(3/inch,2/inch)
 fig = plt.figure(figsize=figsize)
 
-A = gridspec.GridSpec(1, 2, figure=fig, wspace=0.6)
-
-# ax_A =  fig.add_subplot(A[:,:])
-# ax_A.axis('off')
-# ax_A.set_xlabel('Sens - Pred (1/s)', fontsize=fs)
+A = gridspec.GridSpec(1, 1, figure=fig, wspace=0.6)
 
 ax_1 = fig.add_subplot(A[0,0])
-ax_2 = fig.add_subplot(A[0,1])
 
 
-# %% Show Activity of nPE and pPE, compare with (S-P)**2
+# %% Show Activity of nPE and pPE
 
 # data files
 file_for_data = '../results/data/moments/data_mfn_10_PE_neurons_constant_stimuli_P_fixed.pickle'
@@ -59,7 +54,8 @@ else:
         [prediction_initial, stimulus_tested, nPE, pPE] = pickle.load(f)
     
     # plot data
-    plot_nPE_pPE_activity_compare(prediction_initial, stimulus_tested, nPE, pPE, ax1 = ax_1, ax2 = ax_2, ms=3, lw=2, fs=fs)
+    plot_nPE_pPE_activity_compare(prediction_initial, stimulus_tested, nPE, pPE, ax1 = ax_1, ms=3, lw=2, fs=fs+1)
+
 
 # %% save figure
 

### 2023-10-10 scripts/plot_fig_2.py

@@ -29,10 +29,9 @@ figPath = '../results/figures/final/'
 if not os.path.exists(figPath):
     os.mkdir(figPath)
     
-
 # %% Define figure structure
 
-figsize=(18/inch,15/inch)
+figsize=(18/inch,13/inch)
 fig = plt.figure(figsize=figsize)
 
 G = gridspec.GridSpec(3, 1, figure=fig, hspace=1.2)
@@ -69,25 +68,21 @@ ax_D = fig.add_subplot(G2[:,:])
 ax_D.axis('off')
 ax_D.text(-0.05, 1.25, 'D', transform=ax_D.transAxes, fontsize=fs+1)
 
-#ax_D1 = fig.add_subplot(G2[:,0])
 ax_D21 = fig.add_subplot(G2[:,1])
-#ax_D22 = fig.add_subplot(G2[1,1])
 ax_D3 = fig.add_subplot(G2[:,2])
 
 ax_E = fig.add_subplot(G3[:,:])
 ax_E.axis('off')
 ax_E.text(-0.05, 1.25, 'E', transform=ax_E.transAxes, fontsize=fs+1)
 
-#ax_E1 = fig.add_subplot(G3[:,0])
 ax_E21 = fig.add_subplot(G3[:,1])
-#ax_E22 = fig.add_subplot(G3[1,1])
 ax_E3 = fig.add_subplot(G3[:,2])
 
 
 # %% Neuron activity with increasing stimulus mean and variance
 
 # data files
-file_for_data = '../results/data/moments/data_neuron_activity_heatmap_mfn_10.pickle'
+file_for_data = '../results/data/moments/data_heatmap_mfn_10.pickle' 
 
 if not os.path.exists(file_for_data):
     print('Data does not exist yet. Please run corresponding file.')
@@ -96,10 +91,10 @@ else:
     # load data
     with open(file_for_data,'rb') as f:
         [trial_duration, num_values_per_trial, means_tested, variances_tested, 
-         mse_mean, mse_variance, activity_pe_neurons, activity_interneurons] = pickle.load(f)
+         dev_mean, dev_variance, activity_pe_neurons, activity_interneurons] = pickle.load(f)
     
     # plot data
-    end_of_initial_phase = np.int32(trial_duration * 0.5)
+    end_of_initial_phase = np.int32(trial_duration * 0.75)
     plot_neuron_activity(end_of_initial_phase, means_tested, variances_tested, activity_interneurons, None, id_fixed=3,
                          ax1=ax_C1, ax2=ax_C2)
[…]

### 2023-10-10 scripts/run_fig_1.py

@@ -14,7 +14,7 @@ import pickle
 
 from src.functions_simulate import simulate_PE_circuit_P_fixed_S_constant
 from src.plot_data import plot_example_mean, plot_example_variance, plot_mse_heatmap, plot_interneuron_activity_heatmap
-from src.plot_data import plot_neuron_activity
+from src.plot_data import plot_neuron_activity, plot_nPE_pPE_activity_compare
 
 import warnings
 warnings.filterwarnings("ignore")
@@ -22,10 +22,10 @@ warnings.filterwarnings("ignore")
 dtype = np.float32
 
 
-# %% How do the interneurons & PE neurons behave with different input statistics? (later you could combine it with the one above!)
+# %% Illustrate activity of nPE and pPE neurons
 
 run_cell = True
-plot_only = False
+plot_only = True
 
 if run_cell:
     
@@ -56,12 +56,4 @@ if run_cell:
             [prediction_initial, stimulus_tested, nPE, pPE] = pickle.load(f)
         
     ### plot 
-    plt.figure()
-    plt.plot(stimulus_tested - prediction_initial, nPE)
-    plt.plot(stimulus_tested - prediction_initial, pPE)
-    
-    plt.figure()
-    plt.plot(stimulus_tested - prediction_initial, nPE**2)
-    plt.plot(stimulus_tested - prediction_initial, pPE**2)
-    plt.plot(stimulus_tested - prediction_initial, (stimulus_tested - prediction_initial)**2)
-
+    plot_nPE_pPE_activity_compare(prediction_initial, stimulus_tested, nPE, pPE)

### 2023-10-10 scripts/run_fig_2.py

@@ -21,53 +21,10 @@ warnings.filterwarnings("ignore")
 dtype = np.float32
 
 
-# %% How do the interneurons & PE neurons behave with different input statistics? (later you could combine it with the one above!)
-
-run_cell = True
-plot_only = True
-
-if run_cell:
-    
-    ### choose mean-field network to simulate
-    mfn_flag = '10' # valid options are '10', '01', '11
-    
-    ### filename for data
-    file_for_data = '../results/data/moments/data_neuron_activity_heatmap_mfn_' + mfn_flag + '.pickle'
-    
-    ### get data
-    if not plot_only: # simulate respective network
-    
-        ## define parameters ranges tested:
-        means_tested = np.linspace(3,6,7, dtype=dtype)
-        variances_tested = np.linspace(2,9,8, dtype=dtype)
-        
-        ## run simulations
-        [trial_duration, num_values_per_trial, means_tested, variances_tested, mse_mean, mse_variance, 
-         activity_pe_neurons, activity_interneurons] = simulate_pe_uniform_para_sweep(mfn_flag, means_tested, variances_tested, 
-                                                                                      file_for_data, record_interneuron_activity = True,
-                                                                                      record_pe_activity = True)
-    
-    else: # load results from previous simulation
-
-        with open(file_for_data,'rb') as f:
-            [trial_duration, num_values_per_trial, means_tested, variances_tested, 
-             mse_mean, mse_variance, activity_pe_neurons, activity_interneurons] = pickle.load(f)
-        
-    ### plot 
-    end_of_initial_phase = np.int32(trial_duration * 0.5)
-    plot_neuron_activity(end_of_initial_phase, means_tested, variances_tested, activity_interneurons, None, id_fixed=3)
-    plot_neuron_activity(end_of_initial_phase, means_tested, variances_tested, None, activity_pe_neurons, id_fixed=3) 
-
-
 # %% Example: estimating mean and variance through PE neuron activity
 
-# Todo:
-    # choose parameters consistent with the subsequent plots
-    # beautify plotting functions
-    # rerun (you have to because something in the functions changed!)
-
 run_cell = False
-plot_only = True
+plot_only = False
 
 if run_cell:
     
@@ -93,14 +50,9 @@ if run_cell:
 
 
 # %% Systematically run network for different parameterisations of uniform distribution
-# record MSE between running estimate and m- or v-neuron output
[…]

### 2023-10-10 src/default_parameters.py

@@ -23,13 +23,8 @@ dtype = np.float32
 def default_para_mfn(mfn_flag, baseline_activity = dtype([0, 0, 0, 0, 4, 4, 4, 4]), one_column=True):
     
     ### time constants
-    #if one_column:
     tc_var_per_stim = dtype(5000)
     tc_var_pred = dtype(5000)
-    # else:
-    #     tc_var_per_stim = dtype(1000)
-    #     tc_var_pred = dtype(1000)
-    
     tau_pe = [dtype(60), dtype(2)]
     
     ### weights

### 2023-10-10 src/functions_simulate.py

@@ -105,6 +105,8 @@ def simulate_PE_circuit_P_fixed_S_constant(mfn_flag, initial_prediction, constan
     
     ### test all stimuli
     for i, stimulus in enumerate(constant_stimuli):
+        
+        print('Stimulus ', i+1, '/', len(constant_stimuli))
     
         repeats_per_value = trial_duration//num_values_per_trial
         stimuli = random_uniform_from_moments(stimulus, 0, num_values_per_trial)
@@ -358,10 +360,13 @@ def simulate_pe_uniform_para_sweep(mfn_flag, means_tested, variances_tested, fil
       tc_var_per_stim, tc_var_pred, tau_pe, fixed_input] = default_para_mfn(mfn_flag)
     
     ### initialise
-    mse_mean = np.zeros((len(means_tested), len(variances_tested), trial_duration), dtype=dtype)
-    mse_variance = np.zeros((len(means_tested), len(variances_tested), trial_duration), dtype=dtype)
+    dev_mean = np.zeros((len(means_tested), len(variances_tested)), dtype=dtype)
+    dev_variance = np.zeros((len(means_tested), len(variances_tested)), dtype=dtype)
     activity_interneurons = np.zeros((len(means_tested), len(variances_tested), trial_duration, 4), dtype=dtype)
     activity_pe_neurons = np.zeros((len(means_tested), len(variances_tested), trial_duration, 2), dtype=dtype)
+    
+    ### parameter to compute the steady state quantities
+    n_ss = 3 * trial_duration//4
     
     ### parameter sweep
     for i, mean_dist in enumerate(means_tested):
@@ -397,39 +402,39 @@ def simulate_pe_uniform_para_sweep(mfn_flag, means_tested, variances_tested, fil
                 
                 ## compute mean squared error between running average/variance and m or v neuron
                 running_average = np.cumsum(stimuli)/np.arange(1,len(stimuli)+1, dtype=dtype)
-                mse_mean[i, j, :] = (running_average - prediction)**2
+                dev_mean[i, j] = (np.mean(running_average[n_ss:]) - np.mean(prediction[n_ss:])) / np.mean(running_average[n_ss:])
                 
                 mean_running = np.cumsum(stimuli)/np.arange(1,len(stimuli)+1, dtype=dtype)
                 momentary_variance = (stimuli - mean_running)**2
                 running_variance = np.cumsum(momentary_variance)/np.arange(1,len(stimuli)+1, dtype=dtype)
-                mse_variance[i, j, :] = (running_variance - variance)**2
[…]

### 2023-10-10 src/plot_data.py

@@ -87,10 +87,48 @@ def lighten_color(color, amount=0.5):
 
 
 
-def plot_nPE_pPE_activity_compare(P, S, nPE, pPE, fs = 6, lw = 1, ms = 1, ax1 = None, ax2 = None):
+def plot_gain_factors(gains_nPE, gains_pPE, figsize=(10,1), fs = 6):
     
-    if ((ax1 is None) or (ax2 is None)):
-        _, (ax1, ax2) = plt.subplots(1,2)
+    fig = plt.figure(figsize=figsize)
+    
+    G = gridspec.GridSpec(1, 2, figure=fig, hspace=1, width_ratios=[19,1])
+    G1 = gridspec.GridSpecFromSubplotSpec(2, 1, subplot_spec=G[0,0], hspace=0.5)
+
+    ax_B = fig.add_subplot(G1[1,0])
+    ax_A = fig.add_subplot(G1[0,0], sharex=ax_B)
+    plt.setp(ax_A.get_xticklabels(), visible=False)
+    ax_C = fig.add_subplot(G[0,1])
+    
+    ax_B.tick_params(size=2.0) 
+    ax_B.tick_params(axis='both', labelsize=fs)
+    ax_A.tick_params(size=2.0) 
+    ax_A.tick_params(axis='both', labelsize=fs)
+    ax_C.tick_params(size=2.0) 
+    ax_C.tick_params(axis='both', labelsize=fs)
+    
+    max_gain = max(np.nanmax(gains_nPE), np.nanmax(gains_pPE))
+    min_gain = min(np.nanmin(gains_nPE), np.nanmin(gains_pPE))
+    
+    data = pd.DataFrame(gains_nPE)
+    g = sns.heatmap(data.T, xticklabels=20, yticklabels=0, vmin=min_gain, vmax=max_gain, ax = ax_A, 
+                    cbar_ax = ax_C, cbar_kws={'label': r'log(gain)'})
+    g.set_facecolor('#EBEBEB')
+    
+    data = pd.DataFrame(gains_pPE)
+    g = sns.heatmap(data.T, xticklabels=20, yticklabels=0, vmin=min_gain, vmax=max_gain, 
+                    ax = ax_B, cbar = False)
+    g.set_facecolor('#EBEBEB')
+    
+    ax_C.tick_params(labelsize=fs)
+    ax_C.tick_params(size=2.0)
+    ax_C.yaxis.label.set_size(fs)
+    
+    
[…]

### 2023-10-11 records/manuscript_2023.tex
Polish Code and Figure 2S3

@@
 \ref{fig:Fig_3_S2}, we used a linear activation function to investigate the impact of the input-output transfer function on the weighting of sensory inputs and predictions.
-The time constant $\tau_V = 5$ s.
+The time constant $\tau_V$ was 5 s in the mean-field model, and 2 s in the population model (Fig.
+\ref{fig:Fig_2_S2})., XXX and 0.5 s in 2S3 --> put everything for the not mean field model in the figure captions I would recommend, you can here say that it is only for the mean-field model ...
+XXX
 
@@
 %
-where $\lambda$ denotes the speed at which the perfect integrator evolves (lower PE circuit: $\lambda=3\cdot 10^{-3}$  in Fig.
-\ref{fig:Fig_2} and $\lambda=4.5\cdot 10^{-2}$ otherwise, higher PE circuit:  $\lambda = 7\cdot 10^{-4}$).
+where $\lambda$ denotes the speed at which the perfect integrator evolves (lower PE circuit: $\lambda=3\cdot 10^{-3}$ for the mean-field model in Fig.
+\ref{fig:Fig_2},  $\lambda=5\cdot 10^{-2}$ for the population model in Fig.
+\ref{fig:Fig_2_S2}, and $\lambda=4.5\cdot 10^{-2}$ otherwise; higher PE circuit:  $\lambda = 7\cdot 10^{-4}$).
+XXX in Fig.
+\ref{fig:Fig_2_S3} wPE2PE * 100 => I think it is best to put this is the captions, I mean for 2S2 and 2S3 ...
+for mean field network can be shown here and said "if not stated otherwise in the figures/captions"
 
@@
 \textbf{Figure 2 Supplementary Fig.
-2:} Number of time steps were $10^5$ (C-E) or $2\cdot 10^5$ (B) with $400$ (C-E) or $200$ (B) constant values drawn from a uniform distribution with mean of $5$ and variance of $4$.
+2:} Number of time steps were $10^5$ with $200$ constant values drawn from a uniform distribution with mean of $5$ and variance of $4$.
 Time step was $0.1$.
@@
 %
-\textbf{Figure 2 Supplementary Fig. 3:} Number of identical mean-field networks: $1000$.  Simulation time steps = $10^5$.\\
+\textbf{Figure 2 Supplementary Fig.
+3:} Number of time steps were $4000$ (XXX maybe mention here and other above stating "except otherwise stated" or so XXX).
+Number of identical mean-field networks: $1000$.
+Simulation time steps = $10^5$.\\
 %

### 2023-10-11 scripts/plot_fig_2_support_2.py

@@ -98,7 +98,7 @@ else:
     with open(file_data,'rb') as f:
         [num_seeds, SDs, M_steady_state, V_steady_state] = pickle.load(f)
 
-    plot_deviation_in_population_net(SDs, num_seeds, M_steady_state, V_steady_state, 'Uncorrelated deviations (SD)', ax=ax_C, ylim=None)
+    plot_deviation_in_population_net(SDs, num_seeds, M_steady_state, V_steady_state, r'Uncorrelated dev, SD of $\gamma$', ax=ax_C, ylim=None)
     
 
 # %% Correlated changes
@@ -114,7 +114,7 @@ else:
     with open(file_data,'rb') as f:
         [num_seeds, means, M_steady_state, V_steady_state] = pickle.load(f)
 
-    plot_deviation_in_population_net(means, num_seeds, M_steady_state, V_steady_state, 'Correlated deviations (mean)', ax=ax_D, 
+    plot_deviation_in_population_net(means, num_seeds, M_steady_state, V_steady_state, r'Correlated dev, mean of $\gamma$', ax=ax_D, 
                                      plt_ylabel=False, ylim=None)
     
 

### 2023-10-11 scripts/plot_fig_2_support_3.py

@@ -33,7 +33,7 @@ if not os.path.exists(figPath):
     
 # %% Define figure structure
 
-figsize=(18/inch,7/inch)
+figsize=(18/inch,6/inch)
 fig = plt.figure(figsize=figsize)
 
 G = gridspec.GridSpec(6, 3, figure=fig, width_ratios=[6,4,3], hspace=2.5, wspace=0.5)
@@ -51,10 +51,10 @@ ax_B1.text(-0.3, 1., 'B', transform=ax_B1.transAxes, fontsize=fs+1)
 plt.setp(ax_B1.get_xticklabels(), visible=False)
 ax_B2 = fig.add_subplot(G[3:,1])
 
-ax_C1 = fig.add_subplot(G[1:3,2])
+ax_C1 = fig.add_subplot(G[:3,2])
 ax_C1.text(-0.5, 1.1, 'C', transform=ax_C1.transAxes, fontsize=fs+1)
 plt.setp(ax_C1.get_xticklabels(), visible=False)
-ax_C2 = fig.add_subplot(G[3:5,2])
+ax_C2 = fig.add_subplot(G[3:,2])
 
 
 # %% Show examples with transitions
@@ -81,10 +81,10 @@ else:
     # plot data
     m_neuron_after = np.vstack((m_neuron_1, m_neuron_2))
     v_neuron_after = np.vstack((v_neuron_1, v_neuron_2))
-    plot_examples_spatial_M(num_time_steps, m_neuron_before, m_neuron_after, 5, np.array([5,8]), 
-                            labels=['Before ($\mu=5$, $\sigma^2=4$)',r'After ($\mu=5$, $\sigma^2=6$)','After ($\mu=8$, $\sigma^2=4$)'], show_xlabel=False, ax=ax_B1)
-    plot_examples_spatial_V(num_time_steps, v_neuron_before, v_neuron_after, 2**2, np.array([6,2**2]), 
-                            labels=['Before ($\mu=5$, $\sigma^2=4$)',r'After ($\mu=5$, $\sigma^2=6$)','After ($\mu=8$, $\sigma^2=4$)'], ax=ax_B2)
+    plot_examples_spatial_M(num_time_steps, m_neuron_before, m_neuron_after, 4, np.array([4,6]), 
+                            labels=['Stimulus 1',r'Stimulus 2 ($\sigma^2$ changed)','Stimulus 2 ($\mu$ changed)'], show_xlabel=False, ax=ax_B1)
[…]

### 2023-10-11 scripts/run_fig_2.py

@@ -68,7 +68,7 @@ if run_cell:
     
         ## define parameters ranges tested:
         means_tested = np.linspace(3,6,7, dtype=dtype)
-        variances_tested = np.linspace(3,6,7, dtype=dtype) #np.linspace(2,9,8, dtype=dtype)
+        variances_tested = np.linspace(3,6,7, dtype=dtype) 
         
         ## run simulations
         [trial_duration, num_values_per_trial, means_tested, variances_tested, dev_mean, dev_variance, 

### 2023-10-11 scripts/run_fig_2_support_2.py

@@ -10,30 +10,108 @@ Created on Mon Jan  9 09:04:01 2023
 
 import pickle
 import numpy as np
+import seaborn as sns
+import pandas as pd
 import matplotlib.pyplot as plt
 
 from src.default_parameters import Neurons, Activity_Zero, Network, Stimulation
 from src.functions_save import load_network_para
 from src.functions_networks import run_population_net
 
-from src.plot_data import plot_gain_factors
+from src.plot_data import plot_gain_factors, plot_deviation_in_population_net, plot_M_and_V_for_population_example
 
 dtype = np.float32
 
 
-# %% plot gain factors
+# %% Find nPE and pPE neurons and plot gain factors
 
-run_cell = True
+run_cell = False
+plot_only = True
 
 if run_cell:
     
-    import seaborn as sns
-    import pandas as pd
+    if not plot_only: # simulate respective network
     
-    with open('../results/data/population/data_population_network_para4corr_gain.pickle','rb') as f:
+        # get data
+        fln = '../results/data/population/data_population_network_parameters.pickle'
+        
+        with open(fln,'rb') as f:
+            [neurons_visual, inp_ext_soma, inp_ext_dend, weight_name, Dict_w, Dict_t] = pickle.load(f)
+        
+        # parametrisation neurons, network and rates
+        NeuPar = Neurons()
+        RatePar = Activity_Zero(NeuPar)
+        NetPar = Network(NeuPar, Dict_w, Dict_t, weight_name, neurons_visual)
+           
+        # take out connections from PE neurons to M or V
+        NetPar.wME *= 0
+        NetPar.wVarE *= 0
+        
+        
+        ### First S>P
+        StimPar = Stimulation(5, 1e-3, inp_ext_soma, inp_ext_dend, neurons_visual, num_values_per_trial = np.int32(20))
+        
+        RatePar.r_mem0 = np.array([0], dtype=dtype)
+        
+        # run network
+        run_population_net(NeuPar, NetPar, StimPar, RatePar, 0.1, folder='population', fln='test')
+        
+        
+        # load simulation data
+        PathData = '../results/data/population'
[…]

### 2023-10-11 scripts/run_fig_2_support_3.py

@@ -22,11 +22,9 @@ from src.plot_data import plot_examples_spatial_M, plot_examples_spatial_V, plot
 
 dtype = np.float32
 
-
-
 # %% Test several input statistcis
 
-run_cell = False
+run_cell = True
 plot_only = True
 
 if run_cell:
@@ -40,7 +38,7 @@ if run_cell:
     
     ### input statistic parameterisation
     means_tested = np.linspace(3,6,7)
-    stds_tested = np.linspace(np.sqrt(4),np.sqrt(8), 5)
+    stds_tested = np.sqrt(np.linspace(3,6,7, dtype=dtype))
     
     ### initialise
     deviation_mean = np.zeros((len(means_tested), len(stds_tested)))
@@ -54,8 +52,7 @@ if run_cell:
         
                 print('Mean no.: ' , (i+1) , '/' , len(means_tested), ', Std no.: ', (j+1), '/', len(stds_tested))
                 [mean_stimulus, spatial_std, spatial_noise, 
-                 num_time_steps, m_neuron, v_neuron, rates_final] = simulate_spatial_example(mfn_flag, dtype(mean), dtype(std), file_temporary, 
-                                                                                num_sub_nets = np.int32(1000), num_time_steps = np.int32(4000))
+                 num_time_steps, m_neuron, v_neuron, rates_final] = simulate_spatial_example(mfn_flag, dtype(mean), dtype(std), file_temporary)
                                                                                              
                 deviation_mean[i,j] = (m_neuron[-1] - mean) / mean
                 deviation_std[i,j] = (v_neuron[-1] - std**2) / std**2
@@ -70,9 +67,13 @@ if run_cell:
             [means_tested, stds_tested, deviation_mean, deviation_std] = pickle.load(f)
 
     ### plot data
-    
-    plot_deviation_spatial(deviation_mean, means_tested, stds_tested, vmin=0, vmax=5, fs=6, ax=None)
-    plot_deviation_spatial(deviation_std, means_tested, stds_tested, vmin=0, vmax=5, fs=6, show_mean=False, ax=None)
+    x_examples = [4, 4, 6]
+    y_examples = [4, 6, 4]
+    markers_examples = ['o', '^', '^']
+    plot_deviation_spatial(deviation_mean, means_tested, stds_tested, vmin=0.35, vmax=1.2, fs=6, ax=None,
+                           x_examples = x_examples, y_examples = y_examples, markers_examples = markers_examples)
[…]

### 2023-10-11 src/default_parameters.py

@@ -113,7 +113,7 @@ class Neurons(NamedTuple):
     NCells: list = np.array([140,20,20,20], dtype=np.int32)
     tau_inv_E: dtype = dtype(1.0/60.0)
     tau_inv_I: dtype = dtype(1.0/2.0)
-    tau_inv_var: dtype = dtype(1.0/1000) # might need to be adpated
+    tau_inv_var: dtype = dtype(1.0/2000) # might need to be adpated
 
 
 class Network:
@@ -159,9 +159,6 @@ class Network:
         if pPE_true is None:
             pPE_true = np.ones((1, NE))==1
         
-        speed = dtype(0.2)
-        scaling_factor = dtype(1)
-        
         if p_conn is not None:
             sparisty_bool = np.random.choice([0,1], p=[1-p_conn,p_conn], size=NE) == 1
             nPE_true[~sparisty_bool] = False
@@ -173,8 +170,9 @@ class Network:
         gain[pPE_true] *= 1/sum(pPE_true)
         gain[~(nPE_true+pPE_true)] = 0
         
+        speed = dtype(0.05)
         self.wME = speed * gain
-        self.wVarE = abs(gain * scaling_factor)
+        self.wVarE = abs(gain)
                                             
  
                 

### 2023-10-11 src/functions_simulate.py

@@ -124,8 +124,8 @@ def simulate_PE_circuit_P_fixed_S_constant(mfn_flag, initial_prediction, constan
     return nPE, pPE
 
 
-def simulate_spatial_example(mfn_flag, mean_stimulus, spatial_std, file_for_data, num_sub_nets = np.int32(100), seed = 186, 
-                             num_time_steps = np.int32(1000), dist_type = 'uniform', pa=0.8, M_init = None, V_init = None, rates_init = None):
+def simulate_spatial_example(mfn_flag, mean_stimulus, spatial_std, file_for_data, num_sub_nets = np.int32(1000), seed = 186, 
+                             num_time_steps = np.int32(4000), dist_type = 'uniform', pa=0.8, M_init = None, V_init = None, rates_init = None):
     
     ### load default parameters
     VS, VV = int(mfn_flag[0]), int(mfn_flag[1])
@@ -160,12 +160,12 @@ def simulate_spatial_example(mfn_flag, mean_stimulus, spatial_std, file_for_data
     spatial_noise = np.repeat(spatial_noise, 8)
 
     ### connectivity matrices
-    W_PE_to_V = np.tile(np.array([1.015,1.023,0,0,0,0,0,0]), (1,num_sub_nets)) * scaling
-    W_PE_to_P = 100 * np.tile(w_PE_to_P, (1,num_sub_nets)) / num_sub_nets
+    W_PE_to_V = np.tile(np.array([w_PE_to_V[0],w_PE_to_V[1],0,0,0,0,0,0]), (1,num_sub_nets)) * scaling
+    W_PE_to_P = 100 * np.tile(w_PE_to_P, (1,num_sub_nets)) / num_sub_nets # make faster
     W_P_to_PE = np.tile(w_P_to_PE, (num_sub_nets,1))
     W_PE_to_PE = np.kron(np.eye(num_sub_nets,dtype=dtype),w_PE_to_PE)
     
-    tc_var_per_stim /= 10 # !!!!!!!!!!!!!!!!!!!!!!!!!
+    tc_var_per_stim /= 10 # make faster
     
     ### run model
     if rates_init is None:

### 2023-10-20 records/manuscript_2023.tex
improve code and polish Fig 3 and 3S1

@@
 In a trail, $10$ constant values were drawn from a normal distribution (each $500$ time steps long).
-The variance of this distribution was $0$ (C, F)  and $5.3$ XXX (D).
+The variance of this distribution was $0$ (C, F)  and $5$ (D).
 In each trial, the stimulus mean was drawn from an uniform distribution $U(a, b)$ (C: $U(1,9)$, D: $U(5,5)$).
-In F, $U$ was parameterized such that the trail variability was $3$.\\
+In F, $U(a,b)$ was parameterized such that the trail variability was $3$.\\
 %
 \textbf{Figure 3 Supplementary Fig.
-1:}  $120$ trials, each simulated for $5000$ time steps.
+1:}  $100$ trials, each simulated for $5000$ time steps.
 In a trail, $10$ constant values were drawn from a normal distribution $N(\mu_\mathrm{in}, \sigma_\mathrm{in})$, each $500$ time steps long.
 In each trial, the stimulus mean was drawn from an uniform distribution $U(a, b)$.
-Switch of input statistics occurs after 60 trails.
+Switch of input statistics occurs after 50 trails.
 State 1: stimulus $\in N(\mu_\mathrm{in}, 0)$ with $\mu_\mathrm{in} \in U(5,5)$, State 2:  stimulus $\in N(\mu_\mathrm{in}, 3)$ with $\mu_\mathrm{in} \in U(5,5)$, State 3: stimulus $\in N(\mu_\mathrm{in}, 3)$ with $\mu_\mathrm{in} \in U(0,10)$, State 4: stimulus $\in N(\mu_\mathrm{in}, 0)$ with $\mu_\mathrm{in} \in U(0,10)$.\\

### 2023-10-20 scripts/plot_fig_3.py

@@ -59,18 +59,13 @@ ax_B.text(-0.13, 1.25, 'B', transform=ax_B.transAxes, fontsize=fs+1)
 
 ax_B1 = fig.add_subplot(G12[0,0])
 ax_B1.axis('off')
-#ax_B1.set_title('Neuron activity increases \nwith trial variability', fontsize=fs, pad=10)
-# ax_B2 = fig.add_subplot(G12[0,1])
-# ax_B2.set_title('Neuron activity increases \nwith stimulus variability', fontsize=fs, pad=10)
-# ax_B3 = fig.add_subplot(G12[0,2])
-# plt.setp(ax_B2.get_yticklabels(), visible=False)
 
 ax_C = fig.add_subplot(G21[0,:])
 ax_C.axis('off')
 ax_C.text(-0.15, 1.25, 'C', transform=ax_C.transAxes, fontsize=fs+1)
 
 ax_D2 = fig.add_subplot(G21[0,1])
-ax_D2.set_title('Example in which the network relies more strongly on prediction', fontsize=fs, pad=10)
+ax_D2.set_title('Example in which the network relies more strongly on sensory inputs', fontsize=fs, pad=10)
 ax_D3 = fig.add_subplot(G21[0,2])
 ax_D4 = fig.add_subplot(G2[0,1])
 ax_D4.set_title('Sensory weight for different \ntrial and stimulus variabilities', fontsize=fs, pad=10)
@@ -81,39 +76,14 @@ ax_D.axis('off')
 ax_D.text(-0.15, 1.25, 'D', transform=ax_D.transAxes, fontsize=fs+1)
 
 ax_E2 = fig.add_subplot(G31[0,1])
-ax_E2.set_title('Example in which the network relies more strongly on sensory inputs', fontsize=fs, pad=10)
+ax_E2.set_title('Example in which the network relies more strongly on prediction', fontsize=fs, pad=10)
 ax_E3 = fig.add_subplot(G31[0,2])
 ax_E4 = fig.add_subplot(G3[0,1])
 ax_E4.set_title('Sensory weight decreases \nwith shorter trials', fontsize=fs, pad=10)
 ax_E4.text(-0.35, 1.25, 'F', transform=ax_E4.transAxes, fontsize=fs+1)
 
 
-# %% Neuron activities
-
-# data files
-# file_for_data_within = '../results/data/weighting/data_activity_neurons_exploration_within_10.pickle'
[…]

### 2023-10-20 scripts/run_fig_2_support_2.py

@@ -131,8 +131,10 @@ if run_cell:
 
 # %% Sparsity 
 
+# continue here
+
 run_cell = True
-plot_only = False
+plot_only = True
 
 if run_cell:
     
@@ -149,7 +151,7 @@ if run_cell:
             [_, _, gain_factors_nPE_5, gain_factors_pPE_5, nPE_true, pPE_true] = pickle.load(f)
         
         # conditions to be tested
-        p_conns = np.linspace(0.6, 1, 5)
+        p_conns = np.linspace(0.7, 1, 7)
         num_seeds = 10
         
         # initialise arrays
@@ -160,7 +162,7 @@ if run_cell:
             for seed in range(num_seeds):
                 
                 print('p_conn = ', p_conn, ', seed = ', seed)
-            
+                
                 # parametrisation neurons, network and rates
                 NeuPar = Neurons()
                 RatePar = Activity_Zero(NeuPar)
@@ -186,6 +188,9 @@ if run_cell:
                 avg_start = len(r_mem)//2
                 M_steady_state[cond_num,seed] = np.mean(r_mem[-avg_start:])
                 V_steady_state[cond_num,seed] = np.mean(r_var[-avg_start:])
+                
+                print(np.mean(r_mem[-avg_start:]))
+                print(np.mean(r_var[-avg_start:]))
         
         # save data
         with open(file_data,'wb') as f:

### 2023-10-20 src/functions_simulate.py

@@ -34,9 +34,11 @@ def stimuli_moments_from_uniform(n_trials, num_values_per_trial, min_mean, max_m
         inputs_per_stimulus = np.random.normal(mean_stimuli[id_stim], sd_stimuli[id_stim], size=num_values_per_trial)
         stimuli = np.concatenate((stimuli, inputs_per_stimulus))
      
-    results = (dtype(stimuli),)
     if out_trial_means:
+        results = (dtype(stimuli),)
         results += (mean_stimuli,)
+    else:
+        results = dtype(stimuli)
         
     return results
 
@@ -634,7 +636,7 @@ def simulate_activity_neurons(mfn_flag, variability_within, variability_across, 
         
 def simulate_dynamic_weighting_eg(mfn_flag, min_mean_before, max_mean_before, m_sd_before, n_sd_before, 
                                   min_mean_after, max_mean_after, m_sd_after, n_sd_after, seed = np.int32(186),
-                                  n_trials = np.int32(120), trial_duration = np.int32(5000), num_values_per_trial = np.int32(10),
+                                  n_trials = np.int32(100), trial_duration = np.int32(5000), num_values_per_trial = np.int32(10),
                                   file_for_data = None):
     
     ### load default parameters

### 2023-10-20 src/plot_data.py

@@ -2235,7 +2235,7 @@ def plot_weight_over_trial(weight_ctrl, weight_mod, n_trials, fs=6, lw=1, leg_te
 
 def plot_transitions_examples(n_trials, trial_duration, stimuli, alpha, beta, weighted_output, 
                               time_plot = 0, ylim=None, xlim=None, plot_ylable=True, lw=1, 
-                              figsize=(3.5,5), plot_only_weights=False, fs=6, transition_point=60, ax2=None):
+                              figsize=(3.5,5), plot_only_weights=False, fs=6, transition_point=50, ax2=None):
     
     if ax2 is None:
         f1, ax2 = plt.subplots(1, 1, figsize=figsize, tight_layout=True)
@@ -2283,7 +2283,7 @@ def plot_transitions_examples(n_trials, trial_duration, stimuli, alpha, beta, we
     #     ax2.set_ylabel('Weights', color='white', fontsize=fs)
     #     ax2.tick_params(axis='y', colors='white')
     ax2.axvline(transition_point, color='k', ls='--', lw=lw, zorder=10)
-    ax2.set_xlabel('Time (#trials)', fontsize=fs)
+    ax2.set_xlabel('Time (number of trials)', fontsize=fs)
     ax2.set_xlim([time_plot * time[-1],time[-1]])
     ax2.set_ylim([0,1.05])
     if xlim is not None:
@@ -2344,15 +2344,15 @@ def plot_weighting_limit_case_example(n_trials, trial_duration, stimuli, predict
     ax1.plot(time[time > time_plot * time[-1]], weighted_output[time > time_plot * time[-1]], 
              color=color_weighted_output, lw=lw)#, label='weighted \noutput')  
     ax1.set_ylabel('Weighted output (1/s)', fontsize=fs)
-    ax1.set_xlabel('Time (#trials)', fontsize=fs)
+    ax1.set_xlabel('Time (number of trials)', fontsize=fs)
     
     if plot_legend:
         #ax1.legend(loc=0, frameon=False, fontsize=fs, handlelength=1)
         #ax1.text(time_plot * time[-1] + 1, 0.85 * np.max(weighted_output[time > time_plot * time[-1]]), 
         #        'weighted \noutput', color=color_weighted_output, fontsize=fs)
         ax1.plot(np.nan, np.nan, color=color_stimuli_background, ms=6, marker='_', 
[…]

### 2023-10-27 records/manuscript_2023.tex
Improve Figs. 4-5, polish code

@@
 
-% evidence for the model (PE neurons, connectivity PE --> M, predictions or P neurons exist, P neurons may hold the mean, time constants chnage along hierarchy)
+% evidence for the model (PE neurons, connectivity PE --> M, predictions or P neurons exist, P neurons may hold the mean, time constants change along hierarchy)
 
@@
 
+% selectivity? One-dimensional system (wahrscheinlich nicht)
+
 % link to Kalman filter and maybe Bayes surprise factor
+
+----
+% Katha feedback: predictions, postdictions, experiments vorschlagen, limitations, grosserer kontext, abgrenzung zu reward prediction error
 
@@
 Finally, to ensure reproducibility, we give all simulation parameters used for the results shown in the figures.
-
-XXX Please note that xxx model parameters chosen do not change overall results are not particular to get the results xxx.
-Important factors are: 1) long enough to get to steady state (as PE circuits used were learned for steady state, 2) lower M updates faster than higher M, 3) update speed for M and V (see lambda and theta etc.) must be chosen in accordance with trial duration stimulus duration.
-XXX This could also go to the section Simulations xxx
 
@@
 
-In the following, we give all figure-specific parameters not directly visible or mentioned in the Figures and captions.
-Please note that the qualitative results are robust to the choice of the simulation parameters and are here stated merely to ensure the reproducibility of all figures.
+The qualitative results were fairly robust to the choice of the simulation parameters and are here stated merely to ensure the reproducibility of all figures.
+However, we note that we made use of PE circuits that had been trained on steady state inputs \ref{XXX}.
+Hence, we must simulate the network long enough to ensure that the PE neurons reach their steady state.
+Moreover, the lower-level M neuron must evolve faster than the higher-level M neuron as indicated in Fig.
+\ref{XXX}.
+Finally, the time constant of the V neurons must be of the same magnitude as the trial duration.
+
+In the following, we give all figure-specific parameters not directly visible or mentioned in the figures and captions.
 Furthermore, to increase readability, we do not include units for the parameters.
-All units can be deduced from the equations above.\newline\\
-%
-\textbf{Figure 1:} Simulation time steps = $10^5$, constant prediction (fixed) = $5$, input mean $\in [0,10]$, input standard deviation $0$.\\
-%
-\textbf{Figure 2:} Simulation time steps = $10^5$, $200$ constant values (each $500$ time steps long) drawn from a uniform distribution, B and C: input standard deviation fixed at $4.5$ when mean is varied, input mean fixed at $4.5$ when input variance is varied.\\
+All units can be deduced from the equations above.
+We simulated the network in Figures \ref{fig:Fig_1} and \ref{fig:Fig_2} for $10^5$ simulation time steps.
+In Figure \ref{fig:Fig_2}, we presented $200$ constant values, each $500$ time steps long.
+In Figure \ref{fig:Fig_3} (including supporting figures) we simulated $100$ trials, while in Figures \ref{fig:Fig_4} to \ref{fig:Fig_5} (including supporting figures), we simulated $200$ trials, each $5000$ time steps long.
+In a trail, $10$ constant values were drawn from a normal distribution $N(\mu_\mathrm{in}, \sigma^2_\mathrm{in})$, each $500$ time steps long.
+The stimulus mean was drawn from an uniform distribution, $U(a, b)$, with mean $\mu_\mathrm{in}$ and variance $\sigma^2_\mathrm{trial}$.\newline\\
+%
+\textbf{Figure 1:} Constant prediction (fixed) = $5$, input mean $\in [0,10]$, input standard deviation $0$.\\
+%
+\textbf{Figure 2:} Inputs drawn from a uniform distribution. B and C: input standard deviation fixed at $4.5$ when mean is varied, input mean fixed at $4.5$ when input variance is varied.\\
+%
+\textbf{Figure 2 Supplementary Fig. 1:} Inputs drawn from different distributions with mean of $5$ and variance of $4$. Number of repetitions with different seeds: 20.\\
 %
 \textbf{Figure 2 Supplementary Fig.
-1:} Simulation time steps = $10^5$, $200$ constant values (each $500$ time steps long) drawn from a uniform distribution with mean of $5$ and variance of $4$.
-Number of repetitions with different seeds: 20.\\
-%
-\textbf{Figure 2 Supplementary Fig.
-2:} Number of time steps were $10^5$ with $200$ constant values drawn from a uniform distribution with mean of $5$ and variance of $4$.
+2:} Inputs drawn from a uniform distribution with mean of $5$ and variance of $4$.
 Time step was $0.1$.
 The connections from the PE neurons to the M or V neuron were altered by a factor $\gamma$ drawn from a normal distribution.
-If not stated otherwise (in the Figure), the mean of this normal distribution was $1$ and the variance $0$, while the connection probability was $1$.
+If not stated otherwise, the mean of this normal distribution was $1$ and the variance $0$, while the connection probability was $1$.
 Number of repetitions with different seeds: 10.\\
 %
-\textbf{Figure 2 Supplementary Fig.
-3:} Number of time steps were $4000$ (XXX maybe mention here and other above stating "except otherwise stated" or so XXX).
-Number of identical mean-field networks: $1000$.
-Simulation time steps = $10^5$.\\
-%
-\textbf{Figure 3:} $100$ trials, each simulated for $5000$ time steps.
-In a trail, $10$ constant values were drawn from a normal distribution (each $500$ time steps long).
-The variance of this distribution was $0$ (C, F)  and $5$ (D).
-In each trial, the stimulus mean was drawn from an uniform distribution $U(a, b)$ (C: $U(1,9)$, D: $U(5,5)$).
-In F, $U(a,b)$ was parameterized such that the trail variability was $3$.\\
+\textbf{Figure 2 Supplementary Fig. 3:} Number of time steps were $4000$. Number of identical mean-field networks: $1000$. \\
+%
+\textbf{Figure 3:} C: $\sigma^2_\mathrm{in} = 0$ and $U(1,9)$, D:  $\sigma^2_\mathrm{in} = 5$ and $U(5,5)$, F: $\sigma^2_\mathrm{in} = 0$ and $U(a,b)$ was parameterized such that the trail variability was $3$.\\
 %
 \textbf{Figure 3 Supplementary Fig.
-1:}  $100$ trials, each simulated for $5000$ time steps.
-In a trail, $10$ constant values were drawn from a normal distribution $N(\mu_\mathrm{in}, \sigma_\mathrm{in})$, each $500$ time steps long.
-In each trial, the stimulus mean was drawn from an uniform distribution $U(a, b)$.
-Switch of input statistics occurs after 50 trails.
+1:}  Switch of input statistics occurs after 50 trails.
 State 1: stimulus $\in N(\mu_\mathrm{in}, 0)$ with $\mu_\mathrm{in} \in U(5,5)$, State 2:  stimulus $\in N(\mu_\mathrm{in}, 3)$ with $\mu_\mathrm{in} \in U(5,5)$, State 3: stimulus $\in N(\mu_\mathrm{in}, 3)$ with $\mu_\mathrm{in} \in U(0,10)$, State 4: stimulus $\in N(\mu_\mathrm{in}, 0)$ with $\mu_\mathrm{in} \in U(0,10)$.\\
@@
 \textbf{Figure 3 Supplementary Fig.
-2:} $200$ trials, each simulated for $5000$ time steps.
-In a trail, $10$ constant values were drawn from a normal distribution with variance $\sigma^2_\mathrm{in}$, each $500$ time steps long.
-In each trial, the stimulus mean was drawn from an uniform distribution with mean of $5$ and a variance of $\sigma^2_\mathrm{trial}$.
-$\sigma^2_\mathrm{trial} \in \lbrace0, 0.75, 1.5, 2.25, 3\rbrace$, $\sigma^2_\mathrm{in} \in \lbrace 3, 2.25, 1.5,0.75, 0\rbrace$.
-A: scaling factors of $w_\mathrm{M\leftarrow PE}$ were 0.3 and 6.7 XXX.\\
-%
-\textbf{Figure 4:} $100-200$ trials, each simulated for $5000$ time steps.
-In a trail, $10$ constant values were drawn from a normal distribution with variance $\sigma^2_\mathrm{in}$, each $500$ time steps long.
-In each trial, the stimulus mean was drawn from an uniform distribution with mean of $5$ and a variance of $\sigma^2_\mathrm{trial}$.
+2:} $\mu_\mathrm{in} = 5$ and $\sigma^2_\mathrm{trial} \in \lbrace0, 0.75, 1.5, 2.25, 3\rbrace$, $\sigma^2_\mathrm{in} \in \lbrace 3, 2.25, 1.5,0.75, 0\rbrace$.
+A: scaling factors of $w_\mathrm{M\leftarrow PE}$ were 0.3 and 7.\\
+%
+\textbf{Figure 4:} $\mu_\mathrm{in} = 5$.
 A, top: $\sigma^2_\mathrm{in} = 0$ and $\sigma^2_\mathrm{trial} = 1$.
@@
 C: $\sigma^2_\mathrm{in} = 1$ and $\sigma^2_\mathrm{trial} = 1$.
-D: $\sigma^2_\mathrm{in} = 5.3 XXX$ and $\sigma^2_\mathrm{trial} = 0$.
-Additional input (perturbation) was either fixed at $1$ (A, D) or systematically varied between $-1$ and $1$ (C).
+D: $\sigma^2_\mathrm{in} = 5$ and $\sigma^2_\mathrm{trial} = 0$.
+Additional input (perturbation) was either fixed at $0.5$ (A, D) or systematically varied between $-1$ and $1$ (C), and was \textit{on} for the last 50\% of the trials.
 To estimate changes in baseline and gain of nPE and pPE neurons, we fitted a linear function to the PE neuron activity for the input range $[0, 2.5]$.\\
@@
 %
-\textbf{Figure 5:} $200$ trials, each simulated for $5000$ time steps (except in panel E).
-In a trail, $10$ constant values were drawn from a normal distribution (each $500$ time steps long) with standard deviation $\sigma_\mathrm{in}$.
-In each trial, the stimulus mean $\mu_\mathrm{in}$ was drawn from an uniform distribution $U(a, b)$.
-A: $\sigma_\mathrm{in} \in \lbrace1,7\rbrace$, $\mu_\mathrm{in} \in U(15, 25)$.
+\textbf{Figure 5:} A: $\sigma_\mathrm{in} \in \lbrace1,7\rbrace$, $\mu_\mathrm{in} \in U(15, 25)$.
 B: $\sigma_\mathrm{in} \in [0,8]$, $\mu_\mathrm{in} \in U(15, 25)$, and $\sigma_\mathrm{in} = 5$, $\mu_\mathrm{in} \in U(15, b)$ with $b\in [20, 48]$.
@@
 \textbf{Figure 5 Supplementary Fig.
-1:} $200$ trials, each simulated for $5000$ time steps.
-In a trail, $10$ constant values were drawn from a normal distribution (each $500$ time steps long) with standard deviation $\sigma_\mathrm{in}$.
-In each trial, the stimulus mean $\mu_\mathrm{in}$ was drawn from an uniform distribution $U(a,b)$.
-We used two different uniform distributions $U(15, 25)$ and  $U(25, 35)$, and introduced scalar variability so that $\sigma_\mathrm{in}$ is a linear function of $\mu_\mathrm{in}$.
-Specifically, we chose $\sigma_\mathrm{in} = \mu_\mathrm{in} -14$.
+1:} We used two different uniform distributions $U(15, 25)$ and  $U(25, 35)$, and introduced scalar variability so that $\sigma_\mathrm{in}$ is a linear function of $\mu_\mathrm{in}$.
+Specifically, we chose $\sigma_\mathrm{in} = \left[\mu_\mathrm{in} -14\right]_+$.
 
@@
 
-\subsection{Comparison to Kalman filter and Bayes Factor surprise}
-%
-Kalman filter. Initialisation
-%
-\begin{align*}
-x_{0|init} &= 0 \\
-P_{0|init} &= \sigma^2\ I
-\end{align*}
-%
-with x being the system state (in my terms the prediction), P is the covariance matrix of the errors of x (in my terms the var of the predictions) and I is the identity matrix.
-%
-Then the "correction" is given by
-%
-\begin{align*}
-K_k &= P_{k|k-1}\ H_k^T\ \left( H_k\ P_{k|k-1}\ H_k^T + R_k \right)^{-1} \\
-x_k &= x_{k|k-1} + K_k\ \left( z_k - H_k\ x_{k|k-1}\right) \\
-P_k &= \left( I - K_k\ H_k\right)\ P_{k|k-1}
-\end{align*}
-%
-with K the kalman gain matrix, H the observation matrix ($z_k = H_k\ x_k + noise$), R the covaraince of the measurement noise and z a new observation.
+%\subsection{Comparison to Kalman filter and Bayes Factor surprise}
+%%
+%Kalman filter. Initialisation
+%%
+%\begin{align*}
+%x_{0|init} &= 0 \\
+%P_{0|init} &= \sigma^2\ I
+%\end{align*}
+%%
+%with x being the system state (in my terms the prediction), P is the covariance matrix of the errors of x (in my terms the var of the predictions) and I is the identity matrix.
+%%
+%Then the "correction" is given by
+%%
+%\begin{align*}
+%K_k &= P_{k|k-1}\ H_k^T\ \left( H_k\ P_{k|k-1}\ H_k^T + R_k \right)^{-1} \\
+%x_k &= x_{k|k-1} + K_k\ \left( z_k - H_k\ x_{k|k-1}\right) \\
+%P_k &= \left( I - K_k\ H_k\right)\ P_{k|k-1}
+%\end{align*}
+%%
+%with K the kalman gain matrix, H the observation matrix ($z_k = H_k\ x_k + noise$), R the covaraince of the measurement noise and z a new observation.
 The last part of the Kalman filter is the "prediction":
-%
-\begin{align*}
-x_{k|k-1} &= F_{k-1}\ x_{k-1} + B_{k-1}\ u_{k-1} \\
-P_{k|k-1} &= F_{k-1}\ P_{k-1}\ F_{k-1}^T + Q_{k-1}
-\end{align*}
-%
-with F the transition matrix ($x_{k|k-1} = F_{k-1}\ x_{k-1}$, u a deterministic perturbation, B the dynamics of the deterministic perturbation. In our terms
-%
-\begin{align*}
-\alpha = K_k = \frac{P_{k|k-1}}{R_k + P_{k|k-1}}
-\end{align*}
-%
-$P_{k|k-1}$, is however $\sigma_P^2$ in my implementation and $R_k$ is fixed variance of inputs $\sigma_S^2$.
+%%
+%\begin{align*}
+%x_{k|k-1} &= F_{k-1}\ x_{k-1} + B_{k-1}\ u_{k-1} \\
+%P_{k|k-1} &= F_{k-1}\ P_{k-1}\ F_{k-1}^T + Q_{k-1}
+%\end{align*}
+%%
+%with F the transition matrix ($x_{k|k-1} = F_{k-1}\ x_{k-1}$, u a deterministic perturbation, B the dynamics of the deterministic perturbation. In our terms
+%%
+%\begin{align*}
+%\alpha = K_k = \frac{P_{k|k-1}}{R_k + P_{k|k-1}}
+%\end{align*}
+%%
+%$P_{k|k-1}$, is however $\sigma_P^2$ in my implementation and $R_k$ is fixed variance of inputs $\sigma_S^2$.
 Hence, my implementation represents (?) the Kalman filter.
@@
 .
-
-XXX Comparison to Bayes Factor surprise
+%
+%XXX Comparison to Bayes Factor surprise
 

### 2023-10-27 scripts/plot_fig_3_support_1.py

@@ -13,9 +13,6 @@ import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import os.path
-import seaborn as sns
-
-from matplotlib.collections import LineCollection
 
 from src.plot_data import plot_fraction_sensory_heatmap, plot_transitions_examples
 

### 2023-10-27 scripts/plot_fig_3_support_2.py

@@ -13,9 +13,6 @@ import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import os.path
-import seaborn as sns
-
-from matplotlib.collections import LineCollection
 
 from src.plot_data import plot_impact_para
 
@@ -49,11 +46,6 @@ ax_B = fig.add_subplot(G[0,1])
 ax_B.text(-0.2, 1.35, 'B', transform=ax_B.transAxes, fontsize=fs+1)
 plt.setp(ax_B.get_xticklabels(), visible=False)
 
-# ax_B1 = fig.add_subplot(G2[0,0])
-# ax_B1.set_title(r'1 $\longrightarrow$ 2', fontsize=fs, pad=10)
-# ax_B2 = fig.add_subplot(G2[0,1])
-# plt.setp(ax_B2.get_yticklabels(), visible=False)
-# ax_B2.set_title(r'2 $\longrightarrow$ 3', fontsize=fs, pad=10)
 
 # %% Load control case and define variabilties tested
 

### 2023-10-27 scripts/plot_fig_4_support_2.py

@@ -6,30 +6,12 @@ Created on Mon Jan  9 09:04:01 2023
 @author: loreen.hertaeg
 """
 
-# %% Import
-
-import numpy as np
-import pickle
-
-from src.functions_simulate import simulate_neuromod
-from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-
 # %% import
 
-import pickle
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import os.path
-import seaborn as sns
-
-from matplotlib.collections import LineCollection
 
 from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_legend_illustrations
 

### 2023-10-27 scripts/plot_fig_4_support_3.py

@@ -6,30 +6,11 @@ Created on Mon Jan  9 09:04:01 2023
 @author: loreen.hertaeg
 """
 
-# %% Import
-
-import numpy as np
-import pickle
-
-from src.functions_simulate import simulate_neuromod
-from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-
 # %% import
 
-import pickle
-import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import os.path
-import seaborn as sns
-
-from matplotlib.collections import LineCollection
 
 from src.plot_data import plot_influence_interneurons_gain_baseline
 
@@ -50,7 +31,7 @@ if not os.path.exists(figPath):
 
 # %% Define figure structure
 
-figsize=(5/inch,4/inch)
+figsize=(6/inch,5/inch)
 fig = plt.figure(figsize=figsize, tight_layout=True)
 
 G = gridspec.GridSpec(1, 1, figure=fig)

### 2023-10-27 scripts/plot_fig_5.py

@@ -13,9 +13,8 @@ import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import os.path
-import seaborn as sns
 
-from src.plot_data import plot_example_contraction_bias, plot_slope_trail_duration, plot_slope_variability, plot_contraction_bias_illustration
+from src.plot_data import plot_example_contraction_bias, plot_slope_trail_duration, plot_slope_variability
 from src.plot_data import plot_illustration_input_cond, plot_illustration_bias_results, plot_illustration_trial_duration
 
 # %% Universal parameters
@@ -35,54 +34,62 @@ if not os.path.exists(figPath):
     
 # %% Define figure structure
 
-figsize=(8/inch,18/inch)
+figsize=(15/inch,10/inch)
 fig = plt.figure(figsize=figsize)
 
-G = gridspec.GridSpec(2, 1, figure=fig, height_ratios=[1,6], hspace=0.3) #, hspace = 1, wspace=2)
-G_lower = gridspec.GridSpecFromSubplotSpec(4, 1, subplot_spec=G[1,0], hspace=1.5)
+G = gridspec.GridSpec(5, 2, figure=fig, wspace=0.5, height_ratios=[1,1,1,1.3,1.3]) 
 
-G1 = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=G[0,0], width_ratios=[1,2], wspace=0.7)#,  wspace=0.8) , width_ratios=[1,1,1,1.2]
-G2 = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=G_lower[0,0], width_ratios=[2,1.5], wspace=0.3)#,  wspace=0.8)
-G3 = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=G_lower[1,0], width_ratios=[2,1.5], wspace=0.3)
-G4 = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=G_lower[2,0], width_ratios=[2,1.5], wspace=0.3)
-G5 = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=G_lower[3,0], width_ratios=[2,1.5], wspace=0.3)
+ax_A = fig.add_subplot(G[:2,0])
+ax_A.text(-0.25, 1.15,'A', transform=ax_A.transAxes, fontsize=fs+1)
 
-ax_A1 = fig.add_subplot(G1[0,0])
[…]

### 2023-10-27 scripts/plot_fig_5_support_1.py

@@ -6,20 +6,6 @@ Created on Mon Jan  9 09:04:01 2023
 @author: loreen.hertaeg
 """
 
-# %% Import
-
-import numpy as np
-import pickle
-
-from src.functions_simulate import simulate_neuromod
-from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-
 # %% import
 
 import pickle
@@ -27,17 +13,14 @@ import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import os.path
-import seaborn as sns
 
-from matplotlib.collections import LineCollection
-
-from src.plot_data import plot_example_contraction_bias
+from src.plot_data import plot_example_contraction_bias, plot_std_vs_mean
 
 # %% Universal parameters
 
 fs = 6
 inch = 2.54
-
+dtype = np.float32
 
 # %% Define files and paths
 
@@ -50,11 +33,13 @@ if not os.path.exists(figPath):
 
 # %% Define figure structure
 
-figsize=(3/inch,3/inch)
+figsize=(6/inch,2.5/inch)
 fig = plt.figure(figsize=figsize)
 
-G = gridspec.GridSpec(1, 1, figure=fig)#, hspace=0.6)
+G = gridspec.GridSpec(1, 2, figure=fig, wspace=1, width_ratios=[1,2])#, hspace=0.6)
 ax_A = fig.add_subplot(G[0,0])
+ax_B = fig.add_subplot(G[0,1])
+
 
 # %% Scalar variability introduces stronger contraction bias for higher values
 
@@ -64,7 +49,7 @@ min_mean_2, max_mean_2 = dtype(25), dtype(35)
 m_std, n_std = dtype(1), dtype(-14)
     
 file_for_data_1 = '../results/data/behavior/data_contraction_bias_trial_mean_range_' + str(max_mean_1 - min_mean_1) + '_max_mean_' + str(max_mean_1) + '.pickle'
-file_for_data_2 = '../results/data/behavior/data_contraction_bias_trial_imean_range_' + str(max_mean_2 - min_mean_2) + '_max_mean_' + str(max_mean_2) + '.pickle'
+file_for_data_2 = '../results/data/behavior/data_contraction_bias_trial_mean_range_' + str(max_mean_2 - min_mean_2) + '_max_mean_' + str(max_mean_2) + '.pickle'
 
[…]

### 2023-10-27 scripts/run_fig_2.py

@@ -12,8 +12,7 @@ import numpy as np
 import pickle
 
 from src.functions_simulate import simulate_example_pe_circuit, simulate_pe_uniform_para_sweep
-from src.plot_data import plot_example_mean, plot_example_variance, plot_mse_heatmap, plot_interneuron_activity_heatmap
-from src.plot_data import plot_neuron_activity
+from src.plot_data import plot_example_mean, plot_example_variance, plot_mse_heatmap, plot_neuron_activity
 
 import warnings
 warnings.filterwarnings("ignore")

### 2023-10-27 scripts/run_fig_2_support_1.py

@@ -87,8 +87,7 @@ if run_cell:
     sem_variance = np.std(dev_variance * 100,2)/np.sqrt(num_reps)
     
     plot_mse_test_distributions(np.mean(dev_mean * 100,2), SEM = sem_mean, title = 'Mean of stimuli', fs=8, 
-                                mean=mean, std=std, dist_types=dist_types, inset_steady_state=True) # x_lim=[0,50000]
+                                mean=mean, std=std, dist_types=dist_types, inset_steady_state=True)
     plot_mse_test_distributions(np.mean(dev_variance * 100,2), SEM = sem_variance, fs = 8,
-                                title = 'Variance of stimuli', inset_steady_state = True) #  , x_lim=[0,50000]
+                                title = 'Variance of stimuli', inset_steady_state = True)
     
-        

### 2023-10-27 scripts/run_fig_2_support_3.py

@@ -10,13 +10,7 @@ Created on Mon Aug 14 10:50:33 2023
 
 import pickle
 import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-import pandas as pd
 
-from src.default_parameters import Neurons, Activity_Zero, Network, Stimulation
-from src.functions_save import load_network_para
-from src.functions_networks import run_population_net
 from src.functions_simulate import simulate_spatial_example
 from src.plot_data import plot_examples_spatial_M, plot_examples_spatial_V, plot_deviation_spatial
 
@@ -172,13 +166,3 @@ if run_cell:
         with open(file_for_data,'rb') as f:
             [mean_stimulus, spatial_std, spatial_noise, num_time_steps, m_neuron, v_neuron, rates_final] = pickle.load(f)
 
-
-    # plt.figure()
-    # plt.plot(np.arange(num_time_steps), m_neuron)
-    # ax = plt.gca()
-    # ax.axhline(mean_stimulus)
-    
-    # plt.figure()
-    # plt.plot(np.arange(num_time_steps), v_neuron)
-    # ax = plt.gca()
-    # ax.axhline(spatial_std**2)

### 2023-10-27 scripts/run_fig_3_support_2.py

@@ -11,8 +11,8 @@ Created on Mon Jul  3 18:45:22 2023
 import numpy as np
 import pickle
 
-from src.functions_simulate import simulate_sensory_weight_time_course, simulate_impact_para
-from src.plot_data import plot_weight_over_trial, plot_impact_para, plot_schema_inputs_tested
+from src.functions_simulate import simulate_impact_para
+from src.plot_data import plot_impact_para, plot_schema_inputs_tested
 
 import warnings
 warnings.filterwarnings("ignore")
@@ -24,14 +24,13 @@ dtype = np.float32
 
 run_cell = True
 load_only = True
-do_plot = True
+do_plot = False
 
 # define parameters used throughout the file
 mean_trials, m_sd = dtype(5), dtype(0)
 variability_within = np.array([0, 0.75, 1.5, 2.25, 3])
 variability_across = np.array([3, 2.25, 1.5, 0.75, 0])
-last_n = np.int32(50)
-n_trials = np.int32(200)
+last_n = np.int32(30)
 
 if run_cell:
     
@@ -47,7 +46,7 @@ if run_cell:
         [stimuli, n, gain_w_PE_to_P,
          gain_v_PE_to_P, add_input, 
          id_cells_modulated, weight] = simulate_impact_para(mfn_flag, variability_within, variability_across, mean_trials, 
-                                        m_sd, last_n = last_n, n_trials = n_trials, file_for_data = file_for_data)
+                                        m_sd, last_n = last_n, file_for_data = file_for_data)
         
     else:
         
@@ -78,8 +77,7 @@ if run_cell:
         
         print('Test different activation function for V neuron:\n')
         [_, _, _, _, _, _, weight_act] = simulate_impact_para(mfn_flag, variability_within, variability_across, mean_trials, 
-                                                          m_sd, last_n = last_n, n_trials = n_trials, 
-                                                          n = 1, file_for_data = file_for_data)
+                                                          m_sd, last_n = last_n, n = 1, file_for_data = file_for_data)
         
         
     else:
@@ -89,166 +87,13 @@ if run_cell:
             
     
     ### plot data 
-    plot_impact_para(weight_ctrl, weight_act, fs=7, ms=7)
+    plot_impact_para(variability_across, weight_ctrl, weight_act, fs=7, ms=7)
     
     
-# %% Impact of baseline activity (all neuron types)
-    
-run_cell = False
-plot_only = True
-
-if run_cell:
-
[…]

### 2023-10-27 scripts/run_fig_4_parallel_4.py

@@ -11,11 +11,15 @@ Created on Mon Jan  9 09:04:01 2023
 import numpy as np
 import pickle
 import matplotlib.pyplot as plt
+import seaborn as sns
+import pandas as pd
 import os.path
 
-from src.functions_simulate import simulate_moment_estimation_upon_changes_PE, simulate_neuromod_effect_on_neuron_properties
+from src.functions_simulate import simulate_neuromod, simulate_moment_estimation_upon_changes_PE, simulate_neuromod_effect_on_neuron_properties
 from src.functions_simulate import simulate_neuromod_combos, simulate_neuromod_pert
-from src.plot_data import plot_neuromod_impact_inter, plot_changes_upon_input2PE_neurons_new, plot_influence_interneurons_baseline_or_gain
+from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs, plot_neuromod_per_net, plot_points_of_interest_neuromod, plot_bar_neuromod
+from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_changes_upon_input2PE_neurons, plot_bar_neuromod_stacked
+from src.plot_data import plot_neurmod_results, plot_neuromod_impact_inter, plot_changes_upon_input2PE_neurons_new, plot_influence_interneurons_baseline_or_gain
 
 import warnings
 warnings.filterwarnings("ignore")
@@ -26,7 +30,7 @@ dtype = np.float32
 # %% Neuromodulator targeting one IN (with various strengths)
 
 run_cell = True
-plot_only = True
+plot_only = False
 column = 0 
 
 xp, xs, xv = 0, 0.5, 0.5
@@ -72,13 +76,11 @@ if run_cell:
     #plot_changes_upon_input2PE_neurons_new()
     plt.figure()
     plt.plot(pert_strength, alpha[:,:,0])
-    plt.ylim([0,1])
     
     plt.figure()
     plt.plot(pert_strength, alpha[:,:,1])
-    plt.ylim([0,1])
-        
-    
+
+
 # %% Activate fractions of IN neurons in lower/higher PE circuit or both for a specified input statistics
 
 run_cell = False
@@ -89,12 +91,12 @@ if run_cell:
     nums = 11
     
     # run through all combinations
-    for mfn_flag in ['10', '01', '11']: # choose mean-field network to simulate
+    for mfn_flag in ['01']: # choose mean-field network to simulate
[…]

### 2023-10-27 scripts/run_fig_4_parallel_5.py

@@ -11,11 +11,15 @@ Created on Mon Jan  9 09:04:01 2023
 import numpy as np
 import pickle
 import matplotlib.pyplot as plt
+import seaborn as sns
+import pandas as pd
 import os.path
 
-from src.functions_simulate import simulate_moment_estimation_upon_changes_PE, simulate_neuromod_effect_on_neuron_properties
-from src.functions_simulate import simulate_neuromod_combos, simulate_neuromod_pert
-from src.plot_data import plot_neuromod_impact_inter, plot_changes_upon_input2PE_neurons_new, plot_influence_interneurons_baseline_or_gain
+from src.functions_simulate import simulate_neuromod, simulate_moment_estimation_upon_changes_PE, simulate_neuromod_effect_on_neuron_properties
+from src.functions_simulate import simulate_neuromod_combos
+from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs, plot_neuromod_per_net, plot_points_of_interest_neuromod, plot_bar_neuromod
+from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_changes_upon_input2PE_neurons, plot_bar_neuromod_stacked
+from src.plot_data import plot_neurmod_results, plot_neuromod_impact_inter, plot_changes_upon_input2PE_neurons_new, plot_influence_interneurons_baseline_or_gain
 
 import warnings
 warnings.filterwarnings("ignore")
@@ -23,65 +27,9 @@ warnings.filterwarnings("ignore")
 dtype = np.float32
 
 
-# %% Neuromodulator targeting one IN (with various strengths)
-
-run_cell = True
-plot_only = True
-column = 0 
-
-xp, xs, xv = 0, 0.5, 0.5
-
-if run_cell:
-    
-    ### filename for data
-    file_for_data = '../results/data/neuromod/data_neuromod_' + str(xp) + '_' + str(xs) + '_' + str(xv) + '.pickle'
-    
-    ### define perturbations tested
-    pert_strength = np.linspace(0,0.5,6)
-    
-    ### main loop
-    if not plot_only:
-        
-        alpha = np.zeros((len(pert_strength), 3, 2))
-    
-        for k, mfn_flag in enumerate(['10', '01', '11']):
-            
-            for std_mean in [0, 1]: # uncertainty of environement [0, 1]
-                
-                n_std = 1 - std_mean # uncertainty of stimulus [1, 0]
[…]

### 2023-10-27 src/functions_networks.py

@@ -10,7 +10,6 @@ Created on Mon Jan  9 09:31:58 2023
 
 import numpy as np
 import os
-from numba import njit
 
 dtype = np.float32
 
@@ -257,7 +256,6 @@ def run_mfn_circuit_coupled(w_PE_to_P, w_P_to_PE, w_PE_to_PE, v_PE_to_P, v_P_to_
     
 ## functions for population network/s
 
-#@njit(cache=True)
 def drdt(tau_inv_E, tau_inv_I, tau_inv_var, wEP, wED, wDS, wDE, wPE, wPP, wPS, wPV, wSE, wSP, wSS, wSV, wVE, wVP, wVS, wVV,
          wEM, wDM, wPM, wSM, wVM, wME, wVarE, rE, rD, rP, rS, rV, r_mem, r_var, StimSoma_E, StimSoma_P, StimSoma_S, StimSoma_V, StimDend): 
     
@@ -367,7 +365,7 @@ def run_population_net(NeuPar, NetPar, StimPar, RatePar, dt, folder: str, fln: s
     wVarE = NetPar.wVarE
     
     ## Stimulation protocol & Inputs
-    stimuli = StimPar.stimuli # iter(StimPar.stimuli)
+    stimuli = StimPar.stimuli
     neurons_visual = StimPar.neurons_visual
     inp_ext_soma = StimPar.inp_ext_soma
     inp_ext_dend = StimPar.inp_ext_dend 

### 2023-11-03 records/manuscript_2023.tex
write Discussion & Abstract, captions for Figs. 1-4 (including supps)

@@
-\documentclass[10pt,a4paper,draft]{article}
-%\documentclass[10pt,a4paper]{article}
+%\documentclass[10pt,a4paper,draft]{article}
+\documentclass[10pt,a4paper]{article}
 
@@
 {\Large
-\textbf\newline{Weighting sensory inputs and predictions with prediction-error neurons}
+\textbf\newline{Integrating feedforward and feedback inputs with prediction-error neurons}
 }
@@
 \begin{abstract}
-blahhh blahhh blah
+At any moment, our brains receive a stream of feedforward sensory stimuli arising from the world we interact with.
+Simultaneously, neural circuits are shaped by feedback signals carrying predictions about the same feedforward inputs our brains are bombarded with.
+Those feedforward and feedback inputs often do not perfectly match.
+Thus, our brains have the challenging task of integrating these conflicting information according to their reliabilities.
+However, how neural circuits keep track of both the sensory and prediction uncertainties is not well understood.
+Here, we propose a network model whose core is hierarchical prediction-error circuits.
+We show that our network can estimate the variance of the sensory stimuli and the uncertainty in the prediction using the activity of negative and positive prediction-error neurons.
+In line with previous hypotheses, we demonstrate that neural circuits rely strongly on the feedback predictions at the beginning of a new stimulus, and if the environment is stable and the sensory cues are noisy.
+Furthermore, we investigate the causes of a pathological weighting of feedforward and feedback inputs.
+In our network, the uncertainty estimation, and, hence, how much we rely on predictions, can be biased by perturbing the intricate interplay of different inhibitory interneurons.
+Finally, we demonstrate the link to biased perception and unravel how the different types of uncertainty contribute to the contraction bias.
+%Hence, our work suggests a neural circuit level mechanism underlying the brain's astonishing ability to integrate feedforward and feedback signals.
 \end{abstract}
@@
     \includegraphics{../results/figures/final/Fig_1}
-\caption{\footnotesize{\bf Neural network model to flexibly integrate sensory information and predictions.\newline} 
-{\bf (A)} Example illustration of context-dependent integration of information.
-Left: When walking down a staircase that is clearly visible, the brain might rely solely on external sensory information.
-Middle: When walking down the same stairs in the absence of visual information, the brain might rely on predictions formed by previous experience.
-Right: When climbing down an unexplored mountain in foggy conditions, the brain might need to integrate sensory information and predictions at the same time.
+\caption{\footnotesize{\bf Neural network model to track both the uncertainty of sensory inputs and predictions.\newline} 
+{\bf (A)} Example illustration for context-dependent integration of information.
+Left: When walking down an unfamiliar staircase that is visible, the brain might rely solely on external sensory information.
+Middle: When walking down the same stairs without visual information, the brain might rely on predictions formed by previous experience.
+Right: When climbing down an unexplored mountain in foggy conditions, the brain might need to integrate sensory information and predictions simultaneously.
 {\bf (B)} Illustration of a prediction-error (PE) circuit with both negative and positive PE (nPE/pPE) neurons that receive inhibition from three different inhibitory interneuron types: parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons.
@@
     \includegraphics[width=1\linewidth]{../results/figures/final/Fig_2}
-\caption{\footnotesize{\bf Prediction-error neurons provide the basis for estimating mean and variance of sensory stimuli.\newline} 
-{\bf (A)} Illustration of the inputs with which the network shown in \ref{fig:Fig_1}C is stimulated.
-Network is exposed to a sequence of constant stimuli drawn from a uniform distribution.
-Stimulus duration is XXX.
+\caption{\footnotesize{\bf Prediction-error neurons as the basis for estimating mean and variance of sensory stimuli.\newline} 
+{\bf (A)} Illustration of the inputs with which the network (Fig. \ref{fig:Fig_1}C) is stimulated. Network is exposed to a sequence of constant stimuli drawn from a uniform distribution.
 {\bf (B)} PE neuron activity hardly changes with stimulus strength (left) but strongly increases with stimulus variability (right).
-{\bf (C)} Interneuron activity strongly changes with stimulus strength (left) but hardly changes with stimulus variability (right) .
+{\bf (C)} Interneuron activity strongly changes with stimulus strength (left) but hardly changes with stimulus variability (right).
 {\bf (D)} M neuron correctly encodes the mean of the sensory stimuli.
 Left: Illustration of the input synapses onto the M neuron.
-Middle: Activity of the M neuron over time for a uniform distribution with mean XXX and standard deviation XXX.
-Right: Normalised mean-squared error (MSE) between the running average and the M neuron activity for different parametrizations of the stimulus distribution.
+Middle: Activity of the M neuron over time for one example distribution (black start in right panel).
+Right: Normalised absolute difference between the averaged mean and the activity of the M neuron in the steady state for different parametrizations of the stimulus distribution.
 {\bf (E)} V neuron correctly encodes the variance of the sensory stimuli.
 Left: Illustration of the input synapses onto the V neuron.
-Middle: Activity of the V neuron over time for a uniform distribution with mean XXX and standard deviation XXX.
-Right: Normalised mean-squared error (MSE) between the instantaneous variance and the V neuron activity for different parametrizations of the stimulus distribution.
+Middle: Activity of the V neuron over time for one example distribution (black start in right panel).
+Right: Normalised absolute difference between the averaged variance and the activity of the V neuron in the steady state for different parametrizations of the stimulus distribution.
 }
@@
 
-\subsection*{Estimating variances of sensory inputs and predictions requires a hierarchy of PE circuits}
+\subsection*{Estimating the uncertainty of both the sensory input and the prediction requires a hierarchy of PE circuits}
 %
@@
     \includegraphics[width=1\linewidth]{../results/figures/final/Fig_3}
-\caption{\footnotesize{\bf Estimating variances of sensory inputs and predictions with hierarchical PE circuits.\newline} 
+\caption{\footnotesize{\bf Estimating the uncertainty of both the sensory input and the prediction.\newline} 
 {\bf (A)} Illustration of the stimulation protocol.
-Network is exposed to a sequence of stimuli (one stimulus per trial).
-To account for stimulus variability, each stimulus is represented by xxx stimulus values drawn from a normal distribution with mean $\mu_\mathrm{stim}$ and $\sigma_\mathrm{stim}^2$.
-To account for the volatility of the environment, in each trial the stimulus mean $\mu_\mathrm{stim}$ is drawn from a uniform distribution (denoted trial variability).
-Trial duration = xxx.
-{\bf (B)} Neuron activity increases with both stimulus and trial variability.
-Neurons in the lower PE circuit increase more strongly with stimulus variability.
-Neurons in the higher PE circuit increase more strongly with trial variability.
-{\bf (C)} Limit case example in which the stimulus variability is low but the trial variability is high.
+The network is exposed to a sequence of stimuli (one stimulus per trial).
+To account for stimulus variability, each stimulus is represented by $10$ stimulus values drawn from a normal distribution.
+To account for the volatility of the environment, in each trial, the stimulus mean is drawn from a uniform distribution (denoted trial variability).
+{\bf (B)} Illustration of how the weighted output is calculated.
+The sensory weight $\alpha$ lies between zero (system relies perfectly on prediction) and one (system relies solely on the sensory input).
+{\bf (C)} Limit case example in which the stimulus variability is zero but the trial variability is high.
 Left: Illustration of the stimulation protocol.
 Middle: Weighted output follows closely the sensory stimuli.
-Right: Sensory weight (function of the variances, see text) close to 1, indicating that the network ignores the prediction.
-Input statistics: XXX.
-{\bf (D)} Limit case example in which the stimulus variability is high but the trial variability is low.
+Right: Sensory weight (function of the variances, see B) close to 1, indicating that the network ignores the prediction.
+Input statistics shown in E.
+{\bf (D)} Limit case example in which the stimulus variability is high but the trial variability is zero.
 Left: Illustration of the stimulation protocol.
@@
 Right: Sensory weight close to zero, indicating that the network ignores the sensory stimuli.
-Input statistics: XXX.
-{\bf (E)} Predictions are weighted more strongly when the stimulus variability is larger than the trial variability.
-{\bf (F)} Predictions are weighted more strongly at the beginning of a new trial and quickly changing stimuli.
+Input statistics shown in E.
+{\bf (E)} Sensory weight for different input statistics. Predictions are weighted more strongly when the stimulus variability is larger than the trial variability.
+{\bf (F)} Sensory weight throughout a trial for two different trial durations. Predictions are weighted more strongly at the beginning of a new trial. 
 }
@@
 	\centering
-	\makebox[\textwidth][c]{\includegraphics{../results/figures/final/Fig_4}}
+	\makebox[\textwidth][c]{\includegraphics{../results/figures/final/Fig_4.pdf}}
     %\includegraphics{../results/figures/final/Fig_4} % [width=1\linewidth]
@@
 \newline} 
-{\bf (A)} Neuromodulators acting on the three interneurons may shift the weighting of sensory inputs and predictions.
-The changes depend on the type/s of interneurons targeted by the additional excitatory input emulating a neuromodulator (additional input = XXX).
+{\bf (A)} Neuromodulators acting on the interneurons can shift the weighting of sensory inputs and predictions.
+The changes depend on the type of interneuron targeted and the modulation strength (here simulated through an additional excitatory input).
 Considered are two limit cases (upper row: more sensory-driven before modulation, lower row: more prediction-driven before modulation).
-The combination of interneurons targeted is illustrated below.
-The results are shown for three different PE circuits, specified in the main text.
-{\bf (B)} Illustration showing how the sensory weight depends on changes in both the stimulus and trial variability.
-{\bf (C)} The M and V neuron activities depend on the PE neuron activities.
-Hence, perturbing the nPE and pPE neurons must change the estimation of mean and variance.
-While stimulating the lower PE neurons affects both the lower and higher mean and variance estimation, stimulating the higher PE neurons only affects the V and M neurons in the same subnetwork.
-{\bf (D)} Illustration of the mechanisms underlying the biased estimation of mean and variance when PE neurons are perturbed.
-Both changes in the baseline (left) and gain (right) of PE neurons can contribute to the changes observed in (C).
-Illustration based on toy model described in Methods.
-{\bf (E)} Modulated interneurons change the weighting by changing the overall baseline and the overall gain of PE neurons (sum of changes in nPE and pPE neurons).
-Whether and how a neuromodulator changes the sensory weight, hence, depends on the interneuron targeted and the effect this interneuron has on baseline and gain of the PE neurons, which in turn does depend on the network it is embedded in.
+The results are shown for three different PE circuits (denotes by different markers).
+{\bf (B)} When SOM and VIP neurons are equally modulated, the sensory weight remains unaffected. 
+{\bf (C)} The V neurons' activities depend on the PE neurons.
+Hence, perturbing the nPE and pPE neurons changes the uncertainty estimation.
+While stimulating the lower PE neurons affects both the lower and higher-order V neurons (right), stimulating the higher-order PE neurons only affects the V neuron in the same subnetwork (left).
+{\bf (D)} The V neuron activity, and hence the sensory weight, changes as a result of the modulated PE neuron activity.
+The PE neuron activity, on the other hand, changes as a result of the interneurons being modulated.
+The interneurons change the baseline (left) and the gain (right) of the PE neurons.
+Whether an interneuron increases or decreases the estimated variance depends on both factors.
 }
@@
 
-% Enough? This section seems rather short! Should we do more, like making predictions about what would happen with the bias if neuromodulators activate INs? Like going back?
-
 \section*{Discussion}
 
-We solved the brain.
-
-% summary of principal question, hypothesis and our findings
-
-% multisensory integration --> general idea that this can be applied to S and P ... Firth? link to psychiatric disorders
-
-% evidence for the model (PE neurons, connectivity PE --> M, predictions or P neurons exist, P neurons may hold the mean, time constants change along hierarchy)
-
-% alternative implementations (P in weights ...
-=> plasticity), also we dont implement weighting explicitely -- has been shown that divisive inhibiton can be done with INs (look in the one theoretical paper on neural implemnetation of Kalman filter)
-
-% Prediction update not weighted by uncertainty, discuss evidence fo that
-
-% neuromodulators, does it fit the current literature Lawson etc.? How is it different from expected and unexpected uncertainty ..., and then give more details on local vs. global neuromodulators
-
-% selectivity? One-dimensional system (wahrscheinlich nicht)
-
-% link to Kalman filter and maybe Bayes surprise factor
-
-----
-% Katha feedback: predictions, postdictions, experiments vorschlagen, limitations, grosserer kontext, abgrenzung zu reward prediction error
-
-% ################
-
-% At the end, or in between talk about changes in PE neurons directly that you have currently in Fig. 3 SX ... I copy the part of the last paragraph here as inspiration :
-% For some psychiatric disorders, it has been shown that the weighting of sensory inputs and predictions thereof is impaired (XXX), leading to an overweighting of one of these signals.
-Moreover, it has been hypothesised that factors like stress or cognitive load may also influence the processing of feedforward and feedback inputs (XXX).
-...
-Another factor that may contribute to a distorted weighting is the baseline activity of PE neurons that was set to zero in our model, in line with very low spontaneous firing rates of excitatory neurons in rodent primary sensory areas (XXX).
-Increasing this baseline activity for the nPE neuron (Fig.
-SXXX), pPE neuron (Fig.
-SXXX) or both pushes the network to weight sensory stimuli and predictions more equally.
-We speculate that an increase of baseline activity may be a natural result of an increased cognitive load or stress.
-
-% mention Pakan paper
+% XXX At the end: Read my „Todo“ in folder „read_weighting“: IMPORTANT XXX
+%
+% Inspiration and motivation for our work and summary of hypothesis
+Our work has been driven by the puzzling question of how the brain integrates top-down feedback predictions with the sensory feedforward bottom-up inputs it constantly receives during behavior.
+This task may be particularly challenging when the prediction and the sensory information differ \ref{han2023behavior}.
+Conflicting information may be caused by noise in the sensory feedforward inputs or by changes in the environment that could not be predicted.
+A prominent hypothesis is that how much we rely on our predictions and new sensory evidence is determined by an intricate balance between both, based on how reliable they are \ref[see e.g.][]{yon2021precision}.
+
+This idea is consistent with Bayesian theories on the optimal integration of multiple sensory cues (aka multisensory integration).
+\ref{XX ernst and banks 2002} showed that to estimate the height of a bar humans combine visual and haptic information in a fashion that minimizes the variance of the final estimate.
+XXX name other examples in human and other species XXX.
+These behavioral studies were accompanied by neural recordings showing that XX
+
+% Summary of findings and predictions/implications from our model
+
+While multisensory integration is concerned with the weighting of sensory cues coming from different modalities, it is conceivable that the same ideas hold for top-down predictions and bottom-up sensory inputs that are combined in cortical circuits \ref{yon2021precision}.
+Indeed, there is evidence that the brain can choose to rely more on top-down or bottom-up signals \ref{pakan2018impact, han2023behavior}.
+Following the same ideas from Bayesian multisensory integration, the brain must hence have mechanisms to track both the reliability of the sensory signals and the predictions thereof \ref{yon2021precision}.
+XXX We suggested that xxx Here summarise your contributions XXX
+
+XXX put this in here somewhere XXX
+In our model, we assume a hierarchy of PE circuits that are feed-forwardly connected through the lower-order M neuron whose activity encodes the mean of the sensory bottom-up inputs.
+This local prediction is fed back to the lower-order circuit and at the same time feed-forwarded to the higher-order subnetwork.
+With this architecture in place, we show that we rely more strongly on our internal signals when the perceived sensory cues are noisier than the predictions, either because the environment is stable but highly noisy or because the environment changes quickly.
+This implies that revising our inner model of the world, that is, learning from PEs, should be suppressed if the sensory noise is high or the environment switches rapidly \ref{herzfeld2014memory}.
+Moreover, our work suggests that when experimental trials are short, a neural network can better represent the external world by relying on predictions.
+Hence, studying neural signatures of predictions in the brain might require experiments that involve sufficiently short trials.
+
+Furthermore, we show that the weighting of sensory inputs and predictions can be biased through neuromodulators, as has been suggested before \ref{XXX}.
+In our model, those modulatory signals act through interneurons \ref{cardin2019functional} whose activities increase in the presence of neuromodulators.
+When PV neuron activity increases, the network weighs predictions stronger than without modulation.
+In contrast, when VIP neuron activity increases, the network underestimates the uncertainty of the prediction in a sensory-driven regime and underestimates the uncertainty of the sensory input in a prediction-driven regime.
+Hence, the system leans toward weighting sensory inputs and predictions more equally.
+Increasing SOM neuron activity is comparable to modulating VIP neurons.
+However, the equalising effect is less robust and seems to depend on the network specifics (Fig.
+\ref{XX}A).
+Interestingly, when SOM and VIP neuron activity is modulated to the same degree, the weighting remains unaffected, suggesting that the individual contributions cancel (Fig.
+\ref{XX}B).
+We show that these findings can be best explained by changes in the baseline and gain of PE neurons as a result of the modulation of interneuron activity (Fig.
+\ref{XXX}D).
+These results can be tested experimentally by optogenetically or pharmacologically stimulating specific interneuron types.
+
+Finally, we illustrate that the weighted integration of feedforward and feedback inputs can be interpreted as a neural manifestation of the contraction bias.
+We show that the bias is strongly driven by the variance of the sensory cues and the predictions, as well as the trial duration.
+In our model, if the stimulus variance is zero (perfect sensory-driven network regime), the neural contraction bias is independent of the trial variability.
+In that case, the bias reflects the slow transition from the last stimulus to the new one, and, hence, would diminish if the trial is sufficiently long.
+On the other hand, if the trial variance is zero (perfect prediction-driven network regime), the neural contraction bias is independent of the stimulus variability and a function of the trial duration.
+However, we note that in our model, we only consider the neural representation of the stimulus and do not account for other sources of noise, like execution noise, that surely impacts the contraction bias observed in behavioral studies \ref{XXX}.
+
+% Inspiration for the model & assumption (evidence for that)
+
+In our model, we assumed that there are dedicated neurons that encode the variance of the feedforward sensory inputs and the prediction.
+This assumption is consistent with the idea that neurons explicitly encode in their activity the parameters (for instance, mean or variance) of a probability distribution (O’Neill \& Schultz, 2010, XXX).
+However, how variances of signals are represented in the brain is still not comprehensively understood and alternative ideas have been put forward.
+For instance, it is conceivable that the variance is encoded in a population of neurons, each differently tuned to a specific parameter (Knill \& Pouget, 2004; Ma et al., 2006, 2008).
+The neurons' activities represent how close the sensory input is to the preferred (predicted) input of each neuron.
+
+% —> check out: There exist a number of theoretical models that address the potential neural representation of uncertainty (Deneve, 2008; Hinton and Sejnowski, 1983; Hoyer and Hyvarinen, 2003; Ma et al., 2006; Zemel et al., 1998) … in Dekleva 2016 XXXXX
+
+There has been evidence that indeed (population of) neurons can encode (un-)certainty.
+Neurons in the pulvinar (Komura et al., 2013), the parietal cortex (Kiani and Shadlen, 2009) or the frontal cortex (Kepecs et al., 2008, Middlebrooks and Sommer, 2012) encode the uncertainty of perceptual decisions \ref{masset2020behavior}.
+In addition, it has been shown that neurons in the orbitofrontal cortex of rats signal confidence in decisions \ref{masset2020behavior}.
+Moreover, neural correlates of (expected) uncertainty have been found in regions of the prefrontal cortex 43 \ref{soltani2019adaptive} , the rat orbitofrontal cortex [52,53\ref{soltani2019adaptive}] or the dorsal striatum in monkeys78 \ref{soltani2019adaptive}.
+Interestingly, the accuracy of memory recalls is represented (?) in single neurons of the human parietal and temporal lobes (Rutishauser et al., 2015, 2018) … in \ref{masset2020behavior}.
+
+In our model, we assume a hierarchy of predictions that are locally computed in memory neurons.
+These memory neurons are consistent with the idea of internal representation neurons hypothesised in predictive processing theories \ref{bastos2012canonical, keller2018predictive}.
+While it has been hypothesised that these internal representation neurons might be found in deeper L5 neurons \ref{bastos2012canonical, heindorf2022reduction}, there is also evidence that a group of excitatory L2/3 neurons integrates over negative and positive prediction errors \ref{o2022prediction}.
+
+The core hypothesis of our model is the presence of sensory PE neurons.
+Those neurons have been found in different cortical areas in various species \ref{eliades2008neural, keller2009neural, ayaz2019layer, audette2021temporally}.
+Moreover, while first only hypothesized theoretically \ref{rao1999predictive}, the presence of two types of PE neurons, the negative and positive PE neurons, has been confirmed in a number of recent studies \ref{keller2012sensorimotor, attinger2017visuomotor, jordan2020opposing, audette2021temporally, Zmarz2016XXX?necessary}.
+In our model, the nPE neuron inhibits the memory neuron while the pPE neuron excites it, in line with \ref{keller2018predictive}.
+The weights from those PE neurons onto the M neuron are larger for the lower- than the higher PE circuit, so that the lower-order M neuron evolves faster than the higher-order counterpart.
+This assumption is consistent with the observation that time constants increase along the cortical hierarchy (XXX see refs in Siegle 2021 Allen).
+
+However, if the relation of the paces at which the M neurons evolve is strongly modulated, the network either shows a bias towards the sensory cues or the prediction (Figs.
+\ref{XXX}).
+It has been hypothesized that symptoms in psychiatric diseases may be a result of an erroneous uncertainty estimation \ref{yon2021precision}.
+For instance, hallucinations may be the result of an underestimation of the expectation uncertainty or an overestimation of the sensory uncertainty.
+Conversely, a fixation on a stable environment even when the sensory cues indicate a switch in the environment, may be the consequence of an overestimation of the expectation uncertainty or an underestimation of the sensory uncertainty \ref{yon2021precision}.
+
+% Neuromodulation
+
+A popular hypothesis is that neuromodulators shape the weighting of sensory inputs and predictions thereof \ref{yon2021precision}.
+Theoretical work by \ref{You and Dayan} suggests that acetylcholine (ACh) correlates with expected uncertainty, while noradrenaline (NA) correlates with unexpected uncertainty.
+Expected uncertainty is usually considered to be related to known cue-outcome unreliabilities.
+In contrast, unexpected uncertainty relates to the changes in the environment that produce large PEs outside the expected range of uncertainties \ref{You and Dayan}.
+While in our network the stimulus and trail variability can only be loosely interpreted as 'expected' and 'unexpected' uncertainty, we want to XXX broader context and highlight differences XXX
+
+% XXX —> the unexpected uncertainty in my case (trial variability) is not unexpected time-wise only direction-wise.
+One could change that in a SI figure or one later main figure but not sure how much sense that makes (discuss with Katha and CC)
+
+It is assumed that NA increases in more volatile environments and enhances bottom-up processes [20,21 in \ref{lawson2021computational}, yon2021precision].
+In line with this idea, NA blockade impairs cognitive flexibility [24,25\ref{lawson2021computational}].
+In recent work by \ref{lawson2021computational}, it has been shown that humans receiving propranolol (blocking NA) rely more strongly on their expectations and are slower to update these predictions despite new sensory evidence \ref{yon2021precision}.
+A major target for noradrenergic inputs is SOM neurons whose activity increases in the presence of NA (reviewed in Urban-Ciecko 2016: Fanselow, 2008; Urban-Ciecko 2015; Chen 2015; Xiang 1998; reviewed in Hattori 2017: Paspalas 1999; reviewed in Swanson in Maffei 2019: Kawaguchi and Shindou, 1998).
+In our model, activating SOM neurons globally does not enhance bottom-up input.
+In a volatile environment, that is, a sensory-driven regime, the system takes into account predictions slightly more than without SOM modulation.
+However, increasing SOM activity only in the lower-order circuit shows a slight enhancement of the sensory weight (Fig.
+\ref{XXX}).
+
+Similarly to NA, ACh has also been shown to enhance bottom-up, feedforward inputs (Gil et al., 1997; Hasselmo et al., 1996; Hsieh et al., 2000; Kimura et al., 1999; Kobayashi, 2000, reviewed in Yu and Dayan 2005 or Marshall 2016).
+For instance, subjects relied more strongly on prior beliefs when given cholinergic receptor antagonists (reviewed in Bruckner 2022: Marshall 2016).
+A major target for cholinergic input are VIP neurons whose activity increases in the presence of ACh (reviewed in Wester 2014; reviewed in Hattori 2017: Fu 2016, Kuchibhotla 2017, Nelson 2016; reviewed in Swanson in Maffei 2019: Paul et al., 2017).
+In our model, activating VIP neurons globally only enhances bottom-up input in stable environments for noisy stimuli.
+However, increasing VIP activity only in the higher-order circuit does generally enhance sensory bottom-up inputs (Fig.
+\ref{XXX}).
+
+Neuromodulatory systems have been suggested to gate plasticity \ref{XXX}.
+In a recent study by \ref{jordan2023locus}, locus coeruleus (LC) axon activity is shown to correlate with the magnitude of unsigned visuomotor prediction errors.
+The authors hypothesize that LC output modulates the learning rate at which the internal model evolves \ref{jordan2023locus}.
+In our model, we do not consider precision-weighted PEs (but see \ref{KathaXXX}).
+Hence, a sensible extension to our work would be to adjust the weights from PE neurons onto the M neurons by a function of the stimulus and prediction uncertainty, respectively.
+This would allow us to compare our results more closely to work showing that ACh and NA can adjust the rate of incorporating new sensory evidence when environments change (\ref{XXX}).
+
+%—> XXX man sollte vielleicht noch sagen, dass auch PE neurons von neuromod direct beeinflusst sein können (es gab da zwei Beispiele, die weiter unten sind und die du hier aus Versehen gelo2scht hattest).
+—> Zum einen können diese auch zu biases beitragen, zum anderen konnten sie aber auch Effekt durch IN kompensieren.
+
+% —> XXX should we add in one SI precision-weighted PE to compare with Neuromodulation results? (Discuss with Katha and CC). If so, how to choose the weighting?
+% —> XXX From one trial to the next (so basically environment switch), should we check how fast the update happens when neuromods are applied to compare it to data?
+(Discuss with Katha and CC), —> should we also show or at least mention in discussion that by changing PE activity the prediction/mean can also be wrong?
+How is that actually checked in experiments or how do we know that somebody relies on prediction if the prediction might be wrong
+
+% Limitations, Alternations and future steps
+
+As for any computational model, we brush over a number of biological details in order to keep the model simple and interpretable.
+However, those details, while beyond the scope of this study, may be well investigated in future work.
+For instance, evidence suggests that prediction-errors itself are precision-weighted \ref{yon2021precision}.
+xx Kelly Diederen\ref{yon2021precision} xxx show that neural signatures of reward-prediction errors are augmented when the outcome is reliable and attenuated when the outcome is uncertain.
+XXX Katha’s work XXX
+
+Furthermore, while we provide a neuronal circuit model for estimating the mean and the variance of both sensory signals and predictions, we do not model the weighting of inputs explicitly.
+A respective neural circuit model would require divisive inhibition.
+XXX (look in paper for neural implementation of Kalman filter) xxx have shown how this can be done with inhibitory interneurons.
+
+Our model suggests one potential neuronal circuit mechanism for the estimation of sensory inputs and predictions.
+However, in the light of evidence showing that the integration of feedforward and feedback inputs is species- and modality-dependent, it is conceivable that a plethora of neural mechanisms are used in neural circuits.
+For multisensory integration, \ref{wong2023computational} showed that in Drosophila larva the chosen cue-combination strategy varies depending on the type of sensory information available.
+Also, humans put typically more weight on visual than auditory cues ((Battaglia et al., 2003; Burr \& Alais, 2004).
+… in \ref{o2012can}), but trust vestibular information more than visual information about head direction (Butler et al., 2010 \ref{o2012can}), a finding also observed for monkeys (Angelaki et al., 2009 \ref{o2012can}).
+Moreover, Summerfield et al.
+(2011) \ref{o2012can} showed that humans diverge from an optimal Bayesian strategy in very volatile environments and act according to their experience in the last trial.
+Moreover, it has been suggested that the brain may use different strategies to combine signals depending on the task demands \ref{o2012can}.
+While these behavioral results cannot speak to the underlying circuit mechanisms, it is conceivable that neural implementations for the integration of feedforward and feedback inputs may also vary.
+
+% Relation to other work that came before and closing remarks
+
+A number of normative models have been proposed for state estimation and prediction under uncertainty \ref{soltani2019adaptive}, ranging from the classical Kalman filter to more recent models like \textit{Bayes Factor Surprise} \ref{liakoni2021learning}.
+For instance, Bayes factor surprise formularizes the trade-off between integrating new observations in an existing belief system and resetting this belief system with the new evidence.
+The surprise factor captures how much an animal’s current belief deviates from the new observation.
+
+In more recent years, normative model have been squared with biological constraints.
+For instance, in a paper by \ref{kutschireiter2023bayesian}, uncertainty could be integrated into a ring attractor model encoding head direction.
+This \textit{Bayesian ring attractor} model encodes uncertainty in the amplitude of the network activity and matches the performance of a circular Kalman filter when the recurrent connections are tuned appropriately.
+
+XXX Come back to your model XXX highlight the nice thing about it and then find some concluding remarks
+
+% XXXX PLEASE LOOK AT MILIIDGE 2023 that you reviewed!!!
 
@@
 \bibliographystyle{naturemag} %{plainnat}
-\bibliography{../References_Mismatch}
+\bibliography{References_HertaegWilmesClopath_2023}
 
@@
 	\centering
-    \includegraphics[scale=1]{../results/figures/final/Fig_2_S1}% [width=1\linewidth]
-\caption{\footnotesize{\bf Estimation of mean and variance for different stimulus distributions.\newline}  
-Top: The mean-squared error (MSE) between the running average and the activity of the M neuron decreases to a near-zero level for all stimulus distributions tested. 
-Bottom: The MSE between the instantaneous variance and the V neuron decreases to a low level with minor differences between the distributions tested.
-Zoom-in shows the last half of the trial.
-Mean of the stimulus distribution = XXX, Variance of the stimulus distribution = XXX.
+    \includegraphics{../results/figures/final/Fig_2_S1.pdf}% [width=1\linewidth]
+\caption{\footnotesize{\bf Estimating mean and variance of different stimulus distributions.\newline}  
+Top: The normalised absolute difference between the averaged mean and the activity of the M neuron decreases to a near-zero level for all stimulus distributions tested. 
+Bottom: The normalised absolute difference between the averaged variance and the activity of the V neuron decreases with small differences between the distributions tested.
+Parametrisation of the uniform distribution as in Fig.
+\ref{fig:Fig_2}.
 }
@@
 	\centering
+    \includegraphics{../results/figures/final/Fig_2_S2.pdf}% [width=1\linewidth]
+\caption{\footnotesize{\bf Estimating mean and variance of sensory stimuli in a rate-based population network.\newline}  
+{\bf (A)} Illustration of the rate-based population network and the stimuli over time.
+{\bf (B)} M and V neuron activities over time for one example parameterisation.
+{\bf (C)} The normalised absolute difference between the averaged mean and the activity of the M neuron (dark green) or between the averaged variance and the activity of the V neuron (brown) for uncorrelated deviations, that is, increasing SD of $\gamma$ (left), correlated deviations, that is, increasing mean of $\gamma$ (middle), and the network sparsity.
+}
+\label{fig:Fig_2_S2}
+\end{figure}
+
+\begin{figure}[!h]
+	\centering
+    \includegraphics[width=1\linewidth]{../results/figures/final/Fig_2_S3.pdf}% [width=1\linewidth]
+\caption{\footnotesize{\bf Estimating mean and variance of spatial stimuli.\newline}  
+{\bf (A)} Illustration of a network estimating the mean and variance of a stimulus that varies across space.
+To simulate selectivity, the network comprises $1000$ identical, uncoupled mean-field networks each receiving a different input value drawn from a uniform distribution.
+{\bf (B)} Activity of M neuron (top) and V neuron (bottom) for 2 stimuli. The second stimulus does either differ in the mean (orange) or the variance (yellow) from the first stimulus (indicated in C).
+{\bf (C)} The normalised absolute difference between the averaged mean and the activity of the M neuron (dark green, top) or between the averaged variance and the activity of the V neuron (brown, bottom) for a range of different stimulus statistics.
+The examples from B are shown with colored arrows and markers.
+}
+\label{fig:Fig_2_S3}
+\end{figure}
+
+\begin{figure}[!h]
+	\centering
     \includegraphics{../results/figures/final/Fig_3_S1}% [width=1\linewidth]
 \caption{\footnotesize{\bf Dynamic variance estimation allows flexible adaptation to changes in the stimulus statistics and environment. \newline}  
-{\bf (A)} Sensory weight for different input statistics (same as in Fig. \ref{fig:Fig_3}E). Numbers denote specific example states. Arrows denote the transitions between those states.
-{\bf (B)} The sensory weight over time is shown for all transitions in (A). The switch to a new input statistics occurs at trial 60. Parameters are taken from Fig. \ref{fig:Fig_3}.
+{\bf (A)} Sensory weight for different input statistics. Numbers denote specific example statistics. Arrows denote the transitions between those statistics.
+{\bf (B)} The sensory weight over time is shown for all transitions in (A).
+For the sake of clarity, we only show the trials 40 -60.
+The switch to new input statistics occurs at trial 50.
+Parameters are listed in the Supporting Information.
 }
@@
 \caption{\footnotesize{\bf Perturbing the weighting of sensory inputs and predictions by altering network properties. \newline}  
-{\bf (A)} The weights from the lower-level PE neurons to the M neuron are scaled by a factor below 1 (here, xxx) or above 1 (here, xxx), leading to a distorted weighting.
+{\bf (A)} The weights from the PE neurons to the M neuron in the lower-order subnetwork are scaled by a factor the 0.3 or 7, leading to a distorted sensory weight.
 If the update of the M neuron in the lower subnetwork is too slow ($\blacktriangleleft$), the prediction is overrated.
@@
 	\centering
-    \includegraphics{../results/figures/final/Fig_4_S1}% [width=1\linewidth]
-\caption{\footnotesize{\bf Neuromodulators acting locally either on interneurons in the lower or higher PE circuit. \newline}  
-{\bf (A)} Sensory weight changes with neuromodulators acting on interneurons in the lower PE circuit.
-{\bf (B)} Sensory weight changes with neuromodulators acting on interneurons in the lower PE circuit. Simulation parameters, labels and colors as in Fig. \ref{fig:Fig_4}. 
+    \includegraphics{../results/figures/final/Fig_4_S1.pdf}% [width=1\linewidth]
+\caption{\footnotesize{\bf The impact of neuromodulators acting globally on groups of interneurons. \newline}  
+The sensory weight changes when groups of interneurons are targeted by a neuromodulator.
+Whether the sensory weight decreases or increases not also depends on the modulation strength (see Fig.
+\ref{fig:Fig_4}) but also on how strongly which interneuron is targeted.
+As shown in Fig.
+\ref{fig:Fig_4}, the sensory weight is pushed toward 0.5 if the VIP neuron is stimulated.
+The sensory weight generally decreases when PV neurons are the main target.
+Considered are two limit cases (upper row: more sensory-driven before modulation, lower row: more prediction-driven before modulation).
+The results are shown for three mean-field networks (see \ref{fig:Fig_4}).
 }
@@
     \includegraphics{../results/figures/final/Fig_4_S2}% [width=1\linewidth]
-\caption{\footnotesize{\bf Perturbing the interneurons changes the baseline and gain of nPE and pPE neurons.\newline}  
-{\bf (A)} Changes in the baseline activity of nPE and pPE neurons for different interneurons targeted. 3 different mean-field networks are tested.
-{\bf (B)} Same as in (A) but for the gain of nPE and pPE neurons. Simulation parameters, labels, and colors as in Fig. \ref{fig:Fig_4}. 
+\caption{\footnotesize{\bf The impact of neuromodulators acting locally on groups of interneurons. \newline}  
+{\bf (A)} Sensory weight changes with neuromodulators acting on interneurons in the lower PE circuit.
+{\bf (B)} Sensory weight changes with neuromodulators acting on interneurons in the higher PE circuit.
+In general, the changes in the sensory weight is the opposite of the changes seen for neuromodulators acting on the lower-level PE neurons.
+Simulation parameters, labels and colors as in Fig.
+\ref{fig:Fig_4}.
 }
@@
 
+\begin{figure}[!h]
+	\centering
+    \includegraphics{../results/figures/final/Fig_4_S3}% [width=1\linewidth]
+\caption{\footnotesize{\bf Biased mean and variance estimation by changing the baseline and the gain of nPE and pPE.}
+In a toy model, described in XXX, the contribution of gain and baseline to the changes in the mean and variance estimation can be revealed. XXX Equations
+}
+\label{fig:Fig_4_S3}
+\end{figure}
+
+\begin{figure}[!h]
+	\centering
+    \includegraphics{../results/figures/final/Fig_4_S4}% [width=1\linewidth]
+\caption{\footnotesize{\bf The combined changes in baseline and gain of all PE neurons determine the shift in the snesory weight.\newline}  
+Whether and how a neuromodulator changes the sensory weight depends on the interneuron targeted and the effect this interneuron has on the baseline and gain of both PE neurons, which in turn does depend on the network it is embedded in.
+For small inputs, changes in the baseline dominate, while for large inputs, the changes in the gains dominate the shift in the sensory weight.
+}
+\label{fig:Fig_4_S4}
+\end{figure}
+
 \end{document}

### 2023-11-03 scripts/plot_fig_4.py

@@ -11,9 +11,10 @@ Created on Mon Jan  9 09:04:01 2023
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
+import pickle
 import os.path
 
-from src.plot_data import plot_neuromod_impact_inter, plot_illustration_neuromod_results, plot_standalone_colorbar
+from src.plot_data import plot_neuromod_impact, plot_standalone_colorbar
 from src.plot_data import plot_changes_upon_input2PE_neurons_new, plot_influence_interneurons_baseline_or_gain
 
 
@@ -34,79 +35,110 @@ if not os.path.exists(figPath):
 
 # %% Define figure structure
 
-figsize=(18/inch,14/inch)
+figsize=(14/inch,14/inch)
 fig = plt.figure(figsize=figsize)
 
-G = gridspec.GridSpec(2, 1, figure=fig, hspace=0.7) #, hspace=0.6, height_ratios=[1.2,2])
-A = gridspec.GridSpecFromSubplotSpec(2, 4, subplot_spec=G[0,0])#, width_ratios=[3,1], hspace=0.5)
+G = gridspec.GridSpec(2, 1, figure=fig, hspace=0.5, height_ratios=[1,1.2])
+G1 = gridspec.GridSpecFromSubplotSpec(2, 4, subplot_spec=G[0,0], width_ratios=[1,1,1,1.2], wspace=0.8)
+G2 = gridspec.GridSpecFromSubplotSpec(2, 4, subplot_spec=G[1,0], width_ratios=[1,1,1,1.2], wspace=0.8, hspace=0.7)
 
-G_sub = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=G[1,0], width_ratios=[0.7,1], wspace=0.3)
-C_and_D = gridspec.GridSpecFromSubplotSpec(2, 2, subplot_spec=G_sub[0,1], wspace=0.7, hspace=0.5)
+A = gridspec.GridSpecFromSubplotSpec(2, 3, subplot_spec=G1[:,:3], hspace=0.2, wspace=0.4)
+B = gridspec.GridSpecFromSubplotSpec(2, 1, subplot_spec=G1[:,-1])
+C = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=G2[0,1:], wspace=0.5)
[…]

### 2023-11-03 scripts/plot_fig_4_old.py

@@ -6,34 +6,16 @@ Created on Mon Jan  9 09:04:01 2023
 @author: loreen.hertaeg
 """
 
-# %% Import
-
-import numpy as np
-import pickle
-
-from src.functions_simulate import simulate_neuromod
-from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-
 # %% import
 
-import pickle
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import os.path
-import seaborn as sns
 
-from matplotlib.collections import LineCollection
+from src.plot_data import plot_neuromod_impact_inter, plot_illustration_neuromod_results, plot_standalone_colorbar
+from src.plot_data import plot_changes_upon_input2PE_neurons_new, plot_influence_interneurons_baseline_or_gain
 
-from src.plot_data import plot_neuromod_impact_inter, illustrate_sensory_weight_variance, plot_changes_upon_input2PE_neurons
-from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_influence_interneurons_gain_baseline
-from src.plot_data import plot_legend_illustrations
 
 # %% Universal parameters
 
@@ -52,79 +34,52 @@ if not os.path.exists(figPath):
 
 # %% Define figure structure
 
-figsize=(18/inch,18/inch)
+figsize=(18/inch,14/inch)
 fig = plt.figure(figsize=figsize)
 
-G = gridspec.GridSpec(2, 1, figure=fig, hspace=0.6, height_ratios=[1.2,2])
+G = gridspec.GridSpec(2, 1, figure=fig, hspace=0.7) #, hspace=0.6, height_ratios=[1.2,2])
+A = gridspec.GridSpecFromSubplotSpec(2, 4, subplot_spec=G[0,0])#, width_ratios=[3,1], hspace=0.5)
 
[…]

### 2023-11-03 scripts/plot_fig_4_support_1.py

@@ -12,7 +12,8 @@ import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import os.path
 
-from src.plot_data import plot_neuromod_impact_inter
+from src.plot_data import plot_neuromod_impact_inter, plot_illustration_neuromod_results
+
 
 # %% Universal parameters
 
@@ -31,68 +32,45 @@ if not os.path.exists(figPath):
 
 # %% Define figure structure
 
-figsize=(18/inch,14/inch)
+figsize=(14/inch,14/inch)
 fig = plt.figure(figsize=figsize)
 
-G = gridspec.GridSpec(2, 1, figure=fig, hspace=0.6)
+G = gridspec.GridSpec(2, 1, figure=fig, hspace=0.7)
+A = gridspec.GridSpecFromSubplotSpec(2, 3, subplot_spec=G[0,0])
+B = gridspec.GridSpecFromSubplotSpec(1, 5, subplot_spec=G[1,0])
 
-G1 = gridspec.GridSpecFromSubplotSpec(2, 3, subplot_spec=G[0,0], hspace=0.3)
-G2 = gridspec.GridSpecFromSubplotSpec(2, 3, subplot_spec=G[1,0], hspace=0.3)
-
-ax_A = fig.add_subplot(G1[:,:])
+ax_A = fig.add_subplot(A[:,:4])
 ax_A.axis('off')
-ax_A.set_title('Neuromodulators acting on interneurons in lower PE circuit', fontsize=fs, pad=10)
-ax_A.text(-0.1, 1.1, 'A', transform=ax_A.transAxes, fontsize=fs+1)
+ax_A.set_title('Neuromodulators acting on interneurons can bias the weighting of sensory inputs and predictions', fontsize=fs, pad=18)
 
-ax_A12 = fig.add_subplot(G1[0,0])
+ax_A12 = fig.add_subplot(A[0,0])
 plt.setp(ax_A12.get_xticklabels(), visible=False)
-ax_A13 = fig.add_subplot(G1[0,1])
+ax_A13 = fig.add_subplot(A[0,1])
 plt.setp(ax_A13.get_yticklabels(), visible=False)
 plt.setp(ax_A13.get_xticklabels(), visible=False)
-ax_A14 = fig.add_subplot(G1[0,2])
+ax_A14 = fig.add_subplot(A[0,2])
 plt.setp(ax_A14.get_yticklabels(), visible=False)
[…]

### 2023-11-03 scripts/plot_fig_4_support_3.py

@@ -8,11 +8,12 @@ Created on Mon Jan  9 09:04:01 2023
 
 # %% import
 
+import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import os.path
 
-from src.plot_data import plot_influence_interneurons_gain_baseline
+from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_legend_illustrations
 
 # %% Universal parameters
 
@@ -31,16 +32,35 @@ if not os.path.exists(figPath):
 
 # %% Define figure structure
 
-figsize=(6/inch,5/inch)
-fig = plt.figure(figsize=figsize, tight_layout=True)
-
-G = gridspec.GridSpec(1, 1, figure=fig)
-ax_A = fig.add_subplot(G[0,0])
+figsize=(10/inch,8/inch)
+fig = plt.figure(figsize=figsize)
 
 
-# %% How do the interneurons influence the PE neuron properties?
+G = gridspec.GridSpec(3, 3, figure=fig, wspace=0.3, hspace=0.2)
 
-plot_influence_interneurons_gain_baseline(ax=ax_A)
+ax_A = fig.add_subplot(G[:,:2])
+ax_A.axis('off')
+ax_A.set_title('Illustration of changes in mean and variance \nwith changes in baseline (left) and gain (right)', fontsize=fs, pad=10)
+
+ax_A1 = fig.add_subplot(G[0,0])
+ax_A2 = fig.add_subplot(G[1,0])
+ax_A3 = fig.add_subplot(G[2,0])
+
+ax_B1 = fig.add_subplot(G[0,1])
+ax_B2 = fig.add_subplot(G[1,1])
+ax_B3 = fig.add_subplot(G[2,1])
+
+ax_A_legend = fig.add_subplot(G[0,-1])
+ax_A_legend.axis('off')
+
+# %%  How is the variance influenced by BL of nPE and pPE in lower and higher PE circuit?    
+ 
+plot_illustration_changes_upon_baseline_PE(BL = np.linspace(0,3,7), ax1 = ax_A1, ax2 = ax_A2, ax3 = ax_A3)
+
+# %%  How is the variance influenced by gain of nPE and pPE in lower and higher PE circuit?    
+ 
[…]

### 2023-11-03 scripts/run_fig_2_support_1.py

@@ -87,7 +87,7 @@ if run_cell:
     sem_variance = np.std(dev_variance * 100,2)/np.sqrt(num_reps)
     
     plot_mse_test_distributions(np.mean(dev_mean * 100,2), SEM = sem_mean, title = 'Mean of stimuli', fs=8, 
-                                mean=mean, std=std, dist_types=dist_types, inset_steady_state=True)
+                                mean=mean, std=std, dist_types=dist_types)
     plot_mse_test_distributions(np.mean(dev_variance * 100,2), SEM = sem_variance, fs = 8,
-                                title = 'Variance of stimuli', inset_steady_state = True)
+                                title = 'Variance of stimuli')
     

### 2023-11-03 scripts/run_fig_4.py

@@ -10,12 +10,10 @@ Created on Mon Jan  9 09:04:01 2023
 
 import numpy as np
 import pickle
-import matplotlib.pyplot as plt
-import os.path
 
-from src.functions_simulate import simulate_moment_estimation_upon_changes_PE, simulate_neuromod_effect_on_neuron_properties
-from src.functions_simulate import simulate_neuromod_combos, simulate_neuromod_pert
-from src.plot_data import plot_neuromod_impact_inter, plot_changes_upon_input2PE_neurons_new, plot_influence_interneurons_baseline_or_gain
+from src.functions_simulate import simulate_moment_estimation_upon_changes_PE, simulate_neuromod_pert
+from src.functions_simulate import simulate_neuromod_effect_on_neuron_properties
+from src.plot_data import plot_changes_upon_input2PE_neurons_new, plot_influence_interneurons_baseline_or_gain, plot_neuromod_impact
 
 import warnings
 warnings.filterwarnings("ignore")
@@ -29,7 +27,7 @@ run_cell = True
 plot_only = True
 column = 0 
 
-xp, xs, xv = 0, 0.5, 0.5
+xp, xs, xv = 0, 0, 1
 
 if run_cell:
     
@@ -69,81 +67,9 @@ if run_cell:
             
 
     ### plot data    
-    #plot_changes_upon_input2PE_neurons_new()
-    plt.figure()
-    plt.plot(pert_strength, alpha[:,:,0])
-    plt.ylim([0,1])
+    plot_neuromod_impact(pert_strength, alpha, xp, xs, xv)
+                                
     
-    plt.figure()
-    plt.plot(pert_strength, alpha[:,:,1])
-    plt.ylim([0,1])
-        
-    
-# %% Activate fractions of IN neurons in lower/higher PE circuit or both for a specified input statistics
-
-run_cell = False
-
-if run_cell:
-    
-    # define combo activation resolution
-    nums = 11
-    
-    # run through all combinations
-    for mfn_flag in ['10', '01', '11']: # choose mean-field network to simulate
-        
-        print('MFN code: ', mfn_flag)
-        print(' ')
-    
-        for column in range(3): # 0: both, 1: lower level PE circuit, 2: higher level PE circuit
-
-            for std_mean in [0, 1]: # uncertainty of environement [0, 1]
-                
-                n_std = 1 - std_mean # uncertainty of stimulus [1, 0]
-                    
[…]

### 2023-11-03 src/plot_data.py

@@ -148,7 +148,7 @@ def plot_nPE_pPE_activity_compare(P, S, nPE, pPE, fs = 6, lw = 1, ms = 1, ax1 = 
 
 def plot_standalone_colorbar(parent_axes, fs=6):
     
-    ax_cbar = parent_axes.inset_axes((0,1.5,1,0.13))
+    ax_cbar = parent_axes.inset_axes((-0.4,0.5,1,0.1))
     cbar = mpl.colorbar.ColorbarBase(ax_cbar, orientation='horizontal', cmap=cmap_sensory_prediction)
     
     cbar.outline.set_linewidth(0.1)
@@ -588,7 +588,7 @@ def plot_deviation_in_population_net(x, num_seeds, M_steady_state, V_steady_stat
 
     ax.set_xlabel(xlabel, fontsize=fs)
     if plt_ylabel:
-        ax.set_ylabel('Normalised error (%)', fontsize=fs)
+        ax.set_ylabel('Normalised \nerror (%)', fontsize=fs)
     
     ax.tick_params(size=2.0) 
     ax.tick_params(axis='both', labelsize=fs)
@@ -819,8 +819,57 @@ def illustrate_sensory_weight_variance(ax = None, fs=6):
     cbar.ax.tick_params(labelsize=fs)
     cbar.ax.tick_params(size=2.0)
     cbar.ax.yaxis.label.set_size(fs)
-                
+               
+    
+def plot_neuromod_impact(pert_strength, alpha, xp, xs, xv, figsize=(2,2.5), fs=6, lw=1, s=15, show_ylabel=True, show_xlabel = True,
+                            flg_plot_xlabel=True, flg_plot_bars = True, ax1 = None, ax2=None, highlight=True):
+    
+        if ax1 is None:
+            f, (ax1, ax2) = plt.subplots(2,1, tight_layout=True, figsize=figsize, sharey=True, sharex=True)
+           
+        markers = ['o','s','d']
+        
+        ### load data
+        file_for_data = '../results/data/neuromod/data_neuromod_' + str(xp) + '_' + str(xs) + '_' + str(xv) + '.pickle'
+        with open(file_for_data,'rb') as f:
+            [xp, xs, xv, pert_strength, alpha] = pickle.load(f) 
+        
+        # plot
+        for i, mfn_flag in enumerate(['10','01','11']): 
+            
+            ax2.axhline(alpha[0,i,0], color=cmap_sensory_prediction(alpha[0,i,0]), ls=':', zorder=0)
[…]

### 2023-11-03 records/manuscript_2023.tex
caption Figure 5 and supps

@@
-%\documentclass[10pt,a4paper,draft]{article}
-\documentclass[10pt,a4paper]{article}
+\documentclass[10pt,a4paper,draft]{article}
+%\documentclass[10pt,a4paper]{article}
 
@@
 {\Large
-\textbf\newline{Integrating feedforward and feedback inputs with prediction-error neurons}
+\textbf\newline{Knowing what you do not know: Estimating the uncertainty of feedforward and feedback inputs with prediction-error circuits}
 }
@@
 For instance, when you walk down some staircase in your fully lit basement, your brain might entirely rely on the feedforward (bottom-up) input your senses receive (Fig.
-\ref{fig:Fig_1}, left).
+\ref{fig:Fig_1}A, left).
 In contrast, when you walk down the same stairs in complete darkness, your brain might rely entirely on feedback (top-down) signals generated from a staircase model it has formed over previous experiences (Fig.
-\ref{fig:Fig_1}, middle).
+\ref{fig:Fig_1}A, middle).
 
@@
 In addition, it can only draw on a shaky prediction about what to expect (Fig.
-\ref{fig:Fig_1}, right).
+\ref{fig:Fig_1}A, right).
 
@@
 Multisensory integration is supported by several observations showing that xxx (XXX).
-It is conceivable that the same concepts can be employed for the weighting of sensory inputs and predictions thereof (XXX).
+It is conceivable that the same concepts can be employed for the weighting of sensory inputs and predictions thereof \citep{yon2021precision}.
 A central point in the weighting of inputs is the estimation of variances as a measure of uncertainty.
@@
 When those predictions are wrong, the resulting PEs allow the network to revise the model of the world, thereby ensuring that the predictions are more accurate (XXX).
-Experimental evidence suggests that these PEs may be represented in the activity of distinct groups of neurons, termed PE neurons (XXX).
-Moreover, these neurons may come in two types when excitatory neurons exhibit near-zero, spontaneous firing rates (XXX).
+Experimental evidence suggests that these PEs may be represented in the activity of distinct groups of neurons, termed PE neurons \citep{rao1999predictive, keller2018predictive}.
+Moreover, these neurons may come in two types when excitatory neurons exhibit near-zero, spontaneous firing rates \cite{keller2018predictive}.
 Negative PE (nPE) neurons mainly increase their activity when the prediction is \textit{stronger} than the sensory input, while positive PE (pPE) neurons mainly increase their activity when the prediction is \textit{weaker} than the sensory input.
-Indeed, it has been shown that excitatory neurons in layer 2/3 of rodent primary sensory areas can encode negative or positive PEs (XXX).
+Indeed, it has been shown that excitatory neurons in rodent primary sensory areas can encode negative or positive PEs \citep{keller2012sensorimotor, attinger2017visuomotor, jordan2020opposing, audette2021temporally}.
 
@@
 nPE neurons only increase their activity relative to a baseline when the sensory input is weaker than predicted, while pPE neurons only increase their activity relative to a baseline when the sensory input is stronger than predicted.
-Moreover, both nPE and pPE neurons remain at their baseline activity when the sensory input is fully predicted (XXX).
+Moreover, both nPE and pPE neurons remain at their baseline activity when the sensory input is fully predicted (Fig.
+\ref{fig:Fig_1}B).
 Assuming that the prediction equals the mean of the sensory stimulus, the PE neurons, hence, encode the deviation from the mean.
@@
 While the excitatory neurons are simulated as two coupled point compartments to emulate the soma and dendrites of elongated pyramidal cells, respectively, all inhibitory cell types were modeled as point neurons.
-The connectivity of and inputs to the network were chosen such that the excitatory (E) and inhibitory (I) pathways onto the pyramidal cells were balanced because it has been shown that this E/I balance is necessary for nPE and pPE neurons to emerge (XXX, see Methods).
+The connectivity of and inputs to the network were chosen such that the excitatory (E) and inhibitory (I) pathways onto the pyramidal cells were balanced because it has been shown that this E/I balance is necessary for nPE and pPE neurons to emerge \citep[][see Methods]{hertag2020learning, hertag2022prediction}.
 
@@
 \ref{fig:Fig_1}C).
-In accordance with XXX, we assume that the pPE neuron excites the memory neuron, while the nPE neuron inhibits this neuron (for instance, through lateral inhibition, here not modeled explicitly).
+In accordance with \cite{keller2018predictive}, we assume that the pPE neuron excites the memory neuron, while the nPE neuron inhibits this neuron (for instance, through lateral inhibition, here not modeled explicitly).
 The M neuron connects to the apical dendrites of the PE neurons and some of the interneurons (here, VIP and PV neurons, see Methods for more details).
@@
 The brain's flexibility and adaptability are not least because a plethora of neuromodulators influence the activity of neurons in a variety of ways (XXX).
-A prominent target of neuromodulatory inputs is inhibitory neurons (Cardin 2019, XXX).
+A prominent target of neuromodulatory inputs is inhibitory neurons \citep{cardin2019functional, XXX}.
 Moreover, distinct interneuron types are differently (in-)activated by those neuromodulators.
@@
 We, therefore, tested three different mean-field networks that differ with respect to the distribution of sensory inputs and predictions onto the interneurons, and the underlying connectivity.
-The commonality across those networks is that they exhibit an E/I balance of excitatory and inhibitory pathways onto the PE neurons (XXX).
+The commonality across those networks is that they exhibit an E/I balance of excitatory and inhibitory pathways onto the PE neurons \citep{hertag2022prediction}.
 Across the different mean-field networks tested, activating a SOM or VIP neuron individually forces the networks to weigh both inputs more equally.
@@
 	\centering
-    \includegraphics{../results/figures/final/Fig_5}
+    \includegraphics{../results/figures/final/Fig_5.pdf}
 \caption{\footnotesize{\bf Mechanisms underlying the contraction bias.\newline} 
-{\bf (A)} Illustration of the contraction bias.
-The estimated input is shifted towards the mean of the input distribution.
-Hence, a linear curve fitted to the data has a slope below 1 (left).
-The bias is biggest at the end of the stimulus distribution (right).
-{\bf (B)} Contraction bias in the model.
-Left: Example for two different stimulus variabilities.
-Right: As a consequence of the sensory weight, the slope decreases with stimulus variability (bias increases) and increases with trial variability (bias decreases).
-Stimulus statistics: XXX.
-{\bf (C)} The bias is independent of trial variability when the stimulus variability is zero (left).
-Equally, the bias is independent of the stimulus variability when the trial variability is zero (right).
-Stimulus statistics: XXX.
-{\bf (D)} The slope depends on the trial duration.
-If the sensory weight is 1, the slope is independent of the trail duration and the bias vanishes (dashed-dotted line).
-If the sensory weight is 0, the slope depends on the trial duration and only reaches 1 if the trial duration approaches infinity.
-{\bf (E)} To ensure a larger bias for stimuli drawn from the upper end of the stimulus distribution than from the lower end, scalar variability as observed experimentally is needed.
+{\bf (A)} Contraction bias in the model for two different stimulus uncertainties depicted in the inset.
+Bias is defined as the weighted output minus the stimulus mean.
+The absolute value of the slope (see linear fit) is a measure of the bias.
+The larger the slope, the larger the bias.
+{\bf (B)}  As a consequence of the sensory weight, the slope increases with stimulus variability (bias increases) and decreases with trial variability (bias decreases).
+{\bf (C)} Bias is independent of the trial variability when the stimulus variability is zero.
+{\bf (D)} Bias is independent of the stimulus variability when the trial variability is zero.
+{\bf (E)} The slope depends on the trial duration.
 }
@@
 Our work has been driven by the puzzling question of how the brain integrates top-down feedback predictions with the sensory feedforward bottom-up inputs it constantly receives during behavior.
-This task may be particularly challenging when the prediction and the sensory information differ \ref{han2023behavior}.
+This task may be particularly challenging when the prediction and the sensory information differ \citep{han2023behavior}.
 Conflicting information may be caused by noise in the sensory feedforward inputs or by changes in the environment that could not be predicted.
-A prominent hypothesis is that how much we rely on our predictions and new sensory evidence is determined by an intricate balance between both, based on how reliable they are \ref[see e.g.][]{yon2021precision}.
+A prominent hypothesis is that how much we rely on our predictions and new sensory evidence is determined by an intricate balance between both, based on how reliable they are \citep[see e.g.][]{yon2021precision}.
 
 This idea is consistent with Bayesian theories on the optimal integration of multiple sensory cues (aka multisensory integration).
-\ref{XX ernst and banks 2002} showed that to estimate the height of a bar humans combine visual and haptic information in a fashion that minimizes the variance of the final estimate.
+\citep{ernst2002humans} showed that to estimate the height of a bar humans combine visual and haptic information in a fashion that minimizes the variance of the final estimate.
 XXX name other examples in human and other species XXX.
@@
 
-While multisensory integration is concerned with the weighting of sensory cues coming from different modalities, it is conceivable that the same ideas hold for top-down predictions and bottom-up sensory inputs that are combined in cortical circuits \ref{yon2021precision}.
-Indeed, there is evidence that the brain can choose to rely more on top-down or bottom-up signals \ref{pakan2018impact, han2023behavior}.
-Following the same ideas from Bayesian multisensory integration, the brain must hence have mechanisms to track both the reliability of the sensory signals and the predictions thereof \ref{yon2021precision}.
+While multisensory integration is concerned with the weighting of sensory cues coming from different modalities, it is conceivable that the same ideas hold for top-down predictions and bottom-up sensory inputs that are combined in cortical circuits \citep{yon2021precision}.
+Indeed, there is evidence that the brain can choose to rely more on top-down or bottom-up signals \citep{pakan2018impact, han2023behavior}.
+Following the same ideas from Bayesian multisensory integration, the brain must hence have mechanisms to track both the reliability of the sensory signals and the predictions thereof \citep{yon2021precision}.
 XXX We suggested that xxx Here summarise your contributions XXX
@@
 With this architecture in place, we show that we rely more strongly on our internal signals when the perceived sensory cues are noisier than the predictions, either because the environment is stable but highly noisy or because the environment changes quickly.
-This implies that revising our inner model of the world, that is, learning from PEs, should be suppressed if the sensory noise is high or the environment switches rapidly \ref{herzfeld2014memory}.
+This implies that revising our inner model of the world, that is, learning from PEs, should be suppressed if the sensory noise is high or the environment switches rapidly \citep{herzfeld2014memory}.
 Moreover, our work suggests that when experimental trials are short, a neural network can better represent the external world by relying on predictions.
@@
 
-Furthermore, we show that the weighting of sensory inputs and predictions can be biased through neuromodulators, as has been suggested before \ref{XXX}.
-In our model, those modulatory signals act through interneurons \ref{cardin2019functional} whose activities increase in the presence of neuromodulators.
+Furthermore, we show that the weighting of sensory inputs and predictions can be biased through neuromodulators, as has been suggested before \citep{XXX}.
+In our model, those modulatory signals act through interneurons \citep{cardin2019functional} whose activities increase in the presence of neuromodulators.
 When PV neuron activity increases, the network weighs predictions stronger than without modulation.
@@
 However, the equalising effect is less robust and seems to depend on the network specifics (Fig.
-\ref{XX}A).
+\ref{fig:Fig_4}A).
 Interestingly, when SOM and VIP neuron activity is modulated to the same degree, the weighting remains unaffected, suggesting that the individual contributions cancel (Fig.
-\ref{XX}B).
+\ref{fig:Fig_4}B).
 We show that these findings can be best explained by changes in the baseline and gain of PE neurons as a result of the modulation of interneuron activity (Fig.
-\ref{XXX}D).
+\ref{fig:Fig_4}D).
 These results can be tested experimentally by optogenetically or pharmacologically stimulating specific interneuron types.
@@
 On the other hand, if the trial variance is zero (perfect prediction-driven network regime), the neural contraction bias is independent of the stimulus variability and a function of the trial duration.
-However, we note that in our model, we only consider the neural representation of the stimulus and do not account for other sources of noise, like execution noise, that surely impacts the contraction bias observed in behavioral studies \ref{XXX}.
+However, we note that in our model, we only consider the neural representation of the stimulus and do not account for other sources of noise, like execution noise, that surely impacts the contraction bias observed in behavioral studies \citep{XXX}.
 
@@
 In our model, we assumed that there are dedicated neurons that encode the variance of the feedforward sensory inputs and the prediction.
-This assumption is consistent with the idea that neurons explicitly encode in their activity the parameters (for instance, mean or variance) of a probability distribution (O’Neill \& Schultz, 2010, XXX).
+This assumption is consistent with the idea that neurons explicitly encode in their activity the parameters (for instance, mean or variance) of a probability distribution \citep{ONeill Schultz, 2010, XXX}.
 However, how variances of signals are represented in the brain is still not comprehensively understood and alternative ideas have been put forward.
-For instance, it is conceivable that the variance is encoded in a population of neurons, each differently tuned to a specific parameter (Knill \& Pouget, 2004; Ma et al., 2006, 2008).
+For instance, it is conceivable that the variance is encoded in a population of neurons, each differently tuned to a specific parameter \citep{knill2004bayesian, Ma et al., 2006, 2008}.
 The neurons' activities represent how close the sensory input is to the preferred (predicted) input of each neuron.
@@
 There has been evidence that indeed (population of) neurons can encode (un-)certainty.
-Neurons in the pulvinar (Komura et al., 2013), the parietal cortex (Kiani and Shadlen, 2009) or the frontal cortex (Kepecs et al., 2008, Middlebrooks and Sommer, 2012) encode the uncertainty of perceptual decisions \ref{masset2020behavior}.
-In addition, it has been shown that neurons in the orbitofrontal cortex of rats signal confidence in decisions \ref{masset2020behavior}.
-Moreover, neural correlates of (expected) uncertainty have been found in regions of the prefrontal cortex 43 \ref{soltani2019adaptive} , the rat orbitofrontal cortex [52,53\ref{soltani2019adaptive}] or the dorsal striatum in monkeys78 \ref{soltani2019adaptive}.
-Interestingly, the accuracy of memory recalls is represented (?) in single neurons of the human parietal and temporal lobes (Rutishauser et al., 2015, 2018) … in \ref{masset2020behavior}.
+Neurons in the pulvinar (Komura et al., 2013), the parietal cortex (Kiani and Shadlen, 2009) or the frontal cortex (Kepecs et al., 2008, Middlebrooks and Sommer, 2012) encode the uncertainty of perceptual decisions \citep{masset2020behavior}.
+In addition, it has been shown that neurons in the orbitofrontal cortex of rats signal confidence in decisions \citep{masset2020behavior}.
+Moreover, neural correlates of (expected) uncertainty have been found in regions of the prefrontal cortex 43 \citep{soltani2019adaptive} , the rat orbitofrontal cortex [52,53\citep{soltani2019adaptive}] or the dorsal striatum in monkeys78 \citep{soltani2019adaptive}.
+Interestingly, the accuracy of memory recalls is represented (?) in single neurons of the human parietal and temporal lobes (Rutishauser et al., 2015, 2018) … in \citep{masset2020behavior}.
 
 In our model, we assume a hierarchy of predictions that are locally computed in memory neurons.
-These memory neurons are consistent with the idea of internal representation neurons hypothesised in predictive processing theories \ref{bastos2012canonical, keller2018predictive}.
-While it has been hypothesised that these internal representation neurons might be found in deeper L5 neurons \ref{bastos2012canonical, heindorf2022reduction}, there is also evidence that a group of excitatory L2/3 neurons integrates over negative and positive prediction errors \ref{o2022prediction}.
+These memory neurons are consistent with the idea of internal representation neurons hypothesised in predictive processing theories \citep{bastos2012canonical, keller2018predictive}.
+While it has been hypothesised that these internal representation neurons might be found in deeper L5 neurons \citep{bastos2012canonical, heindorf2022reduction}, there is also evidence that a group of excitatory L2/3 neurons integrates over negative and positive prediction errors \citep{o2022prediction}.
 
 The core hypothesis of our model is the presence of sensory PE neurons.
-Those neurons have been found in different cortical areas in various species \ref{eliades2008neural, keller2009neural, ayaz2019layer, audette2021temporally}.
-Moreover, while first only hypothesized theoretically \ref{rao1999predictive}, the presence of two types of PE neurons, the negative and positive PE neurons, has been confirmed in a number of recent studies \ref{keller2012sensorimotor, attinger2017visuomotor, jordan2020opposing, audette2021temporally, Zmarz2016XXX?necessary}.
-In our model, the nPE neuron inhibits the memory neuron while the pPE neuron excites it, in line with \ref{keller2018predictive}.
+Those neurons have been found in different cortical areas in various species \citep{eliades2008neural, keller2009neural, ayaz2019layer, audette2021temporally}.
+Moreover, while first only hypothesized theoretically \ref{rao1999predictive}, the presence of two types of PE neurons, the negative and positive PE neurons, has been confirmed in a number of recent studies \citep{keller2012sensorimotor, attinger2017visuomotor, jordan2020opposing, audette2021temporally, Zmarz2016XXX?necessary}.
+In our model, the nPE neuron inhibits the memory neuron while the pPE neuron excites it, in line with \citep{keller2018predictive}.
 The weights from those PE neurons onto the M neuron are larger for the lower- than the higher PE circuit, so that the lower-order M neuron evolves faster than the higher-order counterpart.
@@
 \ref{XXX}).
-It has been hypothesized that symptoms in psychiatric diseases may be a result of an erroneous uncertainty estimation \ref{yon2021precision}.
+It has been hypothesized that symptoms in psychiatric diseases may be a result of an erroneous uncertainty estimation \citep{yon2021precision}.
 For instance, hallucinations may be the result of an underestimation of the expectation uncertainty or an overestimation of the sensory uncertainty.
-Conversely, a fixation on a stable environment even when the sensory cues indicate a switch in the environment, may be the consequence of an overestimation of the expectation uncertainty or an underestimation of the sensory uncertainty \ref{yon2021precision}.
+Conversely, a fixation on a stable environment even when the sensory cues indicate a switch in the environment, may be the consequence of an overestimation of the expectation uncertainty or an underestimation of the sensory uncertainty \citep{yon2021precision}.
 
@@
 
-A popular hypothesis is that neuromodulators shape the weighting of sensory inputs and predictions thereof \ref{yon2021precision}.
+A popular hypothesis is that neuromodulators shape the weighting of sensory inputs and predictions thereof \citep{yon2021precision}.
 Theoretical work by \ref{You and Dayan} suggests that acetylcholine (ACh) correlates with expected uncertainty, while noradrenaline (NA) correlates with unexpected uncertainty.
 Expected uncertainty is usually considered to be related to known cue-outcome unreliabilities.
-In contrast, unexpected uncertainty relates to the changes in the environment that produce large PEs outside the expected range of uncertainties \ref{You and Dayan}.
+In contrast, unexpected uncertainty relates to the changes in the environment that produce large PEs outside the expected range of uncertainties \citep{You and Dayan}.
 While in our network the stimulus and trail variability can only be loosely interpreted as 'expected' and 'unexpected' uncertainty, we want to XXX broader context and highlight differences XXX
@@
 
-It is assumed that NA increases in more volatile environments and enhances bottom-up processes [20,21 in \ref{lawson2021computational}, yon2021precision].
-In line with this idea, NA blockade impairs cognitive flexibility [24,25\ref{lawson2021computational}].
-In recent work by \ref{lawson2021computational}, it has been shown that humans receiving propranolol (blocking NA) rely more strongly on their expectations and are slower to update these predictions despite new sensory evidence \ref{yon2021precision}.
+It is assumed that NA increases in more volatile environments and enhances bottom-up processes [20,21 in \citep{lawson2021computational}, yon2021precision].
+In line with this idea, NA blockade impairs cognitive flexibility [24,25\citep{lawson2021computational}].
+In recent work by \citep{lawson2021computational}, it has been shown that humans receiving propranolol (blocking NA) rely more strongly on their expectations and are slower to update these predictions despite new sensory evidence \citep{yon2021precision}.
 A major target for noradrenergic inputs is SOM neurons whose activity increases in the presence of NA (reviewed in Urban-Ciecko 2016: Fanselow, 2008; Urban-Ciecko 2015; Chen 2015; Xiang 1998; reviewed in Hattori 2017: Paspalas 1999; reviewed in Swanson in Maffei 2019: Kawaguchi and Shindou, 1998).
@@
 Neuromodulatory systems have been suggested to gate plasticity \ref{XXX}.
-In a recent study by \ref{jordan2023locus}, locus coeruleus (LC) axon activity is shown to correlate with the magnitude of unsigned visuomotor prediction errors.
-The authors hypothesize that LC output modulates the learning rate at which the internal model evolves \ref{jordan2023locus}.
-In our model, we do not consider precision-weighted PEs (but see \ref{KathaXXX}).
+In a recent study by \citep{jordan2023locus}, locus coeruleus (LC) axon activity is shown to correlate with the magnitude of unsigned visuomotor prediction errors.
+The authors hypothesize that LC output modulates the learning rate at which the internal model evolves \citep{jordan2023locus}.
+In our model, we do not consider precision-weighted PEs (but see \citep{KathaXXX}).
 Hence, a sensible extension to our work would be to adjust the weights from PE neurons onto the M neurons by a function of the stimulus and prediction uncertainty, respectively.
@@
 However, in the light of evidence showing that the integration of feedforward and feedback inputs is species- and modality-dependent, it is conceivable that a plethora of neural mechanisms are used in neural circuits.
-For multisensory integration, \ref{wong2023computational} showed that in Drosophila larva the chosen cue-combination strategy varies depending on the type of sensory information available.
+For multisensory integration, \citep{wong2023computational} showed that in Drosophila larva the chosen cue-combination strategy varies depending on the type of sensory information available.
 Also, humans put typically more weight on visual than auditory cues ((Battaglia et al., 2003; Burr \& Alais, 2004).
-… in \ref{o2012can}), but trust vestibular information more than visual information about head direction (Butler et al., 2010 \ref{o2012can}), a finding also observed for monkeys (Angelaki et al., 2009 \ref{o2012can}).
+… in \citep{o2012can}), but trust vestibular information more than visual information about head direction (Butler et al., 2010 \citep{o2012can}), a finding also observed for monkeys (Angelaki et al., 2009 \citep{o2012can}).
 Moreover, Summerfield et al.
-(2011) \ref{o2012can} showed that humans diverge from an optimal Bayesian strategy in very volatile environments and act according to their experience in the last trial.
-Moreover, it has been suggested that the brain may use different strategies to combine signals depending on the task demands \ref{o2012can}.
+(2011) \citep{o2012can} showed that humans diverge from an optimal Bayesian strategy in very volatile environments and act according to their experience in the last trial.
+Moreover, it has been suggested that the brain may use different strategies to combine signals depending on the task demands \citep{o2012can}.
 While these behavioral results cannot speak to the underlying circuit mechanisms, it is conceivable that neural implementations for the integration of feedforward and feedback inputs may also vary.
@@
 
-A number of normative models have been proposed for state estimation and prediction under uncertainty \ref{soltani2019adaptive}, ranging from the classical Kalman filter to more recent models like \textit{Bayes Factor Surprise} \ref{liakoni2021learning}.
+A number of normative models have been proposed for state estimation and prediction under uncertainty \citep{soltani2019adaptive}, ranging from the classical Kalman filter to more recent models like \textit{Bayes Factor Surprise} \citep{liakoni2021learning}.
 For instance, Bayes factor surprise formularizes the trade-off between integrating new observations in an existing belief system and resetting this belief system with the new evidence.
@@
 In more recent years, normative model have been squared with biological constraints.
-For instance, in a paper by \ref{kutschireiter2023bayesian}, uncertainty could be integrated into a ring attractor model encoding head direction.
+For instance, in a paper by \citep{kutschireiter2023bayesian}, uncertainty could be integrated into a ring attractor model encoding head direction.
 This \textit{Bayesian ring attractor} model encodes uncertainty in the amplitude of the network activity and matches the performance of a circular Kalman filter when the recurrent connections are tuned appropriately.
@@
 
+\begin{figure}[!h]
+	\centering
+    \includegraphics{../results/figures/final/Fig_5_S1}% [width=1\linewidth]
+\caption{\footnotesize{\bf Including scalar variability in the model \newline}  
+When scalar variability is included, that is, the stimulus standard deviation depends linearly on the stimulus mean, the bias is larger for stimuli drawn from the upper end of the stimulus distribution than from the lower end.
+}
+\label{fig:Fig_5_S1}
+\end{figure}
+
 \end{document}

### 2023-11-10 records/manuscript_2023.tex
change in Fig. 4, polish all sectons except Discussion

@@
 In line with previous hypotheses, we demonstrate that neural circuits rely strongly on the feedback predictions at the beginning of a new stimulus, and if the environment is stable and the sensory cues are noisy.
-Furthermore, we investigate the causes of a pathological weighting of feedforward and feedback inputs.
+In addition, we investigate the mechanisms underlying a pathological weighting of feedforward and feedback inputs.
 In our network, the uncertainty estimation, and, hence, how much we rely on predictions, can be biased by perturbing the intricate interplay of different inhibitory interneurons.
@@
 To survive in an ever-changing environment, animals must flexibly adapt their behavior based on previously encoded and novel information.
-This context-dependency must also be reflected in the information processing of neural networks underlying intelligent behavior.
-For instance, when you walk down some staircase in your fully lit basement, your brain might entirely rely on the feedforward (bottom-up) input your senses receive (Fig.
+This adaptation is reflected in the information processing of neural networks underlying context-dependent behavior.
+For instance, when you walk down an unknown staircase in a fully lit basement, your brain might entirely rely on the feedforward (bottom-up) input your senses receive (Fig.
 \ref{fig:Fig_1}A, left).
@@
 \ref{fig:Fig_1}A, middle).
-
-The importance of these feedback inputs has been emphasized by observations showing that top-down projections outnumber feedforward connections (XXX) and that they modulate (XXX) or even entirely drive (XXX) neuron activity.
 But how do neural networks switch between a feedforward-dominated and a feedback-dominated processing mode?
-And how do neural networks combine both input streams wisely?
+And how do neural networks in the brain combine both input streams wisely?
 For instance, if you hike down an unexplored mountain in very foggy conditions, your brain receives unreliable visual information.
@@
 Multisensory integration is supported by several observations showing that xxx (XXX).
-It is conceivable that the same concepts can be employed for the weighting of sensory inputs and predictions thereof \citep{yon2021precision}.
-A central point in the weighting of inputs is the estimation of variances as a measure of uncertainty.
+Here, we propose that the same concepts could be employed for the weighting of sensory inputs and predictions thereof \citep{yon2021precision}.
+A central point in the weighting of inputs is the estimation of their variances as a measure of uncertainty.
 However, how the variance of both the sensory input and the prediction can be computed on the circuit level is not resolved yet.
 
-We hypothesised that prediction error (PE) neurons provide the basis for the neural computation of variances.
+We hypothesized that prediction error (PE) neurons provide the basis for the neural computation of variances.
 PEs are an integral part of the theory of predictive processing which states that the brain constantly compares incoming sensory information with predictions.
-When those predictions are wrong, the resulting PEs allow the network to revise the model of the world, thereby ensuring that the predictions are more accurate (XXX).
-Experimental evidence suggests that these PEs may be represented in the activity of distinct groups of neurons, termed PE neurons \citep{rao1999predictive, keller2018predictive}.
-Moreover, these neurons may come in two types when excitatory neurons exhibit near-zero, spontaneous firing rates \cite{keller2018predictive}.
-Negative PE (nPE) neurons mainly increase their activity when the prediction is \textit{stronger} than the sensory input, while positive PE (pPE) neurons mainly increase their activity when the prediction is \textit{weaker} than the sensory input.
+If those predictions are wrong, the resulting PEs allow the network to revise the model of the world, thereby ensuring that the predictions become more accurate (XXX).
+Experimental evidence suggests that these PEs may be represented in the activity of distinct groups of neurons, termed PE neurons \citep{eliades2008neural, keller2009neural, ayaz2019layer, audette2021temporally}.
+Moreover, these neurons may come in two types when excitatory neurons exhibit near-zero, spontaneous firing rates \cite{rao1999predictive, keller2018predictive}.
+Negative PE (nPE) neurons only increase their activity when the prediction is \textit{stronger} than the sensory input, while positive PE (pPE) neurons only increase their activity when the prediction is \textit{weaker} than the sensory input.
 Indeed, it has been shown that excitatory neurons in rodent primary sensory areas can encode negative or positive PEs \citep{keller2012sensorimotor, attinger2017visuomotor, jordan2020opposing, audette2021temporally}.
@@
 Furthermore, we suggest a network model with a hierarchy of PE circuits to estimate the variance of the prediction, in addition to the variance of the sensory inputs.
-In line with multisensory integration, predictions are weighted more strongly than the sensory stimuli when the environment is stable (that is, predictable) but the sensory inputs are noisy.
-Moreover, we show that predictions are integrated more strongly after a change in the environment, even when the new sensory stimulus is reliable.
+We show that in line with the ideas of multisensory integration, predictions are weighted more strongly than the sensory stimuli when the environment is stable (that is, predictable) and the sensory inputs are noisy.
+Moreover, we find that predictions are taken into account more at the beginning of a new trial than at the end, especially when the new sensory stimulus is reliable.
 In addition, we unravel the mechanisms underlying a neuromodulator-induced shift in the weighting of sensory inputs and predictions.
-In our model, these neuromodulators activate groups of inhibitory neurons like parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (XXX).
-In a computational model, these interneurons have been shown to establish a multi-pathway balance of excitation and inhibition that is the basis for nPE and pPE neurons (XXX).
-By breaking this balance, the excitatory neurons change their baseline firing rate and gain, leading to a biased variance estimation.
-Finally, we show that this weighting can be understood as the neural manifestation of the contraction bias (XXX).
+In our model, these neuromodulators activate groups of inhibitory neurons such as parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (XXX).
+These interneurons have been suggested to establish a multi-pathway balance of excitation and inhibition that is the basis for nPE and pPE neurons \citep{hertag2020learning, hertag2022prediction}.
+By breaking this balance, the PE neurons change their baseline firing rate and gain, leading to a biased variance estimation.
+Finally, we show that this weighting can be understood as a neural manifestation of the contraction bias (XXX).
 
@@
 
-\subsection*{nPE and pPE neurons as the basis for estimating mean and variance of sensory stimuli}
-%
-We hypothesise that the distinct response patterns of negative and positive prediction-error (nPE/pPE) neurons represent the backbone for estimating the mean and the variance of sensory stimuli.
-nPE neurons only increase their activity relative to a baseline when the sensory input is weaker than predicted, while pPE neurons only increase their activity relative to a baseline when the sensory input is stronger than predicted.
-Moreover, both nPE and pPE neurons remain at their baseline activity when the sensory input is fully predicted (Fig.
+\subsection*{Prediction-error neurons as the basis for estimating the mean and variance of sensory stimuli}
+%
+We hypothesize that the distinct response patterns of negative and positive prediction-error (nPE/pPE) neurons act as a backbone for estimating the mean and the variance of sensory stimuli.
+An nPE neuron only increases its activity relative to a baseline when the sensory input is weaker than predicted, while a pPE neuron only increases its activity relative to a baseline when the sensory input is stronger than predicted.
+Moreover, both nPE and pPE neurons remain at their baseline activities when the sensory input is fully predicted (Fig.
 \ref{fig:Fig_1}B).
-Assuming that the prediction equals the mean of the sensory stimulus, the PE neurons, hence, encode the deviation from the mean.
-Thus, the squared sum of nPE and pPE neuron activity represents the variance of the feedforward input.
+If the prediction equals the mean of the sensory stimulus, the PE neurons, hence, encode the deviation from the mean.
+Thus, the squared sum of nPE and pPE neuron activity represents the variance of the feedforward input (provided that the PE neurons are silent without sensory stimulation).
 %
@@
 
-To test our hypothesis, we studied a rate-based mean-field network the core of which is a prediction-error (PE) circuit with excitatory nPE and pPE neurons, as well as inhibitory parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (Fig.
+To test our hypothesis, we study a rate-based mean-field network.
+The core network is a prediction-error (PE) circuit with excitatory nPE and pPE neurons, as well as inhibitory parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (Fig.
 \ref{fig:Fig_1}B).
 While the excitatory neurons are simulated as two coupled point compartments to emulate the soma and dendrites of elongated pyramidal cells, respectively, all inhibitory cell types were modeled as point neurons.
-The connectivity of and inputs to the network were chosen such that the excitatory (E) and inhibitory (I) pathways onto the pyramidal cells were balanced because it has been shown that this E/I balance is necessary for nPE and pPE neurons to emerge \citep[][see Methods]{hertag2020learning, hertag2022prediction}.
-
-In addition to this core circuit, we model a memory (M) neuron that perfectly integrates the activity of the PE neurons (Fig.
+The connectivity of and inputs to the network were chosen such that the excitatory (E) and inhibitory (I) pathways onto the pyramidal cells were partially balanced.
+This balance that is only temporarily broken during mismatches has been shown to be necessary for nPE and pPE neurons to emerge \citep[][see Methods]{hertag2020learning, hertag2022prediction}.
+
+We assume that this core circuit is shaped by feedback connections \ref{XXX} that have been hypothesized to carry information about expectations or predictions \ref{XXX}.
+To account for predictions, we model a memory (M) neuron that integrates the activity of the PE neurons (Fig.
 \ref{fig:Fig_1}C).
-In accordance with \cite{keller2018predictive}, we assume that the pPE neuron excites the memory neuron, while the nPE neuron inhibits this neuron (for instance, through lateral inhibition, here not modeled explicitly).
-The M neuron connects to the apical dendrites of the PE neurons and some of the interneurons (here, VIP and PV neurons, see Methods for more details).
-In this network, the M neuron serves as a prediction that is dynamically updated when new sensory information is available.
-We furthermore simulate a downstream neuron (termed V neuron), modeled as a leaky integrator with a squared activation function, that receives excitatory output synapses from the PE neurons.
+Following \cite{keller2018predictive}, we assume that the pPE neuron excites the memory neuron, while the nPE neuron inhibits this neuron (for instance, through lateral inhibition, here not modeled explicitly) [XXX KW: look into Anna V paper, opposing effects tested biorxiv great ref XXX].
+Because feedback connections are shown to target the apical dendrites of pyramidal cells \citep{LarkumXXX} and interneurons located in superficial layers of the cortex \citep{XXX}, the M neuron makes connections with the dendritic compartment of the PE neurons and some of the interneurons (here, VIP and PV neurons, see Methods for more details).
+We, furthermore, simulate a downstream neuron (termed V neuron), modeled as a leaky integrator with a quadratic activation function, that receives excitatory synapses from the PE neurons.
 Hence, in this setting, the V neuron encodes the variance of the sensory stimuli (Fig.
@@
 
-To show that this network can indeed represent mean and variance in the respective neurons, we stimulate it with a sequence of step-wise constant inputs drawn from a uniform distribution (Fig.
-\ref{fig:Fig_2}A), assuming that the sensory stimulus varies over time.
+To show that this network can indeed represent the mean and the variance in the respective neurons, we stimulate it with a sequence of step-wise constant inputs drawn from a uniform distribution (Fig.
+\ref{fig:Fig_2}A).
+We, hence, assume that the sensory stimulus varies over time.
 In line with the distinct response patterns for nPE and pPE neurons, these neurons change only slightly with increasing stimulus mean but increase strongly with input variance (Fig.
 \ref{fig:Fig_2}B).
-This is in contrast to the three interneurons that strongly increase with stimulus mean while they only moderately increase with stimulus variance (Fig.
+In contrast, the three interneurons strongly increase with stimulus mean and only moderately increase with stimulus variance (Fig.
 \ref{fig:Fig_2}C).
-The activity of the memory neuron M gradually approaches the mean of the sensory inputs (Fig.
-\ref{fig:Fig_2}D, middle), while the activity of the V neuron approaches the variance of the inputs (Fig.
+The activity of the M neuron gradually approaches the mean of the sensory inputs (Fig.
+\ref{fig:Fig_2}D, middle), while the activity of the V neuron approaches the variance of those inputs (Fig.
 \ref{fig:Fig_2}E, middle).
-This is true for a wide range of input statistics (Fig.
+We show that this holds for a wide range of input statistics (Fig.
 \ref{fig:Fig_2}D-E, right) and input distributions (Fig.
 \ref{fig:Fig_2_S1}).
-Small deviations from the true mean occur mainly for larger input variances, while the estimated variance is fairly independent of the input statistics tested.
-
-XXX coming soon: Paragraph on network beyond mean-field XXX
-
-In summary, nPE and pPE neurons can be the basis to estimate the mean and the variance of sensory stimuli that vary over time.
+Small deviations from the true mean occur mainly for large input variances, while the estimated variance is fairly independent of the input statistics tested.
+
+We verified our results in a heterogeneous network in which a population of neurons represents each neuron type of the PE circuit, and the synaptic connection strengths from each PE neuron onto the M and V neuron are different (see Methods, Fig.
+\ref{fig:Fig_2_S2}A).
+As before, the network can correctly estimate the mean and the variance of the sensory stimuli (Fig.
+\ref{fig:Fig_2_S2}B).
+Furthermore, we show that the errors with which the M and V neurons encode the stimulus statistics are independent of uncorrelated modulations of those connection strengths (Fig.
+\ref{fig:Fig_2_S2}C) and the sparsity of the network (Fig.
+\ref{fig:Fig_2_S2}E).
+When all connection strengths are collectively shifted to higher values, the error increases for the V neuron, while it remains unaffected for the M neuron.
+
+While our mean-field network was designed to track the mean and the variance of stimuli that vary in time, we reasoned that the same principles apply to stimuli that vary across space.
+To show that, we simulated a population network that consists of unconnected replicates of the mean-field network described above (Fig.
+\ref{fig:Fig_2_S3}A).
+Each mean-field network receives a short, constant input from a different part of the receptive field.
+If the connection strengths from the PE neurons to the M and V neurons are adjusted accordingly (see Methods), the network correctly estimates the stimulus average and spatial uncertainty.
+(Fig.
+\ref{fig:Fig_2_S3}B-C).
+
+In summary, nPE and pPE neurons can serve as a basis to estimate the mean and the variance of sensory stimuli which vary over time and space.
 
@@
 %
-Following the ideas of Bayesian multisensory integration (XXX), the weighting of sensory stimuli and predictions thereof would require knowledge of their variances.
+Following the ideas of Bayesian multisensory integration (XXX), the weighting of sensory stimuli and predictions would require knowledge about their uncertainties.
 As we have shown in the previous section, the variance of the sensory stimulus can be estimated using PE neurons.
-We hypothesise that the same principles apply to computing the variance of the prediction.
-Hence, we augment the network with a \textit{higher} PE circuit that receives output synapses from the M neuron of the \textit{lower} PE circuit (Fig.
+We hypothesize that the same principles apply to computing the variance of the prediction.
+Hence, we augment the network with a \textit{higher} PE circuit that receives feedforward synapses from the M neuron of the \textit{lower} PE circuit (Fig.
 \ref{fig:Fig_1}D).
-Both subnetworks are modeled the same, except that the M neuron in the higher PE circuit evolves more slowly than the one in the lower PE circuit.
+Both subnetworks are identical except for the M neuron in the higher PE circuit which is modeled with slower dynamics than the one in the lower PE circuit.
 %
@@
 
-To test the network's ability to estimate the variances correctly, we stimulated the network with a sequence of inputs.
-In each trial one stimulus is shown to the network.
-To account for the stimulus variance, each stimulus is composed of \textit{n} constant values drawn from a normal distribution with mean $\mu_\mathrm{stim}$ and variance $\sigma_\mathrm{stim}^2$, and presented one after the other.
-To account for potential changes in the environment, in each trial, we draw $\mu_\mathrm{stim}$ from a uniform distribution (Fig.
+To test the network's ability to estimate the variances correctly, we stimulated the network with a sequence of inputs whose mean can vary from trial to trial.
+More precisely, in each trial, the network is presented with a stimulus that is composed of \textit{n} constant, consecutive values drawn from a normal distribution.
+The variance of this distribution represents the stimulus noise.
+To account for potential changes in the environment, we draw the stimulus mean from a uniform distribution (Fig.
 \ref{fig:Fig_3}A).
-Hence, the inputs change on two different time scales, with stimulus variability (faster time scale) and trial variability (slower time scale).
-
-As expected, the neurons' activity increase for both stimulus and trial variances (Fig.
-\ref{fig:Fig_3}B).
-While the neurons in the lower PE circuit increase more strongly with stimulus variability, the neurons in the higher PE circuit increase more strongly with trial variability, indicating that the different subnetworks process different aspects of the inputs.
-We first consider two limit cases.
-In the first limit case, a different but low-variance stimulus is presented in each trial (Fig.
+Hence, the inputs change on two different time scales, where the stimulus variability has a faster time scale than the trial variability.
+
+Following the formalism of multisensory integration (XXX), we assume that the network's output is a weighted sum of the feedforward sensory input and the feedback prediction.
+The weights assigned to each input stream are functions of the uncertainties, that is, the activities of the V neurons.
+The sensory weight captures, hence, how much the network relies on the sensory input (Fig.
+\ref{fig:Fig_2}B).
+To test our network, we first consider two limit cases.
+In the first limit case, a different, low-variance stimulus is presented in each trial (Fig.
 \ref{fig:Fig_3}C, left).
-In line with the ideas of multisensory integration (XXX), the network should therefore follow the sensory inputs closely and ignore the predictions.
+According to the theory, the network should follow the sensory inputs closely and ignore the predictions.
 When we arithmetically calculate the weighted output (Fig.
-\ref{fig:Fig_3}C, middle) based on the feedforward and feedback inputs, and the sensory weight (Fig.
-\ref{fig:Fig_3}C, right), the network correctly represents mostly the sensory input (for more details, see Methods).
-In the second limit case, the same but high-variance stimulus is presented in each trial (Fig.
+\ref{fig:Fig_3}C, middle) and the sensory weight (Fig.
+\ref{fig:Fig_3}C, right), the network indeed shows a clear preference for the sensory input.
+In the second limit case, the same, high-variance stimulus is presented in each trial (Fig.
 \ref{fig:Fig_3}D, left).
@@
 
-In a next step, to validate the network responses more broadly, we systematically varied the trial and stimulus variability independently.
-If both variances are similar, the sensory weight approaches \textit{0.5}, reflecting equal contribution of sensory inputs and predictions to the weighted output.
+To validate the network responses fully, we systematically varied the trial and stimulus variability independently.
+If both variances are similar, the sensory weight approaches \textit{0.5}, reflecting equal contribution of the sensory input and the prediction to the weighted output.
 Only if both variances are zero, the network represents the sensory input perfectly.
-In line with the limit case examples above, if the stimulus variance is larger than the trial variance, the network weights the prediction more strongly than the sensory input.
-This is reversed if the stimulus variance is smaller than the trial variance (Fig.
+In line with the limit case examples above, if the stimulus variance is larger than the trial variance, the network weights the prediction more strongly than the sensory input (Fig.
 \ref{fig:Fig_3}E).
-Because the network dynamically estimates the mean and variances of the sensory input and the prediction, the weighted output and the sensory weight changes accordingly when the input statistics changes (Fig.
+Because the network dynamically estimates the sensory and prediction uncertainty, the sensory weight changes when the input statistics shifts (Fig.
 \ref{fig:Fig_3_S1}).
 
-The first limit case (Fig.
-\ref{fig:Fig_3}C) shows that even in a sensory-driven input regime, the prediction is weighted more at the beginning of a new trial than in the steady state.
-This is further confirmed in simulations in which the trail duration was shortened.
-For those simulations, the prediction even outweighs the sensory input, reflected in a very low sensory weight (Fig.
+Inspecting closely the dynamics of our network, we noticed that the prediction is typically weighted higher at the beginning of a new trial than in the steady state.
+This is particularly pronounced in a sensory-driven input regime (see Fig.
+\ref{fig:Fig_3}C).
+This is further confirmed in simulations in which the trail duration was shortened (Fig.
 \ref{fig:Fig_3}F).
-This suggests that predictions influence neural activity more significantly in experiments that rely on very fast stimulus changes.
-
-It has been shown [speculated?], that sensory inputs or predictions are overrated in some psychiatric disorders (XXX).
+Our model makes therefore the following experimentally testable prediction: Sensory predictions influence neural activity more significantly in experiments that rely on fast stimulus changes.
+
+It has been hypothesized, that some symptoms in psychiatric disorders like autism and schizophrenia can be ascribed to a pathological weighting of sensory inputs and predictions \citep{yon2021precision}.
 We thus wondered which network properties might bias the estimation of the variances, and, consequently, the weighting of different input streams.
-In our network, the M neuron evolves faster in the lower subnetwork than in the higher one.
-We identified the speed at which the M neurons are updated with new information as a decisive factor in the integration of inputs.
+We identified the time scales at which the M neurons incorporate new information as a decisive factor in the integration of inputs.
 To show this, we varied the weights from the PE neurons onto the lower-level M neuron.
-If the M neuron evolves too slowly, the prediction is overrated.
-In contrast, if the M neuron incorporates new information too quickly, the sensory input is overrated (Fig.
+If the weights are too small (M updates too slowly), the system relies too much on feedback predictions.
+In contrast, if the weights are too large (M updates too fast), the system relies too much on the feedforward sensory information (Fig.
 \ref{fig:Fig_3_S2}A).
-While the speeds at which the activity of the M neurons evolve may underlie pathological weighting of inputs, the precise activation function of the V neurons is less pivotal.
-When we replaced the squared activation function with a linear, rectified function, the V neurons do not encode the variance but the averaged absolute deviation of the sensory stimuli.
+While the speeds at which the activity of the M neurons evolve influence the weighting of inputs, the precise activation function of the V neurons is less pivotal.
+When we replaced the quadratic activation function with a linear, rectified function, the V neurons did not encode the variance but the average absolute deviation of the sensory stimuli.
 However, the sensory weight is only slightly shifted to larger values for low trial/high stimulus variability (Fig.
@@
 In summary, we show that the variances of both the sensory inputs and predictions thereof can be dynamically computed in networks comprising a lower and higher PE circuit.
-The model shows that predictions are trusted more strongly at the beginning of a new stimulus, and if sensory inputs are noisy on a short time scale while predictable on longer time scales.
-
-% XXX Do we actually propose that predictions are sent up the hierarchy?
+In such a network, predictions are given more weight at the beginning of a new stimulus, and if the sensory inputs are noisy while the environment is stable.
 
@@
 %
-The brain's flexibility and adaptability are not least because a plethora of neuromodulators influence the activity of neurons in a variety of ways (XXX).
+The brain's flexibility and adaptability are supported by a plethora of neuromodulators which influence the activity of neurons in a variety of ways (XXX).
 A prominent target of neuromodulatory inputs is inhibitory neurons \citep{cardin2019functional, XXX}.
-Moreover, distinct interneuron types are differently (in-)activated by those neuromodulators.
-For instance, it has been shown that XXX (XXX).
+Moreover, distinct interneuron types are differently (in-)activated by those neuromodulators (XXX).
 We, therefore, wondered if and how the weighting of sensory inputs and predictions thereof may be biased when neuromodulators activate distinct interneuron types.
@@
 
-To this end, we modeled the presence of neuromodulators by injecting an additional excitatory input into one or two interneuron types.
+To this end, we modeled the presence of a neuromodulator by injecting an additional excitatory input into an interneuron type.
 We reasoned that the network effect of a neuromodulator not only depends on the interneuron type it targets but also on the inputs this neuron receives and the connections it makes with other neurons in the network.
-We, therefore, tested three different mean-field networks that differ with respect to the distribution of sensory inputs and predictions onto the interneurons, and the underlying connectivity.
+We, therefore, tested different mean-field networks that differ in the distribution of sensory inputs and predictions onto the interneurons, and the underlying connectivity.
 The commonality across those networks is that they exhibit an E/I balance of excitatory and inhibitory pathways onto the PE neurons \citep{hertag2022prediction}.
-Across the different mean-field networks tested, activating a SOM or VIP neuron individually forces the networks to weigh both inputs more equally.
-As a consequence, predictions are overrated in a sensory-driven input regime.
-Similarly, sensory inputs are overrated in a prediction-driven input regime.
-Interestingly, when both interneuron types are activated to the same degree, this effect disappears (Fig.
-\ref{fig:Fig_4}A, left).
-In contrast, stimulating PV neurons biases the network's output towards predictions.
-This effect is even more pronounced when PV and SOM, or PV and VIP neurons are activated simultaneously (Fig.
-\ref{fig:Fig_4}A, middle and right).
-
-In the previous simulations, we assumed that a neuromodulator acts globally, that is, on the interneurons in both the lower and the higher PE circuit.
-While this agrees with experimental data showing that XXX (XXX), we note that neuromodulators may also act more locally.
-The effect of stimulating an interneuron type in the lower PE circuit on the sensory weight is mostly the opposite of activating the same interneuron in the higher PE circuit (Fig.
+
+Across the different mean-field networks tested, increasing the activity of PV neurons biases the network's output toward predictions (Fig.
+\ref{fig:Fig_4}A left).
+In contrast, increasing VIP activity forces the networks to weigh both inputs more equally.
+As a consequence, predictions are overrated in a sensory-driven input regime, and, sensory inputs are overrated in a prediction-driven input regime (Fig.
+\ref{fig:Fig_4}A right).
+Increasing SOM neuron activity, while qualitatively similar to increasing VIP neuron activity, depends on the mean-field network tested and the strength of activation (Fig.
+\ref{fig:Fig_4}A middle).
+
+Neuromodulators are most likely increasing the activity of more than one interneuron type.
+For instance, xxx is known to xxx.
+To account for the co-activation of interneurons, we injected an excitatory input into two interneuron types at the same time and varied the strength with which each interneuron was modulated (Fig.
 \ref{fig:Fig_4_S1}).
-For instance, the sensory inputs are overrated when the higher-level VIP neuron is activated, while the prediction is overrated when the lower-level VIP neuron is activated.
-When VIP and SOM neurons are stimulated equally, the sensory weight remains unchanged, independently of which PE circuit is targeted by the neuromodulator.
-
-What are the mechanisms that give rise to these effects?
-And how do the combined local changes give rise to the global one observed in our network simulations?
-The sensory weight is chosen to be a function of the V neurons of the lower and higher PE circuit.
-Hence, any changes to the sensory weight result from changes to the neurons encoding the variances (Fig.
-\ref{fig:Fig_4}B).
-In our network, the V neurons only receive excitatory output synapses from PE neurons.
+If PV neurons are the major target of a neuromodulator, the network is still biased toward predictions.
+If SOM and VIP neurons are equally stimulated, the weighting of sensory inputs and predictions remains largely unaffected (Fig.
+\ref{fig:Fig_4}B), suggesting that the individual effects cancel out.
+
+What are the network mechanisms underlying these observation?
+The sensory weight is a function of the lower and higher V neuron activity.
+Hence, any changes to the sensory weight result from changes to the neurons encoding the variances.
+In our network, the V neurons only receive excitatory synapses from PE neurons.
 Hence, any changes in the sensory weights upon activation of interneurons must be due to changes in the PE neurons.
@@
 Moreover, the higher-level V neuron is also slightly affected.
-At first, this is counterintuitive because the V neuron in the higher subnetwork does not receive direct synapses from the PE neurons in the lower subnetwork.
-However, the activity of the lower-level M neuron encoding the prediction increases with an excitatory input onto the pPE neuron and decreases with an excitatory input onto the nPE neuron (the opposite is true for an inhibitory input).
-Because neurons in the higher PE circuit receive synapses from the lower-level M neuron, the activity of the higher-level V neuron is also affected.
-In contrast, stimulating either PE neuron in the higher subnetwork increases the activity of the higher-level V neuron but leaves the lower-level M and V neurons unaffected (Fig.
-\ref{fig:Fig_4}C).
-
-Stimulating PE neurons may cause both an increase in the baseline activity and a change in the neuron's gain.
-To disentangle both effects, we illustrate each contribution separately using a mathematically tractable toy model (see Methods for more details).
-The variance estimated in the lower subnetwork increases with both increasing baseline and gain of the lower-level PE neurons.
-In contrast, when the gain of those PE neurons decreases, so does the variance.
-Similarly, the variance estimated in the higher subnetwork is equally influenced by changes in baseline and gain of the higher-level PE neurons.
-Moreover, changes to the baseline and the gain of the lower-level PE neurons increases the higher-level variance as a result of a biased prediction.
-Furthermore, the mean of the sensory stimuli is overpredicted (underpredicted) when the baseline or the gain of the lower-level pPE (nPE) neuron increases, or when the gain of the lower-level nPE (pPE) neuron decreases.
-In summary, if an interneuron causes an additional inhibitory input to a PE neuron, the neuron's gain is reduced.
-If an interneuron causes an additional disinhibitory input to a PE neuron, the neuron's baseline and gain are increased (Fig.
-\ref{fig:Fig_4}D).
-
-This suggests that to understand the effect of neuromodulators on the sensory weight, we need to unravel the effect of interneuron activation on baseline and gain of PE neurons.
-To this end, we stimulated either PV, SOM, or VIP neurons independently for all three mean-field networks and measured the changes to baseline and gain of both PE neurons (Fig.
-\ref{fig:Fig_4}E).
-In all three networks tested, activating PV neurons decreases both quantities, leading to a decrease in the estimated variance (Fig.
-\ref{fig:Fig_4}E \& Fig.
-\ref{fig:Fig_4_S2}).
-Stimulating SOM or VIP neurons decreases the overall gain but increases the baseline activity of the PE neurons.
-Whether and how much the gain of the nPE or pPE neuron is reduced depends on the inputs onto SOM and VIP neurons, and the connectivity they make with other neurons in the network.
-Similarly, how much the baseline is elevated depends on the specifics of the mean-field network (Fig.
-\ref{fig:Fig_4}E \& Fig.
-\ref{fig:Fig_4_S2}).
-Hence, whether stimulating the SOM or VIP neuron decreases or increases the activity of the V neuron depends on the input statistics: for low-mean stimuli, the elevated baseline activity dominates the changes in the variance, while for high-mean stimuli the changes in the gain dominate.
-
-%XXX Summary ...
-effect of neuromodulators acting on INs on sensory weight best understood as combined effect on the baseline and gain of nPE and pPE neurons that in turn determine the activity of the V neuron.
-Those effects not ony depend on the target IN but also the network in which it is embedded, that is the connectivity and the inputs onto INs.
-Moreover, it depends on whether local or global.
-The effects of activation on lower is mainly the opposite of same in higher.
-However, the efefcts don't cancel each other in global modulation.
-Mostly because 1) V higher additionally changes due to changes in prediction but also 2) effects in higher are stronger !?
-and 3) 1/(1+(Vs + D)/(Vp + D)) ...
-that is it cannot cancel per se.
-Even though the underlying reasons for a change in the V neuron depends on the the network characteristics, there are some consistent changes in sensory weigt across MFN ...
-name
-
-% in caption: you need to make sure that in e you used a perturbation of 1 and the mean and std of stimuli were ...
-the lines only illustrate that it depends on the input statistics.
-Of course, the markers would also change with other perturbaton strengths, would they though?
-Think through
-% Mention that for V neuron (higher) the darker line is actually not horizontal (unaffected)!
-% Mention that in BL illustration you need a factor to increase visibility ... (in caption)
-
-\subsection*{The contraction bias as a result of the weighted integration of sensory inputs and predictions}
-% 
-The weighted integration of sensory inputs and predictions thereof manifests in all-day behavior, in the form of a phenomenon called \textit{contraction bias}.
-The contraction bias describes the tendency to overestimate sensory stimuli drawn from the lower end of a stimulus distribution and to underestimate stimuli drawn from the upper end of the same distribution (Fig.
-\ref{fig:Fig_5}A).
-This \textit{bias towards the mean} has been reported in different species and modalities (XXX).
+This is because the lower-level M neuron is also modulated by the lower-level PE neurons and makes feedforward connections to the higher-level PE circuit.
+In contrast, stimulating either PE neuron in the higher subnetwork increases the activity of the higher-level V neuron but leaves the lower-level neurons unaffected.
+
+This suggests that to understand the effect of neuromodulators on the sensory weight, we need to unravel the effect of interneuron activation on PE neurons.
+Increasing interneuron activity leads to changes in the baseline and gain of PE neurons that bias the estimation of mean and variance (Fig.
+\ref{fig:Fig_4_S3}, see Methods).
+In all three networks tested, activating PV neurons decreases both baseline and gain of the PE neurons, leading to a decrease in the estimated variance (Fig.
+\ref{fig:Fig_4}D \& Fig.
+\ref{fig:Fig_4_S4}).
+Stimulating the SOM or VIP neuron decreases the gain in either nPE or pPE neuron.
+However, the baseline of those neurons can either decrease or increase depending on the connectivity with other neurons in the network.
+The summed effect over nPE and pPE neuron (Fig.
+\ref{fig:Fig_4_S4}), hence, suggests that whether the activity of the V neuron increases or decreases depends on the input statistics: for low-mean stimuli, the elevated baseline activity dominates the changes in the variance, while for high-mean stimuli the changes in the gain dominate.
+
+Altogether, we show that neuromodulators increasing the activity of interneurons bias the weighting of sensory inputs and predictions by changing the gain and baseline of PE neurons.
+Whether the sensory weight increases or decreases depends not only on the interneuron it targets but also on the network it is embedded in and the input regime.
+
+\subsection*{Explaining the contraction bias with the weighting of sensory inputs and predictions}
+% 
+We hypothesized that the weighted integration of sensory inputs and predictions thereof manifests in all-day behavior, in the form of a phenomenon called \textit{contraction bias}.
+The contraction bias describes the tendency to overestimate sensory stimuli drawn from the lower end of a stimulus distribution and to underestimate stimuli drawn from the upper end of the same distribution.
+This \textit{bias toward the mean} has been reported in different species and modalities (XXX).
 %
@@
 
-The weighted output of our network can be interpreted as a neural manifestation of the contraction bias (see Methods for a thorough analysis).
-The bias increases with stimulus variance (Fig.
-\ref{fig:Fig_5}B), decreasing the slope of the linear fit modeling the relationship between the true and estimated stimuli (Fig.\ref{fig:Fig_5}B, right; compare with Fig.
-\ref{fig:Fig_5}A).
-In contrast, the bias decreases with trial variance, so that the slope of the linear fit approaches 1 (Fig.
-\ref{fig:Fig_5}B right).
-
-What are the underlying network factors that contribute to this phenomenon?
-To disentangle the potentially different sources of the bias, we first simulated a network without stimulus variability (variance set to zero) for two different trial variabilities.
-In this case, a contraction bias emerges but is independent of the volatility of the environment (Fig.
-\ref{fig:Fig_5}C, left).
-We show mathematically that the bias results from the transient neuron activity before reaching a steady state, and vanishes if the trial duration approaches infinity (see Methods for more details, and Fig.
+We first investigated whether the network's output can be interpreted as a neuronal manifestation of the contraction bias (see Methods for an illustrative analysis).
+To this end, we define the contraction bias as the trial-averaged difference between the weighted output and the sensory stimulus.
+When plotted over the trail-averaged stimuli, the bias is positive for stimuli below the mean of the input distribution and negative for stimuli above the mean (Fig.
+\ref{fig:Fig_5}A), in line with a \textit{bias toward the mean}.
+To measure the amount of bias in the network, we use the slope of the linear fit to the relationship between bias and trial stimulus.
+The larger the absolute slope, the larger the bias.
+
+What are the underlying network factors that contribute to the neuronal contraction bias in the network?
+We have seen that how much the prediction is taken into account is determined by both the lower and higher-level V neurons encoding the variance of the stimulus and the prediction.
+Hence, the bias must be similarly influenced by these factors.
+When we increase the stimulus uncertainty, the bias increases (Fig.
+\ref{fig:Fig_5}B).
+In contrast, when we increase the trial-to-trial uncertainty, the bias decreases (Fig.
+\ref{fig:Fig_5}B).
+
+To further disentangle the different sources of the bias, we first simulated a network without stimulus uncertainty (variance set to zero) for two trial-to-trail variances (volatility of the environment).
+In this case, the emerging contraction bias is independent of the volatility of the environment (Fig.
+\ref{fig:Fig_5}C).
+We show mathematically that the bias results from the network output not yet reaching its new steady state within the trial duration (see Methods).
+In other words, the bias is the difference between the weighted output at the end of the trial and its steady state (the shown stimulus).
+How fast the new steady state is reached depends only on the time constants in the network and not the trial-to-trial variability.
+We next resume the limit case in which the same, high-variance stimulus is shown in every trial.
+In this case, the contraction bias is also largely independent of the stimulus variance (Fig.
 \ref{fig:Fig_5}D).
-We next resume the limit case in which the same but high-variance stimulus is shown in every trial.
-In this case, the weighted output exhibits a contraction bias that is largely independent of the stimulus variance (Fig.
-\ref{fig:Fig_5}C, right and Fig.
-\ref{fig:Fig_5}D).
-As shown mathematically (see Methods), the bias results from the finite trial duration and the tendency to weight the prediction more strongly than the sensory inputs.
-% Maybe show the limit for both cases and show the theoretical values for each T as well ... ?
-
-So far, we assumed that the stimulus variance is independent of the trial mean.
-A consequence of this choice is that the bias on either end of the stimulus distribution is largely the same (but with reversed signs).
+Our mathematical analysis reveals that the bias is well described by the difference between the prediction, that is, mean stimulus over the history of all stimuli shown, and the current stimulus, weighted by a function of the trial duration.
+The analysis of both limit cases suggests that the bias also depends on the trial duration.
+To confirm this, we extended the trial duration for either limit case.
+As expected from the analysis, the bias decreases steadily in the simulations (Fig.
+\ref{fig:Fig_5}E).
+Hence, we predict that the contraction bias can be reduced for sufficiently long trials.
+
+So far, we assumed that the stimulus variance is independent of the stimulus mean.
+A consequence of this choice is that the bias on either end of the input distribution is largely the same (but with reversed signs).
 However, behavioral (neural?) data (XXX) shows that the bias increases for stimuli drawn from the upper end of the distribution, a phenomenon usually attributed to \textit{scalar variability}.
-To capture this in the model, we assume that the stimulus standard deviation linearly increases with the trial mean.
+To capture this in the model, we assume that the stimulus standard deviation linearly increases with the stimulus mean.
 In these simulations, as expected, the bias increases for a stimulus distribution shifted to higher trial means (Fig.
-\ref{fig:Fig_5}E).
-
-In summary, the weighted integration of sensory inputs and predictions can be interpreted as a neural manifestation of the contraction bias.
-Both stimulus and trial variability contribute to the contraction bias but the underlying mechanisms differ.
+\ref{fig:Fig_5_S1}).
+
+In summary, we show that the weighted integration of sensory inputs and predictions can be interpreted as a neural manifestation of the contraction bias.
+While the stimulus and trial-to-trial variability shape the contraction bias, their contributions differ.
+Moreover, we reveal that the trial duration contributes to the bias.
 
@@
 
+% From previous results: In the previous simulations, we assumed that a neuromodulator acts globally, that is, on the interneurons in both the lower and the higher PE circuit.
+While this agrees with experimental data showing that XXX (XXX), we note that neuromodulators may also act more locally.
+The effect of stimulating an interneuron type in the lower PE circuit on the sensory weight is mostly the opposite of activating the same interneuron in the higher PE circuit (Fig.
+\ref{fig:Fig_4_S1}).
+For instance, the sensory inputs are overrated when the higher-level VIP neuron is activated, while the prediction is overrated when the lower-level VIP neuron is activated.
+When VIP and SOM neurons are stimulated equally, the sensory weight remains unchanged, independently of which PE circuit is targeted by the neuromodulator.
+
 A popular hypothesis is that neuromodulators shape the weighting of sensory inputs and predictions thereof \citep{yon2021precision}.
@@
 The mean-field network model consists of a \textit{lower} and \textit{higher} PE circuit (Fig.
-\ref{fig:Fig_1}).
+\ref{fig:Fig_1}C-D).
 Each PE circuit contains an excitatory nPE neuron and pPE neuron ($\mathrm{N}_\mathrm{nPE} = \mathrm{N}_\mathrm{pPE} = 1$), as well as inhibitory neurons.
@@
 %
-We follow the notation that column and row vectors are indicated by small letters with an underscore $\underline{\bullet}$, matrices are denoted by capital letters, and scalars are given by small letters without an underscore.
+We follow the notation that column and row vectors are indicated by letters with an underscore $\underline{\bullet}$, matrices are denoted by capital letters, and scalars are given by small letters without an underscore.
 Furthermore, a time derivative (e.g., $\frac{dx}{dt}$) is denoted by a dot above the letter (e.g., $\dot{x}$).
@@
 %
-All values for neuron and network parameters, details on the model equations, as well as supporting analyses can be found in the supplementary material.
+All values for neuron and network parameters, details on the model equations for the mean-field and the population network, as well as supporting analyses can be found in the supplementary material.
 
@@
 %
-We arithmetically calculated the weighted output of sensory inputs and predictions, $r_\mathrm{out}$, based on Bayesian multisensory integration (XXX),
+We arithmetically calculated the weighted output of sensory inputs and predictions, $r_\mathrm{out}$, based on ideas of Bayesian multisensory integration (XXX),
 %
@@
 The network receives feedforward stimuli $s$ that may vary between trials.
-To account for noise, each stimulus is composed of N$_\mathrm{in}$ piece-wise constant values drawn from a normal distribution with mean $\mu_\mathrm{in}$ and standard deviation $\sigma_\mathrm{in}$.
+To account for noise, each stimulus is composed of N$_\mathrm{in}$ consecutive, piece-wise constant values drawn from a normal distribution with mean $\mu_\mathrm{in}$ and standard deviation $\sigma_\mathrm{in}$.
 To account for changes in the environment, $\mu_\mathrm{in}$ is drawn from a uniform distribution $U(a,b)$ with mean $\mu_\mathrm{trial} = \frac{a+b}{2}$ and standard deviation $\sigma_\mathrm{trial} = \frac{b-a}{\sqrt{12}}$.
-The parameterisation of both distributions varies across the experiments.
+The parameterization of both distributions varies across the experiments.
 All stimulus/input parameters can be found in the supplementary material.
@@
 All simulations were performed in customized Python code written by LH.
-Source code to reproduce the simulations, analyses and figures will be available after publication at \url{XXXX}.
+Source code to reproduce the simulations, analyses, and figures will be available after publication at \url{https://github.com/lhertaeg/weighted_sensory_prediction}.
 Differential equations were numerically integrated using a 2\textsuperscript{nd}-order Runge-Kutta method.
@@
 We then provide the connectivity of the network and the inputs to the neurons for both the mean-field and population model.
-Finally, to ensure reproducibility, we give all simulation parameters used for the results shown in the figures.
+Finally, to ensure reproducibility, we summarize all simulation parameters used for the results shown in the figures.
 
@@
 The equations for the population PE circuit (Fig.
-\ref{fig:Fig_2_S2}) are directly deduced from the mean-field equations and can also be found in XXXref your own paperXXX.
+\ref{fig:Fig_2_S2}) are directly deduced from the mean-field equations and can also be found in \citep{hertag2022prediction}.
 
@@
 \ref{fig:Fig_3_S2}, we used a linear activation function to investigate the impact of the input-output transfer function on the weighting of sensory inputs and predictions.
-The time constant $\tau_V$ was 5 s in the mean-field model, and 2 s in the population model (Fig.
-\ref{fig:Fig_2_S2})., XXX and 0.5 s in 2S3 --> put everything for the not mean field model in the figure captions I would recommend, you can here say that it is only for the mean-field model ...
-XXX
+The time constant $\tau_V$ was 5 s in the mean-field model, 2~s in the heterogeneous population model (Fig.
+\ref{fig:Fig_2_S2}), and 0.5~s in the population model with selectivity (Fig.
+\ref{fig:Fig_2_S3}).
 
@@
 %
-The connectivity between neurons of the PE circuit, both for the mean-field and the population network, were taken from \ref{XXX}.
+The connectivity between neurons of the PE circuit, both for the mean-field and the population network, were taken from \citep{hertag2022prediction}.
 We considered three mean-field networks (see Table \ref{tab:wXM}) that differed in terms of the inputs (feedforward vs.
@@
 %
-where $\lambda$ denotes the speed at which the perfect integrator evolves (lower PE circuit: $\lambda=3\cdot 10^{-3}$ for the mean-field model in Fig.
-\ref{fig:Fig_2},  $\lambda=5\cdot 10^{-2}$ for the population model in Fig.
-\ref{fig:Fig_2_S2}, and $\lambda=4.5\cdot 10^{-2}$ otherwise; higher PE circuit:  $\lambda = 7\cdot 10^{-4}$).
-XXX in Fig.
-\ref{fig:Fig_2_S3} wPE2PE * 100 => I think it is best to put this is the captions, I mean for 2S2 and 2S3 ...
-for mean field network can be shown here and said "if not stated otherwise in the figures/captions"
+where $\lambda$ denotes the speed at which the perfect integrator evolves.
+In the lower PE circuit,  $\lambda=3\cdot 10^{-3}$ for the mean-field model in Fig.
+\ref{fig:Fig_2} and $\lambda=4.5\cdot 10^{-2}$ otherwise.
+In the higher PE circuit, $\lambda = 7\cdot 10^{-4}$.
+For the heterogeneous population model, $\lambda=5\cdot 10^{-2}$ in Fig.
+\ref{fig:Fig_2_S2}, and for the population model with selectivity (Fig.
+\ref{fig:Fig_2_S3}), $ \lambda=3\cdot 10^{-1}$.
 
@@
 \caption{\footnotesize{\bf Gain factors of nPE and pPE neurons in the population model.\newline}
-{The logarithm of the gain factors of nPE (top) and pPE (bottom) neurons in the population model from \ref{XX}.
+{The logarithm of the gain factors of nPE (top) and pPE (bottom) neurons in the population model from \citep{hertag2022prediction}.
 The network contains $67$ nPE neurons and $66$ pPE neurons.
-The remaining excitatory neurons were not classified as PE neurons and were not connected to the $M$ neuron.
-}}
+The remaining excitatory neurons were not classified as PE neurons and were not connected to the $M$ neuron.}}
 \label{fig:Fig_gains}
@@
 However, in Supp Fig.
-\ref{fig:Fig_2_S3} XXX, each mean-field network receives a stimulus $s_i$ drawn from a distribution.
+\ref{fig:Fig_2_S3}, each mean-field network receives a stimulus $s_i$ drawn from a distribution.
 In this case, $\theta$ must be chosen such that deviations from the true variance can be mitigated or fully corrected.
@@
 All simulations were performed in customized Python code written by LH.
-Source code to reproduce the simulations, analyses and figures will be available after publication at \url{XXXX}.
+Source code to reproduce the simulations, analyses and figures will be available after publication at \url{https://github.com/lhertaeg/weighted_sensory_prediction}.
 Differential equations were numerically integrated using a 2\textsuperscript{nd}-order Runge-Kutta method.
@@
 The qualitative results were fairly robust to the choice of the simulation parameters and are here stated merely to ensure the reproducibility of all figures.
-However, we note that we made use of PE circuits that had been trained on steady state inputs \ref{XXX}.
+However, we note that we made use of PE circuits that had been trained on steady state inputs \citep{hertag2022prediction}.
 Hence, we must simulate the network long enough to ensure that the PE neurons reach their steady state.
 Moreover, the lower-level M neuron must evolve faster than the higher-level M neuron as indicated in Fig.
-\ref{XXX}.
+\ref{fig:Fig_3_S2}.
 Finally, the time constant of the V neurons must be of the same magnitude as the trial duration.
@@
 Furthermore, we use the simplified model to discuss the effect of neuromodulators in our network.
-Finally, we reveal the connection between the sensory weight and the contraction bias, and compare our network with the Kalman filter as well as Bayes factor surprise.
-XXX
+Finally, we reveal the connection between the sensory weight and the contraction bias.
 
@@
 
-Similarly, the variance is affected by the gain of the nPE and pPE neuron,
+Likewise, the variance is affected by the gain of the nPE and pPE neuron,
 %
@@
 As shown in sections \ref{sec:impact_gain} and \ref{sec:impact_baseline}, changes in the input-output transfer function of the PE neurons may bias the variance estimation in the network, and, hence, the sensory weight.
-Thus, understanding changes in the sensory weight requires an understanding whether and how different types of interneurons change the PE neurons.
+Thus, understanding changes in the sensory weight requires an understanding of whether and how different types of interneurons change the PE neurons.
 
@@
 
-Together, this suggests that whether the sensory weight decreases, increases or remains the same in the presence of a neuromodulator depends on a number of factors:
+Together, this suggests that whether the sensory weight decreases, increases, or remains the same in the presence of a neuromodulator depends on several factors:
 %
@@
 \item Which interneuron type/s is/are affected by the neuromodulator? And are these interneurons inhibited or excited by the neuromodulator?
-\item How are these interneurons embedded in the network, that is, what is the connectivity and the inputs to those neurons?
+\item How are these interneurons embedded in the network, that is, what are the connectivity and the inputs to those neurons?
 \end{itemize}
 %
-Hence, different neuromodulators may have the same effect on the sensory weight or the same neuromodulator may have different effects depending on brain area, species etc..
+Hence, different neuromodulators may have the same effect on the sensory weight or the same neuromodulator may have different effects depending on brain area, species, etc..
 
@@
 %
-The contraction bias can be formalized through the slope in a linear equation describing the relationship between the estimated and true sensory stimulus.
-If the slope is close to 1, the bias is small.
-In contrast, if the slope is close to zero, the bias is large.
-
-If the prediction encoded in the activity of the M neuron changes very slowly, the bias is well-captured by the sensory weight, 
-%
- \begin{align}
- r_\mathrm{out} = \alpha_\mathrm{S} \cdot s_\mathrm{FF} + \mathrm{constant}
- \end{align}
-%
-However, the prediction is normally a function of the history of sensory stimuli the network received, and cannot be approximated by a constant.
-To investigate how the slope (describing the contraction bias) depends on the sensory weight and other factors, let us resume a toy model in which we assume that the prediction decays exponentially with time constant $\tau$ to a presented constant stimulus value, $s$,
+In the simulations, we define the bias as the trial-averaged difference between the weighted output and the true stimulus.
+For the sake of simplicity, we use $r_\mathrm{out}$ at the end of a trial as a proxy for the trial average in the subsequent analysis.
+Hence,
+%
+\begin{align}
+ \mathrm{bias} = r_\mathrm{out}(T) - s.
+\end{align}
+%
+To investigate how the bias depends on the sensory weight and potentially other factors, let us resume a toy model in which we assume that the prediction decays exponentially with time constant $\tau$ to a presented constant stimulus value, $s$,
 %
@@
 with $P_0$ describing the initial value at time $t=0$.
-Let us assume that within a trial with trial duration $T$, the stimulus value changes $n$ times ($T = n\cdot \Delta t$).
+Let us further assume that within a trial with trial duration $T$, the stimulus value changes $n$ times ($T = n\cdot \Delta t$).
 The prediction during the presentation of the $n$th stimulus value can be expressed as
@@
 %
-Hence, the slope $  m = \left[ \alpha_\mathrm{S}\  \mathrm{e}^{-T/\tau_M} + \left( 1 -   \mathrm{e}^{-T/\tau_M} \right)\right] $ depends on the sensory weight $\alpha_\mathrm{S}$, the trial duration $T$ and time constant $\tau$.
+Hence, the bias in our toy model can be expressed by
+%
+\begin{align*}
+ \mathrm{bias} = \left( 1 -\alpha_\mathrm{S} \right)\cdot \mathrm{e}^{-T/\tau_M}\cdot \left(\mu - s \right).
+\end{align*}
+%
+The absolute slope $\left( 1 -\alpha_\mathrm{S} \right)\cdot \mathrm{e}^{-T/\tau_M}$ indicates how strong the bias is.
+It depends on the sensory weight $\alpha_\mathrm{S}$, the trial duration $T$ and time constant $\tau$.
 Please note that the sensory weight is a function of the trial duration itself (see Fig.
 \ref{fig:Fig_3}F).
-However, for illustration purposes, we take $\alpha_\mathrm{S}$ to be constant here.
-
-In this toy model, if the variance of the prediction is zero (that is, in a prediction-driven input regime), $\alpha_\mathrm{S} \approx 0$, and, hence, the slope is $\left( 1 -   \mathrm{e}^{-T/\tau_M} \right)$.
-Thus, decreasing or increasing the stimulus variance does not have an effect on the bias (see Fig.
+However, for illustration purposes, we take $\alpha_\mathrm{S}$ to be constant.
+
+In this toy model, if the variance of the prediction is zero (that is, in a prediction-driven input regime), $\alpha_\mathrm{S} \approx 0$, and, hence, the bias is $\mathrm{e}^{-T/\tau_M}\cdot \left(\mu - s \right)$.
+Thus, the bias is independent of the stimulus variance (see Fig.
 \ref{fig:Fig_5} D).
 
-Likewise, if the variance of the sensory stimulus is zero (that is, in a stimulus-driven input regime), $\alpha_\mathrm{S} \approx 1$, and, hence, the slope is $1$.
+Likewise, if the variance of the sensory stimulus is zero (that is, in a stimulus-driven input regime), $\alpha_\mathrm{S} \approx 1$, and, hence, the bias approaches $0$ if the neurons reach their steady state.
 Thus, decreasing or increasing the trail variance does not have an effect on the bias (see Fig.
@@
 
-Moreover, the toy model also reveals that if $T\rightarrow \infty$, the slope $m$ goes to 1, suggesting that without further noise in the system, the network can perfectly represent the sensory stimulus (the prediction would approach the current stimulus).
+%Moreover, the toy model also reveals that if $T\rightarrow \infty$, the slope $m$ goes to 1, suggesting that without further noise in the system, the network can perfectly represent the sensory stimulus (the prediction would approach the current stimulus).
 
@@
 {\bf (B)} The precise activation function for the V neurons does not have a major impact on the sensory weight.
-Only for inputs with high stimulus variability, the sensory stimulus is slightly overrated when the squared activation function is replaced by a linear, rectified activation function.
+Only for inputs with high stimulus variability, the sensory stimulus is slightly overrated when the quadratic activation function is replaced by a linear, rectified activation function.
 }

### 2023-11-10 scripts/run_fig_4.py

@@ -24,10 +24,10 @@ dtype = np.float32
 # %% Neuromodulator targeting one IN (with various strengths)
 
 run_cell = True
-plot_only = True
+plot_only = False
 column = 0 
 
-xp, xs, xv = 0, 0, 1
+xp, xs, xv = 1, 0, 0
 
 if run_cell:
     
@@ -35,7 +35,7 @@ if run_cell:
     file_for_data = '../results/data/neuromod/data_neuromod_' + str(xp) + '_' + str(xs) + '_' + str(xv) + '.pickle'
     
     ### define perturbations tested
-    pert_strength = np.linspace(0,0.5,6)
+    pert_strength = np.linspace(0,1,6)
     
     ### main loop
     if not plot_only:
@@ -67,7 +67,7 @@ if run_cell:
             
 
     ### plot data    
-    plot_neuromod_impact(pert_strength, alpha, xp, xs, xv)
+    plot_neuromod_impact(pert_strength, alpha, xp, xs, xv, file_for_data=file_for_data)
                                 
     
 # %% How are the variance neurons influenced by changes in nPE and pPE neurons in the lower and higher PE circuit?

### 2023-11-10 src/plot_data.py

@@ -822,7 +822,7 @@ def illustrate_sensory_weight_variance(ax = None, fs=6):
                
     
 def plot_neuromod_impact(pert_strength, alpha, xp, xs, xv, figsize=(2,2.5), fs=6, lw=1, s=15, show_ylabel=True, show_xlabel = True,
-                            flg_plot_xlabel=True, flg_plot_bars = True, ax1 = None, ax2=None, highlight=True):
+                            flg_plot_xlabel=True, flg_plot_bars = True, ax1 = None, ax2=None, highlight=True, file_for_data = None):
     
         if ax1 is None:
             f, (ax1, ax2) = plt.subplots(2,1, tight_layout=True, figsize=figsize, sharey=True, sharex=True)
@@ -830,7 +830,9 @@ def plot_neuromod_impact(pert_strength, alpha, xp, xs, xv, figsize=(2,2.5), fs=6
         markers = ['o','s','d']
         
         ### load data
-        file_for_data = '../results/data/neuromod/data_neuromod_' + str(xp) + '_' + str(xs) + '_' + str(xv) + '.pickle'
+        if file_for_data is None:
+            file_for_data = '../results/data/neuromod/data_neuromod_' + str(xp) + '_' + str(xs) + '_' + str(xv) + '.pickle'
+            
         with open(file_for_data,'rb') as f:
             [xp, xs, xv, pert_strength, alpha] = pickle.load(f) 
         
@@ -841,10 +843,10 @@ def plot_neuromod_impact(pert_strength, alpha, xp, xs, xv, figsize=(2,2.5), fs=6
             ax1.axhline(alpha[0,i,1], color=cmap_sensory_prediction(alpha[0,i,1]), ls=':', zorder=0)
             
             ax2.scatter(pert_strength, alpha[:,i,0], marker=markers[i], c=alpha[:,i,0], cmap=cmap_sensory_prediction, 
-                        vmin=0, vmax=1, s=s)
+                        vmin=0, vmax=1, s=s, clip_on=False, zorder=10)
             
             ax1.scatter(pert_strength, alpha[:,i,1], marker=markers[i], c=alpha[:,i,1], cmap=cmap_sensory_prediction, 
-                        vmin=0, vmax=1, s=s)
+                        vmin=0, vmax=1, s=s, clip_on=False, zorder=10)
             
 
             

### 2023-11-17 records/manuscript_2023.tex
first complete draft

@@
 %\usepackage[super,sort&compress,comma]{natbib}
-\usepackage[numbers, round, sort&compress, comma]{natbib}
-%\usepackage[round]{natbib}
+%\usepackage[numbers, round, sort&compress, comma]{natbib}
+\usepackage[round]{natbib}
 
@@
 {\Large
-\textbf\newline{Knowing what you do not know: Estimating the uncertainty of feedforward and feedback inputs with prediction-error circuits}
+\textbf\newline{Knowing what you don't know: Estimating the uncertainty of feedforward and feedback inputs with prediction-error circuits} % Uncertainty estimation with prediction-error circuits
 }
@@
 \bigskip
-1 Modelling of Cognitive Processes, TU Berlin, Berlin, Germany.\\
+1 Modeling of Cognitive Processes, TU Berlin, Berlin, Germany.\\
 2 Department of Physiology, University of Bern, Switzerland.\\
@@
 A common hypothesis is that the brain weights different inputs according to their reliabilities.
-A prominent example of this hypothesis is Bayesian multisensory integration (XXX).
+A prominent example of this hypothesis is Bayesian multisensory integration \citep[see, e.g.,][]{deneve2004bayesian}.
 According to this theory, neural networks represent information from multiple modalities by a linear combination of the uncertainty-weighted single-modality estimates.
-Multisensory integration is supported by several observations showing that xxx (XXX).
-Here, we propose that the same concepts could be employed for the weighting of sensory inputs and predictions thereof \citep{yon2021precision}.
+Multisensory integration is supported by several observations showing that animals can combine information from different modalities in a fashion that minimizes the variance of the final estimate \citep{ernst2002humans, battaglia2003bayesian, kording2004bayesian, alais2004ventriloquist, rowland2007bayesian, gu2008neural, fetsch2012neural}.
+Here, we propose that the same concepts could be employed for the weighting of sensory inputs and predictions thereof \citep{kording2004bayesian, yon2021precision}.
 A central point in the weighting of inputs is the estimation of their variances as a measure of uncertainty.
@@
 PEs are an integral part of the theory of predictive processing which states that the brain constantly compares incoming sensory information with predictions.
-If those predictions are wrong, the resulting PEs allow the network to revise the model of the world, thereby ensuring that the predictions become more accurate (XXX).
+If those predictions are wrong, the resulting PEs allow the network to revise the model of the world, thereby ensuring that the predictions become more accurate \citep{keller2018predictive}.
 Experimental evidence suggests that these PEs may be represented in the activity of distinct groups of neurons, termed PE neurons \citep{eliades2008neural, keller2009neural, ayaz2019layer, audette2021temporally}.
-Moreover, these neurons may come in two types when excitatory neurons exhibit near-zero, spontaneous firing rates \cite{rao1999predictive, keller2018predictive}.
+Moreover, these neurons may come in two types when excitatory neurons exhibit near-zero, spontaneous firing rates \citep{rao1999predictive, keller2018predictive}.
 Negative PE (nPE) neurons only increase their activity when the prediction is \textit{stronger} than the sensory input, while positive PE (pPE) neurons only increase their activity when the prediction is \textit{weaker} than the sensory input.
@@
 In addition, we unravel the mechanisms underlying a neuromodulator-induced shift in the weighting of sensory inputs and predictions.
-In our model, these neuromodulators activate groups of inhibitory neurons such as parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (XXX).
+In our model, these neuromodulators activate groups of inhibitory neurons such as parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons \citep{markram2004interneurons, rudy2011three, pfeffer2013inhibition, jiang2015principles, tremblay2016gabaergic, campagnola2022local}.
 These interneurons have been suggested to establish a multi-pathway balance of excitation and inhibition that is the basis for nPE and pPE neurons \citep{hertag2020learning, hertag2022prediction}.
 By breaking this balance, the PE neurons change their baseline firing rate and gain, leading to a biased variance estimation.
-Finally, we show that this weighting can be understood as a neural manifestation of the contraction bias (XXX).
+Finally, we show that this weighting can be understood as a neural manifestation of the contraction bias \citep{hollingworth1910central, jazayeri2010temporal, ashourian2011bayesian, petzschner2011iterative, akrami2018posterior, meirhaeghe2021precise}.
 
@@
 
-We assume that this core circuit is shaped by feedback connections \ref{XXX} that have been hypothesized to carry information about expectations or predictions \ref{XXX}.
+We assume that this core circuit is shaped by feedback connections \citep{larkum2013cellular, harris2015neocortical} that have been hypothesized to carry information about expectations or predictions \citep{mumford1992computational, larkum2013cellular, friston2008hierarchical}.
 To account for predictions, we model a memory (M) neuron that integrates the activity of the PE neurons (Fig.
 \ref{fig:Fig_1}C).
-Following \cite{keller2018predictive}, we assume that the pPE neuron excites the memory neuron, while the nPE neuron inhibits this neuron (for instance, through lateral inhibition, here not modeled explicitly) [XXX KW: look into Anna V paper, opposing effects tested biorxiv great ref XXX].
-Because feedback connections are shown to target the apical dendrites of pyramidal cells \citep{LarkumXXX} and interneurons located in superficial layers of the cortex \citep{XXX}, the M neuron makes connections with the dendritic compartment of the PE neurons and some of the interneurons (here, VIP and PV neurons, see Methods for more details).
+Following \cite{keller2018predictive}, we assume that the pPE neuron excites the memory neuron, while the nPE neuron inhibits this neuron (for instance, through lateral inhibition, here not modeled explicitly).
+Because feedback connections are shown to target the apical dendrites of pyramidal cells \citep{larkum2013cellular} and interneurons located in superficial layers of the cortex \citep[see, e.g.][]{tremblay2016gabaergic}, the M neuron makes connections with the dendritic compartment of the PE neurons and some of the interneurons (here, VIP and PV neurons, see Methods for more details).
 We, furthermore, simulate a downstream neuron (termed V neuron), modeled as a leaky integrator with a quadratic activation function, that receives excitatory synapses from the PE neurons.
@@
 %
-Following the ideas of Bayesian multisensory integration (XXX), the weighting of sensory stimuli and predictions would require knowledge about their uncertainties.
+Following the ideas of Bayesian multisensory integration, the weighting of sensory stimuli and predictions would require knowledge about their uncertainties.
 As we have shown in the previous section, the variance of the sensory stimulus can be estimated using PE neurons.
@@
 
-Following the formalism of multisensory integration (XXX), we assume that the network's output is a weighted sum of the feedforward sensory input and the feedback prediction.
+Following the formalism of multisensory integration \citep[see, e.g.][]{pouget2013probabilistic}, we assume that the network's output is a weighted sum of the feedforward sensory input and the feedback prediction.
 The weights assigned to each input stream are functions of the uncertainties, that is, the activities of the V neurons.
@@
 %
-The brain's flexibility and adaptability are supported by a plethora of neuromodulators which influence the activity of neurons in a variety of ways (XXX).
-A prominent target of neuromodulatory inputs is inhibitory neurons \citep{cardin2019functional, XXX}.
-Moreover, distinct interneuron types are differently (in-)activated by those neuromodulators (XXX).
+The brain's flexibility and adaptability are supported by a plethora of neuromodulators which influence the activity of neurons in a variety of ways \citep{avery2017neuromodulatory}.
+A prominent target of neuromodulatory inputs is inhibitory neurons \citep{cardin2019functional,  hattori2017functions, swanson2019hiring}.
+Moreover, distinct interneuron types are differently (in-)activated by those neuromodulators \citep{wester2014behavioral, hattori2017functions, swanson2019hiring}.
 We, therefore, wondered if and how the weighting of sensory inputs and predictions thereof may be biased when neuromodulators activate distinct interneuron types.
@@
 Neuromodulators are most likely increasing the activity of more than one interneuron type.
-For instance, xxx is known to xxx.
 To account for the co-activation of interneurons, we injected an excitatory input into two interneuron types at the same time and varied the strength with which each interneuron was modulated (Fig.
@@
 The contraction bias describes the tendency to overestimate sensory stimuli drawn from the lower end of a stimulus distribution and to underestimate stimuli drawn from the upper end of the same distribution.
-This \textit{bias toward the mean} has been reported in different species and modalities (XXX).
+This \textit{bias toward the mean} has been reported in different species and modalities \citep{hollingworth1910central, jazayeri2010temporal, ashourian2011bayesian, petzschner2011iterative, akrami2018posterior, meirhaeghe2021precise}.
 %
@@
 A consequence of this choice is that the bias on either end of the input distribution is largely the same (but with reversed signs).
-However, behavioral (neural?) data (XXX) shows that the bias increases for stimuli drawn from the upper end of the distribution, a phenomenon usually attributed to \textit{scalar variability}.
+However, behavioral data \citep[see, e.g.][]{rakitin1998scalar} shows that the bias increases for stimuli drawn from the upper end of the distribution, a phenomenon usually attributed to \textit{scalar variability}.
 To capture this in the model, we assume that the stimulus standard deviation linearly increases with the stimulus mean.
@@
 \section*{Discussion}
-
-% XXX At the end: Read my „Todo“ in folder „read_weighting“: IMPORTANT XXX
 %
@@
 Conflicting information may be caused by noise in the sensory feedforward inputs or by changes in the environment that could not be predicted.
-A prominent hypothesis is that how much we rely on our predictions and new sensory evidence is determined by an intricate balance between both, based on how reliable they are \citep[see e.g.][]{yon2021precision}.
+A prominent hypothesis is that how much we rely on our predictions and new sensory evidence is determined by an intricate balance between both, based on how reliable they are \citep[see e.g.][]{kording2004bayesian, yon2021precision}.
 
 This idea is consistent with Bayesian theories on the optimal integration of multiple sensory cues (aka multisensory integration).
-\citep{ernst2002humans} showed that to estimate the height of a bar humans combine visual and haptic information in a fashion that minimizes the variance of the final estimate.
-XXX name other examples in human and other species XXX.
-These behavioral studies were accompanied by neural recordings showing that XX
-
-% Summary of findings and predictions/implications from our model
-
-While multisensory integration is concerned with the weighting of sensory cues coming from different modalities, it is conceivable that the same ideas hold for top-down predictions and bottom-up sensory inputs that are combined in cortical circuits \citep{yon2021precision}.
-Indeed, there is evidence that the brain can choose to rely more on top-down or bottom-up signals \citep{pakan2018impact, han2023behavior}.
+\cite{ernst2002humans} showed that to estimate the height of a bar humans combine visual and haptic information in a fashion that minimizes the variance of the final estimate.
+Similar studies confirmed that animals can optimally combine multiple sensory information by taking into account their uncertainties \citep{battaglia2003bayesian, kording2004bayesian, alais2004ventriloquist, rowland2007bayesian, gu2008neural, fetsch2012neural}.
+These behavioral studies were accompanied by neural recordings identifying populations of neurons that can form the basis of multisensory integration \citep{wallace1998multisensory, gu2008neural, fetsch2012neural}.
+
+\subsection*{Summary of findings}
+
+While multisensory integration is concerned with the weighting of sensory cues coming from different modalities, it is conceivable that the same ideas hold for top-down predictions and bottom-up sensory inputs that are combined in cortical circuits \citep{kording2004bayesian, yon2021precision}.
+Indeed, there is evidence that the brain can rely more on top-down or bottom-up signals \citep{pakan2018impact, han2023behavior}.
 Following the same ideas from Bayesian multisensory integration, the brain must hence have mechanisms to track both the reliability of the sensory signals and the predictions thereof \citep{yon2021precision}.
-XXX We suggested that xxx Here summarise your contributions XXX
-
-XXX put this in here somewhere XXX
+
+Here, we show that PE neurons can serve as the backbone for estimating the uncertainty of both the feedforward sensory inputs and the feedback predictions (Figs.
+\ref{fig:Fig_2} \& \ref{fig:Fig_3}).
 In our model, we assume a hierarchy of PE circuits that are feed-forwardly connected through the lower-order M neuron whose activity encodes the mean of the sensory bottom-up inputs.
-This local prediction is fed back to the lower-order circuit and at the same time feed-forwarded to the higher-order subnetwork.
+This local prediction is fed back to the lower-order circuit and at the same time feed-forwarded to the higher-order subnetwork (Fig.
+\ref{fig:Fig_1}).
 With this architecture in place, we show that we rely more strongly on our internal signals when the perceived sensory cues are noisier than the predictions, either because the environment is stable but highly noisy or because the environment changes quickly.
@@
 
-Furthermore, we show that the weighting of sensory inputs and predictions can be biased through neuromodulators, as has been suggested before \citep{XXX}.
+Furthermore, we show that the weighting of sensory inputs and predictions can be biased through neuromodulators, as has been suggested before \citep[see, e.g.,][]{yon2021precision}.
 In our model, those modulatory signals act through interneurons \citep{cardin2019functional} whose activities increase in the presence of neuromodulators.
 When PV neuron activity increases, the network weighs predictions stronger than without modulation.
-In contrast, when VIP neuron activity increases, the network underestimates the uncertainty of the prediction in a sensory-driven regime and underestimates the uncertainty of the sensory input in a prediction-driven regime.
-Hence, the system leans toward weighting sensory inputs and predictions more equally.
-Increasing SOM neuron activity is comparable to modulating VIP neurons.
-However, the equalising effect is less robust and seems to depend on the network specifics (Fig.
+In contrast, when VIP neuron activity increases, the network underestimates the uncertainty of the prediction in a sensory-driven regime, and it underestimates the uncertainty of the sensory input in a prediction-driven regime.
+Hence, the system leans toward weighting sensory inputs and predictions more equally (Fig.
 \ref{fig:Fig_4}A).
-Interestingly, when SOM and VIP neuron activity is modulated to the same degree, the weighting remains unaffected, suggesting that the individual contributions cancel (Fig.
+When SOM and VIP neuron activities are modulated to the same degree, the weighting remains unaffected, suggesting that the individual contributions cancel (Fig.
 \ref{fig:Fig_4}B).
-We show that these findings can be best explained by changes in the baseline and gain of PE neurons as a result of the modulation of interneuron activity (Fig.
+We show that these findings can be explained by changes in the baseline and gain of PE neurons arising through the modulation of interneuron activity (Fig.
 \ref{fig:Fig_4}D).
@@
 Finally, we illustrate that the weighted integration of feedforward and feedback inputs can be interpreted as a neural manifestation of the contraction bias.
-We show that the bias is strongly driven by the variance of the sensory cues and the predictions, as well as the trial duration.
-In our model, if the stimulus variance is zero (perfect sensory-driven network regime), the neural contraction bias is independent of the trial variability.
-In that case, the bias reflects the slow transition from the last stimulus to the new one, and, hence, would diminish if the trial is sufficiently long.
-On the other hand, if the trial variance is zero (perfect prediction-driven network regime), the neural contraction bias is independent of the stimulus variability and a function of the trial duration.
-However, we note that in our model, we only consider the neural representation of the stimulus and do not account for other sources of noise, like execution noise, that surely impacts the contraction bias observed in behavioral studies \citep{XXX}.
-
-% Inspiration for the model & assumption (evidence for that)
-
+We show that the bias is strongly driven by the variances of the sensory cue and the prediction, as well as the trial duration.
+While the sensory noise increases the contraction bias, the uncertainty in the prediction or long trials decreases the bias (Fig.
+\ref{fig:Fig_5}.
+However, we note that we only consider a neural representation of the stimulus and do not account for other sources of noise, like execution noise, that surely impacts the contraction bias observed in behavioral studies.
+
+\subsection*{Biological evidence for model choices and assumptions}
+%
 In our model, we assumed that there are dedicated neurons that encode the variance of the feedforward sensory inputs and the prediction.
-This assumption is consistent with the idea that neurons explicitly encode in their activity the parameters (for instance, mean or variance) of a probability distribution \citep{ONeill Schultz, 2010, XXX}.
+This assumption is consistent with the idea that neurons explicitly encode in their activity the parameters (for instance, mean or variance) of a probability distribution \citep{o2010coding, o2012can}.
 However, how variances of signals are represented in the brain is still not comprehensively understood and alternative ideas have been put forward.
-For instance, it is conceivable that the variance is encoded in a population of neurons, each differently tuned to a specific parameter \citep{knill2004bayesian, Ma et al., 2006, 2008}.
+For instance, it is conceivable that the variance is encoded in a population of neurons, each differently tuned to a specific parameter \citep{knill2004bayesian}.
 The neurons' activities represent how close the sensory input is to the preferred (predicted) input of each neuron.
-
-% —> check out: There exist a number of theoretical models that address the potential neural representation of uncertainty (Deneve, 2008; Hinton and Sejnowski, 1983; Hoyer and Hyvarinen, 2003; Ma et al., 2006; Zemel et al., 1998) … in Dekleva 2016 XXXXX
-
-There has been evidence that indeed (population of) neurons can encode (un-)certainty.
-Neurons in the pulvinar (Komura et al., 2013), the parietal cortex (Kiani and Shadlen, 2009) or the frontal cortex (Kepecs et al., 2008, Middlebrooks and Sommer, 2012) encode the uncertainty of perceptual decisions \citep{masset2020behavior}.
-In addition, it has been shown that neurons in the orbitofrontal cortex of rats signal confidence in decisions \citep{masset2020behavior}.
-Moreover, neural correlates of (expected) uncertainty have been found in regions of the prefrontal cortex 43 \citep{soltani2019adaptive} , the rat orbitofrontal cortex [52,53\citep{soltani2019adaptive}] or the dorsal striatum in monkeys78 \citep{soltani2019adaptive}.
-Interestingly, the accuracy of memory recalls is represented (?) in single neurons of the human parietal and temporal lobes (Rutishauser et al., 2015, 2018) … in \citep{masset2020behavior}.
+Similarly, a neuron's response variability has been suggested to be related to the uncertainty of sensory stimuli \citep{hoyer2002interpreting, ma2006bayesian}.
+
+There has been evidence that indeed (population of) neurons can encode (un-)certainty \citep{soltani2019adaptive}.
+For instance, neurons in the parietal cortex in monkeys encode the degree of confidence in a perceptual decision \citep{kiani2009representation}.
+Similarly, the firing rate of neurons in the orbitofrontal cortex have been shown to encode confidence irrespective of sensory modality \citep{masset2020behavior}.
+Neural signatures of uncertainty have been found in regions of the prefrontal cortex \citep{rushworth2008choice}, the rat insular and orbitofrontal cortex \citep{jo2016differential}, or the dorsal striatum in monkeys \citep{white2016neurons}.
+Moreover, the accuracy of memory recalls is encoded in single neurons of the human parietal and temporal lobes \cite{rutishauser2015representation, rutishauser2018single}.
 
 In our model, we assume a hierarchy of predictions that are locally computed in memory neurons.
-These memory neurons are consistent with the idea of internal representation neurons hypothesised in predictive processing theories \citep{bastos2012canonical, keller2018predictive}.
-While it has been hypothesised that these internal representation neurons might be found in deeper L5 neurons \citep{bastos2012canonical, heindorf2022reduction}, there is also evidence that a group of excitatory L2/3 neurons integrates over negative and positive prediction errors \citep{o2022prediction}.
+These memory neurons are consistent with the idea of internal representation neurons hypothesized in predictive processing theories \citep{bastos2012canonical, keller2018predictive}.
+While it has been hypothesized that these internal representation neurons might be deeper L5 neurons \citep{bastos2012canonical, heindorf2022reduction}, there is also evidence that a group of excitatory L2/3 neurons integrates over negative and positive prediction errors \citep{o2022prediction}.
 
@@
 Those neurons have been found in different cortical areas in various species \citep{eliades2008neural, keller2009neural, ayaz2019layer, audette2021temporally}.
-Moreover, while first only hypothesized theoretically \ref{rao1999predictive}, the presence of two types of PE neurons, the negative and positive PE neurons, has been confirmed in a number of recent studies \citep{keller2012sensorimotor, attinger2017visuomotor, jordan2020opposing, audette2021temporally, Zmarz2016XXX?necessary}.
-In our model, the nPE neuron inhibits the memory neuron while the pPE neuron excites it, in line with \citep{keller2018predictive}.
-The weights from those PE neurons onto the M neuron are larger for the lower- than the higher PE circuit, so that the lower-order M neuron evolves faster than the higher-order counterpart.
-This assumption is consistent with the observation that time constants increase along the cortical hierarchy (XXX see refs in Siegle 2021 Allen).
-
-However, if the relation of the paces at which the M neurons evolve is strongly modulated, the network either shows a bias towards the sensory cues or the prediction (Figs.
-\ref{XXX}).
-It has been hypothesized that symptoms in psychiatric diseases may be a result of an erroneous uncertainty estimation \citep{yon2021precision}.
-For instance, hallucinations may be the result of an underestimation of the expectation uncertainty or an overestimation of the sensory uncertainty.
-Conversely, a fixation on a stable environment even when the sensory cues indicate a switch in the environment, may be the consequence of an overestimation of the expectation uncertainty or an underestimation of the sensory uncertainty \citep{yon2021precision}.
-
-% Neuromodulation
-
-% From previous results: In the previous simulations, we assumed that a neuromodulator acts globally, that is, on the interneurons in both the lower and the higher PE circuit.
-While this agrees with experimental data showing that XXX (XXX), we note that neuromodulators may also act more locally.
-The effect of stimulating an interneuron type in the lower PE circuit on the sensory weight is mostly the opposite of activating the same interneuron in the higher PE circuit (Fig.
-\ref{fig:Fig_4_S1}).
-For instance, the sensory inputs are overrated when the higher-level VIP neuron is activated, while the prediction is overrated when the lower-level VIP neuron is activated.
-When VIP and SOM neurons are stimulated equally, the sensory weight remains unchanged, independently of which PE circuit is targeted by the neuromodulator.
-
+Moreover, while first only hypothesized theoretically \citep{rao1999predictive}, the presence of two types of PE neurons, the negative and positive PE neurons, has been confirmed in several recent studies \citep{keller2012sensorimotor, attinger2017visuomotor, jordan2020opposing, audette2021temporally}.
+In our model, the nPE neuron inhibits the memory neuron while the pPE neuron excites it, in line with \cite{keller2018predictive}.
+The weights from those PE neurons onto the M neuron are larger for the lower than the higher PE circuit, so that the lower-order M neuron evolves faster than the higher-order counterpart.
+This assumption is consistent with the observation that time constants increase along the cortical hierarchy \citep{murray2014hierarchy, chaudhuri2015large, runyan2017distinct}.
+
+If the relation of the paces at which the M neurons evolve is strongly modulated, the network either shows a bias towards the sensory cues or the prediction (Figs.
+\ref{fig:Fig_3_S2}).
+It has been hypothesized that symptoms in psychiatric diseases may derive from an erroneous uncertainty estimation \citep{yon2021precision}.
+For instance, hallucinations may arise from an underestimation of the expectation uncertainty or an overestimation of the sensory uncertainty.
+Conversely, a fixation on the environment, even when the sensory cues indicate a switch in the environment, may originate from an overestimation of the expectation uncertainty or an underestimation of the sensory uncertainty \citep{yon2021precision}.
+
+\subsection*{Neuromodulators and uncertainty}
+%
 A popular hypothesis is that neuromodulators shape the weighting of sensory inputs and predictions thereof \citep{yon2021precision}.
-Theoretical work by \ref{You and Dayan} suggests that acetylcholine (ACh) correlates with expected uncertainty, while noradrenaline (NA) correlates with unexpected uncertainty.
-Expected uncertainty is usually considered to be related to known cue-outcome unreliabilities.
-In contrast, unexpected uncertainty relates to the changes in the environment that produce large PEs outside the expected range of uncertainties \citep{You and Dayan}.
-While in our network the stimulus and trail variability can only be loosely interpreted as 'expected' and 'unexpected' uncertainty, we want to XXX broader context and highlight differences XXX
-
-% XXX —> the unexpected uncertainty in my case (trial variability) is not unexpected time-wise only direction-wise.
-One could change that in a SI figure or one later main figure but not sure how much sense that makes (discuss with Katha and CC)
-
-It is assumed that NA increases in more volatile environments and enhances bottom-up processes [20,21 in \citep{lawson2021computational}, yon2021precision].
-In line with this idea, NA blockade impairs cognitive flexibility [24,25\citep{lawson2021computational}].
-In recent work by \citep{lawson2021computational}, it has been shown that humans receiving propranolol (blocking NA) rely more strongly on their expectations and are slower to update these predictions despite new sensory evidence \citep{yon2021precision}.
-A major target for noradrenergic inputs is SOM neurons whose activity increases in the presence of NA (reviewed in Urban-Ciecko 2016: Fanselow, 2008; Urban-Ciecko 2015; Chen 2015; Xiang 1998; reviewed in Hattori 2017: Paspalas 1999; reviewed in Swanson in Maffei 2019: Kawaguchi and Shindou, 1998).
-In our model, activating SOM neurons globally does not enhance bottom-up input.
-In a volatile environment, that is, a sensory-driven regime, the system takes into account predictions slightly more than without SOM modulation.
-However, increasing SOM activity only in the lower-order circuit shows a slight enhancement of the sensory weight (Fig.
-\ref{XXX}).
-
-Similarly to NA, ACh has also been shown to enhance bottom-up, feedforward inputs (Gil et al., 1997; Hasselmo et al., 1996; Hsieh et al., 2000; Kimura et al., 1999; Kobayashi, 2000, reviewed in Yu and Dayan 2005 or Marshall 2016).
-For instance, subjects relied more strongly on prior beliefs when given cholinergic receptor antagonists (reviewed in Bruckner 2022: Marshall 2016).
-A major target for cholinergic input are VIP neurons whose activity increases in the presence of ACh (reviewed in Wester 2014; reviewed in Hattori 2017: Fu 2016, Kuchibhotla 2017, Nelson 2016; reviewed in Swanson in Maffei 2019: Paul et al., 2017).
+Theoretical work by \citep{yu2005uncertainty} suggests that acetylcholine (ACh) correlates with \textit{expected uncertainty}, while noradrenaline (NA) correlates with \textit{unexpected uncertainty}.
+Expected uncertainty is usually interpreted as known cue-outcome unreliabilities.
+In contrast, unexpected uncertainty relates to the changes in the environment that produce large PEs outside the expected range of uncertainties \citep{yu2005uncertainty}.
+While in our network the stimulus and trail-to-trial variability can only be loosely interpreted as 'expected' and 'unexpected' uncertainty, we want to compare the effects of ACh and NA on the weighting with the effects hypothesized in the literature.
+
+It is assumed that NA increases in more volatile environments and enhances bottom-up processes \citep{hasselmo1997noradrenergic, yon2021precision}.
+In line with this idea, NA blockade impairs cognitive flexibility \citep{ridley1981new, janitzky2015optogenetic}.
+In recent work by \cite{lawson2021computational}, it has been shown that humans receiving propranolol (blocking NA) rely more strongly on their expectations and are slower to update these predictions despite new sensory evidence \citep{yon2021precision}.
+A main target for noradrenergic inputs is SOM neurons whose activity increases in the presence of NA \citep[reviewed in, e.g.,][]{urban2016somatostatin, hattori2017functions, swanson2019hiring}.
+In our model, activating SOM neurons does not enhance sensory bottom-up input.
+In a volatile environment, that is, a sensory-driven regime, the system takes into account predictions slightly more than without SOM modulation (Figs.
+\ref{fig:Fig_4}A and \ref{fig:Fig_4_S1}).
+
+However, we note that in our simulations, we assumed that neuromodulators act globally, that is, on the interneurons in both the lower and the higher PE circuit.
+While this agrees with the view that neuromodulators can control network states globally, there is also evidence that they can have a more local, finely adjusted impact on neural circuits \citep{nadim2014neuromodulation}.
+In our model, increasing SOM activity only in the lower-order circuit shows a slight enhancement of the sensory weight (Fig.
+\ref{fig:Fig_4_S2}), that is, the bottom-up inputs.
+This suggests that whether a neuromodulator biases the network toward feedforward bottom-up or feedback top-down inputs depends on its spatial and temporal scale of influence.
+
+Similarly to NA, ACh has also been shown to enhance bottom-up, feedforward inputs \citep[reviewed in, e.g.,][]{yu2005uncertainty, marshall2016pharmacological}.
+For instance, subjects relied more strongly on prior beliefs when given cholinergic receptor antagonists \citep{marshall2016pharmacological}.
+A major target for cholinergic inputs is VIP neurons whose activity increases in the presence of ACh \citep[reviewed in, e.g.,][]{wester2014behavioral, hattori2017functions, swanson2019hiring}.
 In our model, activating VIP neurons globally only enhances bottom-up input in stable environments for noisy stimuli.
-However, increasing VIP activity only in the higher-order circuit does generally enhance sensory bottom-up inputs (Fig.
-\ref{XXX}).
-
-Neuromodulatory systems have been suggested to gate plasticity \ref{XXX}.
-In a recent study by \citep{jordan2023locus}, locus coeruleus (LC) axon activity is shown to correlate with the magnitude of unsigned visuomotor prediction errors.
+However, increasing VIP activity only in the higher-order PE circuit generally enhances sensory bottom-up inputs (Fig.
+\ref{fig:Fig_4_S2}).
+
+\subsection*{Limitations \& future steps}
+%
+As for any computational model, we brush over several biological details to keep the model simple and interpretable.
+However, those details, while beyond the scope of this study, may be well investigated in future work.
+For instance, neuromodulatory systems have been suggested to gate plasticity \citep{pawlak2010timing}.
+In a recent study by \cite{jordan2023locus}, locus coeruleus (LC) axon activity is shown to correlate with the magnitude of unsigned visuomotor prediction errors.
 The authors hypothesize that LC output modulates the learning rate at which the internal model evolves \citep{jordan2023locus}.
-In our model, we do not consider precision-weighted PEs (but see \citep{KathaXXX}).
+In our model, we do not consider precision-weighted PEs \citep[but see][]{wilmes2023uncertainty}.
 Hence, a sensible extension to our work would be to adjust the weights from PE neurons onto the M neurons by a function of the stimulus and prediction uncertainty, respectively.
-This would allow us to compare our results more closely to work showing that ACh and NA can adjust the rate of incorporating new sensory evidence when environments change (\ref{XXX}).
-
-%—> XXX man sollte vielleicht noch sagen, dass auch PE neurons von neuromod direct beeinflusst sein können (es gab da zwei Beispiele, die weiter unten sind und die du hier aus Versehen gelo2scht hattest).
-—> Zum einen können diese auch zu biases beitragen, zum anderen konnten sie aber auch Effekt durch IN kompensieren.
-
-% —> XXX should we add in one SI precision-weighted PE to compare with Neuromodulation results? (Discuss with Katha and CC). If so, how to choose the weighting?
-% —> XXX From one trial to the next (so basically environment switch), should we check how fast the update happens when neuromods are applied to compare it to data?
-(Discuss with Katha and CC), —> should we also show or at least mention in discussion that by changing PE activity the prediction/mean can also be wrong?
-How is that actually checked in experiments or how do we know that somebody relies on prediction if the prediction might be wrong
-
-% Limitations, Alternations and future steps
-
-As for any computational model, we brush over a number of biological details in order to keep the model simple and interpretable.
-However, those details, while beyond the scope of this study, may be well investigated in future work.
-For instance, evidence suggests that prediction-errors itself are precision-weighted \ref{yon2021precision}.
-xx Kelly Diederen\ref{yon2021precision} xxx show that neural signatures of reward-prediction errors are augmented when the outcome is reliable and attenuated when the outcome is uncertain.
-XXX Katha’s work XXX
-
-Furthermore, while we provide a neuronal circuit model for estimating the mean and the variance of both sensory signals and predictions, we do not model the weighting of inputs explicitly.
-A respective neural circuit model would require divisive inhibition.
-XXX (look in paper for neural implementation of Kalman filter) xxx have shown how this can be done with inhibitory interneurons.
-
-Our model suggests one potential neuronal circuit mechanism for the estimation of sensory inputs and predictions.
+This would allow us to compare our results more closely to work showing that ACh and NA can adjust the rate at which new sensory evidence is incorporated when environments change \citep{marshall2016pharmacological, bruckner2022understanding}.
+
+Our model suggests \textit{one} potential neuronal circuit mechanism for the estimation of sensory inputs and predictions.
 However, in the light of evidence showing that the integration of feedforward and feedback inputs is species- and modality-dependent, it is conceivable that a plethora of neural mechanisms are used in neural circuits.
-For multisensory integration, \citep{wong2023computational} showed that in Drosophila larva the chosen cue-combination strategy varies depending on the type of sensory information available.
-Also, humans put typically more weight on visual than auditory cues ((Battaglia et al., 2003; Burr \& Alais, 2004).
-… in \citep{o2012can}), but trust vestibular information more than visual information about head direction (Butler et al., 2010 \citep{o2012can}), a finding also observed for monkeys (Angelaki et al., 2009 \citep{o2012can}).
-Moreover, Summerfield et al.
-(2011) \citep{o2012can} showed that humans diverge from an optimal Bayesian strategy in very volatile environments and act according to their experience in the last trial.
-Moreover, it has been suggested that the brain may use different strategies to combine signals depending on the task demands \citep{o2012can}.
+For multisensory integration, \cite{wong2023computational} showed that in Drosophila larva the chosen cue-combination strategy varies depending on the type of sensory information available.
+Also, humans put typically more weight on visual than auditory cues \citep{battaglia2003bayesian, alais2004ventriloquist}, but trust vestibular information more than visual information about head direction \citep{butler2010bayesian}, a finding also observed for monkeys \citep{fetsch2009dynamic}.
+Moreover, \cite{summerfield2011perceptual} showed that humans diverge from an optimal Bayesian strategy in very volatile environments and act according to their experience in the last trial.
+It has been suggested that the brain may use different strategies to combine signals depending on the task demands \citep{o2012can}.
 While these behavioral results cannot speak to the underlying circuit mechanisms, it is conceivable that neural implementations for the integration of feedforward and feedback inputs may also vary.
 
+Furthermore, while we provide a neuronal circuit model for estimating the mean and variance of both sensory signals and predictions, we do not explicitly model the weighting of inputs.
+A respective neural circuit model would require nested inhibitory interneurons providing divisive inhibition.
+How this subnetwork interacts with the PE circuits, and how the presence of neuromodulators acting on the interneurons directly involved in the weighting, impacts our findings is subject to future work.
+
 % Relation to other work that came before and closing remarks
-
-A number of normative models have been proposed for state estimation and prediction under uncertainty \citep{soltani2019adaptive}, ranging from the classical Kalman filter to more recent models like \textit{Bayes Factor Surprise} \citep{liakoni2021learning}.
-For instance, Bayes factor surprise formularizes the trade-off between integrating new observations in an existing belief system and resetting this belief system with the new evidence.
+\subsection*{Relation to other work \& conclusions}
+%
+Many normative models have been proposed for state estimation and prediction under uncertainty \citep{soltani2019adaptive}, ranging from the classical Kalman filter to more recent models like \textit{Bayes Factor Surprise} \citep{liakoni2021learning}.
+For instance, the Bayes factor surprise formularizes the trade-off between integrating new observations in an existing belief system and resetting this belief system with novel evidence.
 The surprise factor captures how much an animal’s current belief deviates from the new observation.
 
-In more recent years, normative model have been squared with biological constraints.
-For instance, in a paper by \citep{kutschireiter2023bayesian}, uncertainty could be integrated into a ring attractor model encoding head direction.
+In recent years, normative models have been squared with biological constraints.
+For instance, in a paper by \cite{kutschireiter2023bayesian}, uncertainty could be integrated into a ring attractor model encoding head direction.
 This \textit{Bayesian ring attractor} model encodes uncertainty in the amplitude of the network activity and matches the performance of a circular Kalman filter when the recurrent connections are tuned appropriately.
-
-XXX Come back to your model XXX highlight the nice thing about it and then find some concluding remarks
-
-% XXXX PLEASE LOOK AT MILIIDGE 2023 that you reviewed!!!
+In other seminal work, it has been proposed that Bayesian inference in time can be linked to the dynamics of leaky integrate-and-fire neurons with spike-dependent adaptation \citep{deneve2008bayesian}.
+
+Here, we proposed an alternative view in which PE neurons serve as the backbone for estimating both the uncertainty of the feedforward sensory stimuli arising from the external world and the feedback signals carrying predictions about the same feedforward inputs our brains are bombarded with.
+Our work is an important step toward a better understanding of the brain’s ability to integrate these often unreliable feedforward and feedback signals that often do not match perfectly.
 
@@
 %
-We arithmetically calculated the weighted output of sensory inputs and predictions, $r_\mathrm{out}$, based on ideas of Bayesian multisensory integration (XXX),
+We arithmetically calculated the weighted output of sensory inputs and predictions, $r_\mathrm{out}$, based on ideas of Bayesian multisensory integration \citep[see, e.g.][]{pouget2013probabilistic},
 %
@@
 
-\bibliographystyle{naturemag} %{plainnat}
+%\bibliographystyle{naturemag} %{plainnat}
+\bibliographystyle{plainnat}
 \bibliography{References_HertaegWilmesClopath_2023}
@@
 %
-The weighted output $r_\mathrm{out}(t) $ is a linear combination of the current sensory input $s(t)$ and the activity of the memory neuron, $r_\mathrm{M}(t)$, inspired by Bayesian multisensory integration (XXX)
+The weighted output $r_\mathrm{out}(t) $ is a linear combination of the current sensory input $s(t)$ and the activity of the memory neuron, $r_\mathrm{M}(t)$, inspired by Bayesian multisensory integration \citep[see, e.g.][]{pouget2013probabilistic},
 %
@@
 \textbf{Figure 4 Supplementary Fig.
-2:} In the default setting baseline was $0$ and gain of PE neurons was $1$ (see Eqs.
-XXX).
+2:} In the default setting baseline was $0$ and gain of PE neurons was $1$.
 The results (left) were computed for baselines $\in [0, 3]$, while the results (right) were computed for gains $\in [0.5, 1.5]$.
-XXX Here refer to equations XXX.
+The results are based on the Eqs.
+(\ref{eq:prediction_gain}), (\ref{eq:variance_gain}), (\ref{eq:condition_baseline_mean_1}) and  (\ref{eq:condition_baseline_variance_2}).
 \\
@@
 %
-%XXX Comparison to Bayes Factor surprise
+% Comparison to Bayes Factor surprise
 
@@
 \caption{\footnotesize{\bf Biased mean and variance estimation by changing the baseline and the gain of nPE and pPE.}
-In a toy model, described in XXX, the contribution of gain and baseline to the changes in the mean and variance estimation can be revealed. XXX Equations
+In a toy model, described in sections \ref{sec:gain_impact} and \ref{sec:impact_baseline}, the contribution of gain and baseline to the changes in the mean and variance estimation are summarized.
+The results are based on the Eqs.
+(\ref{eq:prediction_gain}), (\ref{eq:variance_gain}), (\ref{eq:condition_baseline_mean_1}) and  (\ref{eq:condition_baseline_variance_2}).
 }

### 2023-11-24 records/manuscript_2023.tex
add first part of feedback

@@
 In our network, the uncertainty estimation, and, hence, how much we rely on predictions, can be biased by perturbing the intricate interplay of different inhibitory interneurons.
-Finally, we demonstrate the link to biased perception and unravel how the different types of uncertainty contribute to the contraction bias.
+We finally show the link to biased perception and unravel how the different types of uncertainty contribute to the contraction bias.
 %Hence, our work suggests a neural circuit level mechanism underlying the brain's astonishing ability to integrate feedforward and feedback signals.
@@
 To test the network's ability to estimate the variances correctly, we stimulated the network with a sequence of inputs whose mean can vary from trial to trial.
-More precisely, in each trial, the network is presented with a stimulus that is composed of \textit{n} constant, consecutive values drawn from a normal distribution.
+More precisely, in each trial, the The network is given a stimulus that is composed of $N_\mathrm{in}$ constant values, each drawn from a normal distribution and presented for $N_\mathrm{step}$ consecutive time steps.
 The variance of this distribution represents the stimulus noise.
@@
 
-What are the network mechanisms underlying these observation?
+What are the network mechanisms underlying these observations?
 The sensory weight is a function of the lower and higher V neuron activity.
@@
 
-While multisensory integration is concerned with the weighting of sensory cues coming from different modalities, it is conceivable that the same ideas hold for top-down predictions and bottom-up sensory inputs that are combined in cortical circuits \citep{kording2004bayesian, yon2021precision}.
+%While multisensory integration is concerned with the weighting of sensory cues coming from different modalities, it is conceivable that the same ideas hold for top-down predictions and bottom-up sensory inputs that are combined in cortical circuits \citep{kording2004bayesian, yon2021precision}.
 Indeed, there is evidence that the brain can rely more on top-down or bottom-up signals \citep{pakan2018impact, han2023behavior}.
@@
 
-Furthermore, we show that the weighting of sensory inputs and predictions can be biased through neuromodulators, as has been suggested before \citep[see, e.g.,][]{yon2021precision}.
+Furthermore, we show that the weighting of sensory inputs and predictions can be biased by neuromodulators, as has been suggested before \citep[see, e.g.,][]{yon2021precision}.
 In our model, those modulatory signals act through interneurons \citep{cardin2019functional} whose activities increase in the presence of neuromodulators.
@@
 \begin{align}
-T \cdot \underline{\dot{h}}_\mathrm{PE}^\mathrm{low} =& -\underline{h}_\mathrm{PE}^\mathrm{low} + W_\mathrm{PE\leftarrow PE} \cdot \underline{r}_\mathrm{PE}^\mathrm{low} + \underline{w}_\mathrm{PE\leftarrow M} \cdot r_\mathrm{M}^\mathrm{low} + \underline{w}_\mathrm{PE\leftarrow FF} \cdot s + \underline{I}_\mathrm{PE} \nonumber\\
-%
-T \cdot \underline{\dot{h}}_\mathrm{PE}^\mathrm{high} =& -\underline{h}_\mathrm{PE}^\mathrm{high} + W_\mathrm{PE\leftarrow PE} \cdot \underline{r}_\mathrm{PE}^\mathrm{high} + \underline{w}_\mathrm{PE\leftarrow M} \cdot r_\mathrm{M}^\mathrm{high} + \underline{w}_\mathrm{PE\leftarrow FF} \cdot r_\mathrm{M}^\mathrm{low} + \underline{I}_\mathrm{PE}.
+T_{c} \cdot \underline{\dot{h}}_\mathrm{PE}^\mathrm{low} =& -\underline{h}_\mathrm{PE}^\mathrm{low} + W_\mathrm{PE\leftarrow PE} \cdot \underline{r}_\mathrm{PE}^\mathrm{low} + \underline{w}_\mathrm{PE\leftarrow M} \cdot r_\mathrm{M}^\mathrm{low} + \underline{w}_\mathrm{PE\leftarrow FF} \cdot s + \underline{I}_\mathrm{PE} \nonumber\\
+%
+T_{c} \cdot \underline{\dot{h}}_\mathrm{PE}^\mathrm{high} =& -\underline{h}_\mathrm{PE}^\mathrm{high} + W_\mathrm{PE\leftarrow PE} \cdot \underline{r}_\mathrm{PE}^\mathrm{high} + \underline{w}_\mathrm{PE\leftarrow M} \cdot r_\mathrm{M}^\mathrm{high} + \underline{w}_\mathrm{PE\leftarrow FF} \cdot r_\mathrm{M}^\mathrm{low} + \underline{I}_\mathrm{PE}.
 %
@@
 The connection strengths between the \textit{pre}-synaptic neuron (population) and the neurons of the PE circuit are denoted by $W_\mathrm{PE\leftarrow pre}$ and $\underline{w}_{\mathrm{PE} \leftarrow pre}$, respectively.
-The activities of the neurons evolve with time constants summarized in $T$.
+The activities of the neurons evolve with time constants summarized in $T_{c}$.
 
@@
 The network receives feedforward stimuli $s$ that may vary between trials.
-To account for noise, each stimulus is composed of N$_\mathrm{in}$ consecutive, piece-wise constant values drawn from a normal distribution with mean $\mu_\mathrm{in}$ and standard deviation $\sigma_\mathrm{in}$.
+To account for noise, each stimulus is composed of N$_\mathrm{in}$ constant values drawn from a normal distribution with mean $\mu_\mathrm{in}$ and standard deviation $\sigma_\mathrm{in}$, and are presented for N$_\mathrm{step}$ consecutive time steps.
 To account for changes in the environment, $\mu_\mathrm{in}$ is drawn from a uniform distribution $U(a,b)$ with mean $\mu_\mathrm{trial} = \frac{a+b}{2}$ and standard deviation $\sigma_\mathrm{trial} = \frac{b-a}{\sqrt{12}}$.
@@
 %
-%We are grateful to Vezha Boboeva, Douglas Feitosa Tom\'e, J\'ulia Gallinaro and Klara Kaleb for helpful comments on earlier versions of this manuscript, and we want to thank all members of the Clopath lab for discussion and support.
-This work was supported by BBSRC BB/N013956/1, BB/N019008/1, Wellcome Trust 200790/Z/16/Z, Simons Foundation 564408 and EPSRC EP/R035806/1.
+This work was supported by Deutsche Forschungsgemeinschaft (DFG) Grant 460088091, Biotechnology and Biological Sciences Research Council (BBSRC) Grants BB/N013956/1 and BB/N019008/1, Wellcome Trust Grant 200790/Z/16/Z, Simons Foundation Grant 564408, and Engineering and Physical Sciences Research Council (EPSRC) Grant EP/R035806/1.
 
@@
 In the simulations, we define the bias as the trial-averaged difference between the weighted output and the true stimulus.
-For the sake of simplicity, we use $r_\mathrm{out}$ at the end of a trial as a proxy for the trial average in the subsequent analysis.
+For the sake of simplicity, we use $r_\mathrm{out}$ at the end of a trial as a proxy for the trial, $T$, average in the subsequent analysis.
 Hence,
@@
 The absolute slope $\left( 1 -\alpha_\mathrm{S} \right)\cdot \mathrm{e}^{-T/\tau_M}$ indicates how strong the bias is.
-It depends on the sensory weight $\alpha_\mathrm{S}$, the trial duration $T$ and time constant $\tau$.
+It depends on the sensory weight $\alpha_\mathrm{S}$, the trial duration $T$ and time constant $\tau_M$.
 Please note that the sensory weight is a function of the trial duration itself (see Fig.
@@
 {\bf (A)} Illustration of the rate-based population network and the stimuli over time.
+The weights from the PE neurons onto the M or V neuron are scaled by a factor $\gamma$ drawn from a normal distribution $N(\mu_\mathrm{\gamma}, \sigma_\mathrm{\gamma})$.
 {\bf (B)} M and V neuron activities over time for one example parameterisation.
-{\bf (C)} The normalised absolute difference between the averaged mean and the activity of the M neuron (dark green) or between the averaged variance and the activity of the V neuron (brown) for uncorrelated deviations, that is, increasing SD of $\gamma$ (left), correlated deviations, that is, increasing mean of $\gamma$ (middle), and the network sparsity.
+{\bf (C)} The normalised absolute difference between the averaged mean and the activity of the M neuron (dark green) or between the averaged variance and the activity of the V neuron (brown) for uncorrelated deviations, that is, increasing $\sigma_\mathrm{\gamma}$ (left), correlated deviations, that is, increasing $\mu_\mathrm{\gamma}$ (middle), and the network sparsity.
 }

### 2023-12-08 records/manuscript_2023_v2.tex
housekeeping

--- 2023-11-24 records/manuscript_2023.tex
@@
 \begin{abstract}
-At any moment, our brains receive a stream of feedforward sensory stimuli arising from the world we interact with.
-Simultaneously, neural circuits are shaped by feedback signals carrying predictions about the same feedforward inputs our brains are bombarded with.
+At any moment, our brains receive a stream of sensory stimuli arising from the world we interact with.
+Simultaneously, neural circuits are shaped by feedback signals carrying predictions about the same inputs we experience.
 Those feedforward and feedback inputs often do not perfectly match.
-Thus, our brains have the challenging task of integrating these conflicting information according to their reliabilities.
-However, how neural circuits keep track of both the sensory and prediction uncertainties is not well understood.
-Here, we propose a network model whose core is hierarchical prediction-error circuits.
-We show that our network can estimate the variance of the sensory stimuli and the uncertainty in the prediction using the activity of negative and positive prediction-error neurons.
-In line with previous hypotheses, we demonstrate that neural circuits rely strongly on the feedback predictions at the beginning of a new stimulus, and if the environment is stable and the sensory cues are noisy.
-In addition, we investigate the mechanisms underlying a pathological weighting of feedforward and feedback inputs.
-In our network, the uncertainty estimation, and, hence, how much we rely on predictions, can be biased by perturbing the intricate interplay of different inhibitory interneurons.
-We finally show the link to biased perception and unravel how the different types of uncertainty contribute to the contraction bias.
+Thus, our brains have the challenging task of integrating these conflicting streams of information according to their reliabilities.
+However, how neural circuits keep track of both the stimulus and prediction uncertainty is not well understood.
+Here, we propose a network model whose core is a hierarchical prediction-error circuit.
+We show that our network can estimate the variance of the sensory stimuli and the uncertainty of the prediction using the activity of negative and positive prediction-error neurons.
+In line with previous hypotheses, we demonstrate that neural circuits rely strongly on feedback predictions if the perceived stimuli are noisy and the underlying generative process, that is, the environment is stable.
+Moreover, we show that predictions modulate neural activity at the onset of a new stimulus, even if this sensory information is reliable.
+In our network, the uncertainty estimation, and, hence, how much we rely on predictions, can be influenced by perturbing the intricate interplay of different inhibitory interneurons.
+We, therefore, investigate the contribution of those inhibitory interneurons to the weighting of feedforward and feedback inputs.
+Finally, we show that our network can be linked to biased perception and unravel how stimulus and prediction uncertainty contribute to the contraction bias.
 %Hence, our work suggests a neural circuit level mechanism underlying the brain's astonishing ability to integrate feedforward and feedback signals.
@@
 Experimental evidence suggests that these PEs may be represented in the activity of distinct groups of neurons, termed PE neurons \citep{eliades2008neural, keller2009neural, ayaz2019layer, audette2021temporally}.
-Moreover, these neurons may come in two types when excitatory neurons exhibit near-zero, spontaneous firing rates \citep{rao1999predictive, keller2018predictive}.
-Negative PE (nPE) neurons only increase their activity when the prediction is \textit{stronger} than the sensory input, while positive PE (pPE) neurons only increase their activity when the prediction is \textit{weaker} than the sensory input.
+Moreover, these neurons may come in two types when excitatory neurons exhibit near-zero, spontaneous firing rates \citep{rao1999predictive, keller2018predictive}: negative PE (nPE) neurons only increase their activity when the prediction is \textit{stronger} than the sensory input, while positive PE (pPE) neurons only increase their activity when the prediction is \textit{weaker} than the sensory input.
 Indeed, it has been shown that excitatory neurons in rodent primary sensory areas can encode negative or positive PEs \citep{keller2012sensorimotor, attinger2017visuomotor, jordan2020opposing, audette2021temporally}.
@@
 By breaking this balance, the PE neurons change their baseline firing rate and gain, leading to a biased variance estimation.
-Finally, we show that this weighting can be understood as a neural manifestation of the contraction bias \citep{hollingworth1910central, jazayeri2010temporal, ashourian2011bayesian, petzschner2011iterative, akrami2018posterior, meirhaeghe2021precise}.
+Finally, we show that this weighting can be understood as a neural manifestation of the contraction bias, that is, the magnitude of the represented sensory input is biased towards the mean of the past stimuli experienced \citep{hollingworth1910central, jazayeri2010temporal, ashourian2011bayesian, petzschner2011iterative, akrami2018posterior, meirhaeghe2021precise}.
 
@@
 Right: When climbing down an unexplored mountain in foggy conditions, the brain might need to integrate sensory information and predictions simultaneously.
-{\bf (B)} Illustration of a prediction-error (PE) circuit with both negative and positive PE (nPE/pPE) neurons that receive inhibition from three different inhibitory interneuron types: parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons.
+{\bf (B)} Top: Illustration of a prediction-error (PE) circuit with both negative and positive PE (nPE/pPE) neurons that receive inhibition from three different inhibitory interneuron types: parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons.
 Local excitatory connections are not shown for clarity.
+Bottom: Responses of an nPE and pPE neuron.
+The nPE neuron only increases its activity relative to a baseline when the sensory input is weaker than predicted, while the pPE neuron only increases its activity relative to a baseline when the sensory input is stronger than predicted.
 {\bf (C)} Illustration of network model that estimates the mean and variance of the external sensory stimuli.
@@
 \caption{\footnotesize{\bf Prediction-error neurons as the basis for estimating mean and variance of sensory stimuli.\newline} 
-{\bf (A)} Illustration of the inputs with which the network (Fig. \ref{fig:Fig_1}C) is stimulated. Network is exposed to a sequence of constant stimuli drawn from a uniform distribution.
+{\bf (A)} Illustration of the inputs with which the network (Fig.
+\ref{fig:Fig_1}C) is stimulated.
+Network is exposed to a sequence of constant stimuli drawn from a uniform distribution.
+The gray shaded boxes symbolize different values from the distribution.
 {\bf (B)} PE neuron activity hardly changes with stimulus strength (left) but strongly increases with stimulus variability (right).
@@
 
-To test our hypothesis, we study a rate-based mean-field network.
-The core network is a prediction-error (PE) circuit with excitatory nPE and pPE neurons, as well as inhibitory parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (Fig.
+To test our hypothesis, we study a rate-based mean-field network: the core network is a prediction-error (PE) circuit with excitatory nPE and pPE neurons, as well as inhibitory parvalbumin-expressing (PV), somatostatin-expressing (SOM), and vasoactive intestinal peptide-expressing (VIP) interneurons (Fig.
 \ref{fig:Fig_1}B).
@@
 Following \cite{keller2018predictive}, we assume that the pPE neuron excites the memory neuron, while the nPE neuron inhibits this neuron (for instance, through lateral inhibition, here not modeled explicitly).
-Because feedback connections are shown to target the apical dendrites of pyramidal cells \citep{larkum2013cellular} and interneurons located in superficial layers of the cortex \citep[see, e.g.][]{tremblay2016gabaergic}, the M neuron makes connections with the dendritic compartment of the PE neurons and some of the interneurons (here, VIP and PV neurons, see Methods for more details).
+Because feedback connections are shown to target the apical dendrites of pyramidal cells \citep{larkum2013cellular} and interneurons located in superficial layers of the cortex \citep[see, e.g.][]{tremblay2016gabaergic}, the memory neuron makes connections with the dendritic compartment of the PE neurons and some of the interneurons (here, VIP and PV neurons, see Methods for more details).
 We, furthermore, simulate a downstream neuron (termed V neuron), modeled as a leaky integrator with a quadratic activation function, that receives excitatory synapses from the PE neurons.
-Hence, in this setting, the V neuron encodes the variance of the sensory stimuli (Fig.
-\ref{fig:Fig_1}C).
+In this setting, the V neuron encodes the variance of the sensory stimuli (see the lower-level subnetwork in Fig.
+\ref{fig:Fig_1}C, the higher-level circuit is described later).
 
@@
 \ref{fig:Fig_2}C).
-The activity of the M neuron gradually approaches the mean of the sensory inputs (Fig.
+The activity of the memory neuron gradually approaches the mean of the sensory inputs (Fig.
 \ref{fig:Fig_2}D, middle), while the activity of the V neuron approaches the variance of those inputs (Fig.
@@
 
-We verified our results in a heterogeneous network in which a population of neurons represents each neuron type of the PE circuit, and the synaptic connection strengths from each PE neuron onto the M and V neuron are different (see Methods, Fig.
+We verified our results in a heterogeneous network in which a population of neurons belongs to each neuron type of the PE circuit, and the synaptic connection strengths from each PE neuron onto the M and V neuron are different (see Methods, Fig.
 \ref{fig:Fig_2_S2}A).
@@
 \ref{fig:Fig_2_S2}E).
-When all connection strengths are collectively shifted to higher values, the error increases for the V neuron, while it remains unaffected for the M neuron.
+When all connection strengths are collectively shifted to higher values, the error increases for the variance neuron, while it remains unaffected for the memory neuron.
 
@@
 We hypothesize that the same principles apply to computing the variance of the prediction.
-Hence, we augment the network with a \textit{higher} PE circuit that receives feedforward synapses from the M neuron of the \textit{lower} PE circuit (Fig.
+To show this, we augment the network with a \textit{higher} PE circuit that receives feedforward synapses from the memory (M) neuron of the \textit{lower} PE circuit (Fig.
 \ref{fig:Fig_1}D).
@@
 To account for stimulus variability, each stimulus is represented by $10$ stimulus values drawn from a normal distribution.
-To account for the volatility of the environment, in each trial, the stimulus mean is drawn from a uniform distribution (denoted trial variability).
+To account for the volatility of the environment, in each trial, the stimulus mean is drawn from a uniform distribution (denoted trial-to-trial variability).
 {\bf (B)} Illustration of how the weighted output is calculated.
 The sensory weight $\alpha$ lies between zero (system relies perfectly on prediction) and one (system relies solely on the sensory input).
-{\bf (C)} Limit case example in which the stimulus variability is zero but the trial variability is high.
+{\bf (C)} Limit case example in which the stimulus variability is zero but the trial-to-trial variability is high.
 Left: Illustration of the stimulation protocol.
@@
 Input statistics shown in E.
-{\bf (D)} Limit case example in which the stimulus variability is high but the trial variability is zero.
+{\bf (D)} Limit case example in which the stimulus variability is high but the trial-to-trial variability is zero.
 Left: Illustration of the stimulation protocol.
@@
 Input statistics shown in E.
-{\bf (E)} Sensory weight for different input statistics. Predictions are weighted more strongly when the stimulus variability is larger than the trial variability.
-{\bf (F)} Sensory weight throughout a trial for two different trial durations. Predictions are weighted more strongly at the beginning of a new trial. 
+{\bf (E)} Sensory weight for different input statistics. Predictions are weighted more strongly when the stimulus variability is larger than the trial-to-trial variability.
+{\bf (F)} Sensory weight, averaged over many trials, for two different trial durations. Gray shading denotes the SEM. Predictions are weighted more strongly at the beginning of a new trial. 
 }
@@
 To test the network's ability to estimate the variances correctly, we stimulated the network with a sequence of inputs whose mean can vary from trial to trial.
-More precisely, in each trial, the The network is given a stimulus that is composed of $N_\mathrm{in}$ constant values, each drawn from a normal distribution and presented for $N_\mathrm{step}$ consecutive time steps.
+More precisely, in each trial, the network is given a stimulus that is composed of $N_\mathrm{in}$ constant values, each drawn from a normal distribution and presented for $N_\mathrm{step}$ consecutive time steps.
 The variance of this distribution represents the stimulus noise.
@@
 \ref{fig:Fig_3}A).
-Hence, the inputs change on two different time scales, where the stimulus variability has a faster time scale than the trial variability.
+Hence, the inputs change on two different time scales, where the stimulus variability has a faster time scale than the trial-to-trial variability.
 
@@
 The weights assigned to each input stream are functions of the uncertainties, that is, the activities of the V neurons.
-The sensory weight captures, hence, how much the network relies on the sensory input (Fig.
+The sensory weight captures how much the network relies on the sensory input (Fig.
 \ref{fig:Fig_2}B).
 To test our network, we first consider two limit cases.
-In the first limit case, a different, low-variance stimulus is presented in each trial (Fig.
+In the first limit case, we show a low-variance stimulus that differs in each trial (low stimulus uncertainty, high trial-to-trial uncertainty, see Fig.
 \ref{fig:Fig_3}C, left).
@@
 \ref{fig:Fig_3}C, right), the network indeed shows a clear preference for the sensory input.
-In the second limit case, the same, high-variance stimulus is presented in each trial (Fig.
+In the second limit case, we show a high-variance stimulus the mean of which does not change from trial to trial (high stimulus uncertainty, low trial-to-trial uncertainty, see Fig.
 \ref{fig:Fig_3}D, left).
@@
 \ref{fig:Fig_3}F).
-Our model makes therefore the following experimentally testable prediction: Sensory predictions influence neural activity more significantly in experiments that rely on fast stimulus changes.
+Our model makes therefore the following experimentally testable prediction: sensory predictions influence neural activity more significantly in experiments that rely on fast stimulus changes.
 
@@
 We thus wondered which network properties might bias the estimation of the variances, and, consequently, the weighting of different input streams.
-We identified the time scales at which the M neurons incorporate new information as a decisive factor in the integration of inputs.
-To show this, we varied the weights from the PE neurons onto the lower-level M neuron.
-If the weights are too small (M updates too slowly), the system relies too much on feedback predictions.
-In contrast, if the weights are too large (M updates too fast), the system relies too much on the feedforward sensory information (Fig.
+We identified the time scales at which the memory neurons incorporate new information as a decisive factor in the integration of inputs.
+To show this, we varied the weights from the PE neurons onto the lower-level memory neuron.
+If the weights are too small (the memory neuron updates too slowly), the system relies too much on feedback predictions.
+In contrast, if the weights are too large (the memory neuron updates too fast), the system relies too much on the feedforward sensory information (Fig.
 \ref{fig:Fig_3_S2}A).
-While the speeds at which the activity of the M neurons evolve influence the weighting of inputs, the precise activation function of the V neurons is less pivotal.
+While the speeds at which the activity of the memory neurons evolve influence the weighting of inputs, the precise activation function of the variance neurons is less pivotal.
 When we replaced the quadratic activation function with a linear, rectified function, the V neurons did not encode the variance but the average absolute deviation of the sensory stimuli.
@@
 {\bf (C)} The V neurons' activities depend on the PE neurons.
-Hence, perturbing the nPE and pPE neurons changes the uncertainty estimation.
-While stimulating the lower PE neurons affects both the lower and higher-order V neurons (right), stimulating the higher-order PE neurons only affects the V neuron in the same subnetwork (left).
+Hence, perturbing the nPE and pPE neuron changes the uncertainty estimation.
+Left: stimulating the pPE (triangle) or nPE (upside down triangle) affects the V neuron of the same subnetwork, denoted by matching marker and line colors (dark brown: lower circuit, light brown: higher circuit).
+Right: while stimulating the lower PE neurons affects the higher-order V neuron, stimulating the higher-order PE neurons does not change the activity of the lower-order V neuron (denoted by not matching marker and line colors).
 {\bf (D)} The V neuron activity, and hence the sensory weight, changes as a result of the modulated PE neuron activity.
@@
 
-To this end, we modeled the presence of a neuromodulator by injecting an additional excitatory input into an interneuron type.
+To this end, we modeled the presence of a neuromodulator by injecting an additional excitatory input into an interneuron type (while a neuromodulator can also suppress neuronal activity, we focus on the more common excitatory effects that have been described).
 We reasoned that the network effect of a neuromodulator not only depends on the interneuron type it targets but also on the inputs this neuron receives and the connections it makes with other neurons in the network.
@@
 What are the network mechanisms underlying these observations?
-The sensory weight is a function of the lower and higher V neuron activity.
+The sensory weight is a function of the lower and higher variance (V) neuron activity.
 Hence, any changes to the sensory weight result from changes to the neurons encoding the variances.
 In our network, the V neurons only receive excitatory synapses from PE neurons.
-Hence, any changes in the sensory weights upon activation of interneurons must be due to changes in the PE neurons.
+As a consequence, any changes in the sensory weights upon activation of interneurons must be due to changes in the PE neurons.
 To disentangle the effect of nPE and pPE neurons, we perturbed those neurons individually in both the lower or higher subnetwork by injecting either an inhibitory or excitatory additional input (Fig.
@@
 Moreover, the higher-level V neuron is also slightly affected.
-This is because the lower-level M neuron is also modulated by the lower-level PE neurons and makes feedforward connections to the higher-level PE circuit.
+This is because the lower-level memory neuron is also modulated by the lower-level PE neurons and makes feedforward connections to the higher-level PE circuit.
 In contrast, stimulating either PE neuron in the higher subnetwork increases the activity of the higher-level V neuron but leaves the lower-level neurons unaffected.
@@
 The summed effect over nPE and pPE neuron (Fig.
-\ref{fig:Fig_4_S4}), hence, suggests that whether the activity of the V neuron increases or decreases depends on the input statistics: for low-mean stimuli, the elevated baseline activity dominates the changes in the variance, while for high-mean stimuli the changes in the gain dominate.
+\ref{fig:Fig_4_S4}) suggests that whether the activity of the V neuron increases or decreases depends on the input statistics: for low-mean stimuli, the elevated baseline activity dominates the changes in the variance, while for high-mean stimuli the changes in the gain dominate.
 
@@
 Bias is defined as the weighted output minus the stimulus mean.
-The absolute value of the slope (see linear fit) is a measure of the bias.
+The absolute value of the slope of the linear fit, \textit{m}, is a measure of the bias.
 The larger the slope, the larger the bias.
-{\bf (B)}  As a consequence of the sensory weight, the slope increases with stimulus variability (bias increases) and decreases with trial variability (bias decreases).
-{\bf (C)} Bias is independent of the trial variability when the stimulus variability is zero.
-{\bf (D)} Bias is independent of the stimulus variability when the trial variability is zero.
+{\bf (B)}  As a consequence of the sensory weight, the slope increases with stimulus variability (bias increases) and decreases with trial-to-trial variability (bias decreases).
+{\bf (C)} Bias is independent of the trial-to-trial variability when the stimulus variability is zero.
+{\bf (D)} Bias is independent of the stimulus variability when the trial-to-trial variability is zero.
 {\bf (E)} The slope depends on the trial duration.
@@
 \ref{fig:Fig_5}E).
-Hence, we predict that the contraction bias can be reduced for sufficiently long trials.
+We, therefore, predict that the contraction bias can be reduced for sufficiently long trials.
 
@@
 
-This idea is consistent with Bayesian theories on the optimal integration of multiple sensory cues (aka multisensory integration).
+This idea is consistent with Bayesian theories on the optimal integration of multiple sensory cues (multisensory integration).
 \cite{ernst2002humans} showed that to estimate the height of a bar humans combine visual and haptic information in a fashion that minimizes the variance of the final estimate.
@@
 
+%Following the same ideas from Bayesian multisensory integration, the brain must hence have mechanisms to track both the reliability of the sensory signals and the predictions thereof \citep{yon2021precision}.
+
 \subsection*{Summary of findings}
-
-%While multisensory integration is concerned with the weighting of sensory cues coming from different modalities, it is conceivable that the same ideas hold for top-down predictions and bottom-up sensory inputs that are combined in cortical circuits \citep{kording2004bayesian, yon2021precision}.
-Indeed, there is evidence that the brain can rely more on top-down or bottom-up signals \citep{pakan2018impact, han2023behavior}.
-Following the same ideas from Bayesian multisensory integration, the brain must hence have mechanisms to track both the reliability of the sensory signals and the predictions thereof \citep{yon2021precision}.
 
@@
 \ref{fig:Fig_2} \& \ref{fig:Fig_3}).
-In our model, we assume a hierarchy of PE circuits that are feed-forwardly connected through the lower-order M neuron whose activity encodes the mean of the sensory bottom-up inputs.
+In our model, we assume a hierarchy of PE circuits that are feed-forwardly connected through the lower-order memory neuron whose activity encodes the mean of the sensory bottom-up inputs.
 This local prediction is fed back to the lower-order circuit and at the same time feed-forwarded to the higher-order subnetwork (Fig.
 \ref{fig:Fig_1}).
-With this architecture in place, we show that we rely more strongly on our internal signals when the perceived sensory cues are noisier than the predictions, either because the environment is stable but highly noisy or because the environment changes quickly.
-This implies that revising our inner model of the world, that is, learning from PEs, should be suppressed if the sensory noise is high or the environment switches rapidly \citep{herzfeld2014memory}.
-Moreover, our work suggests that when experimental trials are short, a neural network can better represent the external world by relying on predictions.
-Hence, studying neural signatures of predictions in the brain might require experiments that involve sufficiently short trials.
+With this architecture in place, we show that we rely more strongly on our internal signals when the perceived sensory cues are noisier than the predictions.
+Moreover, our work suggests that predictions modulate neural activity more at the onset of a new sensory input, even if the stimulus is not noisy.
+%This implies that revising our inner model of the world, that is, learning from PEs, should be suppressed if the sensory noise is high or the environment switches rapidly \citep{herzfeld2014memory}. 
+As a consequence, studying neural signatures of predictions in the brain might require experiments that involve sufficiently short trials.
 
@@
 While the sensory noise increases the contraction bias, the uncertainty in the prediction or long trials decreases the bias (Fig.
-\ref{fig:Fig_5}.
+\ref{fig:Fig_5}).
 However, we note that we only consider a neural representation of the stimulus and do not account for other sources of noise, like execution noise, that surely impacts the contraction bias observed in behavioral studies.
@@
 %
-In our model, we assumed that there are dedicated neurons that encode the variance of the feedforward sensory inputs and the prediction.
+In our model, we assumed that there are dedicated neurons that encode the variance of the feedforward sensory inputs and the prediction \citep[see also][]{wilmes2023uncertainty}.
 This assumption is consistent with the idea that neurons explicitly encode in their activity the parameters (for instance, mean or variance) of a probability distribution \citep{o2010coding, o2012can}.
@@
 
-There has been evidence that indeed (population of) neurons can encode (un-)certainty \citep{soltani2019adaptive}.
+There has been evidence that indeed (population of) neurons can encode uncertainty \citep{soltani2019adaptive}.
 For instance, neurons in the parietal cortex in monkeys encode the degree of confidence in a perceptual decision \citep{kiani2009representation}.
@@
 In our model, the nPE neuron inhibits the memory neuron while the pPE neuron excites it, in line with \cite{keller2018predictive}.
-The weights from those PE neurons onto the M neuron are larger for the lower than the higher PE circuit, so that the lower-order M neuron evolves faster than the higher-order counterpart.
+The weights from those PE neurons onto the memory neuron are larger for the lower than the higher PE circuit, so that the lower-order memory neuron evolves faster than the higher-order counterpart.
 This assumption is consistent with the observation that time constants increase along the cortical hierarchy \citep{murray2014hierarchy, chaudhuri2015large, runyan2017distinct}.
 
-If the relation of the paces at which the M neurons evolve is strongly modulated, the network either shows a bias towards the sensory cues or the prediction (Figs.
+If the relation of the paces at which the memory neurons evolve is strongly modulated, the network either shows a bias towards the sensory cues or the prediction (Figs.
 \ref{fig:Fig_3_S2}).
@@
 A popular hypothesis is that neuromodulators shape the weighting of sensory inputs and predictions thereof \citep{yon2021precision}.
-Theoretical work by \citep{yu2005uncertainty} suggests that acetylcholine (ACh) correlates with \textit{expected uncertainty}, while noradrenaline (NA) correlates with \textit{unexpected uncertainty}.
+Theoretical work by \cite{yu2005uncertainty} suggests that acetylcholine (ACh) correlates with \textit{expected uncertainty}, while noradrenaline (NA) correlates with \textit{unexpected uncertainty}.
 Expected uncertainty is usually interpreted as known cue-outcome unreliabilities.
@@
 The authors hypothesize that LC output modulates the learning rate at which the internal model evolves \citep{jordan2023locus}.
-In our model, we do not consider precision-weighted PEs \citep[but see][]{wilmes2023uncertainty}.
-Hence, a sensible extension to our work would be to adjust the weights from PE neurons onto the M neurons by a function of the stimulus and prediction uncertainty, respectively.
+In our model, we do not consider precision-weighted PEs \citep[but see][]{wilmes2023uncertainty, granier2023precision}.
+Hence, a sensible extension to our work would be to adjust the weights from PE neurons onto the memory neurons by a function of the stimulus and prediction uncertainty, respectively.
 This would allow us to compare our results more closely to work showing that ACh and NA can adjust the rate at which new sensory evidence is incorporated when environments change \citep{marshall2016pharmacological, bruckner2022understanding}.
@@
 In recent years, normative models have been squared with biological constraints.
-For instance, in a paper by \cite{kutschireiter2023bayesian}, uncertainty could be integrated into a ring attractor model encoding head direction.
-This \textit{Bayesian ring attractor} model encodes uncertainty in the amplitude of the network activity and matches the performance of a circular Kalman filter when the recurrent connections are tuned appropriately.
+For instance, \cite{kutschireiter2023bayesian} showed that a \textit{Bayesian ring attractor} model can encode uncertainty in the amplitude of the network activity and matches the performance of a circular Kalman filter when the recurrent connections are tuned appropriately.
 In other seminal work, it has been proposed that Bayesian inference in time can be linked to the dynamics of leaky integrate-and-fire neurons with spike-dependent adaptation \citep{deneve2008bayesian}.
 
-Here, we proposed an alternative view in which PE neurons serve as the backbone for estimating both the uncertainty of the feedforward sensory stimuli arising from the external world and the feedback signals carrying predictions about the same feedforward inputs our brains are bombarded with.
-Our work is an important step toward a better understanding of the brain’s ability to integrate these often unreliable feedforward and feedback signals that often do not match perfectly.
+Here, we proposed an alternative view in which PE neurons serve as the backbone for estimating both the uncertainty of the feedforward sensory stimuli arising from the external world and the feedback signals carrying predictions about the same feedforward inputs our brains receive.
+Our work is an important step toward a better understanding of the brain’s ability to integrate these unreliable feedforward and feedback signals that often do not match perfectly.
 
@@
 Each PE circuit contains an excitatory nPE neuron and pPE neuron ($\mathrm{N}_\mathrm{nPE} = \mathrm{N}_\mathrm{pPE} = 1$), as well as inhibitory neurons.
-The inhibitory neurons comprise PV, SOM and VIP neurons ($\mathrm{N}_\mathrm{SOM} = \mathrm{N}_\mathrm{VIP} = 1$, $\mathrm{N}_\mathrm{PV} = 2$).
+The inhibitory neurons comprise PV, SOM and VIP neurons ($\mathrm{N}_\mathrm{SOM} = \mathrm{N}_\mathrm{VIP} = 1$, $\mathrm{N}_\mathrm{PV} = 2$), further explained in \cite{hertag2022prediction}.
 In addition to the core PE circuit, each subnetwork also includes one memory neuron $M$ and one variance neuron $V$.
@@
 The network receives time-dependent stimuli $s$ and neuron/compartment-specific external background input $\underline{I}_\mathrm{PE}$.
-The connection strengths between the \textit{pre}-synaptic neuron (population) and the neurons of the PE circuit are denoted by $W_\mathrm{PE\leftarrow pre}$ and $\underline{w}_{\mathrm{PE} \leftarrow pre}$, respectively.
+The connection strengths between the \textit{pre}-synaptic population and the neurons of the PE circuit are denoted by $W_\mathrm{PE\leftarrow pre}$ (if $r$ is a vector) or $\underline{w}_{\mathrm{PE} \leftarrow pre}$ (if $r$ is a scalar).
 The activities of the neurons evolve with time constants summarized in $T_{c}$.
 
-The activities of the lower and higher M neuron evolve according to a perfect integrator. The memory neurons receive synapses from both nPE and pPE neurons of the same subnetwork,
+The activities of the lower and higher memory (M) neuron evolve according to a perfect integrator. The M neurons receive synapses from both nPE and pPE neurons of the same subnetwork,
 %
@@
 %
+We thank In\^es C.
+Guerreiro for comments on earlier versions of this manuscript.
 This work was supported by Deutsche Forschungsgemeinschaft (DFG) Grant 460088091, Biotechnology and Biological Sciences Research Council (BBSRC) Grants BB/N013956/1 and BB/N019008/1, Wellcome Trust Grant 200790/Z/16/Z, Simons Foundation Grant 564408, and Engineering and Physical Sciences Research Council (EPSRC) Grant EP/R035806/1.
@@
 \appendix
-
 \tableofcontents
+
+\setcounter{figure}{-1}
+\renewcommand{\thefigure}{S\arabic{figure}}
+
 \section{Detailed Methods}
@@
 In the following, we describe in more detail the equations for the dynamics of the neurons in the prediction-error circuit, as well as the memory and variance neurons.
-We then provide the connectivity of the network and the inputs to the neurons for both the mean-field and population model.
+We then provide the connectivity of the network and the inputs to the neurons for both the mean-field and multi-cell population model.
 Finally, to ensure reproducibility, we summarize all simulation parameters used for the results shown in the figures.
@@
 The inhibitory neurons comprise PV, SOM and VIP neurons ($\mathrm{N}_\mathrm{SOM} = \mathrm{N}_\mathrm{VIP} = 1$, $\mathrm{N}_\mathrm{PV} = 2$).
+As has been shown in \cite{hertag2022prediction}, we need two soma-targeting interneurons, one receiving the sensory input and one receiving the prediction, to obtain a perfect nPE and pPE neuron in the same recurrent network.
+We, therefore, used two PV neurons (as suggested in the original paper).
 In addition to the core PE circuit, each subnetwork also includes one memory neuron $M$ and one variance neuron $V$.
@@
 In Fig.
-\ref{fig:Fig_2_S2}, we replaced this lower mean-field PE circuit with a heterogeneous population model containing $200$ neurons ($\mathrm{N}_\mathrm{SOM} = \mathrm{N}_\mathrm{VIP} = \mathrm{N}_\mathrm{PV} = 20$, $140$ excitatory neurons).
+\ref{fig:Fig_2_S2}, we replaced this lower mean-field PE circuit with a heterogeneous multi-cell population model containing $200$ neurons ($\mathrm{N}_\mathrm{SOM} = \mathrm{N}_\mathrm{VIP} = \mathrm{N}_\mathrm{PV} = 20$, $140$ excitatory neurons).
 In Fig.
@@
 %
-where the weights $w_{\mathrm{nD\leftarrow nE}}$, $w_{\mathrm{nD\leftarrow pE}}$, $w_{\mathrm{pD\leftarrow nE}}$ and $w_{\mathrm{pD\leftarrow pE}}$ denote the recurrent excitatory connections between PCs, including backpropagating activity from the soma to the dendrites.
+where the weights $w_{\mathrm{nD\leftarrow nE}}$, $w_{\mathrm{nD\leftarrow pE}}$, $w_{\mathrm{pD\leftarrow nE}}$ and $w_{\mathrm{pD\leftarrow pE}}$ denote the recurrent excitatory connections between PCs.
 $w_{\mathrm{nD\leftarrow SOM}}$ and $w_{\mathrm{pD\leftarrow SOM}}$ represent the strength of dendritic inhibition from the SOM neuron.
@@
 \ref{fig:Fig_3_S2}, we used a linear activation function to investigate the impact of the input-output transfer function on the weighting of sensory inputs and predictions.
-The time constant $\tau_V$ was 5 s in the mean-field model, 2~s in the heterogeneous population model (Fig.
-\ref{fig:Fig_2_S2}), and 0.5~s in the population model with selectivity (Fig.
+The time constant $\tau_V$ was $5\,s$ in the mean-field model, $2\,s$ in the heterogeneous multi-cell population model (Fig.
+\ref{fig:Fig_2_S2}), and $0.5\,s$ in the network model with selectivity (Fig.
 \ref{fig:Fig_2_S3}).
@@
 \begin{align*}
-w_\mathrm{M\leftarrow nE}\  &=\ \frac{-\lambda}{g_\mathrm{nPE} \cdot N_\mathrm{nPE}} \nonumber\\
-w_\mathrm{M\leftarrow pE}\  &=\ \frac{\lambda}{g_\mathrm{pPE} \cdot N_\mathrm{pPE}}
+w_\mathrm{M\leftarrow nE}\  &=\ \frac{-\lambda^\mathrm{loc}}{g_\mathrm{nPE} \cdot N_\mathrm{nPE}} \nonumber\\
+w_\mathrm{M\leftarrow pE}\  &=\ \frac{\lambda^\mathrm{loc}}{g_\mathrm{pPE} \cdot N_\mathrm{pPE}}
 \end{align*}
 %
-where $\lambda$ denotes the speed at which the perfect integrator evolves.
-In the lower PE circuit,  $\lambda=3\cdot 10^{-3}$ for the mean-field model in Fig.
-\ref{fig:Fig_2} and $\lambda=4.5\cdot 10^{-2}$ otherwise.
-In the higher PE circuit, $\lambda = 7\cdot 10^{-4}$.
-For the heterogeneous population model, $\lambda=5\cdot 10^{-2}$ in Fig.
-\ref{fig:Fig_2_S2}, and for the population model with selectivity (Fig.
-\ref{fig:Fig_2_S3}), $ \lambda=3\cdot 10^{-1}$.
-
-For the mean-field networks ($ N_\mathrm{nPE} = N_\mathrm{pPE} = 1 $), the gain factors are given in Table~\ref{tab:gain_factors_MFN}.
+where $\lambda^\mathrm{loc}$ denotes a weight for the lower or higher-order PE circuit, $\mathrm{loc}\in\lbrace\mathrm{lower}, \mathrm{higher}\rbrace$.
+In the lower PE circuit,  $\lambda^\mathrm{lower}=3\cdot 10^{-3}$ for the mean-field model in Fig.
+\ref{fig:Fig_2} and $\lambda^\mathrm{lower}=4.5\cdot 10^{-2}$ for Figs.
+\ref{fig:Fig_3}-\ref{fig:Fig_5}.
+In the higher PE circuit, $\lambda^\mathrm{higher} = 7\cdot 10^{-4}$.
+
+For the mean-field networks ($ N_\mathrm{nPE} = N_\mathrm{pPE} = 1 $), the gain factors $g_\mathrm{nPE}$ and $g_\mathrm{pPE}$ are given in Table~\ref{tab:gain_factors_MFN}.
 For the population network, the gain factors for all PE neurons are shown in Fig.
@@
     \includegraphics{../results/figures/final/Figure_gains}
-\caption{\footnotesize{\bf Gain factors of nPE and pPE neurons in the population model.\newline}
-{The logarithm of the gain factors of nPE (top) and pPE (bottom) neurons in the population model from \citep{hertag2022prediction}.
+\caption{\footnotesize{\bf Gain factors of nPE and pPE neurons in the multi-cell population model.\newline}
+{The logarithm of the gain factors of nPE (top) and pPE (bottom) neurons in the multi-cell population model from \citep{hertag2022prediction}.
 The network contains $67$ nPE neurons and $66$ pPE neurons.
@@
 
-The memory neuron \textit{M} connects to the post-synaptic neurons \textit{X} in the PE circuit with the connection strength $w_\mathrm{X\leftarrow M}$.
-If a connection exists, we assume $w_\mathrm{X\leftarrow M} = 1$.
+The memory neuron \textit{M} connects to the post-synaptic neurons \textit{X} in the PE circuit with the connection strength $w_\mathrm{X\leftarrow M} = 1$, if a connection exists, $w_\mathrm{X\leftarrow M} = 0$ otherwise.
 In all mean-field networks and the population network, the dendrites of nPE and pPE neurons and one of the two (populations of) PV neurons receive connections from the memory neuron.
@@
 Whether the SOM or VIP neurons are the target of the feedback projections depend on the specific mean-field network (see Table \ref{tab:wXM}).
-In the population model, 30\% of the SOM neurons and 70\% of the VIP neurons receive input from the memory neuron.
+In the multi-cell population model, 30\% of the SOM neurons and 70\% of the VIP neurons receive input from the memory neuron.
 %
@@
 %
-We introduce the factor $\theta$ to compensate for cross-terms that are the result of a quadratic activation function but are not in line with the definition of the variance.
-
+The factor $\theta$ denotes the unscaled weight, and can be chosen to compensate for potential  deviations from the definition of the variance as a result of a quadratic activation function.
 In the mean-field network, $\theta = 1$ because nPE and pPE neuron activity is mutually exclusive, and, hence, the cross-term $\mathrm{nPE}\cdot\mathrm{pPE}$ would be zero (under the assumption that they have a negligible baseline activity).
-This is also true for the population model, in which each nPE neuron only contributes a small fraction to the overall, mean nPE, and each pPE neuron only contributes a small fraction to the overall, mean pPE.
+This is also true for the multi-cell population model.
+Each PE neuron receives the same feedforward stimulus and contributes only a small (scaled) fraction to the overall PE (adding up to the same PE used in the mean-field network).
 
 However, in Supp Fig.
-\ref{fig:Fig_2_S3}, each mean-field network receives a stimulus $s_i$ drawn from a distribution.
+\ref{fig:Fig_2_S3}, each mean-field network receives a different stimulus $s_i$ drawn from a distribution at time $t$.
+While the stimuli below the mean of the distribution activate the nPE neurons (each located in a different mean-field network), the stimuli above the mean of the distribution activate the pPE neurons (each located in a different mean-field network).
+Because the V neuron first sums all the contributions from the PE neurons before applying the non-linearity, its steady state activity is similar to the squared sum of the \textit{averaged} nPE and \textit{averaged} pPE neuron activity.
 In this case, $\theta$ must be chosen such that deviations from the true variance can be mitigated or fully corrected.
@@
 %
-for zero activity at time $t=0$. In the limit of $t\rightarrow \infty$ (steady state), this is exactly the exponential average of the feedforward input, $E(s_\mathrm{FF})$.
+for zero activity at time $t=0$. In the limit of $t\rightarrow \infty$ (steady state), this is the exponential moving average of the feedforward input, $E(s_\mathrm{FF})$.
 
@@
 %
-The gains of the PE neurons, if not equal between nPE and pPE neuron on average, can bias the activity of both the M and V neuron. To show this, we resume the toy model from section \ref{sec:toy}. 
+The gains of the PE neurons, if not equal between the nPE and pPE neuron on average, can bias the activity of both the M and V neuron.
+To show this, we resume the toy model from section \ref{sec:toy}.
 %
@@
 %
-The baselines of the PE neurons, if not equal between nPE and pPE neuron on average, can bias the activity of both the M and V neuron.
+The baselines of the PE neurons, if not equal between the nPE and pPE neuron on average, can bias the activity of both the M and V neuron.
 By means of the toy model from section \ref{sec:toy}, we can write
@@
 Thus, the M neuron encodes the true mean of the feedforward input only if $p_0 = n_0$.
-Hence, the mean is overpredicted if $p_0 > n_0$.
+As a result, the mean is overpredicted if $p_0 > n_0$.
 Likewise, the mean is underpredicted if $p_0 < n_0$ (see Fig.
@@
 %
-Hence, different neuromodulators may have the same effect on the sensory weight or the same neuromodulator may have different effects depending on brain area, species, etc..
+As a result, different neuromodulators may have the same effect on the sensory weight or the same neuromodulator may have different effects depending on brain area, species, etc..
 
@@
 In the simulations, we define the bias as the trial-averaged difference between the weighted output and the true stimulus.
-For the sake of simplicity, we use $r_\mathrm{out}$ at the end of a trial as a proxy for the trial, $T$, average in the subsequent analysis.
+For the sake of simplicity, we use $r_\mathrm{out}$ at the end of a trial, $T$, as a proxy for the trial average in the subsequent analysis.
 Hence,
@@
 
-In this toy model, if the variance of the prediction is zero (that is, in a prediction-driven input regime), $\alpha_\mathrm{S} \approx 0$, and, hence, the bias is $\mathrm{e}^{-T/\tau_M}\cdot \left(\mu - s \right)$.
+In this toy model, if the variance of the prediction is zero (that is, in a prediction-driven input regime), $\alpha_\mathrm{S} \approx 0$, and, consequently, the bias is $\mathrm{e}^{-T/\tau_M}\cdot \left(\mu - s \right)$.
 Thus, the bias is independent of the stimulus variance (see Fig.
@@
 
-Likewise, if the variance of the sensory stimulus is zero (that is, in a stimulus-driven input regime), $\alpha_\mathrm{S} \approx 1$, and, hence, the bias approaches $0$ if the neurons reach their steady state.
+Likewise, if the variance of the sensory stimulus is zero (that is, in a stimulus-driven input regime), $\alpha_\mathrm{S} \approx 1$, and, consequently, the bias approaches $0$ if the neurons reach their steady state.
 Thus, decreasing or increasing the trail variance does not have an effect on the bias (see Fig.
@@
 \section{Supplementary Figures}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}
 
@@
 {\bf (C)} The normalised absolute difference between the averaged mean and the activity of the M neuron (dark green) or between the averaged variance and the activity of the V neuron (brown) for uncorrelated deviations, that is, increasing $\sigma_\mathrm{\gamma}$ (left), correlated deviations, that is, increasing $\mu_\mathrm{\gamma}$ (middle), and the network sparsity.
+To speed up simulations, we chose $\lambda^\mathrm{lower}=5\cdot 10^{-2}$.
 }
@@
 The examples from B are shown with colored arrows and markers.
+To speed up simulations, we chose $\lambda^\mathrm{lower}=3\cdot 10^{-1}$.
 }
@@
 {\bf (B)} The sensory weight over time is shown for all transitions in (A).
-For the sake of clarity, we only show the trials 40 -60.
+For the sake of clarity, we only show the trials 40-60.
 The switch to new input statistics occurs at trial 50.
@@
     \includegraphics{../results/figures/final/Fig_4_S4}% [width=1\linewidth]
-\caption{\footnotesize{\bf The combined changes in baseline and gain of all PE neurons determine the shift in the snesory weight.\newline}  
+\caption{\footnotesize{\bf The combined changes in baseline and gain of all PE neurons determine the shift in the sensory weight.\newline}  
 Whether and how a neuromodulator changes the sensory weight depends on the interneuron targeted and the effect this interneuron has on the baseline and gain of both PE neurons, which in turn does depend on the network it is embedded in.

### 2023-12-08 scripts/run_fig_2_support_2.py

@@ -94,17 +94,11 @@ if run_cell:
     avg_start = len(r_mem)//2
     print(np.mean(r_mem[-avg_start:]))
     print(np.mean(r_var[-avg_start:]))
-
-        
-      
-        
-     
-        
         
 
 # %% Find nPE and pPE neurons and plot gain factors
 
-run_cell = False
+run_cell = True
 plot_only = True
 
 if run_cell:
@@ -210,7 +204,7 @@ if run_cell:
 
 # %% Sparsity 
 
-run_cell = True
+run_cell = False
 plot_only = False
 
 if run_cell:

### 2023-12-13 records/manuscript_2023.tex
small changes for biorxiv submission

@@
 Left: Illustration of the input synapses onto the M neuron.
-Middle: Activity of the M neuron over time for one example distribution (black start in right panel).
+Middle: Activity of the M neuron over time for one example distribution (black star in right panel).
 Right: Normalised absolute difference between the averaged mean and the activity of the M neuron in the steady state for different parametrizations of the stimulus distribution.
@@
 Left: Illustration of the input synapses onto the V neuron.
-Middle: Activity of the V neuron over time for one example distribution (black start in right panel).
+Middle: Activity of the V neuron over time for one example distribution (black star in right panel).
 Right: Normalised absolute difference between the averaged variance and the activity of the V neuron in the steady state for different parametrizations of the stimulus distribution.
@@
 Each mean-field network receives a short, constant input from a different part of the receptive field.
-If the connection strengths from the PE neurons to the M and V neurons are adjusted accordingly (see Methods), the network correctly estimates the stimulus average and spatial uncertainty.
-(Fig.
+If the connection strengths from the PE neurons to the M and V neurons are adjusted accordingly (see Methods), the network correctly estimates the stimulus average and spatial uncertainty (Fig.
 \ref{fig:Fig_2_S3}B-C).
@@
 \ref{fig:Fig_3}A).
-Hence, the inputs change on two different time scales, where the stimulus variability has a faster time scale than the trial variability.
+Hence, the inputs change on two different time scales.
+%, where the stimulus variability has a faster time scale than the trial-to-trial variability.
 
@@
 \ref{fig:Fig_3}C, right), the network indeed shows a clear preference for the sensory input.
-In the second limit case, the same, high-variance stimulus is presented in each trial (Fig.
+In the second limit case, we show a high-variance stimulus, the mean of which does not change from trial to trial (high stimulus uncertainty, low trial-to-trial uncertainty, see Fig.
 \ref{fig:Fig_3}D, left).
@@
 How fast the new steady state is reached depends only on the time constants in the network and not the trial-to-trial variability.
-We next resume the limit case in which the same, high-variance stimulus is shown in every trial.
+We next resume the limit case in which the stimulus uncertainty is high while the trial-to-trial uncertainty is zero.
 In this case, the contraction bias is also largely independent of the stimulus variance (Fig.
@@
 Our mathematical analysis reveals that the bias is well described by the difference between the prediction, that is, mean stimulus over the history of all stimuli shown, and the current stimulus, weighted by a function of the trial duration.
+
 The analysis of both limit cases suggests that the bias also depends on the trial duration.
@@
 \begin{align}
-P_\mathrm{n} = P_\mathrm{0} \cdot \mathrm{e}^{-\Delta t/\tau_M}  + \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} s_i \cdot \mathrm{e}^{-(n-i)\cdot \Delta t/ \tau_M}
+P_\mathrm{n} = P_\mathrm{0} \cdot \mathrm{e}^{-\Delta t/\tau_M}  + \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} s_i \cdot \mathrm{e}^{-(n-i)\cdot \Delta t/ \tau_M}.
 \end{align}
@@
 \begin{align}
-\langle P_\mathrm{n} \rangle_s = \mathrm{e}^{-\Delta t/\tau_M}  \int\limits_{\mu - \frac{\sigma_\mathrm{P}}{12}}^{\mu + \frac{\sigma_\mathrm{P}}{12}} P_\mathrm{0} \ f(P_\mathrm{0})\ dP_0+ \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot \Delta t/ \tau_M} \int\limits_{s - \frac{\sigma_\mathrm{S}}{12}}^{s + \frac{\sigma_\mathrm{S}}{12}} x\ f(x)\ dx
+\langle P_\mathrm{n} \rangle_s = \mathrm{e}^{-\Delta t/\tau_M}  \int\limits_{\mu - \frac{\sigma_\mathrm{P}}{12}}^{\mu + \frac{\sigma_\mathrm{P}}{12}} P_\mathrm{0} \ f(P_\mathrm{0})\ dP_0+ \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot \Delta t/ \tau_M} \int\limits_{s - \frac{\sigma_\mathrm{S}}{12}}^{s + \frac{\sigma_\mathrm{S}}{12}} x\ f(x)\ dx.
 \end{align}
@@
 \begin{align}
-\langle P_\mathrm{n} \rangle_s = \mu \cdot \mathrm{e}^{-T/\tau_M} + \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot \Delta t/ \tau_M} \cdot s
+\langle P_\mathrm{n} \rangle_s = \mu \cdot \mathrm{e}^{-T/\tau_M} + \left( 1 -   \mathrm{e}^{-\Delta t/\tau_M} \right) \sum_{i=1}^{n} \mathrm{e}^{-(n-i)\cdot \Delta t/ \tau_M} \cdot s.
 \end{align}
@@
 \begin{align*}
-\langle P_\mathrm{n} \rangle_s =  \mu \cdot \mathrm{e}^{-T/\tau_M} + \left( 1 -   \mathrm{e}^{-T/\tau_M} \right) \cdot s
+\langle P_\mathrm{n} \rangle_s =  \mu \cdot \mathrm{e}^{-T/\tau_M} + \left( 1 -   \mathrm{e}^{-T/\tau_M} \right) \cdot s.
 \end{align*}
@@
 \begin{align*}
- r_\mathrm{out} = \left[ \alpha_\mathrm{S}\  \mathrm{e}^{-T/\tau_M} + \left( 1 -   \mathrm{e}^{-T/\tau_M} \right)\right] \cdot s + \left( 1 -\alpha_\mathrm{S} \right)\ \mathrm{e}^{-T/\tau_M}\ \mu
+ r_\mathrm{out} = \left[ \alpha_\mathrm{S}\  \mathrm{e}^{-T/\tau_M} + \left( 1 -   \mathrm{e}^{-T/\tau_M} \right)\right] \cdot s + \left( 1 -\alpha_\mathrm{S} \right)\ \mathrm{e}^{-T/\tau_M}\ \mu.
 \end{align*}
@@
     \includegraphics{../results/figures/final/Fig_2_S2.pdf}% [width=1\linewidth]
-\caption{\footnotesize{\bf Estimating mean and variance of sensory stimuli in a rate-based population network.\newline}  
-{\bf (A)} Illustration of the rate-based population network and the stimuli over time.
+\caption{\footnotesize{\bf Estimating mean and variance of sensory stimuli in a rate-based multi-cell population network.\newline}  
+{\bf (A)} Illustration of the rate-based multi-cell population network and the stimuli over time.
 The weights from the PE neurons onto the M or V neuron are scaled by a factor $\gamma$ drawn from a normal distribution $N(\mu_\mathrm{\gamma}, \sigma_\mathrm{\gamma})$.
@@
 \caption{\footnotesize{\bf Dynamic variance estimation allows flexible adaptation to changes in the stimulus statistics and environment. \newline}  
-{\bf (A)} Sensory weight for different input statistics. Numbers denote specific example statistics. Arrows denote the transitions between those statistics.
+{\bf (A)} Illustration of the sensory weight for different input statistics. Numbers denote specific examples. Arrows denote the transitions between those statistics.
 {\bf (B)} The sensory weight over time is shown for all transitions in (A).
-For the sake of clarity, we only show the trials 40 -60.
+For the sake of clarity, we only show the trials 40-60.
 The switch to new input statistics occurs at trial 50.
@@
 \caption{\footnotesize{\bf Perturbing the weighting of sensory inputs and predictions by altering network properties. \newline}  
-{\bf (A)} The weights from the PE neurons to the M neuron in the lower-order subnetwork are scaled by a factor the 0.3 or 7, leading to a distorted sensory weight.
+{\bf (A)} The weights from the PE neurons to the M neuron in the lower-order subnetwork are scaled by a factor 0.3 or 7, leading to a distorted sensory weight.
 If the update of the M neuron in the lower subnetwork is too slow ($\blacktriangleleft$), the prediction is overrated.
@@
 Considered are two limit cases (upper row: more sensory-driven before modulation, lower row: more prediction-driven before modulation).
-The results are shown for three mean-field networks (see \ref{fig:Fig_4}).
+The results are shown for three mean-field networks (see Fig.
+\ref{fig:Fig_4}).
 }
@@
 In a toy model, described in sections \ref{sec:gain_impact} and \ref{sec:impact_baseline}, the contribution of gain and baseline to the changes in the mean and variance estimation are summarized.
-The results are based on the Eqs.
+The results shown are based on the Eqs.
 (\ref{eq:prediction_gain}), (\ref{eq:variance_gain}), (\ref{eq:condition_baseline_mean_1}) and  (\ref{eq:condition_baseline_variance_2}).
@@
     \includegraphics{../results/figures/final/Fig_4_S4}% [width=1\linewidth]
-\caption{\footnotesize{\bf The combined changes in baseline and gain of all PE neurons determine the shift in the snesory weight.\newline}  
+\caption{\footnotesize{\bf The combined changes in baseline and gain of all PE neurons determine the shift in the sensory weight.\newline}  
 Whether and how a neuromodulator changes the sensory weight depends on the interneuron targeted and the effect this interneuron has on the baseline and gain of both PE neurons, which in turn does depend on the network it is embedded in.
 For small inputs, changes in the baseline dominate, while for large inputs, the changes in the gains dominate the shift in the sensory weight.
+Gray lines denote different mean inputs, illustrating that the same interneuron can decrease or increase the variance depending on the input regime.
 }

### 2024-01-15 records/manuscript_2023.tex
delete old files

@@
 %
-All values for neuron and network parameters, details on the model equations for the mean-field and the population network, as well as supporting analyses can be found in the supplementary material.
+All values for neuron and network parameters, details on the model equations for the mean-field and the multi-cell population network, as well as supporting analyses can be found in the supplementary material.
 
@@
 %
-The connectivity between neurons of the PE circuit, both for the mean-field and the population network, were taken from \citep{hertag2022prediction}.
+The connectivity between neurons of the PE circuit, both for the mean-field and the multi-cell population network, were taken from \citep{hertag2022prediction}.
 We considered three mean-field networks (see Table \ref{tab:wXM}) that differed in terms of the inputs (feedforward vs.
@@
 For the mean-field networks ($ N_\mathrm{nPE} = N_\mathrm{pPE} = 1 $), the gain factors $g_\mathrm{nPE}$ and $g_\mathrm{pPE}$ are given in Table~\ref{tab:gain_factors_MFN}.
-For the population network, the gain factors for all PE neurons are shown in Fig.
+For the multi-cell population network, the gain factors for all PE neurons are shown in Fig.
 \ref{fig:Fig_gains}.
@@
 The memory neuron \textit{M} connects to the post-synaptic neurons \textit{X} in the PE circuit with the connection strength $w_\mathrm{X\leftarrow M} = 1$, if a connection exists, $w_\mathrm{X\leftarrow M} = 0$ otherwise.
-In all mean-field networks and the population network, the dendrites of nPE and pPE neurons and one of the two (populations of) PV neurons receive connections from the memory neuron.
+In all mean-field networks and the multi-cell population network, the dendrites of nPE and pPE neurons and one of the two (populations of) PV neurons receive connections from the memory neuron.
 Furthermore, we assume that the \textit{M} neuron does not excite the soma of PCs.
@@
 In the case of the mean-field network, these inputs were set such that the baseline firing rates are $r_\mathrm{pE}=r_\mathrm{pD}=r_\mathrm{nE}=r_\mathrm{nD}=0\, s^{-1}$ and $r_\mathrm{P} = r_\mathrm{S}=r_\mathrm{V}=4\, s^{-1}$.
-In the case of the population network, we set the external inputs of all neuron types to $5\, s^{-1}$, while the background inputs to the dendrites were computed such that the dendrites are inactive during baseline.
+In the case of the multi-cell population network, we set the external inputs of all neuron types to $5\, s^{-1}$, while the background inputs to the dendrites were computed such that the dendrites are inactive during baseline.
 
@@
 In general, for the three mean-field networks tested, we chose $w_i = 1 - w_\mathrm{X\leftarrow M}$  (see Table \ref{tab:wXM}).
-In the population network, 70\% of the SOM neurons and 30\% of the VIP neurons receive the feedforward input.
+In the multi-cell population network, 70\% of the SOM neurons and 30\% of the VIP neurons receive the feedforward input.
 

### 2024-01-15 scripts/plot_fig_5_old.py

@@ -1,295 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Jan  9 09:04:01 2023
-
-@author: loreen.hertaeg
-"""
-
-# %% import
-
-import pickle
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-import os.path
-import seaborn as sns
-
-from src.plot_data import plot_example_contraction_bias, plot_slope_trail_duration, plot_slope_variability, plot_contraction_bias_illustration
-
-
-# %% Universal parameters
-
-fs = 6
-inch = 2.54
-dtype = np.float32
-
-# %% Define files and paths
-
-figure_name = 'Fig_5.png'
-figPath = '../results/figures/final/'
-
-if not os.path.exists(figPath):
-    os.mkdir(figPath)
-    
-    
-# %% Define figure structure
-
-figsize=(18/inch,10/inch)
-fig = plt.figure(figsize=figsize)
-
-G = gridspec.GridSpec(2, 4, figure=fig, hspace = 1, wspace=2, width_ratios=[1,1,1,1])
-
-G1 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=G[0,:2],  wspace=0.8)
-G2 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=G[0,2:],  wspace=0.8)
-G_lower = gridspec.GridSpecFromSubplotSpec(1, 4, subplot_spec=G[1,:], width_ratios=[1,1,1,1.2],  wspace=1)
-G3 = gridspec.GridSpecFromSubplotSpec(1, 3, subplot_spec=G_lower[0,:3],  wspace=0.8)
-G31 = gridspec.GridSpecFromSubplotSpec(1, 2, subplot_spec=G3[0,:2],  wspace=0.5)
-G4 = gridspec.GridSpecFromSubplotSpec(1, 1, subplot_spec=G_lower[0,3])
-
-ax_A = fig.add_subplot(G1[0,:])
-ax_A.axis('off')
-ax_A.set_title('Illustration of contraction bias', fontsize=fs, pad=10)
-ax_A.text(-0.2, 1.15,'A', transform=ax_A.transAxes, fontsize=fs+1)
-
-ax_A1 = fig.add_subplot(G1[0,0])
[…]

### 2024-01-15 scripts/run_fig_2_support_2.py

@@ -20,7 +20,7 @@ from src.plot_data import plot_gain_factors, plot_deviation_in_population_net, p
 dtype = np.float32
 
 
-# %% find mistake
+# %% check 
 
 run_cell = False
 
@@ -204,7 +204,7 @@ if run_cell:
 
 # %% Sparsity 
 
-run_cell = False
+run_cell = True
 plot_only = False
 
 if run_cell:

### 2024-01-15 scripts/run_fig_4_parallel_1.py

@@ -1,256 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Jan  9 09:04:01 2023
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-import pickle
-import matplotlib.pyplot as plt
-import seaborn as sns
-import pandas as pd
-import os.path
-
-from src.functions_simulate import simulate_neuromod, simulate_moment_estimation_upon_changes_PE, simulate_neuromod_effect_on_neuron_properties
-from src.functions_simulate import simulate_neuromod_combos, simulate_neuromod_pert
-from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs, plot_neuromod_per_net, plot_points_of_interest_neuromod, plot_bar_neuromod
-from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_changes_upon_input2PE_neurons, plot_bar_neuromod_stacked
-from src.plot_data import plot_neurmod_results, plot_neuromod_impact_inter, plot_changes_upon_input2PE_neurons_new, plot_influence_interneurons_baseline_or_gain
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-
-# %% Neuromodulator targeting one IN (with various strengths)
-
-run_cell = True
-plot_only = False
-column = 0 
-
-xp, xs, xv = 1, 0, 0
-
-if run_cell:
-    
-    ### filename for data
-    file_for_data = '../results/data/neuromod/data_neuromod_' + str(xp) + '_' + str(xs) + '_' + str(xv) + '.pickle'
-    
-    ### define perturbations tested
-    pert_strength = np.linspace(0,0.5,6)
-    
-    ### main loop
-    if not plot_only:
-        
-        alpha = np.zeros((len(pert_strength), 3, 2))
-    
-        for k, mfn_flag in enumerate(['10', '01', '11']):
-            
-            for std_mean in [0, 1]: # uncertainty of environement [0, 1]
-                
-                n_std = 1 - std_mean # uncertainty of stimulus [1, 0]
-                print('MFN: ', mfn_flag, ', Input case', std_mean)
-                
-                [_, _, _, _, alpha_after_pert] = simulate_neuromod_pert(mfn_flag, pert_strength, std_mean, n_std, column, 
-                                                                        xp, xs, xv, file_for_data = 'Test')
[…]

### 2024-01-15 scripts/run_fig_4_parallel_3.py

@@ -1,256 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Jan  9 09:04:01 2023
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-import pickle
-import matplotlib.pyplot as plt
-import seaborn as sns
-import pandas as pd
-import os.path
-
-from src.functions_simulate import simulate_neuromod, simulate_moment_estimation_upon_changes_PE, simulate_neuromod_effect_on_neuron_properties
-from src.functions_simulate import simulate_neuromod_combos, simulate_neuromod_pert
-from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs, plot_neuromod_per_net, plot_points_of_interest_neuromod, plot_bar_neuromod
-from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_changes_upon_input2PE_neurons, plot_bar_neuromod_stacked
-from src.plot_data import plot_neurmod_results, plot_neuromod_impact_inter, plot_changes_upon_input2PE_neurons_new, plot_influence_interneurons_baseline_or_gain
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-
-# %% Neuromodulator targeting one IN (with various strengths)
-
-run_cell = True
-plot_only = False
-column = 0 
-
-xp, xs, xv = 0, 0, 1
-
-if run_cell:
-    
-    ### filename for data
-    file_for_data = '../results/data/neuromod/data_neuromod_' + str(xp) + '_' + str(xs) + '_' + str(xv) + '.pickle'
-    
-    ### define perturbations tested
-    pert_strength = np.linspace(0,0.5,6)
-    
-    ### main loop
-    if not plot_only:
-        
-        alpha = np.zeros((len(pert_strength), 3, 2))
-    
-        for k, mfn_flag in enumerate(['10', '01', '11']):
-            
-            for std_mean in [0, 1]: # uncertainty of environement [0, 1]
-                
-                n_std = 1 - std_mean # uncertainty of stimulus [1, 0]
-                print('MFN: ', mfn_flag, ', Input case', std_mean)
-                
-                [_, _, _, _, alpha_after_pert] = simulate_neuromod_pert(mfn_flag, pert_strength, std_mean, n_std, column, 
-                                                                        xp, xs, xv, file_for_data = 'Test')
[…]

### 2024-01-15 scripts/run_fig_4_parallel_5.py

@@ -1,202 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Jan  9 09:04:01 2023
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-import pickle
-import matplotlib.pyplot as plt
-import seaborn as sns
-import pandas as pd
-import os.path
-
-from src.functions_simulate import simulate_neuromod, simulate_moment_estimation_upon_changes_PE, simulate_neuromod_effect_on_neuron_properties
-from src.functions_simulate import simulate_neuromod_combos
-from src.plot_data import plot_heatmap_neuromod, plot_combination_activation_INs, plot_neuromod_per_net, plot_points_of_interest_neuromod, plot_bar_neuromod
-from src.plot_data import plot_illustration_changes_upon_baseline_PE, plot_illustration_changes_upon_gain_PE, plot_changes_upon_input2PE_neurons, plot_bar_neuromod_stacked
-from src.plot_data import plot_neurmod_results, plot_neuromod_impact_inter, plot_changes_upon_input2PE_neurons_new, plot_influence_interneurons_baseline_or_gain
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-
-# %% Activate fractions of IN neurons in lower/higher PE circuit or both for a specified input statistics
-
-run_cell = True
-
-if run_cell:
-    
-    # define combo activation resolution
-    nums = 11
-    
-    # run through all combinations
-    for mfn_flag in ['11']: # choose mean-field network to simulate
-        
-        print('MFN code: ', mfn_flag)
-        print(' ')
-    
-        for column in range(2,-1,-1): # 0: both, 1: lower level PE circuit, 2: higher level PE circuit
-
-            for std_mean in [0, 1]: # uncertainty of environement [0, 1]
-                
-                n_std = 1 - std_mean # uncertainty of stimulus [1, 0]
-                    
-                for IN_combo_flg in range(3): # 0: PV - SOM, 1: SOM - VIP, 2: VIP - PV
-        
-                    # filename for data & define activation combos
-                    identifier = '_column_' + str(column) + '_acrossvar_' + str(std_mean) + '_withinvar_' + str(n_std)
-                    print(identifier)
-        
-                    if IN_combo_flg==0:
[…]

### 2024-01-15 scripts/run_figures_conference.py

@@ -1,245 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Tue Nov 15 10:27:24 2022
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-import seaborn as sns
-import matplotlib.pyplot as plt
-import pandas as pd
-import pickle
-
-from src.mean_field_model import default_para, stimuli_moments_from_uniform, run_mean_field_model, alpha_parameter_exploration
-from src.mean_field_model import run_mean_field_model_one_column
-
-from src.toy_model import random_uniform_from_moments, random_lognormal_from_moments, random_gamma_from_moments, random_binary_from_moments
-from src.plot_toy_model import plot_limit_case, plot_alpha_para_exploration_ratios, plot_fraction_sensory_comparsion, plot_alpha_para_exploration
-from src.plot_toy_model import plot_manipulation_results
-
-from src.plot_results_mfn import plot_prediction, plot_variance, plot_mse, plot_manipulations, plot_mse_heatmap, plot_limit_case_example
-from src.plot_results_mfn import plot_diff_heatmap, plot_transitions_examples, plot_deviation_vs_effect_size, plot_deviation_vs_PE_II
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-inch = 2.54
-
-# %% Example: PE neurons can estimate/establish the mean and variance of a normal random variable 
-# mean and varaince are drawn from a unimodal distribution
-
-flag = 0
-
-if flag==1:
-    
-    ### define/load parameters
-    input_flg = '10' # 10, 01, 11
-    
-    VS, VV = int(input_flg[0]), int(input_flg[1])
-    filename = '../results/data/moments/Data_Optimal_Parameters_MFN_' + input_flg + '.pickle'
-    file_data4plot = '../results/data/moments/Data_example_MFN_' + input_flg + '.pickle'
-    
-    ### load data for plotting
-    with open(file_data4plot,'rb') as f:
-        [n_trials, trial_duration, stimuli, prediction, variance] = pickle.load(f)  
-           
-    ### plot prediction and variance (in comparison to the data) 
-    fig_size = (6 / inch, 3 / inch)
-    plot_prediction(n_trials, stimuli, trial_duration, prediction, figsize=fig_size, lw=1, fs=7, 
[…]

### 2024-01-15 scripts/run_moments_perturbation.py

@@ -1,749 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Jul 25 08:58:09 2022
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-import pickle
-
-from src.mean_field_model import default_para, stimuli_moments_from_uniform, run_mean_field_model, alpha_parameter_exploration
-from src.mean_field_model import run_mean_field_model_one_column, run_pe_circuit_mfn
-
-from src.toy_model import random_uniform_from_moments, random_lognormal_from_moments, random_gamma_from_moments
-from src.plot_toy_model import plot_limit_case, plot_alpha_para_exploration_ratios, plot_fraction_sensory_comparsion, plot_alpha_para_exploration
-from src.plot_toy_model import plot_manipulation_results
-
-from src.plot_results_mfn import plot_prediction, plot_variance, plot_mse, plot_manipulations, plot_deviations_upon_perturbations
-from src.plot_results_mfn import plot_heatmap_perturbation_all, plot_deviation_vs_effect_size, plot_deviation_vs_PE
-from src.plot_results_mfn import plot_deviation_vs_PE_II
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-# %% Perturbing the E/I balance in PE circuits - example
-
-flag = 0
-flg_plot_only = 0
-
-if flag==1:
-    
-    ### load and define parameters
-    input_flg = '10' # 10, 01, 11
-    
-    VS, VV = int(input_flg[0]), int(input_flg[1])
-    filename = '../results/data/moments/Data_Optimal_Parameters_MFN_' + input_flg + '.pickle'
-    file_data4plot = '../results/data/perturbations/data_perturbation_example_' + input_flg + '.pickle'
-    
-    if flg_plot_only==0:
-    
-        [w_PE_to_P, w_P_to_PE, w_PE_to_PE, v_PE_to_P, v_P_to_PE, v_PE_to_PE, 
-         tc_var_per_stim, tc_var_pred, tau_pe, fixed_input] = default_para(filename, VS=VS, VV=VV)
-    
-        if input_flg=='10':
-            nPE_scale = 1.015
-            pPE_scale = 1.023
-        elif input_flg=='01':
-            nPE_scale = 1.7 # 1.72
-            pPE_scale = 1.7 # 1.68
[…]

### 2024-01-15 scripts/run_neuron_properties.py

@@ -1,700 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Thu Nov 24 08:56:15 2022
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-import pickle
-
-from src.mean_field_model import default_para, stimuli_moments_from_uniform, run_mean_field_model, alpha_parameter_exploration
-from src.plot_toy_model import plot_limit_case, plot_alpha_para_exploration_ratios, plot_fraction_sensory_comparsion, plot_alpha_para_exploration
-from src.plot_toy_model import plot_manipulation_results
-# from src.mean_field_model import stimuli_moments_from_uniform, run_toy_model, default_para, alpha_parameter_exploration
-# from src.mean_field_model import random_uniform_from_moments, random_lognormal_from_moments, random_gamma_from_moments
-# from src.mean_field_model import stimuli_from_mean_and_std_arrays
-from src.plot_results_mfn import plot_limit_case_example, plot_transitions_examples, heatmap_summary_transitions, plot_transition_course
-from mpl_toolkits.axes_grid1.inset_locator import inset_axes
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-# %% erase after testing
-
-import matplotlib.pyplot as plt
-import matplotlib as mpl
-import seaborn as sns
-from matplotlib.colors import LinearSegmentedColormap, ListedColormap
-    
-color_sensory = '#D76A03'
-color_prediction = '#19535F'
-
-cmap_sensory_prediction = LinearSegmentedColormap.from_list(name='cmap_sensory_prediction', 
-                                                        colors=['#19535F','#fefee3','#D76A03'])
-
-# %% No squared activation function
-
-# Results: 
-    # var(S)<var(P): weighting with abs will lead to decaresed sensory weight
-    # var(S)>var(P): weighting with abs will lead to increased sensory weight
-    # In principle, the weighting in general is fine, that is, when var(S)<var(P), system is sensory driven, while when var(S)>var(P), the system is more prediction driven
-    # it just seems that there is a tendency to taking the "other" more into account
[…]

### 2024-01-15 scripts/run_on_desktop.py

@@ -1,648 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Wed Aug 31 13:10:26 2022
-
-@author: loreen.hertaeg
-"""
-
-
-# %% Import
-
-import numpy as np
-import pickle
-
-import sys
-sys.path.append('../src')
-
-from mean_field_model import default_para, stimuli_moments_from_uniform, run_mean_field_model, alpha_parameter_exploration
-#from plot_toy_model import plot_limit_case, plot_alpha_para_exploration_ratios, plot_fraction_sensory_comparsion, plot_alpha_para_exploration
-#from plot_toy_model import plot_manipulation_results
-#from plot_results_mfn import plot_limit_case_example, plot_transitions_examples, heatmap_summary_transitions, plot_transition_course
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-
-# %% Test: Exc/Inh perturbation, whole range
-
-flag = 1
-
-if flag==1:
-    
-    for column in [1,2]:
-
-        ### define MFN networks that will be simulated
-        input_flgs = ['10', '01', '11']
-        
-        for input_flg in input_flgs:
-        
-            ### load and define parameters
-            #input_flg = '10'
-            filename = '../results/data/moments/Data_Optimal_Parameters_MFN_' + input_flg + '.pickle'
-            file_data4plot = '../results/data/weighting_perturbation/test_weighting_perturbations_' + input_flg + '_column_' + str(column) + '.pickle'
-            
-            [w_PE_to_P, w_P_to_PE, w_PE_to_PE, v_PE_to_P, v_P_to_PE, v_PE_to_PE, 
-             tc_var_per_stim, tc_var_pred, tau_pe, fixed_input] = default_para(filename)
-            
-            if input_flg=='10':
-                nPE_scale = 1.015
-                pPE_scale = 1.023
-            elif input_flg=='01':
-                nPE_scale = 1.7 # 1.72
-                pPE_scale = 1.7 # 1.68
-            elif input_flg=='11':
-                nPE_scale = 2.49
-                pPE_scale = 2.53
-                
-            w_PE_to_P[0,0] *= nPE_scale * 15 # !!!!!!!!!!! to make it faster
-            w_PE_to_P[0,1] *= pPE_scale * 15 # !!!!!!!!!!!
[…]

### 2024-01-15 scripts/run_on_desktop_1.py

@@ -1,261 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Wed Aug 31 13:10:26 2022
-
-@author: loreen.hertaeg
-"""
-
-
-# %% Import
-
-import numpy as np
-import pickle
-
-import sys
-sys.path.append('../src')
-
-from mean_field_model import default_para, stimuli_moments_from_uniform, run_mean_field_model, alpha_parameter_exploration
-#from plot_toy_model import plot_limit_case, plot_alpha_para_exploration_ratios, plot_fraction_sensory_comparsion, plot_alpha_para_exploration
-#from plot_toy_model import plot_manipulation_results
-#from plot_results_mfn import plot_limit_case_example, plot_transitions_examples, heatmap_summary_transitions, plot_transition_course
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-
-# %% Test: Exc/Inh perturbation, VIP in MFN1, +/- 1 perturbation, whole range
-
-flag = 1
-
-if flag==1:
-    
-    ### define MFN networks that will be simulated
-    input_flgs = ['10', '01', '11']
-    
-    for input_flg in input_flgs:
-    
-        ### load and define parameters
-        # input_flg = '10'
-        filename = '../results/data/moments/Data_Optimal_Parameters_MFN_' + input_flg + '.pickle'
-        file_data4plot = file_data4plot = '../results/data/weighting_perturbation/test_weighting_perturbations_' + input_flg + '.pickle'
-        
-        [w_PE_to_P, w_P_to_PE, w_PE_to_PE, v_PE_to_P, v_P_to_PE, v_PE_to_PE, 
-         tc_var_per_stim, tc_var_pred, tau_pe, fixed_input] = default_para(filename)
-        
-        if input_flg=='10':
-            nPE_scale = 1.015
-            pPE_scale = 1.023
-        elif input_flg=='01':
-            nPE_scale = 1.7 # 1.72
-            pPE_scale = 1.7 # 1.68
-        elif input_flg=='11':
-            nPE_scale = 2.49
-            pPE_scale = 2.53
-            
-        w_PE_to_P[0,0] *= nPE_scale * 15 # !!!!!!!!!!! to make it faster
-        w_PE_to_P[0,1] *= pPE_scale * 15 # !!!!!!!!!!!
[…]

### 2024-01-15 scripts/run_weighting_mean_field.py

@@ -1,846 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Wed May  4 08:01:46 2022
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-import pickle
-
-from src.mean_field_model import default_para, stimuli_moments_from_uniform, run_mean_field_model, alpha_parameter_exploration
-from src.plot_toy_model import plot_limit_case, plot_alpha_para_exploration_ratios, plot_fraction_sensory_comparsion, plot_alpha_para_exploration
-from src.plot_toy_model import plot_manipulation_results
-# from src.mean_field_model import stimuli_moments_from_uniform, run_toy_model, default_para, alpha_parameter_exploration
-# from src.mean_field_model import random_uniform_from_moments, random_lognormal_from_moments, random_gamma_from_moments
-# from src.mean_field_model import stimuli_from_mean_and_std_arrays
-from src.plot_results_mfn import plot_limit_case_example, plot_transitions_examples, heatmap_summary_transitions, plot_transition_course
-from src.plot_results_mfn import  plot_alpha_over_trial
-
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-# %% erase after testing
-
-import matplotlib.pyplot as plt
-import matplotlib as mlp
-import seaborn as sns
-
-# %% Note!!!
-
-# Here the stimuli are drawn from a normal distribution with mean and std 
-# for each stimulus presentation, mean and std are drawn from a uniform distribution
-
-# %% limit cases
-
-flag = 0
-flg_limit_case = 0 # 0 = mean the same, std large; 1 = mean varies, std = 0
-
-flg_plot_only = 1
-
-if flag==1:
-    
-    ### file to save data
-    file_data4plot = '../results/data/weighting/data_example_limit_case_' + str(flg_limit_case) + '.pickle'
-    
-    if flg_plot_only==0:
-        
-        ### load and define parameters
-        input_flg = '10'
-        filename = '../results/data/moments/Data_Optimal_Parameters_MFN_' + input_flg + '.pickle'
-        
-        [w_PE_to_P, w_P_to_PE, w_PE_to_PE, v_PE_to_P, v_P_to_PE, v_PE_to_PE, 
-         tc_var_per_stim, tc_var_pred, tau_pe, fixed_input] = default_para(filename)
-        
[…]

### 2024-01-15 scripts/run_weighting_perturbation_I.py

@@ -1,1006 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Aug 29 08:54:46 2022
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-import pickle
-
-from src.mean_field_model import default_para, stimuli_moments_from_uniform, run_mean_field_model, alpha_parameter_exploration
-from src.plot_toy_model import plot_limit_case, plot_alpha_para_exploration_ratios, plot_fraction_sensory_comparsion, plot_alpha_para_exploration
-from src.plot_toy_model import plot_manipulation_results
-# from src.mean_field_model import stimuli_moments_from_uniform, run_toy_model, default_para, alpha_parameter_exploration
-# from src.mean_field_model import random_uniform_from_moments, random_lognormal_from_moments, random_gamma_from_moments
-# from src.mean_field_model import stimuli_from_mean_and_std_arrays
-from src.plot_results_mfn import plot_limit_case_example, plot_transitions_examples, heatmap_summary_transitions, plot_transition_course
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-# %% erase after testing
-
-import matplotlib.pyplot as plt
-import matplotlib as mpl
-import seaborn as sns
-from matplotlib.colors import LinearSegmentedColormap, ListedColormap
-    
-color_sensory = '#D76A03'
-color_prediction = '#19535F'
-
-cmap_sensory_prediction = LinearSegmentedColormap.from_list(name='cmap_sensory_prediction', 
-                                                        colors=['#19535F','#fefee3','#D76A03'])
-
-# %% Test some settings (not for publication later)
-
-flag = 0
-flg_plot_only = 0
-
-if flag==1:
-    
-    ### file to save data
-    #file_data4plot = '../results/data/weighting/data_example_limit_case_' + str(flg_limit_case) + '.pickle'
-    
-    if flg_plot_only==0:
-        
-        ### load and define parameters
-        input_flg = '10'
-        filename = '../results/data/moments/Data_Optimal_Parameters_MFN_10.pickle'
-        
-        [w_PE_to_P, w_P_to_PE, w_PE_to_PE, v_PE_to_P, v_P_to_PE, v_PE_to_PE, 
[…]

### 2024-01-15 scripts/run_weighting_perturbation_II.py

@@ -1,975 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Sun Sep 4 14:43:46 2022
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-import pickle
-
-from src.mean_field_model import default_para, stimuli_moments_from_uniform, run_mean_field_model, alpha_parameter_exploration
-from src.plot_toy_model import plot_limit_case, plot_alpha_para_exploration_ratios, plot_fraction_sensory_comparsion, plot_alpha_para_exploration
-from src.plot_toy_model import plot_manipulation_results
-# from src.mean_field_model import stimuli_moments_from_uniform, run_toy_model, default_para, alpha_parameter_exploration
-# from src.mean_field_model import random_uniform_from_moments, random_lognormal_from_moments, random_gamma_from_moments
-# from src.mean_field_model import stimuli_from_mean_and_std_arrays
-from src.plot_results_mfn import plot_limit_case_example, plot_transitions_examples, heatmap_summary_transitions, plot_transition_course
-
-import warnings
-warnings.filterwarnings("ignore")
-
-dtype = np.float32
-
-# %% erase after testing
-
-import matplotlib.pyplot as plt
-import matplotlib as mpl
-import seaborn as sns
-from matplotlib.colors import LinearSegmentedColormap, ListedColormap
-    
-color_sensory = '#D76A03'
-color_prediction = '#19535F'
-
-cmap_sensory_prediction = LinearSegmentedColormap.from_list(name='cmap_sensory_prediction', 
-                                                        colors=['#19535F','#fefee3','#D76A03'])
-
-# %% summary statistics with one example perturbation (VIP activated) affecting one 'column'
-
-flag = 0
-flg_plot_only = 1
-column = 2 # 1 or 2
-
-if flag==1:
-    
-    ### file to save data
-    file_data4plot = '../results/data/weighting_perturbation/data_weighting_heatmap_perturbation_example_column_' + str(column) + '.pickle'
-    
-    if flg_plot_only==0:
-        
-        ### load and define parameters
-        input_flg = '10'
-        filename = '../results/data/moments/Data_Optimal_Parameters_MFN_10.pickle'
-        
-        [w_PE_to_P, w_P_to_PE, w_PE_to_PE, v_PE_to_P, v_P_to_PE, v_PE_to_PE, 
[…]

### 2024-01-15 src/Default_values.py

@@ -1,106 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Mar 28 14:11:50 2022
-
-@author: loreen.hertaeg
-"""
-
-import numpy as np
-
-dtype = np.float32
-
-# %% Define default parameters
-
-def Default_NeuronPara():
-    
-    NumCells = np.array([140,20,20,20], dtype=np.int32)
-    tau_E_inv = dtype(1/60)
-    tau_I_inv = dtype(1/2)
-    
-    return NumCells, tau_E_inv, tau_I_inv
-
-
-def Default_Connectivity(wED: dtype = dtype(1), wEP: dtype = dtype(-2), 
-                         wDE: dtype = dtype(0.5), wDS: dtype = dtype(-0.5), 
-                         wPE: dtype = dtype(1.2), wPP: dtype = dtype(-1),  
-                         wPS: dtype = dtype(-0.3), wPV: dtype = dtype(-0.3), 
-                         wSV: dtype = dtype(-0.6), wVS: dtype = dtype(-0.7)): 
-                         
-                         
-    weights_mean = np.array([[0,wED,wEP,0,0],[wDE,0,0,wDS,0],[wPE,0,wPP,wPS,wPV],
-                             [1,0,0,0,wSV],[1,0,0,wVS,0]], dtype=dtype)                   
-    
-    conn_prob = np.array([[0,1,0.6,0,0],[0.1,0,0,0.55,0],[0.45,0,0.5,0.6,0.5],
-                          [0.35,0,0,0,0.5],[0.1,0,0,0.45,0]], dtype=dtype)
-    
-    weight_name = np.array([['wEE','wED','wEP','wES','wEV'],['wDE','wDD','wDP','wDS','wDV'],['wPE','wPD','wPP','wPS','wPV'],
-                            ['wSE','wSD','wSP','wSS','wSV'],['wVE','wVD','wVP','wVS','wVV']])
-
-    temp_name = np.array([['TEE','TED','TEP','TES','TEV'],['TDE','TDD','TDP','TDS','TDV'],['TPE','TPD','TPP','TPS','TPV'],
-                          ['TSE','TSD','TSP','TSS','TSV'],['TVE','TVD','TVP','TVS','TVV']])
-
-    return weights_mean, conn_prob, weight_name, temp_name
-
-
-def Default_LearningRates(eEP: dtype = dtype(1e-3), eDS: dtype = dtype(1e-4), 
[…]

### 2024-01-15 src/Functions_Network.py

@@ -1,1091 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon June 14 2021
-
-@author: loreen.hertaeg
-"""
-
-# %% import packages
-
-import numpy as np
-import tables
-import os
-from numba import njit
-from typing import NamedTuple
-
-dtype = np.float32
-
-# %% Classes
-
-class Neurons(NamedTuple):
-        
-        NCells: list = np.array([140,20,20,20], dtype=np.int32)
-        tau_inv_E: dtype = dtype(1.0/60.0)
-        tau_inv_I: dtype = dtype(1.0/2.0)
-
-
-class Network:
-    def __init__(self, Neurons, weights_mean, conn_prob, weight_name, temp_name, uni_low: dtype = dtype(0.5), 
-                 uni_high: dtype = dtype(1.5), flag_hetero: int = np.int32(1), flag_conn: int = np.int32(1)):
-            
-            ### Connection probabilty (either all-to-all or specified by conn_prob)
-            if flag_conn==0:
-                conn_prob = 1*(weights_mean!=0)
-            else:
-                conn_prob = conn_prob
-                
-            if not isinstance(uni_high, np.ndarray):
-                uni_high = uni_high * np.ones_like(weights_mean)
-                
-            if not isinstance(uni_low, np.ndarray):
-                uni_low = uni_low * np.ones_like(weights_mean)
-            
-            NUnits = np.insert(Neurons.NCells, 0, Neurons.NCells[0])
-            NCon = np.round(conn_prob * NUnits).astype(np.int32)
-            self.NCon = NCon
-            self.weight_name = weight_name
-            self.weights_mean = weights_mean
-            
-            ### Connection strengths between neuons
-            for i in range(25):
-                
-                m,n = np.unravel_index(i,(5,5))
-                Mtx = np.zeros((NUnits[m],NUnits[n]), dtype=dtype)
-                
-                if NCon[m,n]>0:
-                    
-                    # establish connections
-                    if m==n:
-                        for l in range(NUnits[m]):
[…]

### 2024-01-15 src/Functions_Network_all.py

@@ -1,740 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Feb  7 13:12:21 2022
-
-@author: loreen.hertaeg
-"""
-
-import os
-import tables
-import numpy as np
-from numba import njit
-from typing import NamedTuple
-
-dtype = np.float32
-
-# %% classes
-
-class Neurons(NamedTuple):
-        
-        num_cells_0: list = np.array([80,20], dtype=np.int32)       # E, I
-        num_cells_1: list = np.array([80, 80, 20], dtype=np.int32)  # E, D, I
-        num_cells_2: list = np.array([80,20], dtype=np.int32)
-        num_cells_3: list = np.array([80, 80, 20], dtype=np.int32)
-        num_cells_4: list = np.array([80,20], dtype=np.int32)
-        num_cells_5: list = np.array([80, 80, 20], dtype=np.int32)
-        num_cells_6: list = np.array([80,20], dtype=np.int32)
-        
-        tci_0: list = np.array([1/20, 1/10], dtype=dtype)
-        tci_1: list = np.array([1/20, 1/20, 1/10], dtype=dtype)
-        tci_2: list = np.array([1/20, 1/10], dtype=dtype)
-        tci_3: list = np.array([1/20, 1/20, 1/10], dtype=dtype)
-        tci_4: list = np.array([1/20, 1/10], dtype=dtype)
-        tci_5: list = np.array([1/20, 1/20, 1/10], dtype=dtype)
-        tci_6: list = np.array([1/20, 1/10], dtype=dtype)
-        
-        leak_0: list = np.array([1, 1], dtype=dtype)
-        leak_1: list = np.array([1, 1, 1], dtype=dtype)
-        leak_2: list = np.array([1, 1], dtype=dtype)
-        leak_3: list = np.array([1, 1, 1], dtype=dtype)
-        leak_4: list = np.array([1, 1], dtype=dtype)
-        leak_5: list = np.array([1, 1, 1], dtype=dtype)
-        leak_6: list = np.array([1, 1], dtype=dtype)
-        
-        rheo_0: list = np.array([0, 0], dtype=dtype)
-        rheo_1: list = np.array([0, 0, 0], dtype=dtype)
-        rheo_2: list = np.array([0, 0], dtype=dtype)
-        rheo_3: list = np.array([0, 0, 0], dtype=dtype)
[…]

### 2024-01-15 src/Functions_PE.py

@@ -1,951 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon June 14 2021
-
-@author: loreen.hertaeg
-"""
-
-# %% import packages
-
-import numpy as np
-import tables
-import os
-from numba import njit
-from typing import NamedTuple
-
-dtype = np.float32
-
-# %% Classes
-
-class Neurons(NamedTuple):
-        
-        NCells: list = np.array([140,20,20,20], dtype=np.int32)
-        tau_inv_E: dtype = dtype(1.0/60.0)
-        tau_inv_I: dtype = dtype(1.0/2.0)
-
-
-class Network:
-    def __init__(self, Neurons, weights_mean, conn_prob, weight_name, temp_name, uni_low: dtype = dtype(0.5), 
-                 uni_high: dtype = dtype(1.5), flag_hetero: int = np.int32(1), flag_conn: int = np.int32(1)):
-            
-            ### Connection probabilty (either all-to-all or specified by conn_prob)
-            if flag_conn==0:
-                conn_prob = 1*(weights_mean!=0)
-            else:
-                conn_prob = conn_prob
-                
-            if not isinstance(uni_high, np.ndarray):
-                uni_high = uni_high * np.ones_like(weights_mean)
-                
-            if not isinstance(uni_low, np.ndarray):
-                uni_low = uni_low * np.ones_like(weights_mean)
-            
-            NUnits = np.insert(Neurons.NCells, 0, Neurons.NCells[0])
-            NCon = np.round(conn_prob * NUnits).astype(np.int32)
-            self.NCon = NCon
-            self.weight_name = weight_name
-            self.weights_mean = weights_mean
-            
-            ### Connection strengths between neuons
-            for i in range(25):
-                
-                m,n = np.unravel_index(i,(5,5))
-                Mtx = np.zeros((NUnits[m],NUnits[n]), dtype=dtype)
-                
-                if NCon[m,n]>0:
-                    
-                    # establish connections
-                    if m==n:
[…]

### 2024-01-15 src/Functions_PredNet.py

@@ -1,432 +1,0 @@
- #!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Tue Jun 29 08:34:59 2021
-
-@author: loreen.hertaeg
-"""
-
-# %% import packages
-
-import numpy as np
-import os
-from numba import njit
-from typing import NamedTuple
-
-dtype = np.float32
-
-# %% Classes
-
-class Neurons(NamedTuple):
-        
-        NCells: list = np.array([140,20,20,20], dtype=np.int32)
-        tau_inv_E: dtype = dtype(1.0/60.0)
-        tau_inv_I: dtype = dtype(1.0/2.0)
-        
-        
-        
-class Network:
-    def __init__(self, Neurons, NetPar_PE, mLE, bool_nPE, bool_pPE, neurons_motor):
-        
-        # Pre-processing
-        NE, NP, NS, NV = np.int32(Neurons.NCells)
-        N_nPE = sum(bool_nPE)
-        N_pPE = sum(bool_pPE)
-        
-        P_Pred_flag = neurons_motor[NE:(NE+NP)]
-        S_Pred_flag = neurons_motor[(NE+NP):(NE+NP+NS)]
-        V_Pred_flag = neurons_motor[(NE+NP+NS):]
-        
-        # PE network connectivity (PE-PE)
-        self.wEP = NetPar_PE.wEP
-        self.wED = NetPar_PE.wED
-        self.wDE = NetPar_PE.wDE
-        self.wDS = NetPar_PE.wDS 
-        self.wPE = NetPar_PE.wPE
-        self.wPP = NetPar_PE.wPP
-        self.wPS = NetPar_PE.wPS
-        self.wPV = NetPar_PE.wPV
-        self.wSE = NetPar_PE.wSE
-        self.wSP = NetPar_PE.wSP
-        self.wSS = NetPar_PE.wSS
-        self.wSV = NetPar_PE.wSV
-        self.wVE = NetPar_PE.wVE
-        self.wVP = NetPar_PE.wVP
-        self.wVS = NetPar_PE.wVS
-        self.wVV = NetPar_PE.wVV
-        
-        # Connectivity between line attractor and PE (L-PE)
-        wLE = np.zeros((1,NE), dtype=dtype)
-        wLE[0,:] = mLE * bool_pPE/N_pPE - mLE * bool_nPE/N_nPE
[…]

### 2024-01-15 src/functions_analyse.py

@@ -1,9 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Jan  9 09:31:58 2023
-
-@author: loreen.hertaeg
-"""
-
-# %% Import

### 2024-01-15 src/functions_save.py

@@ -1,26 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Jan  9 09:31:58 2023
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import os
-import pickle
-
-
-# %% functions
-
-
-def load_network_para(folder):
-    
-    path = '../results/data/' + folder
-    filename = path + '/data_population_network_parameters.pickle'
-    
-    with open(filename,'rb') as f:
-        NeuPar, NetPar, InPar, StimPar, SimPar, RatePar, LearnPar = pickle.load(f)
-        
-    return NeuPar, NetPar, InPar, StimPar, SimPar, RatePar, LearnPar

### 2024-01-15 src/plot_results_mfn.py

@@ -1,872 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Tue Jun 21 11:20:53 2022
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-import pickle
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-import seaborn as sns
-import pandas as pd
-import matplotlib
-
-from matplotlib.colors import LinearSegmentedColormap, ListedColormap
-import matplotlib.ticker as ticker
-from mpl_toolkits.axes_grid1.inset_locator import inset_axes
-from matplotlib.legend import Legend
-
-from src.toy_model import random_uniform_from_moments, random_lognormal_from_moments, random_gamma_from_moments, random_binary_from_moments
-
-dtype = np.float32
-
-# %% set colors
-
-Col_Rate_E = '#9E3039'
-Col_Rate_nE = '#955F89'
-Col_Rate_nD = '#BF9BB8' 
-Col_Rate_pE = '#CB9173' 
-Col_Rate_pD = '#DEB8A6' 
-Col_Rate_PVv = '#508FCE' 
-Col_Rate_PVm = '#2B6299' 
-Col_Rate_SOM = '#79AFB9' 
-Col_Rate_VIP = '#39656D' 
-
-color_sensory = '#D76A03'
-color_prediction = '#19535F'
-
-cmap_sensory_prediction = LinearSegmentedColormap.from_list(name='cmap_sensory_prediction', 
-                                                            colors=['#19535F','#fefee3','#D76A03'])
-
-# %% functions
-
-
-def plot_alpha_over_trial(fraction_course, n_trials):
-    
-    fig, ax = plt.subplots(1, 1, tight_layout=True, figsize=(4,4))
-    colors = sns.color_palette('magma', n_colors=5)
-    
-    for id_stim in range(5):
-        
-        alpha_split = np.array_split(fraction_course[:,id_stim], n_trials)
-        alpha_avg_trial = np.mean(alpha_split,0)
-        alpha_std_trial = np.std(alpha_split,0)
-        sem = alpha_std_trial/np.sqrt(n_trials)
-        trial_fraction = np.linspace(0,1,len(alpha_avg_trial))
-        
-        ax.plot(trial_fraction, alpha_avg_trial, color=colors[id_stim])
[…]

### 2024-01-15 src/plot_toy_model.py

@@ -1,188 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Apr 25 15:14:44 2022
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-import pickle
-import matplotlib.pyplot as plt
-import seaborn as sns
-import pandas as pd
-
-from matplotlib.colors import LinearSegmentedColormap
-import matplotlib.ticker as ticker
-from mpl_toolkits.axes_grid1.inset_locator import inset_axes
-
-dtype = np.float32
-
-# %% set colors
-
-Col_Rate_E = '#9E3039'
-Col_Rate_nE = '#955F89'
-Col_Rate_nD = '#BF9BB8' 
-Col_Rate_pE = '#CB9173' 
-Col_Rate_pD = '#DEB8A6' 
-Col_Rate_PVv = '#508FCE' 
-Col_Rate_PVm = '#2B6299' 
-Col_Rate_SOM = '#79AFB9' 
-Col_Rate_VIP = '#39656D' 
-
-color_sensory = '#D76A03'
-color_prediction = '#19535F'
-
-cmap_sensory_prediction = LinearSegmentedColormap.from_list(name='cmap_sensory_prediction', 
-                                                            colors=['#19535F','#fefee3','#D76A03'])
-
-# %%
-
-
-def plot_manipulation_results(manipulations, fraction_of_sensory_input_in_output, flg_limit_case):
-
-    plt.figure()
-    ax = plt.subplot(1,1,1)    
-
-    label_text = ['nE', 'pE', 'nD', 'pD', 'PVv', 'PVm', 'SOM', 'VIP']
-    colors_cells = [Col_Rate_nE, Col_Rate_pE, Col_Rate_nD, Col_Rate_pD, Col_Rate_PVv, Col_Rate_PVm, Col_Rate_SOM, Col_Rate_VIP]
-    linestyles = ['-', '-', '--', '--', '-', '-', '-', '-']
-
-    for  id_cell in range(8):
-        ax.plot(manipulations, fraction_of_sensory_input_in_output[:,id_cell], label=label_text[id_cell], 
-                color=colors_cells[id_cell], ls = linestyles[id_cell])
-    
-    ax.legend(loc=0)
-    ax.set_ylabel('Fraction sensory')
-    ax.set_xlabel('manipulation strength')
-    if flg_limit_case==0:
[…]

### 2024-01-15 src/toy_model.py

@@ -1,217 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Apr 25 14:51:18 2022
-
-@author: loreen.hertaeg
-"""
-
-# %% Import
-
-import numpy as np
-
-# %%
-
-def alpha_parameter_exploration(stimuli, stimulus_duration, last_n, para_tested_first, para_tested_second, 
-                                tc_var_per_stim, eta_prediction, eta_mean_prediction, tc_var_pred, flg_para):
-    
-    ### number of parameters and stimuli tested
-    num_para_first = len(para_tested_first)
-    num_para_second = len(para_tested_second)
-    n_stimuli = np.int32(len(stimuli)/stimulus_duration)
-    
-    ### initialise
-    fraction_sensory_median = np.zeros((num_para_first, num_para_second))
-    
-    ### run networks over all parameter to be tested
-    if ((flg_para==12) or (flg_para==124)):
-        
-        for row, para_first in enumerate(para_tested_first):
-            for col, para_second in enumerate(para_tested_second):
-                
-                _, _, _, _, alpha, _, _ = run_toy_model(para_first, para_second, eta_mean_prediction, tc_var_pred, stimuli)
-                fraction_sensory_median[row, col] = np.median(alpha[(n_stimuli - last_n) * stimulus_duration:])
-    
-    elif flg_para==13:
-        
-        for row, para_first in enumerate(para_tested_first):
-            for col, para_second in enumerate(para_tested_second):
-                
-                _, _, _, _, alpha, _, _ = run_toy_model(para_first, eta_prediction, para_second, tc_var_pred, stimuli)
-                fraction_sensory_median[row, col] = np.median(alpha[(n_stimuli - last_n) * stimulus_duration:])
-                
-    elif flg_para==14:
-        
-        for row, para_first in enumerate(para_tested_first):
-            for col, para_second in enumerate(para_tested_second):
-                
-                _, _, _, _, alpha, _, _ = run_toy_model(para_first, eta_prediction, eta_mean_prediction, para_second, stimuli)
-                fraction_sensory_median[row, col] = np.median(alpha[(n_stimuli - last_n) * stimulus_duration:])
-                
-    elif flg_para==23:
-        
[…]
